### 2016-03-01 draft_working.tex
added 2 figures

%% LyX 2.1.4 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[10pt,english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{geometry}
\geometry{verbose,tmargin=2cm,bmargin=2cm,lmargin=2cm,rmargin=2cm}
\usepackage{color}
\usepackage{babel}
\usepackage{float}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{esint}
\usepackage[unicode=true,
 bookmarks=true,bookmarksnumbered=true,bookmarksopen=false,
 breaklinks=false,pdfborder={0 0 0},backref=false,colorlinks=true]
 {hyperref}
\usepackage{breakurl}

\makeatletter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\renewcommand\[{\begin{equation}}
\renewcommand\]{\end{equation}}

\makeatother

\begin{document}

\title{Adding Sensitivity to 21cm Inteferometric Probes of Reionization}

\maketitle

\section{Introduction}

The epoch of reionization represents the last key stage of the universe's
early evolution. Specifically, study of this event marks the intersection
of early cosmology and astrophysics. Understanding this event is important
not only as a scientific goal of its own, but also potentially provides
crucial information to fundamental nature of inflation, neutrino mass
and phenonmenology of the first stars and galaxies alike.

Arguably the most promising observational probe of the epoch of reionization
comes from measurement of the ``spin-flip'' transition of neutral
hydrogen of characteristic wavelength 21cm. While probes relying on
emission and scattering of free electrons are limited to the lower
redshifts after the completion of reionization, the 21cm line directly
probes the abundance and distribution of neutral hydrogen, and thus
would potentially shed light on the entire cosmic dawn. Thus radio
interferometry techniques to measure the 21cm power spectrum has been
a top priority in astronomy. Cuurrent generation instruments include
the Precision Array for Probing the Epoch of Reionization (PAPER),
Murchison Widefield Array (MWA), Low Frequency Array (LOFAR), and
next generation instruments such as the Hydrogen Epoch of Reionization
Array (HERA) and the Square Kilometer Array Low (SKA low) are under
construction.

One of the main challenges of interferometric observations of cosmic
21cm line is the foreground contaminations. Both sources within our
galaxy, and to a lesser extent sources outside of our galaxy emit
radio contaminations up to five orders of magnitude higher than the
reionization signal. There are two common methods to deal with the
foreground contamination. In the first, individual sources are identified
and removed, sometimes called the foreground removal technique. The
other technique, commonly refered to as foreground avoidance, makes
use of the fact that most foreground contamination is a smooth spectrum,
and thus is contrained to a Fourier domain ``foreground wedge''.
The main challenge of using the latter technique is the sensitivity,
or the signal to noise ratio. This is the motivation for the design
of the maximum redundancy arrays such as MWA, PAPER, and HERA. While
traditional imaging arrays such as the Very Large Array (VLA) focus
on Fourier mode uv coverage, the redundant arrays have many tightly
packed antennas at equal spacing. Since baselines of the same length
and orientation measure the same Fourier modes on the sky, the maximum
redundancy arrays is able to increase the signal to noise ratio of
the Fourier mode in question. Ali2015 provided the newest upper limit
to the power spectrum measurements with the PAPER-64 data. Some models
expect that the sensitivity required for detection is a factor of
2 away.

The earth's rotation causes the baseline to sweep through different
areas of the sky during the day. This effect is used extensively to
improve uv coverage in imaging with minimum redundancy arrays. In
a redundant array, baselines that are slightly different rotate into
each other at a time delay. Here we present a method to improve the
sensitivity of the maximum redundancy arrays by considering the uv
redundancy of different baselines at a time lag. In section 2 we present
the basic search algorithm for this time delay. In section 3 we present
a necessary correction and comparison with simulated data.

\section{Time delayed redundancy}

Baselines of the same length and orientation are traditionally called
``redundant baselines'', because they measure the same Fourier mode
in the sky. In order to eliminate confusion, we shall introduce slightly
different terminology. We'll call baselines that are the same length
and orientation ``equivalent baselines''. Two equivalent baselines
will be redundant with each other simultaneously. Some non-equivalent
baselines can also be redundant if their respected time series is
shifted with respect to one another by a certain delay. We call the
latter ``near equivalent baselines''. It is our goal to a) identify
the near equivalent baselines that give good delayed redundancy, b)
to find this delay for a given pair of baselines, and c) to quantify
the sensitivity improvement associated with cross multiplying such
a pair of nonequivalent baselines.

\subsection{Cross multiplication of near redundant data}

To begin, we first derive theoretical expectations of cross correlation
of two near-equivalent baselines. Given a point source on the sky,
each baseline measures a track of the source in uv plane. We identify
redundancy of near equivalent baselines for that point source as crossings
of the tracks. Since processed data are measured at discrete time
intervals (in lst-binned PAPER-64 data it's 42.9 seconds ), we do
not have exact overlap of two data points in uv space. We thus need
to find a way to cross multiply data points that are separated by
a short distance in uv space and recover the power spectrum.

We take the visibility as 
\[
\begin{aligned}V_{\nu}(b) & =\int d\Omega A(\hat{r})I(\hat{r})\exp\left[-2\pi i\frac{\nu}{c}b\cdot\hat{r}\right],\\
 & \approx\frac{2k_{B}}{\lambda^{2}}\int d\Omega A(\hat{r})T(\hat{r})\exp\left[-2\pi i\frac{\nu}{c}b\cdot\hat{r}\right],\\
 & =\frac{2k_{B}}{\lambda^{2}}\int\frac{d^{3}r}{X^{2}Y}A(r)T(r)e^{-ikr}.
\end{aligned}
\]
Here $\lambda$ is a mean wavelength, $\boldsymbol{b}$ is the baseline
length, $\hat{\boldsymbol{r}}$ is a direction in the sky, $A$ is
the beam, and $I$ is the intensity, which has been related to $T$,
the brightness temperature. In the last line $(r_{x},r_{y},r_{z})=(Xl,Xm,Y\nu)$,
and $(Xk_{x},Xk_{y},Yk_{z})=2\pi(u,v,\eta)$.

We define the delay transformed visibility:
\[
\begin{aligned}V(b) & =\int d\nu V_{\nu}(b)\exp\left[-2\pi i\nu\eta\right],\\
 & =\frac{2k_{B}}{\lambda^{2}}\int d\Omega d\nu A(\hat{r},\nu)T(\hat{r},\nu)\exp\left[-2\pi i\left(\frac{\nu}{c}b\cdot\hat{r}+\nu\eta\right)\right]
\end{aligned}
\]

\textbackslash{}The beam reception pattern $A$ is dimensionless,
normalized to 1 at its peak (zenith), and we assume it to be the same
for all baselines. We introduce the Fourier temperature and window
function

\[
\begin{aligned}\tilde{T}(k) & =\int\frac{d^{3}r}{X^{2}Y}T(r)e^{-ikr},\\
W(k) & =\int d^{3}rA(r)e^{-ikr}.
\end{aligned}
\]
By the convolution theorem, we have
\[
\begin{aligned}\tilde{V}(k) & \approx\frac{2k_{B}}{\lambda^{2}}\int\frac{d^{3}r}{X^{2}Y}A(r)T(r)e^{-ikr},\\
 & =\frac{2k_{B}}{\lambda^{2}}\frac{1}{(2\pi)^{3}}\int d^{3}k_{1}W(k_{1})\tilde{T}(k+k_{1}).
\end{aligned}
\]
With implicit bounds of integrals from $-\infty$ to $\infty$, we
have
\begin{equation}
\begin{aligned}\langle\tilde{V}^{*}(k-k_{r})\tilde{V}(k)\rangle & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}\frac{1}{(2\pi)^{6}}\int dk_{1}dk_{2}W^{*}(k_{1})W(k_{2})\langle T^{*}(k+k_{1}-k_{r})T(k+k_{2})\rangle,\\
 & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}\frac{1}{(2\pi)^{6}}\int dk_{1}dk_{2}W^{*}(k_{1})W(k_{2})\delta_{D}(k_{2}-k_{1}-k_{r})\hat{P_{21}}(k+k_{2}),\\
 & \approx\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}\frac{\hat{P_{21}}(k)}{(2\pi)^{6}}\int dk_{1}W^{*}(k_{1})W(k_{1}+k_{r}),\\
 & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}\frac{\hat{P_{21}}(k)}{(2\pi)^{3}}\int d^{3}r|A(r)|^{2}e^{-ik_{r}r},\\
 & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}\frac{\hat{P_{21}}(k)}{(2\pi)^{3}}\int d^{3}r|a(r)\phi(r_{z})|^{2}e^{-ik_{r}r}
\end{aligned}
\label{eq:(7)}
\end{equation}
where the first to second line follows from assumption of Gaussian
random sky, and the second to third line follows from the assumption
that the 3D power spectrum varies neglegibly over the region of interest
so that $\hat{P_{21}}(k+k_{2})\approx\hat{P_{21}}(k)$. In the last
line we have separated the 2D spatial beam patter from the frequency
bandpass $\phi$. So in conclusion the cross correlation of visibilities
seperated by $\Delta u$ apart in uv space is proportional to the
power spectrum times the Fourier transform of the beam pattern squared.

\begin{figure}[H]
\includegraphics[scale=0.5]{64layout}

\caption{The PAPER 64 layout. Top panel shows the antenna positions drawn to
scale, bottom panel show the antenna labels and distances. }
\end{figure}

We shall use the 64-antenna PAPER array to demonstrate. The PAPER
64 array is located in the Karoo desert in South Africa (30:43:17.5
S, 21:25:41.8 E). The layout pattern with antenna labels are show
in Fig. 1. I show denote a baseline with the antconfigWe see the antenna
spacing in North-South directions are comparatively close (4m), so
that baselines such as 0\_44 and 0\_7 are very close to equivalent.

\section{Implementation}

An example of rotation tracks is shown in the uv plane below

\begin{figure}[H]
\includegraphics[scale=0.5]{tracks}

\caption{The PAPER 64 layout. Top panel shows the antenna positions drawn to
scale, bottom panel show the antenna labels and distances. }
\end{figure}

The beam squared and its Fourier transform are given as

\begin{figure}[H]
\includegraphics[scale=0.4]{64beam}

\caption{The PAPER 64 layout. Top panel shows the antenna positions drawn to
scale, bottom panel show the antenna labels and distances. }
\end{figure}

\end{document}

### 2016-06-15 Rotation paper/RotationDraft Kopie.tex
updated tex

--- 2016-03-01 draft_working.tex
@@
 \[
-\begin{aligned}V(b) & =\int d\nu V_{\nu}(b)\exp\left[-2\pi i\nu\eta\right],\\
- & =\frac{2k_{B}}{\lambda^{2}}\int d\Omega d\nu A(\hat{r},\nu)T(\hat{r},\nu)\exp\left[-2\pi i\left(\frac{\nu}{c}b\cdot\hat{r}+\nu\eta\right)\right]
-\end{aligned}
+\begin{aligned}\end{aligned}
 \]
 
-\textbackslash{}The beam reception pattern $A$ is dimensionless,
-normalized to 1 at its peak (zenith), and we assume it to be the same
-for all baselines. We introduce the Fourier temperature and window
-function
+The beam reception pattern $A$ is dimensionless, normalized to 1
+at its peak (zenith), and we assume it to be the same for all baselines.
+We introduce the Fourier temperature and window function
 
@@
 \begin{figure}[H]
-\includegraphics[scale=0.5]{64layout}
+\includegraphics[scale=0.5]{figure_1}
 
@@
 that baselines such as 0\_44 and 0\_7 are very close to equivalent. 
-
-\section{Implementation}
-
-An example of rotation tracks is shown in the uv plane below
-
-\begin{figure}[H]
-\includegraphics[scale=0.5]{tracks}
-
-\caption{The PAPER 64 layout. Top panel shows the antenna positions drawn to
-scale, bottom panel show the antenna labels and distances. }
-\end{figure}
-
-The beam squared and its Fourier transform are given as
-
-\begin{figure}[H]
-\includegraphics[scale=0.4]{64beam}
-
-\caption{The PAPER 64 layout. Top panel shows the antenna positions drawn to
-scale, bottom panel show the antenna labels and distances. }
-\end{figure}
-
 \end{document}

### 2016-06-15 Rotation paper/draft_working.tex
updated tex

--- 2016-03-01 draft_working.tex
@@
 \usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{graphicx}
@@
  {hyperref}
-\usepackage{breakurl}
 
@@
 
-\textbackslash{}The beam reception pattern $A$ is dimensionless,
-normalized to 1 at its peak (zenith), and we assume it to be the same
-for all baselines. We introduce the Fourier temperature and window
-function
+We'd like to relate this to the 3 dimensional power spectrum. We start
+by noting that 
+\[
+\begin{aligned}r & =\frac{c}{H_{0}}\int_{0}^{z}\frac{dz'}{E(z')}-\frac{c(1+z)^{2}}{\nu_{21}H_{0}E(z)}\left(\nu-\nu_{21}\right),\\
+ & \equiv D_{c}-Y\Delta\nu,
+\end{aligned}
+\]
+where $\nu_{21}=1420MHz$ is the 21cm transition rest frequency, and
+\[
+E(z)=\sqrt{\Omega_{m}(1+z)^{3}+\Omega_{\Lambda}}.
+\]
+So
+\[
+\nu=\frac{D_{c}-r}{Y}+\nu_{21}.
+\]
+Thus $d\nu=-dr/Y$ and we can rewrite the delayed transformed visibility
+as 
+\[
+\begin{aligned}V(b) & =\frac{1}{Y}\int_{H}d^{3}rA(r)I(r)\exp\left[2\pi i\frac{r}{Y}\left(\frac{b\cdot\hat{r}}{c}+\eta\right)\right]\exp\left[-2\pi i\left(\frac{b}{c}\cdot\hat{r}+\eta\right)\left(\frac{D_{c}}{Y}+\nu_{21}\right)\right]\end{aligned}
+.
+\]
+$A$ is a patch of shell
+
+The beam reception pattern $A$ is dimensionless, normalized to 1
+at its peak (zenith), and we assume it to be the same for all baselines.
+We introduce the Fourier temperature and window function
 
@@
 \]
+Here 
+\[
+k=2\pi\frac{b}{cY}.
+\]
 By the convolution theorem, we have
 \[
-\begin{aligned}\tilde{V}(k) & \approx\frac{2k_{B}}{\lambda^{2}}\int\frac{d^{3}r}{X^{2}Y}A(r)T(r)e^{-ikr},\\
+\begin{aligned}\tilde{V}(k) & =\frac{2k_{B}}{\lambda^{2}}\int\frac{d^{3}r}{Y}A(r)T(r)e^{-ikr},\\
  & =\frac{2k_{B}}{\lambda^{2}}\frac{1}{(2\pi)^{3}}\int d^{3}k_{1}W(k_{1})\tilde{T}(k+k_{1}).
@@
 \]
+With a time offset, the sky has moved from coordinates $r$ to $\Gamma r$.
 With implicit bounds of integrals from $-\infty$ to $\infty$, we
 have
-\begin{equation}
-\begin{aligned}\langle\tilde{V}^{*}(k-k_{r})\tilde{V}(k)\rangle & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}\frac{1}{(2\pi)^{6}}\int dk_{1}dk_{2}W^{*}(k_{1})W(k_{2})\langle T^{*}(k+k_{1}-k_{r})T(k+k_{2})\rangle,\\
- & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}\frac{1}{(2\pi)^{6}}\int dk_{1}dk_{2}W^{*}(k_{1})W(k_{2})\delta_{D}(k_{2}-k_{1}-k_{r})\hat{P_{21}}(k+k_{2}),\\
- & \approx\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}\frac{\hat{P_{21}}(k)}{(2\pi)^{6}}\int dk_{1}W^{*}(k_{1})W(k_{1}+k_{r}),\\
- & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}\frac{\hat{P_{21}}(k)}{(2\pi)^{3}}\int d^{3}r|A(r)|^{2}e^{-ik_{r}r},\\
- & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}\frac{\hat{P_{21}}(k)}{(2\pi)^{3}}\int d^{3}r|a(r)\phi(r_{z})|^{2}e^{-ik_{r}r}
+
+\[
+\begin{aligned}\langle V^{*}(k)V(k')\rangle & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\langle T^{*}(r)T(\Gamma r')\rangle\right)A^{*}(r)A(r')\Phi(b,\eta),\\
+ & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\kappa)T(\kappa')\rangle e^{-i(\kappa r-\kappa'\Gamma r')}\right)A^{*}(r)A(r')\Phi(b,\eta),\\
+ & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\kappa(r-\Gamma r')}\right)A^{*}(r)A(r')\Phi(b,\eta),\\
+ & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\xi(r-\Gamma r')A^{*}(r)A(r')\Phi(b,\eta),\\
+ & \approx\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k)\int d^{3}rd^{3}r'\delta_{D}^{(3)}(r-\Gamma r')A^{*}(r)A(r')\exp\left[-i\frac{2\pi}{c}\left(b\cdot\nu\hat{r}-b'\cdot\nu'\hat{r'}\right)\right]\exp\left[-i2\pi\eta\left(\nu\hat{r}-\nu'\hat{r'}\right)\right],\\
+ & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k)\int d^{3}r|A^{*}(r)A(\Gamma r)|\exp\left[-i2\pi\nu_{r}\left\{ \hat{r}\cdot(\frac{b}{c}+\eta c\hat{r})-\hat{\Gamma r}\cdot(\frac{b'}{c}+\eta'\hat{\Gamma r})\right\} \right],
 \end{aligned}
-\label{eq:(7)}
-\end{equation}
-where the first to second line follows from assumption of Gaussian
-random sky, and the second to third line follows from the assumption
-that the 3D power spectrum varies neglegibly over the region of interest
+\]
+where the phase term is 
+\[
+\Phi(b,\eta)=\text{\ensuremath{\exp\left[-i\frac{2\pi}{c}\left(b\cdot\nu\hat{r}-b'\cdot\nu'\hat{r'}\right)\right]\exp\left[-i2\pi\eta\left(\nu\hat{r}-\nu'\hat{r'}\right)\right]},}
+\]
+The first to second line follows from assumption of Gaussian random
+sky, and the second to third line follows from the assumption that
+the 3D power spectrum varies neglegibly over the region of interest
 so that $\hat{P_{21}}(k+k_{2})\approx\hat{P_{21}}(k)$. In the last
-line we have separated the 2D spatial beam patter from the frequency
-bandpass $\phi$. So in conclusion the cross correlation of visibilities
-seperated by $\Delta u$ apart in uv space is proportional to the
-power spectrum times the Fourier transform of the beam pattern squared.
+line we have written $\nu_{r}$ to remind us that $\nu$ and $r$
+are related by Eq. 4. Since $\Gamma$ is a sky rotation, it doesn't
+affect $\nu$, hence we have taken $\nu_{r}$ outside the parenthesis.
+Notice also that we have omitted a phase factor $ $ because it would
+drop out in the last step.
+
+Now plugging in result from Eq. 4, we see
+
+\[
+\boxed{\langle V^{*}(k)V(k')\rangle=P(k)\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\Psi\int d^{3}r|A^{*}(r)A(\Gamma r)|\exp\left[i\frac{2\pi}{cY}\vec{r}\cdot\left(b-b'\right)\right],}
+\]
+where $\Psi=\exp\left[-2\pi i\frac{b-b'}{c}\cdot\hat{r}\left(\frac{D_{c}}{Y}+\nu_{21}\right)\right]$
+is a phase that's fixed.  So in conclusion the cross correlation of
+visibilities seperated by $\Delta u$ apart in uv space is proportional
+to the power spectrum times the Fourier transform of the cross multiplied
+beam pattern. One can then in principle combine information from different
+baseline pairs if one corrects for the phase and normalization.
 
@@
 
-\section{Implementation}
+\section{Technique}
 
-An example of rotation tracks is shown in the uv plane below
+For a given point source in the sky, each baseline at a particular
+instant corresponds to a point in UV space. Thus as the earth rotates,
+the point source traces out a track in UV plane for a given baseline.
+As an example, we show the rotation tracks for a source that passes
+through zenith in Fig. \ref{fig:Tracks-of-PAPER64} below for the
+baselines of PAPER 64:
 
@@
 
-\caption{The PAPER 64 layout. Top panel shows the antenna positions drawn to
-scale, bottom panel show the antenna labels and distances. }
+\caption{Tracks of PAPER64 for a hypothetical source that passes through zenith.
+Color scale simply represents different tracks. Spacing between dots
+are characteristic of time resolution of compressed data. As the earth
+rotates, tracks are traced out counterclockwise. \label{fig:Tracks-of-PAPER64}}
 \end{figure}
@@
 
+In practice, we first implement a search algorithm that grids the
+uv plane and search for intersections, then confirm the time delay
+by plotting a) cross correlated visibilities from simulated random
+sky, and b) theoretical proportion factor $\int d^{3}r|A^{*}(r)A(\Gamma r)|\exp\left[-i\frac{2\pi}{c}\nu\hat{r}\cdot\left(b-b'\right)\right]$
+from Eq. 10.
+
+\subsection{Grid search algorithm}
+
+The gridding algorithm sets up grids on the UV-plane, creates a point
+source that passes through zenith, and searches for intersections
+of the UV tracks. The search compares each track to each other track
+to locate the closest approaches. To optimize performance, the algorithm
+first sets up coarse grids, locate the cells that contain data points
+from more than one tracks, then fine grains the cell to determine
+the point of closest approach. Fig {[}insert{]} illustrates this procedure.
+This algorithm is mainly used to determine which baselines to cross
+correlate, and give a rough estimate of the time offset. We indeed
+find that baselines 0\_44 and 0\_7 give the closest approach wiht
+a time offset of roughly 0.03 sidereal days. In reality, a wide beam
+(e.g. Fig. 2) samples a patch of sky, not just a point source. And
+thus the real UV representation of a baseline is the track convolved
+with an appropriate point spread function. Accounting for these factors
+accurately in this context is difficult. Thus for accurate determination
+of the time offset, we resort to two other methods. It is also worth
+pointing out that the near-redundant pairs is not only useful at a
+particular time corresponding to the crossing. The search algorithm
+uses a fiducial point source in the sky to find the offset, and all
+time-stamps separated by this offset corresponds to UV-track crossing
+of a point in the sky, and thus all data of the two baselines can
+be utilized.
+
+\subsection{Gaussian sky simulation}
+
+Having found the pairs of baselines to cross-correlate, and a rough
+estimate of the time offset, we make a more accurate determination
+of the time offset. Here we take N random realizations of the sky
+on a healpix map\footnote{We use functionalities in the python package AIPY for healpix mapping
+as well as coordinate transforms. }. For each realization, each pixel is given a Gaussian random value
+of brightness temperature. We then rotate the baseline positions by
+a simple rotation matrix, recording the visibilities, for each baseline,
+as given in Eq. 1 \footnote{As a caveat, there are two obvious ways to achieve the rotation. One
+can either fix the sky and rotate the baselines, or the other way
+around. We found however, that we must not physically rotate the sky
+map, for the numerical roundoffs due to finite resolutions of the
+map turns out to be significant. Thus we let the sky, represented
+by the healpix map, be fixed, and rotate the baselines. }. The resulting visibilities for the two baselines are then convolved
+via the Fourier convolution theorem, to obtain values of the cross
+correlation as a function of time-offset. The peak of the curve then
+corresponds to maximum redundancy. The accuracy of this result is
+limited by (simulated) cosmic variance and finite spacial resolution,
+and hence can be beat down by averaging over a large number of universes.
+A numerical estimate of the error of the peak height with $N=1$ is
+$\lesssim20\%$, and thus with $N=2500,$ we achieve an error of peak
+height $<0.5\%$. An example is shown in Fig. {[}insert{]}
+
+\subsection{Theoretical determination}
+
+To test our understanding of the time-offset, and to verify our derivation
+in Eq. 10 makes sense, we must check the curve obtained in Fig. against
+one from the multiplicative factor in Eq. 10. In particular, we'd
+really like to compare the peak location as well as height. For the
+latter, the most useful information comes in its comparison with that
+of an equivalent baseline. Thus in Fig. {[}insert{]} below we show
+on the left comparison of the convolution of an equivalent baseline,
+normalized to 1 at the peak, whose location is at 0 offset as expected.
+On the right we show comparison of the baseline pair found in the
+grid search. The peaks are normalized with the same factor as the
+peak height on the left, i.e. the two plots have thus the same scale
+on the vertical axis. The height of the plot on the right thus quantifies
+the added sensitivity.
+
+\begin{figure}[H]
+\includegraphics[scale=0.25]{redun_agree}
+
+\includegraphics[scale=0.3]{simucorroff}
+
+\caption{The PAPER 64 layout. Top panel shows the antenna positions drawn to
+scale, bottom panel show the antenna labels and distances. }
+\end{figure}
+
+\section{Application to PAPER 64}
+
+\subsection{Expected Sensitivity contribution}
+
+The latest release of PAPER 64 data uses baselines sep-1,1, sep0,1
+and sep1,1 {[}cite Ali2015{]}, and achieved a $2\sigma$ upper limit
+of $22.4mK^{2}$. There, the three sets of equivalent baselines are
+only cross multiplied by itself. Assuming that each baseline delivers
+the same quality of data, the relative contribution to sensitivity
+can be estimated 
 \end{document}

### 2016-07-01 Rotation paper/draft_working.tex
draft saved

--- 2016-03-01 draft_working.tex
@@
-%% LyX 2.1.4 created this file.  For more info, see http://www.lyx.org/.
-%% Do not edit unless you really know what you are doing.
-\documentclass[10pt,english]{article}
-\usepackage[T1]{fontenc}
-\usepackage[latin9]{inputenc}
-\usepackage{geometry}
-\geometry{verbose,tmargin=2cm,bmargin=2cm,lmargin=2cm,rmargin=2cm}
-\usepackage{color}
-\usepackage{babel}
+\documentclass[preprint2,numberedappendix,tighten,twocolappendix]{aastex6}  % USE THIS TO MAKE BIB, THEN FORMAT USING EMULATEAPJ
+%\documentclass[twocolumn,apj,numberedappendix]{emulateapj}
+\shorttitle{Adding Sensitivity to 21cm Inteferometric Probes of Reionization}
+\shortauthors{Zhang\&Parsons}
 \usepackage{float}
@@
 \usepackage{graphicx}
-\usepackage{esint}
-\usepackage[unicode=true,
- bookmarks=true,bookmarksnumbered=true,bookmarksopen=false,
- breaklinks=false,pdfborder={0 0 0},backref=false,colorlinks=true]
- {hyperref}
-\usepackage{breakurl}
-
-\makeatletter
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
+\usepackage[figuresright]{rotating}
+%\usepackage{rotating}
+\usepackage{natbib}
+%\usepackage{pdflscape}
+%\usepackage{lscape}
+\usepackage{ctable}
+\citestyle{aa}
 \renewcommand\[{\begin{equation}}
@@
 
-\makeatother
-
 \begin{document}
 
-\title{Adding Sensitivity to 21cm Inteferometric Probes of Reionization}
+\title{Adding Sensitivity to 21cm Inteferometric Probes of Reionization by Optimizing Choice of Baselines}
 
-\maketitle
+%\maketitle
+\author{
+Yunfan Gerry Zhang \altaffilmark{1},
+Aaron R. Parsons\altaffilmark{1}\altaffilmark{2}
+}
+
+\altaffiltext{1}{Astronomy Dept., U. California, Berkeley, CA}
+\altaffiltext{2}{Radio Astronomy Lab., U. California, Berkeley, CA}
+%\altaffiltext{3}{Berkeley Center for Cosmological Physics, Berkeley, CA}
+%\altaffiltext{3}{Dept. of Physics and Astronomy, U. Pennsylvania, Philadelphia, PA}
+%\altaffiltext{8}{School of Earth and Space Exploration, Arizona State U., Tempe, AZ}
+
+\begin{abstract}
+We introduce an accurate method to extract additional sensitivity from data of redundant radio arrays. 
+Using the 64 element version of Precision Array for Probing the Epoch of Reionization (PAPER-64), we illustrate 
+how our method finds optimal baselines to cross correlate, including baselines that are slightly different in length and orientation. 
+We predict the sensitivity improvements of including each baseline pairs. The extremely natural and simple implementation of this method
+would increase the sensitivity of PAPER-64 by up to [~50] percent. 
+\end{abstract}
 
@@
 
-The epoch of reionization represents the last key stage of the universe's
-early evolution. Specifically, study of this event marks the intersection
-of early cosmology and astrophysics. Understanding this event is important
-not only as a scientific goal of its own, but also potentially provides
-crucial information to fundamental nature of inflation, neutrino mass
-and phenonmenology of the first stars and galaxies alike.
+The epoch of reionization, or cosmic dawn represents the last key
+stage of the universe's early evolution. Study of this event stands at
+the intersection of cosmology and astrophysics. Understanding this
+event is important not only as a scientific goal
+of its own, but also because it potentially provides crucial information
+to fundamental nature of inflation, neutrino mass and phenomenology
+of the first stars and galaxies alike (e.g. \cite{LiuOpticalDepth}).
 
@@
 comes from measurement of the ``spin-flip'' transition of neutral
-hydrogen of characteristic wavelength 21cm. While probes relying on
-emission and scattering of free electrons are limited to the lower
-redshifts after the completion of reionization, the 21cm line directly
-probes the abundance and distribution of neutral hydrogen, and thus
-would potentially shed light on the entire cosmic dawn. Thus radio
-interferometry techniques to measure the 21cm power spectrum has been
-a top priority in astronomy. Cuurrent generation instruments include
-the Precision Array for Probing the Epoch of Reionization (PAPER),
-Murchison Widefield Array (MWA), Low Frequency Array (LOFAR), and
-next generation instruments such as the Hydrogen Epoch of Reionization
-Array (HERA) and the Square Kilometer Array Low (SKA low) are under
-construction.
+hydrogen of characteristic wavelength 21cm \cite{Furlanetto2006181,PritchardLoeb}.
+While probes such as quasar spectra rely on emission and scattering
+of free electrons, and are thus limited to the lower redshifts after
+the completion of reionization, the 21cm line directly probes the
+abundance and distribution of neutral hydrogen, and thus would potentially
+shed light on all stages of reionization. Radio interferometry efforts
+to measure the 21cm power spectrum has been a top priority in recent years of astronomy.
+Current generation instruments include the Precision Array for Probing
+the Epoch of Reionization (PAPER) \cite{Ali2015,paper32}, Murchison
+Widefield Array (MWA) \cite{MWA}, Low Frequency Array (LOFAR) \cite{LOFAR},
+and next generation instruments such as the Hydrogen Epoch of Reionization
+Array (HERA) (e.g. \cite{HERA,HERAconfiguration,HERABEAM1,HERADISH2})
+and the Square Kilometer Array Low (SKA low) (e.g. \cite{SKA1}) are
+under construction.
 
 One of the main challenges of interferometric observations of cosmic
-21cm line is the foreground contaminations. Both sources within our
-galaxy, and to a lesser extent sources outside of our galaxy emit
-radio contaminations up to five orders of magnitude higher than the
+21cm line is the foreground contamination. Both sources within our
+galaxy, and to a lesser distant sources outside of our galaxy emit
+radio contaminations (via for example synchrotron processes) up to five orders of magnitude stronger than the
 reionization signal. There are two common methods to deal with the
-foreground contamination. In the first, individual sources are identified
-and removed, sometimes called the foreground removal technique. The
-other technique, commonly refered to as foreground avoidance, makes
-use of the fact that most foreground contamination is a smooth spectrum,
-and thus is contrained to a Fourier domain ``foreground wedge''.
-The main challenge of using the latter technique is the sensitivity,
-or the signal to noise ratio. This is the motivation for the design
-of the maximum redundancy arrays such as MWA, PAPER, and HERA. While
-traditional imaging arrays such as the Very Large Array (VLA) focus
-on Fourier mode uv coverage, the redundant arrays have many tightly
-packed antennas at equal spacing. Since baselines of the same length
-and orientation measure the same Fourier modes on the sky, the maximum
-redundancy arrays is able to increase the signal to noise ratio of
-the Fourier mode in question. Ali2015 provided the newest upper limit
-to the power spectrum measurements with the PAPER-64 data. Some models
-expect that the sensitivity required for detection is a factor of
-2 away.
+foreground contamination. In the first, sometimes called the foreground
+removal technique, individual sources are identified and removed in the image domain.
+The other technique, commonly referred to as 'foreground avoidance',
+makes use of the fact that most common foreground contaminations have a smooth
+spectrum, and thus is constrained to a Fourier domain ``foreground
+wedge'' \cite{wedge1,wedge2}. Thus contamination can be avoided
+by restricting our observation to ``outside the wedge'', where the data is currently noise dominated. The main
+challenge of using the avoidance technique is thus the sensitivity, or
+the signal to noise ratio. This is the motivation for the design of
+the maximum redundancy arrays such as PAPER and HERA. While traditional
+imaging arrays focus on Fourier mode uv coverage, the redundant arrays
+have many tightly packed antennas at equal spacing. Since baselines
+of the same length and orientation measure the same Fourier modes
+on the sky, the maximum redundancy arrays is able to increase the
+signal to noise ratio by averaging over measurements of the same baseline.
+\cite{Ali2015} provides the newest upper limit to the power spectrum
+measurements with the PAPER-64 data. From some models we roughly expect
+that the sensitivity required for detection is a factor of 10 away.
+This paper explores a new technique to add sensitivity to the analysis
+of redundant arrays, by making use of the earth's rotation. This technique
+can be applied to redundant arrays such as PAPER and HERA to increase
+the sensitivity by up to a factor of [1.5].
 
-The earth's rotation causes the baseline to sweep through different
-areas of the sky during the day. This effect is used extensively to
-improve uv coverage in imaging with minimum redundancy arrays. In
-a redundant array, baselines that are slightly different rotate into
-each other at a time delay. Here we present a method to improve the
-sensitivity of the maximum redundancy arrays by considering the uv
-redundancy of different baselines at a time lag. In section 2 we present
-the basic search algorithm for this time delay. In section 3 we present
-a necessary correction and comparison with simulated data.
-
-\section{Time delayed redundancy}
+The earth's rotation causes the baselines to pick up different modes of the sky with time. 
+This effect is used extensively to improve uv coverage in imaging with minimum redundancy arrays. In
+a maximally redundant array, baselines that are slightly different
+rotate into each other at a time delay. Here we present a method to
+improve the sensitivity of the maximum redundancy arrays by considering
+the uv redundancy of different baselines at a time lag.
 
@@
 ``redundant baselines'', because they measure the same Fourier mode
-in the sky. In order to eliminate confusion, we shall introduce slightly
-different terminology. We'll call baselines that are the same length
-and orientation ``equivalent baselines''. Two equivalent baselines
-will be redundant with each other simultaneously. Some non-equivalent
-baselines can also be redundant if their respected time series is
-shifted with respect to one another by a certain delay. We call the
-latter ``near equivalent baselines''. It is our goal to a) identify
-the near equivalent baselines that give good delayed redundancy, b)
-to find this delay for a given pair of baselines, and c) to quantify
-the sensitivity improvement associated with cross multiplying such
-a pair of nonequivalent baselines.
+in the sky. In order to eliminate confusion and ambiguity, we shall
+introduce slightly different terminology. We'll call baselines that
+are the same length and orientation ``equivalent baselines'', inspired
+by the mathematical notion of equivalency classes. Two equivalent
+baselines will be redundant with each other simultaneously at all
+times. Some non-equivalent baselines can also be redundant if their
+respected time series is shifted with respect to one another by a
+certain delay. In other words, one baseline can be ``rotated into''
+another. We call these ``near-equivalent baselines''. It is our
+goal to a) identify the near-equivalent baselines that give good
+redundancy, b) to find the optimal time offset for a given pair of baselines, and
+c) to quantify the sensitivity improvement associated with cross multiplying
+such a pair of near-equivalent baselines. In section
+2 we explain the theoretical basis for this cross multiplication.
+In section 3 we present details of the numerical implementation of
+this technique. In section 4 we discuss the expected sensitivity improvement
+with this method and with section 5 we conclude.
 
-\subsection{Cross multiplication of near redundant data}
+\section{Method}
 
-To begin, we first derive theoretical expectations of cross correlation
-of two near-equivalent baselines. Given a point source on the sky,
-each baseline measures a track of the source in uv plane. We identify
-redundancy of near equivalent baselines for that point source as crossings
-of the tracks. Since processed data are measured at discrete time
-intervals (in lst-binned PAPER-64 data it's 42.9 seconds ), we do
-not have exact overlap of two data points in uv space. We thus need
-to find a way to cross multiply data points that are separated by
-a short distance in uv space and recover the power spectrum.
+\subsection{Rough Idea}
+\begin{figure}[H]
+\includegraphics[width=\linewidth]{64layout}
 
-We take the visibility as 
+\caption{The PAPER-64 layout. Each blue dot corresponds to the location of
+an antenna. Top panel shows the antenna positions drawn to scale,
+bottom panel show the antenna labels and distances.\label{fig:The-PAPER-64}
+The numbering of the antennas the bottom panel are original labels
+during instrument assembly and does not bear significant
+meaning. Baselines corresponding to the same separations are called
+equivalent. In the bottom panel, the two baselines labeled with the
+red arrow are equivalent, with separation denoted by sep1,0, for the
+antennas are separated by 1 unit east and 0 unit north. Similarly,
+the baselines labeled with the yellow and green arrows are examples
+of sep1,1 and sep-1,1, respectively. Note sep1,0 and sep-1,0 for example
+are the same baselines and should not be counted twice. }
+\end{figure}
+
+We shall use the 64-antenna PAPER array to demonstrate. 
+The PAPER-64 array is located in the Karoo desert in South Africa (30:43:17.5
+S, 21:25:41.8 E). The layout pattern with antenna labels are show
+in Fig. \ref{fig:The-PAPER-64}. We see the antenna spacing in North-South
+directions are comparatively close (4m), so that baselines such as
+0\_44 and 0\_7 are very close to equivalent. In the bottom panel,
+the two baselines labeled with the red arrow are equivalent, with
+separation denoted by sep1,0, for the antennas are separated by 1
+unit east and 0 unit north. Similarly, the baselines labeled with
+the yellow and green arrows are examples of sep1,1 and sep-1,1, respectively.
+Note sep1,0 and sep-1,0 for example are the same baselines and should
+not be counted twice. Turns out antennas in purely north-south baselines
+are close enough to induce cross talks, and hence are not suitable
+for use. The original PAPER-64 analysis thus used three classes of baselines,
+sep1,0, sep1,1 and sep-1,1 \cite{Ali2015}. There each of these classes
+of baselines are cross multiplied to itself. We shall see that in addition these
+baseline classes can be cross multiplied with a time offset.
+
+Given a point source on the sky, each baseline maps the
+source to a point in uv plane. As the earth rotates with respect to
+the source, the point traces out tracks in the uv-plane. 
+The uv tracks of PAPER-64 over 7.2 hours, at 0.15GHz, for a source that passes through zenith, are shown in Fig. \ref{fig:Tracks-of-PAPER-64}.
+
+\begin{figure}[H]
+\includegraphics[width=\linewidth]{TRACKSpt3days}
+\caption{Tracks of PAPER-64 for a hypothetical source that passes through zenith.
+These tracks are traced out over 0.3 sidereal days, or roughly 7.2
+hours. Color represents different baselines. Frequency is $\nu=0.15\text{GHz}$.
+As the earth rotates, tracks are traced out counterclockwise. \label{fig:Tracks-of-PAPER-64}}
+\end{figure}
+
+Roughly speaking, we can identify
+redundancy of near-equivalent baselines as crossings
+of the uv tracks. Although this is a valid quick method to identify some baselines to cross-multiply, we find
+that it is not accurate enough for time offset determination. The reason is, for wide beam antennas such as PAPER-64, the sources are not point-like. 
+In other words, when tracks of
+one point of the sky are crossing in uv-plane, those corresponding to another is not. To accurately
+constrain redundancy we must take into account of finite beam width. A more rigorous treatment is thus required.  
+We show the beam of PAPER-64 in Fig.\ref{fig:Beam-of-PAPER-64.} for reference:
+
+\begin{figure}[H]
+\includegraphics[width=\linewidth]{BEAM}
+
+\caption{Beam of PAPER-64. Shown here is the values of $A$ (Eq. \eqref{eq:Vis1})
+for frequency of $\nu=150\text{MHz}$. Notice $A$ is a ``baseline's
+beam''. Since all antennas in PAPER-64 have the same beam, $A$ here
+is just the square of an antenna beam, and is normalized to 1 at the
+center. The circles centered around zenith (center of beam) here are
+spaced 10 degrees apart. Such a wide beam limits the accuracy of the uv track-crossing as a method of redundancy search. \label{fig:Beam-of-PAPER-64.}}
+\end{figure}
+
+\subsection{Formalism}
+Below we shall derive theoretical expectations of cross multiplications
+of two near-equivalent baselines. More precisely we shall relate
+the product of visibilities of two different baselines to the power spectrum.
+
+We take the visibility as commonly defined in the literature (e.g.
+\cite{first-paper}): 
+\begin{equation}
+\begin{aligned}V_{\nu}(b) & =\int d\Omega A_{\nu}(\hat{s})\phi(\nu)I_{\nu}(\hat{s})\exp\left[-2\pi i\frac{\nu}{c}b\cdot\hat{s}\right],\\
+ & \approx\frac{2k_{B}}{\lambda^{2}}\int d\Omega A_{\nu}(\hat{s})\phi(\nu)T(\hat{s})\exp\left[-2\pi i\frac{\nu}{c}b\cdot\hat{s}\right],
+\end{aligned}
+\label{eq:Vis1}
+\end{equation}
+Here $\lambda$ is a mean wavelength, $\boldsymbol{b}$ is the baseline
+length, $\hat{\boldsymbol{s}}$ and $\Omega$ are a direction in the
+sky and its corresponding solid angle. $A_{\nu}$ is the (frequency
+dependent) beam, and $I$ is the specific intensity, which has been
+related to $T$, the brightness temperature in the Rayleigh-Jeans
+limit. $\phi(\nu)$ is the frequency band profile. For PAPER-64, for example,
+the bands studied are roughly a top-hat from 144MHz to 154 MHz.
+
+We define the delay transformed visibility \cite{delay-transform}:
+\small
+\begin{equation}
+\begin{aligned}V(b,\tau) & =\int d\nu V_{\nu}(b)\phi(\nu)\exp\left[-2\pi i\nu\tau\right],\\
+ & =\frac{2k_{B}}{\lambda^{2}}\int d\Omega d\nu A(\hat{s},\nu)\phi(\nu)T(\hat{s},\nu)\exp\left[-2\pi i\nu\left(\frac{b\cdot\hat{s}}{c}+\tau\right)\right]
+\end{aligned}
+.\label{eq:Vb1}
+\end{equation}
+\normalsize
+
+Eq. \eqref{eq:Vb1} expresses the delay-transformed visibility as
+an integral over observation coordinates $\hat{s}$ and $\nu$. Ultimately,
+we would like to relate the data, collected with coordinates $\hat{s}$
+and $\nu$, to the power spectrum, written with cosmological coordinates
+$\boldsymbol{r}$ and $\boldsymbol{k}$. We start by noticing that
+{[}Gerry: change all r and k into bold{]} 
 \[
-\begin{aligned}V_{\nu}(b) & =\int d\Omega A(\hat{r})I(\hat{r})\exp\left[-2\pi i\frac{\nu}{c}b\cdot\hat{r}\right],\\
- & \approx\frac{2k_{B}}{\lambda^{2}}\int d\Omega A(\hat{r})T(\hat{r})\exp\left[-2\pi i\frac{\nu}{c}b\cdot\hat{r}\right],\\
- & =\frac{2k_{B}}{\lambda^{2}}\int\frac{d^{3}r}{X^{2}Y}A(r)T(r)e^{-ikr}.
+\begin{aligned}r & =\frac{c}{H_{0}}\int_{0}^{z}\frac{dz'}{E(z')},\\
+ & \approx\frac{c}{H_{0}}\int_{0}^{z_{0}}\frac{dz'}{E(z')}-\frac{c(1+z)^{2}}{\nu_{21}H_{0}E(z)}\left(\nu-\nu_{0}\right),\\
+ & \equiv D_{c}-Y\Delta\nu,
 \end{aligned}
 \]
-Here $\lambda$ is a mean wavelength, $\boldsymbol{b}$ is the baseline
-length, $\hat{\boldsymbol{r}}$ is a direction in the sky, $A$ is
-the beam, and $I$ is the intensity, which has been related to $T$,
-the brightness temperature. In the last line $(r_{x},r_{y},r_{z})=(Xl,Xm,Y\nu)$,
-and $(Xk_{x},Xk_{y},Yk_{z})=2\pi(u,v,\eta)$.
+where $\nu_{21}=1420MHz$ is the 21cm transition rest frequency, $\nu_{0}$
+a reference central frequency with corresponding redshift $z_{0}$,
+and 
+\[
+E(z)=\sqrt{\Omega_{m}(1+z)^{3}+\Omega_{\Lambda}}.
+\]
+Inverting for $\nu$:
+\begin{equation}
+\nu=\frac{D_{c}-r}{Y}+\nu_{21}.\label{eq:nur}
+\end{equation}
+Thus $d\nu=-dr/Y$ and $d^{3}r=-X^{2}Yd\Omega d\nu$.
 
-We define the delay transformed visibility:
+We can thus rewrite the delayed transformed visibility as 
+\small
 \[
-\begin{aligned}V(b) & =\int d\nu V_{\nu}(b)\exp\left[-2\pi i\nu\eta\right],\\
- & =\frac{2k_{B}}{\lambda^{2}}\int d\Omega d\nu A(\hat{r},\nu)T(\hat{r},\nu)\exp\left[-2\pi i\left(\frac{\nu}{c}b\cdot\hat{r}+\nu\eta\right)\right]
+\boxed{\begin{aligned}V(b,\tau) & =\frac{1}{X^{2}Y}\int_{H}d^{3}rA(\vec{r})\phi(r)I(\vec{r})\exp\left[-2\pi i\left(\frac{b}{c}\cdot\hat{r}+\tau\right)\nu_{r}\right]\end{aligned}
+}.
+\]
+\normalsize
+Here $(r_{x},r_{y},r_{z})=(X\hat{s}_{x},X\hat{s}_{y},Y\nu)$, and
+$(Xk_{x},Xk_{y},Yk_{z})=\frac{2\pi}{c}(b_{x},b_{y},\eta)$ relate
+the cosmological coordinates $r$ and $k$ to the measured coordinates.
+We have written $\nu_{r}$ to remind us that $\nu$ and $r$ are related
+by Eq. \eqref{eq:nur}. The beam reception pattern $A$ is dimensionless,
+normalized to 1 at its peak (zenith), and we assume it to be the same
+for all baselines.
+
+With a time offset, the beam pattern has moved relative to the sky.
+Here we choose to fix the sky, and denote the rotated coordinates
+of the beam pattern with the operator $\Gamma$. With implicit bounds
+of integrals from $-\infty$ to $\infty$, we have:
+\begin{widetext}
+\begin{equation}
+\begin{aligned} & \langle V^{*}(b,\tau)V(b',\tau')\rangle\\
+ & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\langle T^{*}(r)T(r')\rangle\right)A^{*}(r)A(\Gamma r')\Phi_{b,\tau}(r,\Gamma r'),\\
+ & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\kappa)T(\kappa')\rangle e^{-i(\kappa r-\kappa'r')}\right)A^{*}(r)A(\Gamma r')\Phi_{b,\tau}(r,\Gamma r'),\\
+ & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\kappa(r-r')}\right)A^{*}(r)A(\Gamma r')\Phi_{b,\tau}(r,\Gamma r'),\\
+ & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\xi(r-r')A^{*}(r)A(\Gamma r')\Phi_{b,\tau}(r,\Gamma r'),\\
+ & \approx\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}rd^{3}r'\delta_{D}^{(3)}(r-r')A^{*}(r)A(\Gamma r')\Phi_{b,\tau}(r,\Gamma r'),\\
+ & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}r|A^{*}(r)A(\Gamma r)||\phi(\nu_{r})|^{2}\exp\left[-i2\pi\nu_{r}\left(\hat{r}\cdot\frac{b}{c}-\hat{\Gamma r}\cdot\frac{b'}{c}\right)\right],\\
+ & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{s},\nu)A(\hat{\Gamma s},\nu)||\phi(\nu)|^{2}\exp\left[-i2\pi\nu\left(\hat{s}\cdot\frac{b}{c}-\hat{\Gamma s}\cdot\frac{b'}{c}\right)\right],
 \end{aligned}
-\]
+\label{eq:main}
+\end{equation}
 
-\textbackslash{}The beam reception pattern $A$ is dimensionless,
-normalized to 1 at its peak (zenith), and we assume it to be the same
-for all baselines. We introduce the Fourier temperature and window
-function
+where in transition from cosmological coordinates back to observing coordinates we have written $\hat{r}\equiv\hat{s}$, and 
+\begin{equation}
+\Phi_{b,\tau}(r,\Gamma r')=|{\phi^{*}}(\ensuremath{\nu_{r}})\phi(\nu_{r'})|\exp\left[-i\frac{2\pi}{c}\left(b\cdot\nu_{r}\hat{r}-b'\cdot\nu_{r'}\Gamma\hat{r'}\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)\right].
+\end{equation}
+\end{widetext}
+The second to third line of Eq.(\ref{eq:main}) follows from assumption of Gaussian random
+sky, and the third to fourth line follows from the assumption that
+the 3D power spectrum varies negligibly over the region of interest
+so that $\hat{P_{21}}(k+k_{2})\approx\hat{P_{21}}(k)$. Since $\Gamma$
+is a sky rotation, it doesn't affect $\nu$, hence we have taken $\nu_{r}$
+outside the parenthesis. Notice that the phase factor $\exp\left[-i2\pi\eta\left(\nu-\nu'\right)\right]$
+drops out in the end.
 
-\[
-\begin{aligned}\tilde{T}(k) & =\int\frac{d^{3}r}{X^{2}Y}T(r)e^{-ikr},\\
-W(k) & =\int d^{3}rA(r)e^{-ikr}.
-\end{aligned}
-\]
-By the convolution theorem, we have
-\[
-\begin{aligned}\tilde{V}(k) & \approx\frac{2k_{B}}{\lambda^{2}}\int\frac{d^{3}r}{X^{2}Y}A(r)T(r)e^{-ikr},\\
- & =\frac{2k_{B}}{\lambda^{2}}\frac{1}{(2\pi)^{3}}\int d^{3}k_{1}W(k_{1})\tilde{T}(k+k_{1}).
-\end{aligned}
-\]
-With implicit bounds of integrals from $-\infty$ to $\infty$, we
-have
+Finally, since the beam pattern and bandwidth are given in $\hat{s}$
+and $\nu$, we convert the integral back to these coordinates to get
+the general relation between the delay-transformed visibilities and
+the power spectrum:
+
 \begin{equation}
-\begin{aligned}\langle\tilde{V}^{*}(k-k_{r})\tilde{V}(k)\rangle & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}\frac{1}{(2\pi)^{6}}\int dk_{1}dk_{2}W^{*}(k_{1})W(k_{2})\langle T^{*}(k+k_{1}-k_{r})T(k+k_{2})\rangle,\\
- & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}\frac{1}{(2\pi)^{6}}\int dk_{1}dk_{2}W^{*}(k_{1})W(k_{2})\delta_{D}(k_{2}-k_{1}-k_{r})\hat{P_{21}}(k+k_{2}),\\
- & \approx\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}\frac{\hat{P_{21}}(k)}{(2\pi)^{6}}\int dk_{1}W^{*}(k_{1})W(k_{1}+k_{r}),\\
- & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}\frac{\hat{P_{21}}(k)}{(2\pi)^{3}}\int d^{3}r|A(r)|^{2}e^{-ik_{r}r},\\
- & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}\frac{\hat{P_{21}}(k)}{(2\pi)^{3}}\int d^{3}r|a(r)\phi(r_{z})|^{2}e^{-ik_{r}r}
-\end{aligned}
-\label{eq:(7)}
+\boxed{\begin{aligned} & \langle V^{*}(b,\tau)V(b',\tau')\rangle\\
+ & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{s},\nu)A(\hat{\Gamma s},\nu)||\phi(\nu)|^{2}\\
+ & \qquad \qquad \qquad \qquad \exp\left[-i2\pi\nu\left(\hat{s}\cdot\frac{b}{c}-\hat{\Gamma s}\cdot\frac{b'}{c}\right)\right].\end{aligned}}
+\label{eq:final}
 \end{equation}
-where the first to second line follows from assumption of Gaussian
-random sky, and the second to third line follows from the assumption
-that the 3D power spectrum varies neglegibly over the region of interest
-so that $\hat{P_{21}}(k+k_{2})\approx\hat{P_{21}}(k)$. In the last
-line we have separated the 2D spatial beam patter from the frequency
-bandpass $\phi$. So in conclusion the cross correlation of visibilities
-seperated by $\Delta u$ apart in uv space is proportional to the
-power spectrum times the Fourier transform of the beam pattern squared.
+So roughly speaking the cross multiplications of visibilities at a time delay
+in uv-space is proportional to the power spectrum times the Fourier
+transform of the cross multiplied beam pattern. We can then in principle
+combine information from different baseline pairs if we correct for
+the phase and normalization. As a check, when applied to equivalent baselines,
+$b=b'$, $\hat{s}=\hat{\Gamma s}$, and Eq.(\ref{eq:final}) reduces to Eq.(B9) of \cite{paper32}.
+
+If this result is robust, then we can achieve all our goals stated in the introduction, i.e. to identify 
+candidate baseline pairs, to find the optimal time offset, and to quantify the sensibility, by computing the integral
+in Eq.(\ref{eq:final}) for various time offsets. Thus we shall first test its robustness numerically.
+
+\section{Analysis}
+\subsection{Numerical Test \label{sec:Techniquet}}
+
+To test the robustness of Eq.(\ref{eq:final}), we need to compare the amplitude and phase of the integral
+for a pair of baselines with products of simulated visibilities of those baselines.
+
+For the simulation, we take N random realizations of the sky
+on a healpix map (\cite{Heal}, \cite{HealPrimer}) \footnote{We use functionalities in the python package AIPY for healpix mapping
+as well as coordinate transforms. }. For each realization, each pixel is given a Gaussian random value
+of brightness temperature. We then rotate the baseline positions with
+the appropriate rotation matrix, recording the visibilities, for each
+baseline \footnote{As a caveat, there are two obvious ways to achieve the rotation. One
+can either fix the sky and rotate the baselines, or the other way
+around. We found however, that we must not physically rotate the sky
+map, for the numerical round-offs due to finite resolutions of the
+map turns out to be significant. Thus we let the sky, represented
+by the healpix map, be fixed, and rotate the baselines. }. The resulting visibilities for the two baselines are then convolved
+via the Fourier convolution theorem, to obtain values of the cross
+correlation as a function of time-offset. The peak of the curve then
+corresponds to maximum redundancy. The accuracy of this result is
+limited by (simulated) cosmic variance and finite spacial resolution,
+and hence can be beat down by averaging over a large number of universes.
+A numerical estimate of the error of the peak height with $N=1$ is
+$\lesssim20\%$, and thus with $N=2500,$ we achieve an error of peak
+height $<0.5\%$.
+
+To test our understanding of the time-offset, and to verify our derivation
+in Eq. \eqref{eq:main}, we must compare the simulated visibility cross multiplication with the integral in
+Eq. \eqref{eq:main}, for a pair of near-equivalent baselines. 
+This comparison is show in Fig. \ref{fig:numerics}. 
+Since the sensitivity contribution is inferred from height of the peak of the cross-multiplied visibilities as a function of time. 
+We first compare the simulation to the integral for a pair of equivalent baselines, then normalize the peaks 
+of the near-equivalent baselines to that of the equivalent ones. 
+In Fig. \ref{fig:numerics}  we show on the top comparison of
+the convolution of an equivalent baseline, normalized to 1 at the
+peak, whose location is at zero time offset as expected. On the bottom we show
+comparison of a near-equivalent baseline pair, of classes sep1,0:1,1 and sep1,1:1,1. The peaks
+are normalized with the same factor as the peak height on the top,
+i.e. the two plots have thus the same scale on the vertical axis.
+The height of the plot on the right thus quantifies the added sensitivity.
 
 \begin{figure}[H]
-\includegraphics[scale=0.5]{64layout}
-
-\caption{The PAPER 64 layout. Top panel shows the antenna positions drawn to
-scale, bottom panel show the antenna labels and distances. }
+\includegraphics[width=\linewidth]{redun_agree}
+\label{fig:numerics}
+\caption{[Placeholder Plot]}
 \end{figure}
 
-We shall use the 64-antenna PAPER array to demonstrate. The PAPER
-64 array is located in the Karoo desert in South Africa (30:43:17.5
-S, 21:25:41.8 E). The layout pattern with antenna labels are show
-in Fig. 1. I show denote a baseline with the antconfigWe see the antenna
-spacing in North-South directions are comparatively close (4m), so
-that baselines such as 0\_44 and 0\_7 are very close to equivalent.
+At the optimal time separation, the integral in Eq. \eqref{eq:final}
+is maximized. Thus as a final check we expect the two beams to have to same fringe pattern
+(frequency and phase). Due to the time delay, however, the beam center
+would be slightly shifted with respect to each other. This we show
+in Fig. \ref{fig:Beam-fringe-pattern}. The left and middle panels show the beam fringe
+patterns for baselines sep0,1 and sep1,1, delayed by 0.0325 days,
+and the right panel shows their cross product. The fringe pattern
+indeed cancels out as we expect.
 
-\section{Implementation}
+\begin{widetext}
+\begin{figure}[H]
+\includegraphics[scale=0.5]{fringe_res}
 
-An example of rotation tracks is shown in the uv plane below
+\caption{Beam fringe pattern of sep1.0(left), sep1,1 at a time delay (middle),
+and their conjugate product. Frequency of $\nu=0.15\text{GHz}$ is
+chosen. \label{fig:Beam-fringe-pattern}}
+\end{figure}
+\end{widetext}
+
+\subsection{Expected Sensitivity Contributions}
+
+Having verified Eq. \eqref{eq:main}, we can thus predict the sensitivity contributions of a
+particular baseline pair simply by computing the integral. 
+Having computed all of the baseline pairs, we find that baseline pairs that are mirror images of each other 
+give the same amount of redundancies (peak height), with the opposite time offset, as expected from symmetry. 
+For example, sep1,0:1,1 is mirror image of sep1,0:1,-1 and thus these two baseline pairs
+give the same sensibility contribution. Thus in our results we shall only show a subset of representative baseline pairs 
+that give different informations.
+
+In the top panel of Fig. \eqref{fig:sensplot} below
+we show the peak heights and locations for a variety of baseline combinations.
+We see that baseline pairs that have crossings at a smaller time delay
+tend to have higher correlations. In other words, correlation peaks
+that are closer to zero time lag are higher. This is expected since
+a) the longer the time delay, the more the antennas have moved with respect
+to the sky and hence the less overlaps in patch of sky surveyed, b) smaller optimal
+time-offset corresponds to smaller differences in orientation, and hence in length of
+baselines in PAPER-64.
 
 \begin{figure}[H]
-\includegraphics[scale=0.5]{tracks}
+\includegraphics[width=\linewidth]{sensitivity}
+\label{fig:sensplot}
+\caption{Relative sensitivity contributions of selected baseline combinations in PAPER-64. In the legend m,n:p,q denotes cross
+multiplying PAPER-64 baselines of east-west, north-south separations (m,n) and (p,q) respectively. The top
+panel shows the peak height (degree of correlation) of each baseline
+combination, while the bottom panel multiplies the heights by the
+corresponding multiplicities as in Eq. \eqref{eq:sensul}, and in some cases an extra factor of 2 (explained in text). 
+In weighting by the multiplicity (bottom), we have chosen to fix the sensitivity 
+contribution of 1,0:1,0 to unity. }
 
-\caption{The PAPER 64 layout. Top panel shows the antenna positions drawn to
-scale, bottom panel show the antenna labels and distances. }
 \end{figure}
 
-The beam squared and its Fourier transform are given as
+To determine that actual relative contribution to sensitivity of these
+baseline pairs, we have to take into account of the multiplicities of
+these baselines, or the size of the equivalency classes in a mathematical
+language. By these we mean how many physical antenna pairs have the
+same length and orientation. Looking at Fig. \ref{fig:The-PAPER-64}
+we see for example sep1,0 will have higher redundancy than sep2,0,
+or sep1,1. The latest release of PAPER-64 data uses baselines sep-1,1,
+sep0,1 and sep1,1 \cite{Ali2015}, and achieved a $2\sigma$ upper
+limit of $22.4mK^{2}$. There, the three sets of equivalent baselines
+are only cross multiplied by itself. Assuming that each baseline delivers
+the same quality of data (i.e. they have the same height of correlation
+peaks, which is in our normalization equal to unity), the relative
+contribution to sensitivity can be estimated.
 
-\begin{figure}[H]
-\includegraphics[scale=0.4]{64beam}
+First we can average of the visibilities of the baselines of the same
+separations. Since PAPER-64 has 8 by 8 antenna configuration, there
+are $(8-|m|)\times(8-|n|)$ copies of the baseline sepm,n. This means
+that if we add visibility measurements of all these equivalent baselines,
+we get a factor of $\sqrt{(8-|m|)\times(8-|n|)}$ reduction in noise
+level. The sensitivity contribution of sepm,n, being the signal to
+noise ratio of the power spectrum, is thus proportional to $(8-|m|)\times(8-|n|)$.
+Next we can average over the power-spectrum taken from baselines of
+the different separations. Thus the total cumulative sensitivity is
+given by 
+\begin{equation}
+\frac{P_{S}}{P_{N}}\propto\sqrt{\sum_{(n,m)}\left((8-|m|)\times(8-|n|)\right)^{2}},\label{eq:sensequiv}
+\end{equation}
+where the sum is only over pairs of m and n. If we now include power
+spectrum from cross-multiplications of near-equivalent baselines of
+types sepm,n and sepm',n', we get a general cumulative sensitivity:
+\small
+\begin{equation}
+\frac{P_{S}}{P_{N}}\propto\sqrt{\sum_{(n,m,n',m')}\Omega_{mnm'n'}\left((8-|m|)(8-|n|)(8-|m'|)(8-|n'|)\right)},\label{eq:sensul}
+\end{equation}
+\normalsize
+where $\Omega$ is the correlation peak height show in top panel of Fig. \eqref{fig:sensplot},
+and the sum is again only over tuples of m,n,m',n'. This form now
+includes the equivalent baselines in Eq. \eqref{eq:sensequiv}, with
+$\Omega_{m=m',n=n'}=1$. 
+Shown in the bottom panel of Fig. \eqref{fig:sensplot} is the peak heights weighted
+by this multiplicity factor (before taking the square root), with one difference. The difference is, 
+since we have "folded over" the negative time delays, we have combined baseline pairs that are identical modulus parity, 
+or in other words are mirror images of each other, by summing over them to get an extra factor of two. For example, 
+instead of plotting 1,1:1,1 and 1,-1:1,-1 separately, we plot 1,1:1,1 with twice the multiplicity. Similarly, baseline pairs such as 
+1,0:1,1 also get the factor of 2 because they are identical to 1,0:1,-1. Baseline pairs such as 1,0:1,0, or 1,1:1,-1 do not get
+the factor of 2 because their mirror images are themselves.
 
-\caption{The PAPER 64 layout. Top panel shows the antenna positions drawn to
-scale, bottom panel show the antenna labels and distances. }
-\end{figure}
+\section{Conclusion}
+
+\bibliographystyle{apj}
+%\nocite{*}
+\bibliography{draft_working}
 

### 2016-11-16 Rotation paper/draft_working.tex
update

@@
 \altaffiltext{2}{Radio Astronomy Lab., U. California, Berkeley, CA}
-%\altaffiltext{3}{Berkeley Center for Cosmological Physics, Berkeley, CA}
-%\altaffiltext{3}{Dept. of Physics and Astronomy, U. Pennsylvania, Philadelphia, PA}
-%\altaffiltext{8}{School of Earth and Space Exploration, Arizona State U., Tempe, AZ}
 
@@
 We introduce an accurate method to extract additional sensitivity from data of redundant radio arrays. 
-Using the 64 element version of Precision Array for Probing the Epoch of Reionization (PAPER-64), we illustrate 
+Using the latest and final 128-element version of Precision Array for Probing the Epoch of Reionization (PAPER-64), we illustrate 
 how our method finds optimal baselines to cross correlate, including baselines that are slightly different in length and orientation. 
 We predict the sensitivity improvements of including each baseline pairs. The extremely natural and simple implementation of this method
-would increase the sensitivity of PAPER-64 by up to [~50] percent. 
+would increase the sensitivity of PAPER-128 by up to [~50] percent. 
 \end{abstract}
@@
 imaging arrays focus on Fourier mode uv coverage, the redundant arrays
-have many tightly packed antennas at equal spacing. Since baselines
+focus on redundant measurements of the same Fourier mode
+with many tightly packed antennas at equal spacing. Since baselines
 of the same length and orientation measure the same Fourier modes
@@
 \begin{figure}[H]
-\includegraphics[width=\linewidth]{64layout}
+\includegraphics[width=\linewidth]{antpos128}
 
-\caption{The PAPER-64 layout. Each blue dot corresponds to the location of
+\caption{The PAPER-128 layout. Each blue dot corresponds to the location of
 an antenna. Top panel shows the antenna positions drawn to scale,
-bottom panel show the antenna labels and distances.\label{fig:The-PAPER-64}
+bottom panel show the antenna labels and distances, excluding the outlier
+antennas.\label{fig:AntPos}
 The numbering of the antennas the bottom panel are original labels
@@
 of sep1,1 and sep-1,1, respectively. Note sep1,0 and sep-1,0 for example
-are the same baselines and should not be counted twice. }
+are the same baselines and should not be counted twice.}
 \end{figure}
 
-We shall use the 64-antenna PAPER array to demonstrate. 
-The PAPER-64 array is located in the Karoo desert in South Africa (30:43:17.5
+We shall use the 128-element PAPER array to demonstrate. 
+The PAPER array is located in the Karoo desert in South Africa (30:43:17.5
 S, 21:25:41.8 E). The layout pattern with antenna labels are show
-in Fig. \ref{fig:The-PAPER-64}. We see the antenna spacing in North-South
+in Fig. \ref{fig:AntPos}. We see the antenna spacing in North-South
 directions are comparatively close (4m), so that baselines such as
@@
 are close enough to induce cross talks, and hence are not suitable
-for use. The original PAPER-64 analysis thus used three classes of baselines,
-sep1,0, sep1,1 and sep-1,1 \cite{Ali2015}. There each of these classes
+for use. The original PAPER-64 analysis used three classes of baselines, here
+equivalent to 
+sep2,0, sep2,1 and sep-2,1 \cite{Ali2015}. There each of these classes
 of baselines are cross multiplied to itself. We shall see that in addition these
@@
 the source, the point traces out tracks in the uv-plane. 
-The uv tracks of PAPER-64 over 7.2 hours, at 0.15GHz, for a source that passes through zenith, are shown in Fig. \ref{fig:Tracks-of-PAPER-64}.
+The uv tracks of PAPER-128 over 7.2 hours, at 0.15GHz, for a source that passes through zenith, are shown in Fig. \ref{fig:Tracks}.
 
 \begin{figure}[H]
-\includegraphics[width=\linewidth]{TRACKSpt3days}
-\caption{Tracks of PAPER-64 for a hypothetical source that passes through zenith.
-These tracks are traced out over 0.3 sidereal days, or roughly 7.2
+\includegraphics[width=\linewidth]{tracks128}
+\caption{Tracks of PAPER-128 (grid only, excluding outliers) for a hypothetical source that passes through zenith.
+These tracks are traced out over 0.2 sidereal days, or roughly 4.8
 hours. Color represents different baselines. Frequency is $\nu=0.15\text{GHz}$.
-As the earth rotates, tracks are traced out counterclockwise. \label{fig:Tracks-of-PAPER-64}}
+As the earth rotates, tracks are traced out counterclockwise. \label{fig:Tracks}}
 \end{figure}
@@
 of the uv tracks. Although this is a valid quick method to identify some baselines to cross-multiply, we find
-that it is not accurate enough for time offset determination. The reason is, for wide beam antennas such as PAPER-64, the sources are not point-like. 
+that it is not accurate enough for time offset determination. The reason is, for wide beam antennas such as PAPER, the sources are not point-like. 
 In other words, when tracks of
@@
 constrain redundancy we must take into account of finite beam width. A more rigorous treatment is thus required.  
-We show the beam of PAPER-64 in Fig.\ref{fig:Beam-of-PAPER-64.} for reference:
+We show the beam of PAPER antennas in Fig.\ref{fig:Beam} for reference:
 
@@
 
-\caption{Beam of PAPER-64. Shown here is the values of $A$ (Eq. \eqref{eq:Vis1})
+\caption{Beam of PAPER antennas. Shown here is the values of $A$ (Eq. \eqref{eq:Vis1})
 for frequency of $\nu=150\text{MHz}$. Notice $A$ is a ``baseline's
-beam''. Since all antennas in PAPER-64 have the same beam, $A$ here
+beam''. Since all antennas in PAPER have the same beam, $A$ here
 is just the square of an antenna beam, and is normalized to 1 at the
 center. The circles centered around zenith (center of beam) here are
-spaced 10 degrees apart. Such a wide beam limits the accuracy of the uv track-crossing as a method of redundancy search. \label{fig:Beam-of-PAPER-64.}}
+spaced 10 degrees apart. Such a wide beam limits the accuracy of the uv track-crossing as a method of redundancy search. \label{fig:Beam}}
 \end{figure}
@@
 
+\subsection{Validation with PAPER-128 Data}
+
 \section{Conclusion}

### 2017-03-02 Rotation paper/draft_working.tex
saving

@@
 We introduce an accurate method to extract additional sensitivity from data of redundant radio arrays.
-Using the latest and final 128-element version of Precision Array for Probing the Epoch of Reionization (PAPER-64), we illustrate 
-how our method finds optimal baselines to cross correlate, including baselines that are slightly different in length and orientation. 
-We predict the sensitivity improvements of including each baseline pairs. The extremely natural and simple implementation of this method
-would increase the sensitivity of PAPER-128 by up to [~50] percent. 
+Our method efficiently finds optimal baselines to cross correlate, including baselines that are slightly different in length and orientation.
+Using the latest and final 128-element version of Precision Array for Probing the Epoch of Reionization (PAPER-128), as well as the planned 37, 128, 240 and 350 element versions of the Hydrogen Epoch of Reionization Array (HERA) now under development, we illustrate how our method applies to different array configurations and predict the sensitivity improvements of including each baseline pairs.
+The extremely natural and simple implementation of this method
+would increase the sensitivity of PAPER-128 by up to [~50] percent, and increase the out-rigger imaging sensitivity of HERA-350 by a factor of 2. 
 \end{abstract}
@@
 the intersection of cosmology and astrophysics. Understanding this
-event is important not only as a scientific goal
-of its own, but also because it potentially provides crucial information
+event is important not only because it is a scientific goal
+of its own, but also because it can provide crucial information
 to fundamental nature of inflation, neutrino mass and phenomenology
-of the first stars and galaxies alike (e.g. \cite{LiuOpticalDepth}).
+of the first stars and galaxies (e.g. \cite{LiuOpticalDepth}).
 
@@
 of the same length and orientation measure the same Fourier modes
-on the sky, the maximum redundancy arrays is able to increase the
+on the sky, a maximum redundancy array is able to increase the
 signal to noise ratio by averaging over measurements of the same baseline.
@@
 rotate into each other at a time delay. Here we present a method to
-improve the sensitivity of the maximum redundancy arrays by considering
+improve the sensitivity of the maximum redundancy arrays by including
 the uv redundancy of different baselines at a time lag.
@@
 baselines will be redundant with each other simultaneously at all
-times. Some non-equivalent baselines can also be redundant if their
+times. Non-equivalent baselines can also be partially redundant if their
 respected time series is shifted with respect to one another by a
@@
 this technique. In section 4 we discuss the expected sensitivity improvement
-with this method and with section 5 we conclude.
+with this method for HERA and PAPER-128 and with section 5 we conclude.
 
@@
 \subsection{Rough Idea}
+
 \begin{figure}[H]
@@
 are the same baselines and should not be counted twice.}
+
 \end{figure}
@@
 Note sep1,0 and sep-1,0 for example are the same baselines and should
-not be counted twice. Turns out antennas in purely north-south baselines
+not be counted twice. Antennas in purely north-south baselines
 are close enough to induce cross talks, and hence are not suitable
@@
 equivalent to 
-sep2,0, sep2,1 and sep-2,1 \cite{Ali2015}. There each of these classes
+sep2,0, sep2,1 and sep-2,1 \cite{Ali2015} of PAPER-128. There each of these classes
 of baselines are cross multiplied to itself. We shall see that in addition these
@@
 the source, the point traces out tracks in the uv-plane. 
-The uv tracks of PAPER-128 over 7.2 hours, at 0.15GHz, for a source that passes through zenith, are shown in Fig. \ref{fig:Tracks}.
+The uv tracks of PAPER-128 over 4.8 hours, at 0.15GHz, for a source that passes through zenith, are shown in Fig. \ref{fig:Tracks}.
 
@@
 constrain redundancy we must take into account of finite beam width. A more rigorous treatment is thus required. 
-We show the beam of PAPER antennas in Fig.\ref{fig:Beam} for reference:
+We show sample beams of HERA and PAPER antennas in Fig.\ref{fig:Beam} for reference.
 
 \begin{figure}[H]
-\includegraphics[width=\linewidth]{BEAM}
+\includegraphics[width=1.2\linewidth]{Beams}
 
-\caption{Beam of PAPER antennas. Shown here is the values of $A$ (Eq. \eqref{eq:Vis1})
-for frequency of $\nu=150\text{MHz}$. Notice $A$ is a ``baseline's
-beam''. Since all antennas in PAPER have the same beam, $A$ here
-is just the square of an antenna beam, and is normalized to 1 at the
-center. The circles centered around zenith (center of beam) here are
-spaced 10 degrees apart. Such a wide beam limits the accuracy of the uv track-crossing as a method of redundancy search. \label{fig:Beam}}
+\caption{Sample beam response of HERA (left) and PAPER (right) antennas, both
+for frequency of $\nu=150\text{MHz}$ and Stokes $I$ polarization.
+Notice $A$ in this paper is a ``baseline's beam'', equivalent to squares of the antenna beams shown here.
+The circles centered around zenith (center of beam) here are
+spaced 10 degrees apart. The finite size of the beams limits the accuracy of the uv track-crossing as a method of redundancy search. \label{fig:Beam}}
 \end{figure}
@@
 related to $T$, the brightness temperature in the Rayleigh-Jeans
-limit. $\phi(\nu)$ is the frequency band profile. For PAPER-64, for example,
+limit. Note that $A$ here is the baseline beam, i.e. product of two antenna beams, examples of which are shown in Fig.\ref{fig:Beam}. $\phi(\nu)$ is the frequency band profile. For PAPER, for example,
 the bands studied are roughly a top-hat from 144MHz to 154 MHz.
@@
 To test the robustness of Eq.(\ref{eq:final}), we need to compare the amplitude and phase of the integral
-for a pair of baselines with products of simulated visibilities of those baselines.
+for a pair of baselines with products of simulated visibilities of those baselines. Without loss of generality, we shall use a single frequency channel of $150$MHz for the following analysis.
 
@@
 In Fig. \ref{fig:numerics}  we show on the top comparison of
-the convolution of an equivalent baseline, normalized to 1 at the
+the convolution of an equivalent baseline of PAPER-128, normalized to 1 at the
 peak, whose location is at zero time offset as expected. On the bottom we show
@@
 i.e. the two plots have thus the same scale on the vertical axis.
-The height of the plot on the right thus quantifies the added sensitivity.
+The height of the plot thus quantifies the added sensitivity. We see at a peak of around 0.055 sidereal days, or 1.32 hours, the two baselines are maximally redundant.
 
@@
 \label{fig:numerics}
-\caption{[Placeholder Plot]}
+\caption{Numerical comparisons of the visibility correlation peaks to the $\Omega$ factor in Eq.(\ref{eq:final}).
+We generated 24000 instances of Gaussian random sky on healpix maps, computed visibilities and cross correlated them to find the correlation.
+We see that the theory and simulation line up in both amplitude and phase.
+}
 \end{figure}
@@
 At the optimal time separation, the integral in Eq. \eqref{eq:final}
-is maximized. Thus as a final check we expect the two beams to have to same fringe pattern
+is maximized. Thus as another check we expect the two beams to have to same fringe pattern
 (frequency and phase). Due to the time delay, however, the beam center
@@
 in Fig. \ref{fig:Beam-fringe-pattern}. The left and middle panels show the beam fringe
-patterns for baselines sep0,1 and sep1,1, delayed by 0.0325 days,
+patterns for baselines sep2,0 and sep2,1, delayed by 0.0325 sidereal days,
 and the right panel shows their cross product. The fringe pattern
@@
 
-\caption{Beam fringe pattern of sep1.0(left), sep1,1 at a time delay (middle),
+\caption{Beam fringe pattern of sep2,0(left), sep2,1 at a time delay (middle),
 and their conjugate product. Frequency of $\nu=0.15\text{GHz}$ is
@@
 particular baseline pair simply by computing the integral. 
+Intuitively we expect the sensitivity to depend on both the uv coverage of the baseline and the patch of sky inside the beam.
+A larger beam like PAPER-beam would tolerate larger time-offsets because more sky area can coincide in the two beams.
+To keep the flow of the discussion, we shall focus on results of PAPER-128, and present the results for HERA in the Appendix.
 Having computed all of the baseline pairs, we find that baseline pairs that are mirror images of each other 
 give the same amount of redundancies (peak height), with the opposite time offset, as expected from symmetry. 
-For example, sep1,0:1,1 is mirror image of sep1,0:1,-1 and thus these two baseline pairs
+For example, sep1,0:1,1 is mirror image of sep1,0:1,-1 and these two baseline pairs
 give the same sensibility contribution. Thus in our results we shall only show a subset of representative baseline pairs 
@@
 language. By these we mean how many physical antenna pairs have the
-same length and orientation. Looking at Fig. \ref{fig:The-PAPER-64}
+same length and orientation. Looking at Fig. \ref{fig:sensplot}
 we see for example sep1,0 will have higher redundancy than sep2,0,
-or sep1,1. The latest release of PAPER-64 data uses baselines sep-1,1,
-sep0,1 and sep1,1 \cite{Ali2015}, and achieved a $2\sigma$ upper
+or sep1,1. The latest release of PAPER-64 data uses the 128-equivalent baselines sep2,1,
+sep2,0 and sep2,-1 \cite{Ali2015}, and achieved a $2\sigma$ upper
 limit of $22.4mK^{2}$. There, the three sets of equivalent baselines
@@
 First we can average of the visibilities of the baselines of the same
-separations. Since PAPER-64 has 8 by 8 antenna configuration, there
-are $(8-|m|)\times(8-|n|)$ copies of the baseline sepm,n. This means
+separations. Since the core of PAPER-128 has 16 by 7 antenna configuration, there
+are $(16-|m|)\times(7-|n|)$ copies of the baseline sepm,n. This means
 that if we add visibility measurements of all these equivalent baselines,
-we get a factor of $\sqrt{(8-|m|)\times(8-|n|)}$ reduction in noise
+we get a factor of $\sqrt{(16-|m|)\times(7-|n|)}$ reduction in noise
 level. The sensitivity contribution of sepm,n, being the signal to
-noise ratio of the power spectrum, is thus proportional to $(8-|m|)\times(8-|n|)$.
+noise ratio of the power spectrum, is thus proportional to $(16-|m|)\times(7-|n|)$.
 Next we can average over the power-spectrum taken from baselines of
@@
 \begin{equation}
-\frac{P_{S}}{P_{N}}\propto\sqrt{\sum_{(n,m)}\left((8-|m|)\times(8-|n|)\right)^{2}},\label{eq:sensequiv}
+\frac{P_{S}}{P_{N}}\propto\sqrt{\sum_{(n,m)}\left((16-|m|)\times(7-|n|)\right)^{2}},\label{eq:sensequiv}
 \end{equation}
@@
 \begin{equation}
-\frac{P_{S}}{P_{N}}\propto\sqrt{\sum_{(n,m,n',m')}\Omega_{mnm'n'}\left((8-|m|)(8-|n|)(8-|m'|)(8-|n'|)\right)},\label{eq:sensul}
+\frac{P_{S}}{P_{N}}\propto\sqrt{\sum_{(n,m,n',m')}\Omega_{mnm'n'}\left((16-|m|)(7-|n|)(16-|m'|)(7-|n'|)\right)},\label{eq:sensul}
 \end{equation}
@@
 \subsection{Validation with PAPER-128 Data}
+Unlike the simulated clean EOR signal, real data are dominated by foreground and thermal noise, neither of which turns out to exhibit the correlation pattern seen in Fig.
+\ref{fig:numerics}.
+Foreground exhibits long-range correlations and would lead to high responses at wide range of time-offsets.
+Instrumental noise fo different baselines/antenna-pairs, on the other hand, are uncorrelated and thus exhibit responses consistent with zero at all time offsets.
 

### 2017-03-04 Rotation paper/draft_working.tex
save

@@
 improve the sensitivity of the maximum redundancy arrays by including
-the uv redundancy of different baselines at a time lag.
+the uv redundancy of different baselines at a time lag. It's worth noting 
+as recent progress the authors of \cite{wterm} proposes an algorithmic search approach to extract information from near-equivalent baselines.
+We here propose a simple and elegant theoretical formalism that in addition allows us to identify baselines to cross-correlate and predict the weight of sensitivity.
 
@@
 
-\subsection{Validation with PAPER-128 Data}
+\subsection{Notes on Data, Foreground and Noise}
 Unlike the simulated clean EOR signal, real data are dominated by foreground and thermal noise, neither of which turns out to exhibit the correlation pattern seen in Fig.
@@
 Foreground exhibits long-range correlations and would lead to high responses at wide range of time-offsets.
-Instrumental noise fo different baselines/antenna-pairs, on the other hand, are uncorrelated and thus exhibit responses consistent with zero at all time offsets.
+Instrumental noise for different baselines/antenna-pairs, on the other hand, are uncorrelated and thus exhibit responses consistent with zero at all time offsets.
+To illustrate, we use the second observing season of PAPER-128 data, taken from Jan.
+21st to March 7th of 2014.
+To illustrate the effect of noise and foreground, we use the data before fringe-rate and delay filtering.
 
@@
 
+\appendix
+\section{\\Hera Analysis}
+\begin{widetext}
+\begin{figure}[H]
+\includegraphics[scale=0.55]{HeraAntpos}
+
+\caption{Planned Hydrogen Epoch of Reionization Array antenna configurations\label{fig:HeraAntpos}}
+\end{figure}
+\end{widetext}
 \bibliographystyle{apj}

### 2017-03-12 Rotation paper/draft_working.tex
sync

@@
 is a sky rotation, it doesn't affect $\nu$, hence we have taken $\nu_{r}$
-outside the parenthesis. Notice that the phase factor $\exp\left[-i2\pi\eta\left(\nu-\nu'\right)\right]$
-drops out in the end.
+outside the parenthesis. Notice that the phase factor $\exp\left[-i2\pi\tau\left(\nu-\nu'\right)\right]$
+drops out in the end. This is significant because this means that correlation is peaked at the same time-offset for all delay channels.
 
@@
 To test the robustness of Eq.(\ref{eq:final}), we need to compare the amplitude and phase of the integral
-for a pair of baselines with products of simulated visibilities of those baselines. Without loss of generality, we shall use a single frequency channel of $150$MHz for the following analysis.
+for a pair of baselines with products of simulated visibilities of those baselines.
+For computational simplicity, we shall use a single frequency channel of $150$MHz for the comparison.
+We may use this simplification without compromising generality because Eq.
+(\ref{eq:final}) is valid for a range of frequency channels if an only if it's valid for each single channels (i.e.
+when $\phi(\nu)$ is proportional to a delta function).
 
@@
 \subsection{Expected Sensitivity Contributions}
-
 Having verified Eq. \eqref{eq:main}, we can thus predict the sensitivity contributions of a
 particular baseline pair simply by computing the integral.
+Having verified the validity of Eq.(\ref{eq:final}) from a single channel, we must point out that the peak height, or sensitivity contribution do depend on the frequency channels and shape of bandpass.
+This is because different frequency channels differ slighly in their phases at peak correlation and average incoherently.
+In Fig.\ref{fig:freqdiff} we compare two channels 0.16 GHz and 0.17 GHz.
+Top pannels shows the equivalent baseline pairs sep1,0 against itself, and bottom shows sep1,0 against sep1,1.
+We see that although the magnitude, drawn by the wide curves match up, the phases do not.
+This means the wider the frequency profile, the more destructive the interference, and lower the sensitivity contribution.
+\begin{figure}[H]
+\includegraphics[width=\linewidth]{freqdiff}
+\label{fig:freqdiff}
+\caption{Comparisons of the peak phases of two different frequencies.
+Top panels shows equivalent baselines, bottom shows a pair of near-equivalents.
+The dot-dashed lines of each color are the imaginary parts, and the solid lines are the real parts.
+For near-equivalent case, we also draw the magnitude by the broad envelops.
+We see that although the magnitude line up for two frequencies, the phases do not.
+This means there will be destructive interference when cross multiplying delay-space visibilities.
+}
+\end{figure}
+
 Intuitively we expect the sensitivity to depend on both the uv coverage of the baseline and the patch of sky inside the beam.

### 2017-03-15 Rotation paper/draft_working.tex
save

--- 2017-03-02 Rotation paper/draft_working.tex
@@
 the uv redundancy of different baselines at a time lag.
+Recently the authors of \cite{wterm} proposed an algorithmic search approach to extract information from all near-equivalent baselines.
+We here propose a simple and elegant theoretical formalism that in addition allows us to identify good baselines to cross-correlate and predict the weight of sensitivity with speed and accuracy.
 
@@
 \[
-\boxed{\begin{aligned}V(b,\tau) & =\frac{1}{X^{2}Y}\int_{H}d^{3}rA(\vec{r})\phi(r)I(\vec{r})\exp\left[-2\pi i\left(\frac{b}{c}\cdot\hat{r}+\tau\right)\nu_{r}\right]\end{aligned}
-}.
+\begin{aligned}V(b,\tau) & =\frac{1}{X^{2}Y}\int_{H}d^{3}rA(\vec{r})\phi(r)I(\vec{r})\exp\left[-2\pi i\left(\frac{b}{c}\cdot\hat{r}+\tau\right)\nu_{r}\right]\end{aligned}.
 \]
@@
 \begin{equation}
-\boxed{\begin{aligned} & \langle V^{*}(b,\tau)V(b',\tau')\rangle\\
+\begin{aligned} & \langle V^{*}(b,\tau)V(b',\tau')\rangle\\
  & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{s},\nu)A(\hat{\Gamma s},\nu)||\phi(\nu)|^{2}\\
- & \qquad \qquad \qquad \qquad \exp\left[-i2\pi\nu\left(\hat{s}\cdot\frac{b}{c}-\hat{\Gamma s}\cdot\frac{b'}{c}\right)\right].\end{aligned}}
+ & \qquad \qquad \qquad \qquad \exp\left[-i2\pi\nu\left(\hat{s}\cdot\frac{b}{c}-\hat{\Gamma s}\cdot\frac{b'}{c}\right)\right].\end{aligned}
 \label{eq:final}
@@
 
+\subsection{Rephasing \label{sec:rephs}}
+Before we discuss tests, there is one subtlety when applying Eq.
+\ref{eq:final}.
+The optimal time offset is given by the $\Gamma$ that maximizes the integral.
+However, first integrating over the spatial dimension $\Omega$, we see the phase term at peak time-offset is inevitably frequency dependent.
+This frequency dependence of the phase would lead to destructive interference when we integrate over frequency.
+
+The physical origin of this frequency dependent phase is due to the two visibilities having different phase centers.
+By default the two visibilities are both phased to zenith at the same time.
+When they are cross-correlated, they must be rephased to the account for the movement of the zenith.
+Thus in practice we must rephase the visibilities before delay transform.
+
+In Fig.\ref{fig:freqdiff} we compare two channels 0.16 GHz and 0.17 GHz.
+Top panels show the equivalent baseline pairs sep1,0 against itself, and bottom panels show sep1,0 against sep1,1.
+The left two panels have zero rephasing, and the right two are rephased to a time offset of 0.055 sidereal days, the optimal time offset for the given baseline pair.
+We see that in bottom left panel, although the magnitude, drawn by the wide curves match up, the phases do not.
+This means the wider the frequency profile, the more destructive the interference, and lower the sensitivity contribution.
+In the rephased case, the phases of the near-equivalent case match up.
+We should use the correponding rephasing for each baseline pair maximum sensitivity.
+
+\begin{figure}[H]
+\includegraphics[width=1.1\linewidth]{rephs}
+\label{fig:freqdiff}
+\caption{Comparisons of the peak phases of two different frequencies.
+Top panels shows equivalent baselines, bottom shows a pair of near-equivalents.
+The dot-dashed lines of each color are the imaginary parts, and the solid lines are the real parts.
+For near-equivalent case, we also draw the magnitude by the broad envelops.
+}
+\end{figure}
+
 \section{Analysis}
@@
 To test the robustness of Eq.(\ref{eq:final}), we need to compare the amplitude and phase of the integral
-for a pair of baselines with products of simulated visibilities of those baselines. Without loss of generality, we shall use a single frequency channel of $150$MHz for the following analysis.
+for a pair of baselines with products of simulated visibilities of those baselines.
+For computational simplicity, we shall use a single frequency channel of $150$MHz for the comparison.
+We may use this simplification without compromising generality because Eq.
+(\ref{eq:final}) is valid for a range of frequency channels if an only if it's valid for each single channels, provided that different channels are combined with rephasing as desscribed in Sec.
+\ref{sec:rephs}.
 
@@
 and their conjugate product. Frequency of $\nu=0.15\text{GHz}$ is
-chosen. \label{fig:Beam-fringe-pattern}}
+chosen and only the real components are shown. \label{fig:Beam-fringe-pattern}}
 \end{figure}
@@
 we see for example sep1,0 will have higher redundancy than sep2,0,
-or sep1,1. The latest release of PAPER-64 data uses the 128-equivalent baselines sep2,1,
+or sep1,1. The latest release of PAPER-128 data uses the 128-equivalent baselines sep2,1,
 sep2,0 and sep2,-1 \cite{Ali2015}, and achieved a $2\sigma$ upper
@@
 
-\subsection{Validation with PAPER-128 Data}
-Unlike the simulated clean EOR signal, real data are dominated by foreground and thermal noise, neither of which turns out to exhibit the correlation pattern seen in Fig.
+%\subsection{Notes on Data, Foreground and Noise}
+%Unlike the simulated clean EOR signal, real data are dominated by foreground %and thermal noise, neither of which turns out to exhibit the correlation %pattern seen in Fig.
 \ref{fig:numerics}.
-Foreground exhibits long-range correlations and would lead to high responses at wide range of time-offsets.
-Instrumental noise fo different baselines/antenna-pairs, on the other hand, are uncorrelated and thus exhibit responses consistent with zero at all time offsets.
+Foreground exhibits long-range %correlations and would lead to high responses at wide range of time-offsets.
+%Instrumental noise for different baselines/antenna-pairs, on the other hand, are uncorrelated and thus exhibit responses consistent with zero at all time offsets.
+%To illustrate, we use the second observing season of PAPER-128 data, taken from Jan.
+21st to March 7th of 2014.
+To illustrate the effect of noise and foreground, we use the data before fringe-rate and delay filtering.
 
 \section{Conclusion}
+Redundant arrays are designed to maximize sensitivity.
+Current generations of redundant radio arrays, such as those probing the power spectrum of the epoch of reionization could benefit from data analysis techniques that improve the sensitivity.
+We present an intuitive analysis of cross-multiplying baselines that are close in length and orientation to each other.
+Given an antenna array configuration, our method quickly identifies the best baseline pairs to cross-multiply and predict the expected sensitivity contribution.
+With the predicted result one can improve existing power-spectrum pipelines through 3 simple steps.
+1).
+Rephase the visibilities prior to delay transforming by the zenith displacement (or more accurately by the phase predicted by our numerical analysis).
+2).
+Shift the visibilities in time.
+3) Cross multiply the visibilities of the two baselines  to form the power-spectrum.
+4) Finally combine the different baseline pairs by appropriate inverse-covariance weighting that takes into account the predicted sensitivity contributions of each case.
 
+\appendix
+\section{\\Hera Analysis}
+\begin{widetext}
+\begin{figure}[H]
+\includegraphics[scale=0.55]{HeraAntpos}
+
+\caption{Planned Hydrogen Epoch of Reionization Array antenna configurations\label{fig:HeraAntpos}}
+\end{figure}
+\end{widetext}
 \bibliographystyle{apj}

### 2017-03-17 Rotation paper/draft_working.tex
save

@@
 The extremely natural and simple implementation of this method
-would increase the sensitivity of PAPER-128 by up to [~50] percent, and increase the out-rigger imaging sensitivity of HERA-350 by a factor of 2. 
+would increase the sensitivity of PAPER-128 by about to $10\%$, and increase the out-rigger imaging sensitivity of HERA-350 by a factor of 2. 
 \end{abstract}
@@
 of its own, but also because it can provide crucial information
-to fundamental nature of inflation, neutrino mass and phenomenology
+to fundamental physics of inflation, neutrino mass and phenomenology
 of the first stars and galaxies (e.g. \cite{LiuOpticalDepth}).
@@
 While probes such as quasar spectra rely on emission and scattering
-of free electrons, and are thus limited to the lower redshifts after
-the completion of reionization, the 21cm line directly probes the
+of free electrons, and are thus limited to the lower redshifts towards the end of reionization, the 21cm line directly probes the
 abundance and distribution of neutral hydrogen, and thus would potentially
@@
 the Epoch of Reionization (PAPER) \cite{Ali2015,paper32}, Murchison
-Widefield Array (MWA) \cite{MWA}, Low Frequency Array (LOFAR) \cite{LOFAR},
-and next generation instruments such as the Hydrogen Epoch of Reionization
+Widefield Array (MWA) \cite{MWA}, Low Frequency Array (LOFAR) \cite{LOFAR}. Next generation instruments such as the Hydrogen Epoch of Reionization
 Array (HERA) (e.g. \cite{HERA,HERAconfiguration,HERABEAM1,HERADISH2})
 and the Square Kilometer Array Low (SKA low) (e.g. \cite{SKA1}) are
-under construction.
+in planning and construction.
 
 One of the main challenges of interferometric observations of cosmic
-21cm line is the foreground contamination. Both sources within our
+21cm line is foreground contamination. Both sources within our
 galaxy, and to a lesser distant sources outside of our galaxy emit
@@
 the signal to noise ratio. This is the motivation for the design of
-the maximum redundancy arrays such as PAPER and HERA. While traditional
-imaging arrays focus on Fourier mode uv coverage, the redundant arrays
-focus on redundant measurements of the same Fourier mode
+the maximum redundancy arrays such as PAPER and HERA. Traditional
+imaging arrays are designed to image localized bright sources and thus favor non-redundant patterns for uv coverage. The redundant arrays on the other hand
+focus on high-sensitivity measurements of the same Fourier mode
 with many tightly packed antennas at equal spacing. Since baselines
@@
 \cite{Ali2015} provides the newest upper limit to the power spectrum
-measurements with the PAPER-64 data. From some models we roughly expect
+measurements with the 64-element version of PAPER(henceforth referred to as PAPER-64). From some models we roughly expect
 that the sensitivity required for detection is a factor of 10 away.
@@
 Recently the authors of \cite{wterm} proposed an algorithmic search approach to extract information from all near-equivalent baselines.
-We here propose a simple and elegant theoretical formalism that in addition allows us to identify good baselines to cross-correlate and predict the weight of sensitivity with speed and accuracy.
+We here present a simple and elegant theoretical formalism that in addition allows us to identify good baselines to cross-correlate and predict the weight of sensitivity with speed and accuracy.
 
@@
 2 we explain the theoretical basis for this cross multiplication.
-In section 3 we present details of the numerical implementation of
-this technique. In section 4 we discuss the expected sensitivity improvement
-with this method for HERA and PAPER-128 and with section 5 we conclude.
+In section 3 we present numerical tests of
+this technique as well as the expected sensitivity improvement
+with this method for HERA and PAPER-128 and with section 4 we conclude.
 
@@
 are close enough to induce cross talks, and hence are not suitable
-for use. The original PAPER-64 analysis used three classes of baselines, here
+for use. The original PAPER-64 analysis (\cite{Ali2015}) used three classes of baselines, here
 equivalent to 
@@
 the source, the point traces out tracks in the uv-plane. 
-The uv tracks of PAPER-128 over 4.8 hours, at 0.15GHz, for a source that passes through zenith, are shown in Fig. \ref{fig:Tracks}.
+We show in Fig. \ref{fig:Tracks} uv tracks of PAPER-128 over 4.8 hours, at 0.15GHz, for a source that passes through zenith.
 
@@
 of the uv tracks. Although this is a valid quick method to identify some baselines to cross-multiply, we find
-that it is not accurate enough for time offset determination. The reason is, for wide beam antennas such as PAPER, the sources are not point-like. 
-In other words, when tracks of
-one point of the sky are crossing in uv-plane, those corresponding to another is not. To accurately
-constrain redundancy we must take into account of finite beam width. A more rigorous treatment is thus required. 
+that it is not accurate enough for time offset determination.
+The reason is that UV tracks are traces of a point source, which though adequate for traditional narrow beam imaging, are not applicable for wide-beam receivers like that of PAPER.
+Here the uv-imprint of the sky for a given baseline and frequency is really a cloud that's the track convolved with the appropriate point-spread function. To accurately
+constrain redundancy, a more rigorous treatment is thus required. 
 We show sample beams of HERA and PAPER antennas in Fig.\ref{fig:Beam} for reference.
@@
 \begin{equation}
-\begin{aligned}V_{\nu}(b) & =\int d\Omega A_{\nu}(\hat{s})\phi(\nu)I_{\nu}(\hat{s})\exp\left[-2\pi i\frac{\nu}{c}b\cdot\hat{s}\right],\\
- & \approx\frac{2k_{B}}{\lambda^{2}}\int d\Omega A_{\nu}(\hat{s})\phi(\nu)T(\hat{s})\exp\left[-2\pi i\frac{\nu}{c}b\cdot\hat{s}\right],
+\begin{aligned}V_{\nu}(\boldsymbol{b}) & =\int d\Omega A_{\nu}(\hat{\boldsymbol{s}})\phi(\nu)I_{\nu}(\hat{\boldsymbol{s}})\exp\left[-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right],\\
+ & \approx\frac{2k_{B}}{\lambda^{2}}\int d\Omega A_{\nu}(\hat{\boldsymbol{s}})\phi(\nu)T(\hat{\boldsymbol{s}})\exp\left[-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right],
 \end{aligned}
@@
 related to $T$, the brightness temperature in the Rayleigh-Jeans
-limit. Note that $A$ here is the baseline beam, i.e. product of two antenna beams, examples of which are shown in Fig.\ref{fig:Beam}. $\phi(\nu)$ is the frequency band profile. For PAPER, for example,
-the bands studied are roughly a top-hat from 144MHz to 154 MHz.
+limit. Note that $A$ here is the baseline beam, i.e. product of two antenna beams, examples of which are shown in Fig.\ref{fig:Beam}. $\phi(\nu)$ is the frequency bandpass profile. In practice,
+power spectrum measurements are typically taken from a few ten MHz centered around the corresponding redshift of interest (e.g. 150 MHz for z=9.5).
 
@@
 \begin{equation}
-\begin{aligned}V(b,\tau) & =\int d\nu V_{\nu}(b)\phi(\nu)\exp\left[-2\pi i\nu\tau\right],\\
- & =\frac{2k_{B}}{\lambda^{2}}\int d\Omega d\nu A(\hat{s},\nu)\phi(\nu)T(\hat{s},\nu)\exp\left[-2\pi i\nu\left(\frac{b\cdot\hat{s}}{c}+\tau\right)\right]
+\begin{aligned}V(\boldsymbol{b},\tau) & =\int d\nu V_{\nu}(\boldsymbol{b})\phi(\nu)\exp\left[-2\pi i\nu\tau\right],\\
+ & =\frac{2k_{B}}{\lambda^{2}}\int d\Omega d\nu A(\hat{\boldsymbol{s}},\nu)\phi(\nu)T(\hat{\boldsymbol{s}},\nu)\exp\left[-2\pi i\nu\left(\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}+\tau\right)\right]
 \end{aligned}
@@
 Eq. \eqref{eq:Vb1} expresses the delay-transformed visibility as
-an integral over observation coordinates $\hat{s}$ and $\nu$. Ultimately,
-we would like to relate the data, collected with coordinates $\hat{s}$
+an integral over observation coordinates $\hat{\boldsymbol{s}}$ and $\nu$. Ultimately,
+we would like to relate the data, collected with coordinates $\hat{\boldsymbol{s}}$
 and $\nu$, to the power spectrum, written with cosmological coordinates
@@
 \[
-\begin{aligned}V(b,\tau) & =\frac{1}{X^{2}Y}\int_{H}d^{3}rA(\vec{r})\phi(r)I(\vec{r})\exp\left[-2\pi i\left(\frac{b}{c}\cdot\hat{r}+\tau\right)\nu_{r}\right]\end{aligned}.
+\begin{aligned}V(\boldsymbol{b},\tau) & =\frac{1}{X^{2}Y}\int_{H}d^{3}rA(\boldsymbol{r})\phi(r)I(\boldsymbol{r})\exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\hat{\boldsymbol{r}}+\tau\right)\nu_{r}\right]\end{aligned}.
 \]
 \normalsize
-Here $(r_{x},r_{y},r_{z})=(X\hat{s}_{x},X\hat{s}_{y},Y\nu)$, and
-$(Xk_{x},Xk_{y},Yk_{z})=\frac{2\pi}{c}(b_{x},b_{y},\eta)$ relate
+Here $(r_{x},r_{y},r_{z})=(X\hat{\boldsymbol{s}}_{x},X\hat{\boldsymbol{s}}_{y},Y\nu)$, and
+$(Xk_{x},Xk_{y},Yk_{z})=\frac{2\pi}{c}(b_{x},b_{y},\tau)$ relate
 the cosmological coordinates $r$ and $k$ to the measured coordinates.
@@
 \begin{equation}
-\begin{aligned} & \langle V^{*}(b,\tau)V(b',\tau')\rangle\\
- & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\langle T^{*}(r)T(r')\rangle\right)A^{*}(r)A(\Gamma r')\Phi_{b,\tau}(r,\Gamma r'),\\
- & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\kappa)T(\kappa')\rangle e^{-i(\kappa r-\kappa'r')}\right)A^{*}(r)A(\Gamma r')\Phi_{b,\tau}(r,\Gamma r'),\\
- & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\kappa(r-r')}\right)A^{*}(r)A(\Gamma r')\Phi_{b,\tau}(r,\Gamma r'),\\
- & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\xi(r-r')A^{*}(r)A(\Gamma r')\Phi_{b,\tau}(r,\Gamma r'),\\
- & \approx\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}rd^{3}r'\delta_{D}^{(3)}(r-r')A^{*}(r)A(\Gamma r')\Phi_{b,\tau}(r,\Gamma r'),\\
- & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}r|A^{*}(r)A(\Gamma r)||\phi(\nu_{r})|^{2}\exp\left[-i2\pi\nu_{r}\left(\hat{r}\cdot\frac{b}{c}-\hat{\Gamma r}\cdot\frac{b'}{c}\right)\right],\\
- & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{s},\nu)A(\hat{\Gamma s},\nu)||\phi(\nu)|^{2}\exp\left[-i2\pi\nu\left(\hat{s}\cdot\frac{b}{c}-\hat{\Gamma s}\cdot\frac{b'}{c}\right)\right],
+\begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau')\rangle\\
+ & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle\right)A^{*}(\boldsymbol{r})A(\Gamma r')\Phi_{b,\tau}(\boldsymbol{r},\Gamma r'),\\
+ & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\boldsymbol{\kappa})T(\boldsymbol{\kappa'})\rangle e^{-i(\boldsymbol{\kappa}\cdot \boldsymbol{r}-\boldsymbol{\kappa'}\cdot\boldsymbol{r'})}\right)A^{*}(\boldsymbol{r})A(\Gamma r')\Phi_{b,\tau}(\boldsymbol{r},\Gamma r'),\\
+ & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\boldsymbol{\kappa}\cdot(\boldsymbol{r}-\boldsymbol{r'})}\right)A^{*}(\boldsymbol{r})A(\Gamma r')\Phi_{b,\tau}(\boldsymbol{r},\Gamma r'),\\
+ & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\xi(\boldsymbol{r}-\boldsymbol{r'})A^{*}(\boldsymbol{r})A(\Gamma r')\Phi_{b,\tau}(\boldsymbol{r},\Gamma r'),\\
+ & \approx\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}rd^{3}r'\delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})A^{*}(\boldsymbol{r})A(\Gamma r')\Phi_{b,\tau}(\boldsymbol{r},\Gamma r'),\\
+ & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}r|A^{*}(\boldsymbol{r})A(\Gamma r)||\phi(\nu_{r})|^{2}\exp\left[-i2\pi\nu_{r}\left(\hat{\boldsymbol{r}}\cdot\frac{\boldsymbol{b}}{c}-\hat{\Gamma r}\cdot\frac{\boldsymbol{b'}}{c}\right)\right],\\
+ & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\hat{\Gamma s},\nu)||\phi(\nu)|^{2}\exp\left[-i2\pi\nu\left(\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b}}{c}-\hat{\Gamma s}\cdot\frac{\boldsymbol{b'}}{c}\right)\right],
 \end{aligned}
@@
 
-where in transition from cosmological coordinates back to observing coordinates we have written $\hat{r}\equiv\hat{s}$, and 
+where in transition from cosmological coordinates back to observing coordinates we have written $\hat{\boldsymbol{r}}\equiv\hat{\boldsymbol{s}}$, and 
 \begin{equation}
-\Phi_{b,\tau}(r,\Gamma r')=|{\phi^{*}}(\ensuremath{\nu_{r}})\phi(\nu_{r'})|\exp\left[-i\frac{2\pi}{c}\left(b\cdot\nu_{r}\hat{r}-b'\cdot\nu_{r'}\Gamma\hat{r'}\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)\right].
+\Phi_{b,\tau}(\boldsymbol{r},\Gamma r')=|{\phi^{*}}(\ensuremath{\nu_{r}})\phi(\nu_{r'})|\exp\left[-i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\Gamma\hat{r'}\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)\right].
 \end{equation}
@@
 outside the parenthesis. Notice that the phase factor $\exp\left[-i2\pi\tau\left(\nu-\nu'\right)\right]$
-drops out in the end. This is significant because this means that correlation is peaked at the same time-offset for all delay channels.
+drops out in the end. This means that correlation is peaked at the same time-offset for all delay channels.
 
-Finally, since the beam pattern and bandwidth are given in $\hat{s}$
+Finally, since the beam pattern and bandwidth are given in $\hat{\boldsymbol{s}}$
 and $\nu$, we convert the integral back to these coordinates to get
@@
 \begin{equation}
-\begin{aligned} & \langle V^{*}(b,\tau)V(b',\tau')\rangle\\
- & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{s},\nu)A(\hat{\Gamma s},\nu)||\phi(\nu)|^{2}\\
- & \qquad \qquad \qquad \qquad \exp\left[-i2\pi\nu\left(\hat{s}\cdot\frac{b}{c}-\hat{\Gamma s}\cdot\frac{b'}{c}\right)\right].\end{aligned}
+\begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau')\rangle\\
+ & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\hat{\Gamma s},\nu)||\phi(\nu)|^{2}\\
+ & \qquad \qquad \qquad \qquad \exp\left[-i2\pi\nu\left(\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b}}{c}-\hat{\Gamma s}\cdot\frac{\boldsymbol{b'}}{c}\right)\right].\end{aligned}
 \label{eq:final}
 \end{equation}
+
+In other words the power spectrum estimate from visibilities of a baseline pair is given by 
+\begin{equation}
+ P(k_{b,\tau}) = \left(\frac{\lambda^{2}}{2k_{B}}\right)^{2} \frac{\langle V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau')\rangle}{\Theta}, 
+\end{equation}
+where the weight
+\begin{equation}
+\Theta =\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\hat{\Gamma s},\nu)||\phi(\nu)|^{2} e^{-2\pi i\nu\left(\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b}}{c}-\hat{\Gamma s}\cdot\frac{\boldsymbol{b'}}{c}\right)}.
+\label{eq:opp}
+\end{equation}
+
 So roughly speaking the cross multiplications of visibilities at a time delay
@@
 the phase and normalization. As a check, when applied to equivalent baselines,
-$b=b'$, $\hat{s}=\hat{\Gamma s}$, and Eq.(\ref{eq:final}) reduces to Eq.(B9) of \cite{paper32}.
+$\boldsymbol{b}=\boldsymbol{b'}$, $\hat{\boldsymbol{s}}=\hat{\Gamma s}$, and Eq.(\ref{eq:final}) reduces to Eq.(B9) of \cite{paper32}.
 
 If this result is robust, then we can achieve all our goals stated in the introduction, i.e. to identify 
-candidate baseline pairs, to find the optimal time offset, and to quantify the sensibility, by computing the integral
-in Eq.(\ref{eq:final}) for various time offsets. Thus we shall first test its robustness numerically.
+candidate baseline pairs, to find the optimal time offset, and to quantify the sensibility, by computing the weight $\Theta$ from
+Eq.(\ref{eq:opp}) for various time offsets. Thus we shall first test its robustness numerically.
 
@@
 \ref{eq:final}.
-The optimal time offset is given by the $\Gamma$ that maximizes the integral.
-However, first integrating over the spatial dimension $\Omega$, we see the phase term at peak time-offset is inevitably frequency dependent.
+The optimal time offset is given by the $\Gamma$ that maximizes the absolute value of the weight $\Theta$.
+However, $\Theta$ is in general complex.
+First integrating over the spatial dimension $\Omega$, we see the phase term at peak time-offset is inevitably frequency dependent.
 This frequency dependence of the phase would lead to destructive interference when we integrate over frequency.
 
-The physical origin of this frequency dependent phase is due to the two visibilities having different phase centers.
+The physical origin of this frequency dependent phase lies the two visibilities having different phase centers.
 By default the two visibilities are both phased to zenith at the same time.
@@
 The left two panels have zero rephasing, and the right two are rephased to a time offset of 0.055 sidereal days, the optimal time offset for the given baseline pair.
-We see that in bottom left panel, although the magnitude, drawn by the wide curves match up, the phases do not.
+We see that in bottom left panel, although the magnitude, drawn by the wide curves match up for the two frequencies, the phases do not.
 This means the wider the frequency profile, the more destructive the interference, and lower the sensitivity contribution.
-In the rephased case, the phases of the near-equivalent case match up.
-We should use the correponding rephasing for each baseline pair maximum sensitivity.
+In the rephased case, the phases of the near-equivalent case match up and can be added without compromising sensitivity.
+We should thus rephase the data separately for each set of baseline pairs.
 
@@
 
-To test the robustness of Eq.(\ref{eq:final}), we need to compare the amplitude and phase of the integral
+To test the robustness of Eq.(\ref{eq:final}), we need to compare the amplitude and phase of the integral weight $\Theta$ 
 for a pair of baselines with products of simulated visibilities of those baselines.
@@
 
-To test our understanding of the time-offset, and to verify our derivation
-in Eq. \eqref{eq:main}, we must compare the simulated visibility cross multiplication with the integral in
-Eq. \eqref{eq:main}, for a pair of near-equivalent baselines. 
-This comparison is show in Fig. \ref{fig:numerics}. 
+The comparison is show in Fig. \ref{fig:numerics}. 
 Since the sensitivity contribution is inferred from height of the peak of the cross-multiplied visibilities as a function of time. 
-We first compare the simulation to the integral for a pair of equivalent baselines, then normalize the peaks 
+We first compare the simulation to the analytical weight $\Theta$ for a pair of equivalent baselines, then normalize the peaks 
 of the near-equivalent baselines to that of the equivalent ones. 
@@
 \begin{figure}[H]
-\includegraphics[width=\linewidth]{redun_agree}
+\includegraphics[width=\linewidth]{corr}
 \label{fig:numerics}
 \caption{Numerical comparisons of the visibility correlation peaks to the $\Omega$ factor in Eq.(\ref{eq:final}).
-We generated 24000 instances of Gaussian random sky on healpix maps, computed visibilities and cross correlated them to find the correlation.
-We see that the theory and simulation line up in both amplitude and phase.
-}
+We generated 2400 instances of Gaussian random sky on healpix maps, computed visibilities and cross correlated them to find the correlation.
+Top panel shows the equivalent baseline pairs sep1,0 against itselt, bottom show set1,0:1:1.
+Evaluations of weight $\Theta$ are shown with solid curves, and the visibility correlations of simulated random sky is given with dashed lines.
+In both cases blue denotes real part, green the imaginary part, and red the magnitude.
+We see that in both cases the theory and simulation line up in both amplitude and phase.}
 \end{figure}
@@
 
-\begin{widetext}
+%\begin{widetext}
 \begin{figure}[H]
-\includegraphics[scale=0.5]{fringe_res}
-
+%\includegraphics[scale=0.5]{fringe_res}
+\includegraphics[width=\linewidth]{fringe_res}
 \caption{Beam fringe pattern of sep2,0(left), sep2,1 at a time delay (middle),
@@
 \end{figure}
-\end{widetext}
+%\end{widetext}
 
@@
 
-Having verified Eq. \eqref{eq:main}, we can thus predict the sensitivity contributions of a
+Having verified Eq. \eqref{eq:final}, we can thus predict the sensitivity contributions of a
 particular baseline pair simply by computing the integral.
@@
 time-offset corresponds to smaller differences in orientation, and hence in length of
-baselines in PAPER-64.
+baselines in PAPER.
 
@@
 \label{fig:sensplot}
-\caption{Relative sensitivity contributions of selected baseline combinations in PAPER-64. In the legend m,n:p,q denotes cross
-multiplying PAPER-64 baselines of east-west, north-south separations (m,n) and (p,q) respectively. The top
+\caption{Relative sensitivity contributions of selected baseline combinations in PAPER-128. In the legend m,n:p,q denotes cross
+multiplying PAPER-128 baselines of east-west, north-south separations (m,n) and (p,q) respectively. The top
 panel shows the peak height (degree of correlation) of each baseline
@@
 
+Finally we combine the different power spectrum measurements by inverse covariance weighting:
+\begin{equation}
+ P(k_{\tau}) = \frac{\sum_{bb'}P(k_{b,\tau})\Theta_{bb'}^2}{\sum_{bb'}\Theta_{bb'}^2}.
+\end{equation}
+So the total noise covariance scales as 
+\begin{equation}
+\sigma_N^2 \sim \frac{\sum_{bb'}\Theta_{bb'}}{\sum_{bb'}\Theta_{bb'}^2}.
+\end{equation}
+With the baseline pairs show in Fig. \ref{fig:sensplot}, we see we get about $10\%$ improvement.
+
 %\subsection{Notes on Data, Foreground and Noise}

### 2017-03-18 Rotation paper/draft_working.tex
save

@@
 the baselines labeled with the yellow and green arrows are examples
-of sep1,1 and sep-1,1, respectively. Note sep1,0 and sep-1,0 for example
+of sep1,1 and sep1,-1, respectively. Note sep1,0 and sep-1,0 for example
 are the same baselines and should not be counted twice.}
@@
 unit east and 0 unit north. Similarly, the baselines labeled with
-the yellow and green arrows are examples of sep1,1 and sep-1,1, respectively.
+the yellow and green arrows are examples of sep1,1 and sep1,-1, respectively.
 Note sep1,0 and sep-1,0 for example are the same baselines and should
@@
 equivalent to 
-sep2,0, sep2,1 and sep-2,1 \cite{Ali2015} of PAPER-128. There each of these classes
+sep2,0, sep2,1 and sep2,-1 \cite{Ali2015} of PAPER-128. There each of these classes
 of baselines are cross multiplied to itself. We shall see that in addition these
@@
  P(k_{b,\tau}) = \left(\frac{\lambda^{2}}{2k_{B}}\right)^{2} \frac{\langle V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau')\rangle}{\Theta}, 
+ \label{eq:opp}
 \end{equation}
@@
 \Theta =\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\hat{\Gamma s},\nu)||\phi(\nu)|^{2} e^{-2\pi i\nu\left(\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b}}{c}-\hat{\Gamma s}\cdot\frac{\boldsymbol{b'}}{c}\right)}.
-\label{eq:opp}
+\label{eq:Theta}
 \end{equation}
@@
 candidate baseline pairs, to find the optimal time offset, and to quantify the sensibility, by computing the weight $\Theta$ from
-Eq.(\ref{eq:opp}) for various time offsets. Thus we shall first test its robustness numerically.
+Eq.(\ref{eq:Theta}) for various time offsets. Thus we shall first test its robustness numerically.
 
@@
 
+\begin{widetext}
 \begin{figure}[H]
-\includegraphics[width=\linewidth]{sensitivity}
+\includegraphics[scale=0.5]{sensitivity}
 \label{fig:sensplot}
@@
 combination, while the bottom panel multiplies the heights by the
-corresponding multiplicities as in Eq. \eqref{eq:sensul}, and in some cases an extra factor of 2 (explained in text). 
+corresponding multiplicities as in Eq. \eqref{eq:sensul}, and in some cases an extra factor of $\sqrt{2}$ (explained in text). 
 In weighting by the multiplicity (bottom), we have chosen to fix the sensitivity 
 contribution of 1,0:1,0 to unity. }
-
 \end{figure}
+\end{widetext}
 
@@
 same length and orientation. Looking at Fig. \ref{fig:sensplot}
-we see for example sep1,0 will have higher redundancy than sep2,0,
-or sep1,1. The latest release of PAPER-128 data uses the 128-equivalent baselines sep2,1,
+we see for example sep1,0 will have higher multiplicity than sep2,0,
+or sep1,1. The latest release of PAPER-64 data uses the 128-equivalent baselines sep2,1,
 sep2,0 and sep2,-1 \cite{Ali2015}, and achieved a $2\sigma$ upper
@@
 
-First we can average of the visibilities of the baselines of the same
-separations. Since the core of PAPER-128 has 16 by 7 antenna configuration, there
-are $(16-|m|)\times(7-|n|)$ copies of the baseline sepm,n. This means
+First we can average of the visibilities of the equivalent baselines. Since the core of PAPER-128 has 16 by 7 antenna configuration, there
+are $M=(16-|m|)\times(7-|n|)$ copies of the baseline sepm,n. This means
 that if we add visibility measurements of all these equivalent baselines,
-we get a factor of $\sqrt{(16-|m|)\times(7-|n|)}$ reduction in noise
-level. The sensitivity contribution of sepm,n, being the signal to
-noise ratio of the power spectrum, is thus proportional to $(16-|m|)\times(7-|n|)$.
-Next we can average over the power-spectrum taken from baselines of
-the different separations. Thus the total cumulative sensitivity is
-given by 
+we get a factor of $\sqrt{M}$ reduction in noise
+level $\sigma_N$ of the visibility.
+The sensitivity contribution of sepm,n, cross multiplied with sepm',n'  thus roughly speaking scales as $\sqrt{\left((16-|m|)(7-|n|)(16-|m'|)(7-|n'|)\right)}=\sqrt{MM'}$.
+For cross-multiplications of near-equivalent baselines of
+types sepm,n and sepm',n', we get a effective weight:
+
 \begin{equation}
-\frac{P_{S}}{P_{N}}\propto\sqrt{\sum_{(n,m)}\left((16-|m|)\times(7-|n|)\right)^{2}},\label{eq:sensequiv}
+\widetilde{\Theta}_{bb'} \propto \Theta_{bb'}\times\sqrt{MM'}.\label{eq:sensul}
 \end{equation}
-where the sum is only over pairs of m and n. If we now include power
-spectrum from cross-multiplications of near-equivalent baselines of
-types sepm,n and sepm',n', we get a general cumulative sensitivity:
-\small
+
+We are being sloppy here for simplicity.
+As we derive in Appendix \ref{sec:appB}, this weight should be corrected by a factor proportional to the signal to noise ratio of a single-baseline visibility $\rho_0=\sigma_S^2/\sigma_N^2$:
+
 \begin{equation}
-\frac{P_{S}}{P_{N}}\propto\sqrt{\sum_{(n,m,n',m')}\Omega_{mnm'n'}\left((16-|m|)(7-|n|)(16-|m'|)(7-|n'|)\right)},\label{eq:sensul}
+\label{eq:tildereal}
+\widetilde{\Theta}_{bb'}=\frac{\Theta_{bb'}\sqrt{M_bM_{b'}}}{\sqrt{1 + \rho_0 \left(M_b+M_{b'} \right)}}.
 \end{equation}
-\normalsize
-where $\Omega$ is the correlation peak height show in top panel of Fig. \eqref{fig:sensplot},
-and the sum is again only over tuples of m,n,m',n'. This form now
-includes the equivalent baselines in Eq. \eqref{eq:sensequiv}, with
-$\Omega_{m=m',n=n'}=1$. 
+
 Shown in the bottom panel of Fig. \eqref{fig:sensplot} is the peak heights weighted
-by this multiplicity factor (before taking the square root), with one difference. The difference is, 
-since we have "folded over" the negative time delays, we have combined baseline pairs that are identical modulus parity, 
-or in other words are mirror images of each other, by summing over them to get an extra factor of two. For example, 
+by the multiplicity factor.
+Points that have zero time delay are the equivalent baseline pairs and their weighted correlation values simply reflect the multiplicity factor.
+For clarity of presentation we have "folded over" the negative time delays and combined baseline pairs that are identical modulus parity,
+or in other words are mirror images of each other, by summing over them to get an extra factor of $\sqrt{2}$. For example, 
 instead of plotting 1,1:1,1 and 1,-1:1,-1 separately, we plot 1,1:1,1 with twice the multiplicity. Similarly, baseline pairs such as 
 1,0:1,1 also get the factor of 2 because they are identical to 1,0:1,-1. Baseline pairs such as 1,0:1,0, or 1,1:1,-1 do not get
-the factor of 2 because their mirror images are themselves.
+the factor of $\sqrt{2}$ because their mirror images are themselves.
 
-Finally we combine the different power spectrum measurements by inverse covariance weighting:
+Finally we estimate the power spectrum by inverse covariance weighting:
 \begin{equation}
- P(k_{\tau}) = \frac{\sum_{bb'}P(k_{b,\tau})\Theta_{bb'}^2}{\sum_{bb'}\Theta_{bb'}^2}.
+\begin{aligned}
+ P(k_{\tau}) &= \frac{\sum_{bb'}P(k_{b,\tau})/\sigma_P^2(bb')}{\sum_{bb'}\sigma_P^2(bb')}, \\
+ &= \frac{\sum_{bb'}P(k_{b,\tau})\widetilde{\Theta}_{bb'}^2}{\sum_{bb'}\widetilde{\Theta}_{bb'}^2},
+ \end{aligned}
 \end{equation}
-So the total noise covariance scales as 
+where the sum is over classes of baseline pairs. 
+Thus the total signal to noise ratio is given by
 \begin{equation}
-\sigma_N^2 \sim \frac{\sum_{bb'}\Theta_{bb'}}{\sum_{bb'}\Theta_{bb'}^2}.
+\rho \propto \frac{\sum_{bb'}^N\widetilde{\Theta}_{bb'}^2}{\sum_{bb'}^N\widetilde{\Theta}_{bb'}},
 \end{equation}
-With the baseline pairs show in Fig. \ref{fig:sensplot}, we see we get about $10\%$ improvement.
+
+where $N$ is the total classes of baseline pairs used.
+When all the $\widetilde{\Theta}$ are unity, $\rho$ improves by $1/\sqrt{N}$ scaling.
+When $\Theta<1$ it's clear that adding more baselines is not always going to lead to improvement.
+Thus for given antenna array we should restrict antenna pairs with $\widetilde{\Theta}$ above a certain $\widetilde{\Theta}_{min}$ for maximum sensitivity.
+
+To move forward we must give an estimate of $\rho_0$ in Eq.\ref{eq:tildereal}.
 
@@
 
-\caption{Planned Hydrogen Epoch of Reionization Array antenna configurations\label{fig:HeraAntpos}}
+\caption{Planned Hydrogen Epoch of Reionization Array antenna configurations. For HERA350, only the 320 elements in the core are shown. \label{fig:HeraAntpos}}
 \end{figure}
 \end{widetext}
+
+\section{\\Derivation of Noise Covariance \label{sec:appB}}
+
+Now to be a little more careful about the scaling relations above, we combine the different power spectrum measurements by inverse covariance weighting \footnote{In practice techniques such as bootstrapping is often used, see for example \citep{Ali2015}}.
+We shall separate the visibility and power spectrum into signal and noise contributions:
+\begin{equation}
+\begin{aligned}
+V &= V_S+V_N,\\
+P &= P_S+P_N.
+\end{aligned}
+\end{equation}
+We shall denote the noise covariance of power spectrum and visibility
+\begin{equation}
+\begin{aligned}
+\sigma_V^2 &= \langle |V_N|^2 \rangle,\\
+\sigma_P^2 &= \langle P_N^2 \rangle.
+\end{aligned}
+\end{equation}
+One may notice that we have used a single covariance for the complex quantity visibility.
+It's simple to show that the same result holds if we use a separate real and imaginary components, as long as they are independent of each other.
+In fact, for simplicity and without loss of generality we shall treat the visibility as a real quantity in the rest of this derivation.
+Note that though we can assume $\langle V_N^{odd-power}\rangle=0$, the same is not true for $P_N$.
+
+Then the variance of $P$ constructed with visibilities $V_1$ and $V_2$ from two baseline classes can be estimated\footnote{We assume all noise terms to be independent for simplicity, in practice the correlation of different measurements ifrom equivalent baselines are alleviated by grouping the baselines in the class and the days of observation, as in \cite{Ali2015}}:
+\begin{equation}
+\begin{aligned}
+\sigma_P^2 &= \langle P^2\rangle -\langle P \rangle^2,\\
+&\propto \langle \frac{(V_{1S}+V_{1N})^2 (V_{2S}+V_{2N})^2}{\Theta^2} \rangle - \langle \frac{(V_{1S}+V_{1N}) (V_{2S}+V_{2N})}{\Theta} \rangle ^2,\\
+&= \frac{1}{\Theta^2} \left( V_{1S}^2\sigma_{V2}^2+V_{2S}^2\sigma_{V1}^2+\langle V_{1N}^2 V_{2N}^2\rangle\right), \\
+&= \frac{1}{\Theta^2} \left[ V_{S}^2(\sigma_{V2}^2+\sigma_{V1}^2) + \sigma_{V1}^2 \sigma_{V2}^2\right], 
+\end{aligned}
+\end{equation}
+
+where in the second last line we have substituted visibility noise covariance. In the final line we used Wick's theorem and the fact that the signal from two visibilities are equal.
+
+Recall from the discussion on multiplicities we can write
+\begin{equation}
+\sigma_V^2=\frac{\sigma_0^2}{M},
+\end{equation}
+where $\sigma_0$ is some single-baseline noise level. Letting $\rho_0=V_S^2/\sigma_0^2$ be the signal to noise ratio for a single baseline, we can write
+
+\begin{equation}
+\begin{aligned}
+\sigma_P^2 & \propto  \frac{\sigma_0^4}{\Theta^2} \left[ \rho_0 \left(\frac{1}{M_1}+\frac{1}{M_2} \right) + \frac{1}{M_1 M_2}\right], \\
+&\propto \frac{1}{\widetilde{\Theta}_{12}^2},
+\end{aligned}
+\end{equation}
+where we have defined a slightly modified version of the effective weight (compare with Eq. \ref{eq:sensul}):
+\begin{equation}
+\widetilde{\Theta}_{12}=\frac{\Theta_{12}\sqrt{M_1M_{2}}}{\sqrt{1 + \rho_0 \left(M_1+M_{2} \right)}}.
+\end{equation}
 \bibliographystyle{apj}

### 2017-03-19 Rotation paper/draft_working.tex
save

@@
 The extremely natural and simple implementation of this method
-would increase the sensitivity of PAPER-128 by about to $10\%$, and increase the out-rigger imaging sensitivity of HERA-350 by a factor of 2. 
+would increase the sensitivity of PAPER-128 by about to $30\%$, and increase the out-rigger imaging sensitivity of HERA-350 by a factor of 2. 
 \end{abstract}
@@
 In Fig.\ref{fig:freqdiff} we compare two channels 0.16 GHz and 0.17 GHz.
-Top panels show the equivalent baseline pairs sep1,0 against itself, and bottom panels show sep1,0 against sep1,1.
-The left two panels have zero rephasing, and the right two are rephased to a time offset of 0.055 sidereal days, the optimal time offset for the given baseline pair.
-We see that in bottom left panel, although the magnitude, drawn by the wide curves match up for the two frequencies, the phases do not.
-This means the wider the frequency profile, the more destructive the interference, and lower the sensitivity contribution.
+The top two panels have zero rephasing, and the bottom two are rephased to a time offset of 0.055 sidereal days, the optimal time offset for the given baseline pair.
+The first and third panels show the equivalent baseline pairs sep1,0 against itself, and second and fourth panels show sep1,0 against sep1,1.
+We see that in the un-rephased case in the second panel, although the magnitude of correlations match up for the two frequencies, the phases do not.
+This means summing over frequency leads to destructive interference and signal loss.
+The wider the frequency profile, the more destructive the interference would be.
 In the rephased case, the phases of the near-equivalent case match up and can be added without compromising sensitivity.
@@
 \caption{Comparisons of the peak phases of two different frequencies.
-Top panels shows equivalent baselines, bottom shows a pair of near-equivalents.
-The dot-dashed lines of each color are the imaginary parts, and the solid lines are the real parts.
-For near-equivalent case, we also draw the magnitude by the broad envelops.
+First and third panels shows equivalent baselines, second and fourth show a pair of near-equivalents.
+For visual simplicity only the real parts are shown.
 }
@@
 \begin{figure}[H]
-\includegraphics[width=\linewidth]{corr}
+\includegraphics[width=\linewidth]{opp}
 \label{fig:numerics}
@@
 
-%\begin{widetext}
+\begin{widetext}
 \begin{figure}[H]
-%\includegraphics[scale=0.5]{fringe_res}
-\includegraphics[width=\linewidth]{fringe_res}
+\includegraphics[scale=0.5]{fringe_res}
+%\includegraphics[width=\linewidth]{fringe_res}
 \caption{Beam fringe pattern of sep2,0(left), sep2,1 at a time delay (middle),
@@
 \end{figure}
-%\end{widetext}
+\end{widetext}
 
-\subsection{Expected Sensitivity Contributions}
+\subsection{Expected Sensitivity Contributions \label{sec:sensitivity}}
 
@@
 
-In the top panel of Fig. \eqref{fig:sensplot} below
+In the top panel of Fig. \ref{fig:sensplot}
 we show the peak heights and locations for a variety of baseline combinations.
@@
 \begin{figure}[H]
-\includegraphics[scale=0.5]{sensitivity}
+\includegraphics[scale=0.6]{sensitivity}
 \label{fig:sensplot}
@@
 In weighting by the multiplicity (bottom), we have chosen to fix the sensitivity 
-contribution of 1,0:1,0 to unity. }
+contribution of 1,0:1,0 to unity. The same symbols and colors correspond in the top and bottom panel, with half of the legend shown in each. }
 \end{figure}
@@
 sep2,0 and sep2,-1 \cite{Ali2015}, and achieved a $2\sigma$ upper
-limit of $22.4mK^{2}$. There, the three sets of equivalent baselines
+limit of $(22.4mK)^{2}$. There, the three sets of equivalent baselines
 are only cross multiplied by itself. Assuming that each baseline delivers
@@
 For cross-multiplications of near-equivalent baselines of
-types sepm,n and sepm',n', we get a effective weight:
+types sepm,n and sepm',n', we get an effective weight:
 
@@
 
-We are being sloppy here for simplicity.
-As we derive in Appendix \ref{sec:appB}, this weight should be corrected by a factor proportional to the signal to noise ratio of a single-baseline visibility $\rho_0=\sigma_S^2/\sigma_N^2$:
-
-\begin{equation}
-\label{eq:tildereal}
-\widetilde{\Theta}_{bb'}=\frac{\Theta_{bb'}\sqrt{M_bM_{b'}}}{\sqrt{1 + \rho_0 \left(M_b+M_{b'} \right)}}.
-\end{equation}
-
-Shown in the bottom panel of Fig. \eqref{fig:sensplot} is the peak heights weighted
+Shown in the bottom panel of Fig. \ref{fig:sensplot} is the peak heights weighted
 by the multiplicity factor.
@@
 where the sum is over classes of baseline pairs. 
-Thus the total signal to noise ratio is given by
+We define the estimator sensitivity to be the inverse of the power-spectrum noise variance:
 \begin{equation}
-\rho \propto \frac{\sum_{bb'}^N\widetilde{\Theta}_{bb'}^2}{\sum_{bb'}^N\widetilde{\Theta}_{bb'}},
+\rho \propto \frac{1}{\sigma_P^2} = \sum_{bb'}^N\widetilde{\Theta}^2_{bb'},
 \end{equation}
 
-where $N$ is the total classes of baseline pairs used.
-When all the $\widetilde{\Theta}$ are unity, $\rho$ improves by $1/\sqrt{N}$ scaling.
-When $\Theta<1$ it's clear that adding more baselines is not always going to lead to improvement.
-Thus for given antenna array we should restrict antenna pairs with $\widetilde{\Theta}$ above a certain $\widetilde{\Theta}_{min}$ for maximum sensitivity.
+So far we have been sloppy in Eq.
+\ref{eq:sensul} for simplicity.
+As we derive in Appendix \ref{sec:appB}, this weight should be corrected by a factor proportional to the signal to noise ratio of a single-baseline visibility $\rho_0=\sigma_S^2/\sigma_N^2$:
+
+\begin{equation}
+\label{eq:tildereal}
+\widetilde{\Theta}_{bb'}=\frac{\Theta_{bb'}\sqrt{M_bM_{b'}}}{\sqrt{1 + \rho_0 \left(M_b+M_{b'} \right)}}.
+\end{equation}
 
 To move forward we must give an estimate of $\rho_0$ in Eq.\ref{eq:tildereal}.
+Assuming a reionization signal of $\Delta_{21cm}^2\sim 30mK^2$, observation at 150MHz ($z=8.5$), 120 days of observation with PAPER antennas, we have roughly
+(See Eq.(20) in \cite{first-paper})
+\begin{equation}
+\rho_0 \sim 0.001\left[\frac{L}{40m}\right] \left[\frac{0.1hMpc^{-1}}{k}\right]^3, 
+\end{equation}
+where $L$ is the baseline length.
+We assumed for simplicity only a single $L$ even in the near-equivalent case because only baselines of nearly equal length would have high redundancy (see Fig.\ref{fig:pairplot}).
 
+As expected, baseline-pairs that have smaller $\widetilde{\Theta}$ contribute less to the sensitivity.
+To further quantify the sensitivity contribution of a given array we run our algorithm over all possible baseline-pairs of three arrays, PAPER128, HERA128 and HERA243.
+In Fig.
+\ref{fig:osens} we plot $\rho$ against the minimum $\widetilde{\Theta}$.
+$\rho(\widetilde{\Theta}_{min})$ is thus the sensitivity of the array when baseline-pairs that have $\widetilde{\Theta}>\widetilde{\Theta}_{min}$ are included.
+The dashed lines represent the values when only the equivalent baseline-pairs are used.
+We see that in all cases using the near-equivalent baselines lead to more and more significant improvements with lower $\widetilde{\Theta_{min}}$, or in other words when worse baseline pairs are used.
+The maximum benefits are expected to be around $30\%$.
+Note that here we normalized $\rho$ such that the contribution of the top single baseline equivalent-pair class, such as (sep0,1:0,1 in the PAPER128 case) are 1.
+This plot therefore does not compare sensitivity across the different arrays.
+The stepwise pattern in the PAPER128 case is characteristic of the rectangular grid; as we step to lower $\widetilde{\Theta}$ large groups of baseline pair classes get included in "batch".
+
+In Fig.
+\ref{fig:pairplot} we present a comparison of 4 different properties for baselines that contribute well to the sensitivity ($\widetilde{\Theta}>0.1$, where again the weight for the top class of equivalent pair is normalized to 1).
+The $dT$ vs.
+$\Theta$ and $dT$ vs.
+$\widetilde{\Theta}$ plots are familiar from Fig.
+\ref{fig:sensplot}.
+Interestingly the near-equivalent baselines form a "U" shape in $dT, \Theta$ plane [TODO].
+Note these near equivalent baselines have around the same $\Theta$ as in PAPER128, but much lower $\widetilde{\Theta}$.
+This is because they are longer baselines with lower multiplicity.
+
+%\begin{centered}
+\begin{figure}[H]
+\includegraphics[width=\linewidth]{osens}
+\label{fig:osens}
+\caption{Sensitivity of redundant arrays as a function of the minimum effective weight.
+Dashed lines represent when only equivalent baseline pairs are used, while solid lines indicate use of both the equivalent and near-equivalent baselines are used.
+The y-axis is normalized independently for each array such that the contribution of the top single equivalent pair class is unity, and thus does not indicate a comparison across the different arrays shown.
+}
+\end{figure}
+%\end{centered}
+
+\begin{widetext}
+\begin{figure}[H]
+\includegraphics[scale=0.5]{pairplot}
+\label{fig:pairplot}
+\caption{Pairplots of the top contributing baseline pairs in three arrays.
+Plotted properties are optimal time delay $dT$, peak height $\Theta$, effective weight $\widetilde{\Theta}$ and baseline length $L$.
+Only those points with $\widetilde{\Theta}>0.1$ are shown (again the weight for the top class of equivalent pair is normalized to 1).
+Only one baseline length is shown since all top contributing pairs have very similar lengths as expected.}
+\end{figure}
+\end{widetext}
 %\subsection{Notes on Data, Foreground and Noise}
@@
 
-Now to be a little more careful about the scaling relations above, we combine the different power spectrum measurements by inverse covariance weighting \footnote{In practice techniques such as bootstrapping is often used, see for example \citep{Ali2015}}.
+In this appendix we give a brief derivation of the effective weight quoted in \ref{sec:sensitivity}.
+We combine the different power spectrum measurements by inverse variance weighting \footnote{In practice techniques such as bootstrapping is often used, see for example \citep{Ali2015}}.
 We shall separate the visibility and power spectrum into signal and noise contributions:
@@
 \end{equation}
-We shall denote the noise covariance of power spectrum and visibility
+We shall denote the noise variance of power spectrum and visibility
 \begin{equation}

### 2017-03-20 Rotation paper/draft_working.tex
save

@@
 
+\begin{figure}[h]
+\includegraphics[width=\linewidth]{opp}
+\label{fig:numerics}
+\caption{Numerical comparisons of the visibility correlation peaks to the $\Omega$ factor in Eq.(\ref{eq:final}).
+We generated 2400 instances of Gaussian random sky on healpix maps, computed visibilities and cross correlated them to find the correlation.
+Top panel shows the equivalent baseline pairs sep1,0 against itselt, bottom show set1,0:1:1.
+Evaluations of weight $\Theta$ are shown with solid curves, and the visibility correlations of simulated random sky is given with dashed lines.
+In both cases blue denotes real part, green the imaginary part, and red the magnitude.
+We see that in both cases the theory and simulation line up in both amplitude and phase.}
+\end{figure}
+
 The comparison is show in Fig. \ref{fig:numerics}. 
@@
 The height of the plot thus quantifies the added sensitivity. We see at a peak of around 0.055 sidereal days, or 1.32 hours, the two baselines are maximally redundant.
-
-\begin{figure}[H]
-\includegraphics[width=\linewidth]{opp}
-\label{fig:numerics}
-\caption{Numerical comparisons of the visibility correlation peaks to the $\Omega$ factor in Eq.(\ref{eq:final}).
-We generated 2400 instances of Gaussian random sky on healpix maps, computed visibilities and cross correlated them to find the correlation.
-Top panel shows the equivalent baseline pairs sep1,0 against itselt, bottom show set1,0:1:1.
-Evaluations of weight $\Theta$ are shown with solid curves, and the visibility correlations of simulated random sky is given with dashed lines.
-In both cases blue denotes real part, green the imaginary part, and red the magnitude.
-We see that in both cases the theory and simulation line up in both amplitude and phase.}
-\end{figure}
 
@@
 
-\subsection{Expected Sensitivity Contributions \label{sec:sensitivity}}
+\subsection{ Sensitivity \label{sec:sensitivity}}
 
@@
 \begin{figure}[H]
-\includegraphics[scale=0.6]{sensitivity}
+\centering
+\includegraphics[width=\textwidth]{sensitivity}
 \label{fig:sensplot}
+
 \caption{Relative sensitivity contributions of selected baseline combinations in PAPER-128. In the legend m,n:p,q denotes cross
@@
 where $L$ is the baseline length.
-We assumed for simplicity only a single $L$ even in the near-equivalent case because only baselines of nearly equal length would have high redundancy (see Fig.\ref{fig:pairplot}).
+We only need a single characteristic $L$ even in the near-equivalent case because only baselines of nearly equal length would have high redundancy.As expected, baseline-pairs that have smaller $\widetilde{\Theta}$ contribute less to the sensitivity.
 
-As expected, baseline-pairs that have smaller $\widetilde{\Theta}$ contribute less to the sensitivity.
-To further quantify the sensitivity contribution of a given array we run our algorithm over all possible baseline-pairs of three arrays, PAPER128, HERA128 and HERA243.
+\subsection{Array Configuration Comparisons \label{sec:arrconf}}
+To further quantify the sensitivity contribution of a given array we run our algorithm over all possible baseline-pairs of  PAPER128, HERA37, HERA128, HERA243 and HERA350.
+The HERA antenna configurations are shown in Fig.
+\ref{fig: HeraAntpos}.The
+hexagonal design is the densest pattern of antenna-packing.
+The lack of short baselines that are visually close to each other means that we expect to see only longer near-equivalent baselines.
+These longer baselines have lower multiplicities, but also there would be many more classes of baselines, especially given the gap in the larger versions.
 In Fig.
@@
 The dashed lines represent the values when only the equivalent baseline-pairs are used.
-We see that in all cases using the near-equivalent baselines lead to more and more significant improvements with lower $\widetilde{\Theta_{min}}$, or in other words when worse baseline pairs are used.
-The maximum benefits are expected to be around $30\%$.
+We see that in all cases using the near-equivalent baselines lead to more and more significant improvements with lower $\widetilde{\Theta}_{min}$, or in other words when worse baseline pairs are used.
+The small HERA37, with no gap (like in HERA350) or short near-equivalent baselines (like in PAPER 128), will not benefit much from the near-equivalent baselines.
+The maximum benefits for other cases are expected to be around $10\% to 30\%$.
+PAPER128 is designed with highly redundant near-equivalent baselines, and thus  these baselines start contributing at higher $\widetilde{\Theta}_{min}$, but the gapped HERA configurations will benefit even more from near-equivalent baselines due to there being more classes of baseline pairs.
 Note that here we normalized $\rho$ such that the contribution of the top single baseline equivalent-pair class, such as (sep0,1:0,1 in the PAPER128 case) are 1.
 This plot therefore does not compare sensitivity across the different arrays.
-The stepwise pattern in the PAPER128 case is characteristic of the rectangular grid; as we step to lower $\widetilde{\Theta}$ large groups of baseline pair classes get included in "batch".
-
-In Fig.
-\ref{fig:pairplot} we present a comparison of 4 different properties for baselines that contribute well to the sensitivity ($\widetilde{\Theta}>0.1$, where again the weight for the top class of equivalent pair is normalized to 1).
-The $dT$ vs.
-$\Theta$ and $dT$ vs.
-$\widetilde{\Theta}$ plots are familiar from Fig.
-\ref{fig:sensplot}.
-Interestingly the near-equivalent baselines form a "U" shape in $dT, \Theta$ plane [TODO].
-Note these near equivalent baselines have around the same $\Theta$ as in PAPER128, but much lower $\widetilde{\Theta}$.
-This is because they are longer baselines with lower multiplicity.
-
+The stepwise pattern case is characteristic of a regular grid; as we step to lower $\widetilde{\Theta}$ large groups of baseline pair classes get included in "batch".
 %\begin{centered}
@@
 
+In Fig.
+\ref{fig:pairplot} we present a comparison of 4 different properties for baselines that contribute well to the sensitivity ($\widetilde{\Theta}>0.1$, where again the weight for the top class of equivalent pair is normalized to 1).
+The $dT$ vs.
+$\Theta$ and $dT$ vs.
+$\widetilde{\Theta}$ plots are familiar from Fig.
+\ref{fig:sensplot}.
+We only show 3 of the mentioned arrays for visual clarity.
+The other two results are similar except for intuitive differences.
+We make a few observations (these are in no way an exhaustive list):
+\begin{itemize}
+
+\item Mapping of equivalent baseline pairs: the equivalent baselines are located in key points in the plots.
+They are the vertical bars in the lower left four plots and the highest peak in the first three histograms, the upper "diagonal edge" in the $\widetilde{\Theta}$ vs.
+$dT$ and the dot of $\Theta=1$.
+
+\item Longer baselines roughly peak at lower time-offset, and lower multiplicity.
+This is intuitive for PAPER since the for the same north-south distance the rotation angle is smaller with longer east-west distance.
+
+\item Interestingly for HERA the near-equivalent baselines form a "U" shape in $dT, \Theta$ plane, somewhat counter-intuitive at first especially given out last point.
+Turns out the "U" pattern is characteristic of the hexagonal array-layout.
+
+\item The top near-equivalent pairs in HERA the same $\Theta$ as in PAPER128, but much lower $\widetilde{\Theta}$.
+This is because they are longer baselines with lower multiplicity.
+In the end these baseline classes still lead to high contributions to total sensitivity (Fig.
+\ref{fig:osens}) simply because there are a lot more baseline pairs for HERA.
+
+\end{itemize}
+
 \begin{widetext}
 \begin{figure}[H]
-\includegraphics[scale=0.5]{pairplot}
+\includegraphics[width=\textwidth]{pairplot}
 \label{fig:pairplot}
@@
 Only those points with $\widetilde{\Theta}>0.1$ are shown (again the weight for the top class of equivalent pair is normalized to 1).
-Only one baseline length is shown since all top contributing pairs have very similar lengths as expected.}
+Only one baseline length is shown since all top contributing pairs have very similar lengths as expected.
+The scatter plots are shown with transparency so that darker regions indicate degeneracies.
+}
 \end{figure}
@@
 3) Cross multiply the visibilities of the two baselines  to form the power-spectrum.
-4) Finally combine the different baseline pairs by appropriate inverse-covariance weighting that takes into account the predicted sensitivity contributions of each case.
+4) Finally combine the different baseline pairs by appropriate inverse-variance weighting that takes into account the predicted sensitivity contributions of each case.
 
 \appendix
-\section{\\Hera Analysis}
+\section{\\Hera Antenna Configurations}
 \begin{widetext}
 \begin{figure}[H]
-\includegraphics[scale=0.55]{HeraAntpos}
+\includegraphics[width=0.9\textwidth]{HeraAntpos}
 
-\caption{Planned Hydrogen Epoch of Reionization Array antenna configurations. For HERA350, only the 320 elements in the core are shown. \label{fig:HeraAntpos}}
+\caption{Planned Hydrogen Epoch of Reionization Array antenna configurations.
+HERA 37 is expected to complete and start collecting data in summer of 2017, and the other three are planned configurations in the next phases.
+For HERA350, only the 320 elements in the core are shown.
+\label{fig:HeraAntpos}}
 \end{figure}
@@
 
-where in the second last line we have substituted visibility noise covariance. In the final line we used Wick's theorem and the fact that the signal from two visibilities are equal.
+where in the second last line we have substituted visibility noise variance. In the final line we used Wick's theorem and the fact that the signal from two visibilities are equal.
 

### 2017-03-26 Rotation paper/draft_working.tex
clean up

@@
-\documentclass[preprint2,numberedappendix,tighten,twocolappendix]{aastex6}  % USE THIS TO MAKE BIB, THEN FORMAT USING EMULATEAPJ
-%\documentclass[twocolumn,apj,numberedappendix]{emulateapj}
+%\documentclass[preprint2,numberedappendix,tighten,twocolappendix]{aastex6}  % USE THIS TO MAKE BIB, THEN FORMAT USING EMULATEAPJ
+\documentclass[twocolumn,apj,numberedappendix]{emulateapj}
 \shorttitle{Adding Sensitivity to 21cm Inteferometric Probes of Reionization}
@@
 \renewcommand\]{\end{equation}}
+\graphicspath{{../figures/}}
 
@@
 The extremely natural and simple implementation of this method
-would increase the sensitivity of PAPER-128 by about to $30\%$, and increase the out-rigger imaging sensitivity of HERA-350 by a factor of 2. 
+would increase the sensitivity of PAPER-128 and different configurations of HERA by about to $20\%$ to $60\%$, with HERA-350 benefitting the most from the technique. 
 \end{abstract}
@@
 can be applied to redundant arrays such as PAPER and HERA to increase
-the sensitivity by up to a factor of [1.5].
+the sensitivity by up to a factor of 1.3.
 
@@
 bottom panel show the antenna labels and distances, excluding the outlier
-antennas.\label{fig:AntPos}
+antennas.
 The numbering of the antennas the bottom panel are original labels
@@
 meaning. Baselines corresponding to the same separations are called
-equivalent. In the bottom panel, the two baselines labeled with the
-red arrow are equivalent, with separation denoted by sep1,0, for the
+equivalent. In the bottom panel, the two baselines (5,77) and (40,85) are an example of equivalent pair, with separation denoted by sep1,0, for the
 antennas are separated by 1 unit east and 0 unit north. Similarly,
-the baselines labeled with the yellow and green arrows are examples
+the baselines (5,69) and (5,85) are examples
 of sep1,1 and sep1,-1, respectively. Note sep1,0 and sep-1,0 for example
 are the same baselines and should not be counted twice.}
-
+\label{fig:AntPos}
 \end{figure}
 
-We shall use the 128-element PAPER array to demonstrate. 
+We shall use the 128-element PAPER array to demonstrate, and extend our results to several HERA configurations in Sec \ref{sec:arrconf}. 
 The PAPER array is located in the Karoo desert in South Africa (30:43:17.5
@@
 directions are comparatively close (4m), so that baselines such as
-0\_44 and 0\_7 are very close to equivalent. In the bottom panel,
-the two baselines labeled with the red arrow are equivalent, with
-separation denoted by sep1,0, for the antennas are separated by 1
-unit east and 0 unit north. Similarly, the baselines labeled with
-the yellow and green arrows are examples of sep1,1 and sep1,-1, respectively.
+0\_44 and 0\_7 are very close to equivalent. In the bottom panel, the two baselines (5,77) and (40,85) are an example of equivalent pair, with separation denoted by sep1,0, for the
+antennas are separated by 1 unit east and 0 unit north. Similarly,
+the baselines (5,69) and (5,85) are examples
+of sep1,1 and sep1,-1, respectively.
 Note sep1,0 and sep-1,0 for example are the same baselines and should
@@
 hours. Color represents different baselines. Frequency is $\nu=0.15\text{GHz}$.
-As the earth rotates, tracks are traced out counterclockwise. \label{fig:Tracks}}
+As the earth rotates, tracks are traced out counterclockwise. }
+\label{fig:Tracks}
 \end{figure}
@@
 \includegraphics[width=1.1\linewidth]{rephs}
-\label{fig:freqdiff}
+
 \caption{Comparisons of the peak phases of two different frequencies.
@@
 }
+\label{fig:freqdiff}
 \end{figure}
@@
 \includegraphics[width=\linewidth]{opp}
-\label{fig:numerics}
+
 \caption{Numerical comparisons of the visibility correlation peaks to the $\Omega$ factor in Eq.(\ref{eq:final}).
@@
 We see that in both cases the theory and simulation line up in both amplitude and phase.}
+\label{fig:numerics}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{sensitivity}
-\label{fig:sensplot}
 
@@
 contribution of 1,0:1,0 to unity. The same symbols and colors correspond in the top and bottom panel, with half of the legend shown in each. }
+\label{fig:sensplot}
 \end{figure}
@@
 For clarity of presentation we have "folded over" the negative time delays and combined baseline pairs that are identical modulus parity,
-or in other words are mirror images of each other, by summing over them to get an extra factor of $\sqrt{2}$. For example, 
+or in other words are mirror images of each other
+%\footnote{The parity symmetry is partially broken unless the array is located at the poles.
+But we are ignoring this here for illustrative purposes.
+See Fig.
+\ref{fig:pairplot} for a more accurate result.
+}
+, by summing over them to get an extra factor of $\sqrt{2}$. For example, 
 instead of plotting 1,1:1,1 and 1,-1:1,-1 separately, we plot 1,1:1,1 with twice the multiplicity. Similarly, baseline pairs such as 
@@
 \subsection{Array Configuration Comparisons \label{sec:arrconf}}
-To further quantify the sensitivity contribution of a given array we run our algorithm over all possible baseline-pairs of  PAPER128, HERA37, HERA128, HERA243 and HERA350.
-The HERA antenna configurations are shown in Fig.
-\ref{fig: HeraAntpos}.The
+We run our algorithm over all possible baseline-pairs of  PAPER128, HERA37, HERA128, HERA243 and HERA350. The HERA antenna configurations are shown in Fig. \ref{fig:HeraAntpos} in the Appendix. The  
 hexagonal design is the densest pattern of antenna-packing.
-The lack of short baselines that are visually close to each other means that we expect to see only longer near-equivalent baselines.
-These longer baselines have lower multiplicities, but also there would be many more classes of baselines, especially given the gap in the larger versions.
+The larger arrays are designed with a "gap" dividing the antennas into three different regions.
+The gaps are designed so as to improve UV-coverage and ease calibration without compromising sensitivity, but also produces much more near-equivalent baselines than the versions without the gap.
+The motivations behind the designs are explained in \cite{HERAconfiguration}.
+The lack of short baselines that are visually close to each other, as well as the shorter beam (Fig.\ref{fig:Beam}) means that we expect to see only longer near-equivalent baselines.
+The lower multiplicities per class of baselines is made up with the larger number of classes of baseline-pairs, especially given the gap in the larger versions.
+
+Having gotten a sense of the sensitivity from a given pair of baselines, we'd like to study the cumulative sensitivity of the array depending on which baseline pairs we include.
+Evidently we should prefer the pairs with larger $\widetilde{\Theta}$.
 In Fig.
 \ref{fig:osens} we plot $\rho$ against the minimum $\widetilde{\Theta}$.
-$\rho(\widetilde{\Theta}_{min})$ is thus the sensitivity of the array when baseline-pairs that have $\widetilde{\Theta}>\widetilde{\Theta}_{min}$ are included.
+$\rho(\widetilde{\Theta}_{min})$ is the sensitivity of the array when baseline-pairs that have $\widetilde{\Theta}>\widetilde{\Theta}_{min}$ are included.
 The dashed lines represent the values when only the equivalent baseline-pairs are used.
@@
 The small HERA37, with no gap (like in HERA350) or short near-equivalent baselines (like in PAPER 128), will not benefit much from the near-equivalent baselines.
-The maximum benefits for other cases are expected to be around $10\% to 30\%$.
-PAPER128 is designed with highly redundant near-equivalent baselines, and thus  these baselines start contributing at higher $\widetilde{\Theta}_{min}$, but the gapped HERA configurations will benefit even more from near-equivalent baselines due to there being more classes of baseline pairs.
-Note that here we normalized $\rho$ such that the contribution of the top single baseline equivalent-pair class, such as (sep0,1:0,1 in the PAPER128 case) are 1.
-This plot therefore does not compare sensitivity across the different arrays.
-The stepwise pattern case is characteristic of a regular grid; as we step to lower $\widetilde{\Theta}$ large groups of baseline pair classes get included in "batch".
+The maximum benefits for other cases are expected to be around $20\%$ to $60\%$.
+PAPER128 is designed with highly redundant near-equivalent baselines, and thus  these baselines start contributing at higher $\widetilde{\Theta}_{min}$, but the gapped HERA configurations will benefit even more from near-equivalent baselines at low $\widetilde{\Theta}_{min}$ due to there being more classes of such pairs.
+Note that here we normalized $\rho$ such that the contribution of the top equivalent baseline pair, such as (sep0,1:0,1 in the PAPER128 case) are 1.
+This plot therefore does not compare the absolute sensitivity across the different arrays.
+The stepwise pattern is characteristic of a regular grid; as we step to lower $\widetilde{\Theta}$ large groups of baseline pair classes get included in "batch".
 %\begin{centered}
@@
 \includegraphics[width=\linewidth]{osens}
-\label{fig:osens}
 \caption{Sensitivity of redundant arrays as a function of the minimum effective weight.
 Dashed lines represent when only equivalent baseline pairs are used, while solid lines indicate use of both the equivalent and near-equivalent baselines are used.
-The y-axis is normalized independently for each array such that the contribution of the top single equivalent pair class is unity, and thus does not indicate a comparison across the different arrays shown.
+The y-axis is normalized independently for each array such that the contribution of the top single equivalent pair class is unity, and thus does not indicate a comparison of absolute sensitivity across the different arrays shown.
 }
+\label{fig:osens}
 \end{figure}
@@
 In Fig.
-\ref{fig:pairplot} we present a comparison of 4 different properties for baselines that contribute well to the sensitivity ($\widetilde{\Theta}>0.1$, where again the weight for the top class of equivalent pair is normalized to 1).
+\ref{fig:pairplot} we present a comparison of 4 different properties for baselines that contribute well to the sensitivity ($\widetilde{\Theta}>0.05$, where again the weight for the top class of equivalent pair is normalized to 1).
 The $dT$ vs.
@@
 We only show 3 of the mentioned arrays for visual clarity.
-The other two results are similar except for intuitive differences.
-We make a few observations (these are in no way an exhaustive list):
+The other two results are similar barring intuitive differences.
+We make a few observations:
 \begin{itemize}
@@
 \item Longer baselines roughly peak at lower time-offset, and lower multiplicity.
-This is intuitive for PAPER since the for the same north-south distance the rotation angle is smaller with longer east-west distance.
+This is intuitive for PAPER since  for the same north-south separation the rotation angle is smaller with longer east-west separation.
 
-\item Interestingly for HERA the near-equivalent baselines form a "U" shape in $dT, \Theta$ plane, somewhat counter-intuitive at first especially given out last point.
-Turns out the "U" pattern is characteristic of the hexagonal array-layout.
+\item Interestingly for HERA the near-equivalent baselines form multiple "U" shapes in $dT, \Theta$ plane, somewhat counter-intuitive especially given out last point.
+This is characteristic of the hexagonal layout.
+Baselines of equal length are more redundant at the right time delay than ones that are slightly different in length, though the latter has a shorter optimal time-delay.
 
@@
 In the end these baseline classes still lead to high contributions to total sensitivity (Fig.
-\ref{fig:osens}) simply because there are a lot more baseline pairs for HERA.
+\ref{fig:osens}) because there are a lot more such baseline pairs for HERA.
 
@@
 
-\begin{widetext}
-\begin{figure}[H]
-\includegraphics[width=\textwidth]{pairplot}
-\label{fig:pairplot}
-\caption{Pairplots of the top contributing baseline pairs in three arrays.
-Plotted properties are optimal time delay $dT$, peak height $\Theta$, effective weight $\widetilde{\Theta}$ and baseline length $L$.
-Only those points with $\widetilde{\Theta}>0.1$ are shown (again the weight for the top class of equivalent pair is normalized to 1).
-Only one baseline length is shown since all top contributing pairs have very similar lengths as expected.
-The scatter plots are shown with transparency so that darker regions indicate degeneracies.
-}
-\end{figure}
-\end{widetext}
 %\subsection{Notes on Data, Foreground and Noise}
@@
 4) Finally combine the different baseline pairs by appropriate inverse-variance weighting that takes into account the predicted sensitivity contributions of each case.
+We showed that such techniques could lead to $20\%$ to $60\%$ increase in sensitivity for PAPER and HERA.
+\begin{widetext}
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{pairplot}
+
+\caption{Pairplots of the top contributing baseline pairs in three arrays.
+Plotted properties are optimal time delay $dT$, peak height $\Theta$, effective weight $\widetilde{\Theta}$ and baseline length $L$.
+Only those points with $\widetilde{\Theta}>0.05$ are shown (the weight for the top class of equivalent pair is normalized to 1).
+Only one baseline length is shown since all top contributing pairs have very similar lengths as expected.
+The scatter plots are shown with transparency so that darker regions indicate degeneracies.
+Scatter points of the 3 arrays overlap, in order indicated by the legend.
+}
+\label{fig:pairplot}
+\end{figure}
+\end{widetext}
+
+\pagebreak
 
@@
 \includegraphics[width=0.9\textwidth]{HeraAntpos}
-
+\label{fig:HeraAntpos}
 \caption{Planned Hydrogen Epoch of Reionization Array antenna configurations.
@@
 For HERA350, only the 320 elements in the core are shown.
-\label{fig:HeraAntpos}}
+}
 \end{figure}

### 2017-03-30 Rotation paper/draft_working.tex
save

@@
 Yunfan Gerry Zhang \altaffilmark{1},
-Aaron R. Parsons\altaffilmark{1}\altaffilmark{2}
+Adrian C. Liu \altaffilmark{1, 2},
+Aaron R. Parsons\altaffilmark{1, 2}
 }
@@
 \begin{abstract}
-We introduce an accurate method to extract additional sensitivity from data of redundant radio arrays.
-Our method efficiently finds optimal baselines to cross correlate, including baselines that are slightly different in length and orientation.
-Using the latest and final 128-element version of Precision Array for Probing the Epoch of Reionization (PAPER-128), as well as the planned 37, 128, 240 and 350 element versions of the Hydrogen Epoch of Reionization Array (HERA) now under development, we illustrate how our method applies to different array configurations and predict the sensitivity improvements of including each baseline pairs.
-The extremely natural and simple implementation of this method
+The observational effort to measure the primordial 21cm power spectrum with radio interferometers require high sensitivity, or signal to noise ratio.
+We introduce a straightforward and accurate method to extract additional sensitivity from data of redundant radio arrays.
+Our method efficiently finds optimal baselines to cross correlate, including baselines that are slightly different in length and orientation, and quantifies the sensitivity contributions of given baselines.
+Using the configurations and beams of the latest and final 128-element version of Precision Array for Probing the Epoch of Reionization (PAPER-128), as well as the planned 37, 128, 240 and 350 element versions of the Hydrogen Epoch of Reionization Array (HERA) now under development, we illustrate how our method applies to different arrays and predict the sensitivity improvements of including each baseline pairs.
+We show that inclusion of non-perfectly redundant baselines
 would increase the sensitivity of PAPER-128 and different configurations of HERA by about to $20\%$ to $60\%$, with HERA-350 benefitting the most from the technique. 
@@
 
-The epoch of reionization, or cosmic dawn represents the last key
-stage of the universe's early evolution. Study of this event stands at
+The epoch of reionization, or cosmic dawn, represents the last key
+stage of our universe's early evolution. Study of this event stands at
 the intersection of cosmology and astrophysics. Understanding this
@@
 to fundamental physics of inflation, neutrino mass and phenomenology
-of the first stars and galaxies (e.g. \cite{LiuOpticalDepth}).
+of the first stars and galaxies (e.g. \citealt{LiuOpticalDepth, Liu2016b, Mao2008, DEw21cm, Bull2015, Oyama20131186} etc.).
 
@@
 comes from measurement of the ``spin-flip'' transition of neutral
-hydrogen of characteristic wavelength 21cm \cite{Furlanetto2006181,PritchardLoeb}.
+hydrogen of characteristic wavelength 21cm \citep{Furlanetto2006181,PritchardLoeb}.
 While probes such as quasar spectra rely on emission and scattering
@@
 abundance and distribution of neutral hydrogen, and thus would potentially
-shed light on all stages of reionization. Radio interferometry efforts
+shed light on all stages of reionization. Radio interferometric efforts
 to measure the 21cm power spectrum has been a top priority in recent years of astronomy.
 Current generation instruments include the Precision Array for Probing
-the Epoch of Reionization (PAPER) \cite{Ali2015,paper32}, Murchison
-Widefield Array (MWA) \cite{MWA}, Low Frequency Array (LOFAR) \cite{LOFAR}. Next generation instruments such as the Hydrogen Epoch of Reionization
-Array (HERA) (e.g. \cite{HERA,HERAconfiguration,HERABEAM1,HERADISH2})
-and the Square Kilometer Array Low (SKA low) (e.g. \cite{SKA1}) are
-in planning and construction.
+the Epoch of Reionization (PAPER) \citep{Ali2015,paper32}, Murchison
+Widefield Array (MWA) \citep{Bowman2013, Tingay2013}, Low Frequency Array (LOFAR) \citep{LOFAR}. Next generation instruments such as the Hydrogen Epoch of Reionization
+Array (HERA) (e.g. \citealt{HERA,HERAconfiguration,HERABEAM1,HERADISH2}) currently under construction
+and the Square Kilometer Array Low (SKA-low) (e.g. \citealt{SKA1}) in planning.
 
-One of the main challenges of interferometric observations of cosmic
-21cm line is foreground contamination. Both sources within our
-galaxy, and to a lesser distant sources outside of our galaxy emit
-radio contaminations (via for example synchrotron processes) up to five orders of magnitude stronger than the
+One of the main challenges of observations of cosmic
+21cm line, interferometric or global, is foreground contamination. Both sources within our
+galaxy, and to a lesser extent distant sources outside of our galaxy emit
+radio contamination (via for example synchrotron processes) up to five orders of magnitude stronger than the
 reionization signal. There are two common methods to deal with the
-foreground contamination. In the first, sometimes called the foreground
+foreground contamination in power spectrum measurements. In the first, sometimes called the foreground
 removal technique, individual sources are identified and removed in the image domain.
-The other technique, commonly referred to as 'foreground avoidance',
-makes use of the fact that most common foreground contaminations have a smooth
-spectrum, and thus is constrained to a Fourier domain ``foreground
-wedge'' \cite{wedge1,wedge2}. Thus contamination can be avoided
-by restricting our observation to ``outside the wedge'', where the data is currently noise dominated. The main
+The other technique, commonly referred to as ``foreground avoidance'',
+makes use of the fact that most common foreground contaminants have smooth
+spectra, and thus is constrained to a Fourier domain ``foreground
+wedge'' \citep{wedge1,wedge2}. Thus, contamination can be avoided
+by restricting our observation to ``outside the wedge'', where theoretically foreground contamination is limited and noise becomes dominant. The main
 challenge of using the avoidance technique is thus the sensitivity, or
@@
 the maximum redundancy arrays such as PAPER and HERA. Traditional
-imaging arrays are designed to image localized bright sources and thus favor non-redundant patterns for uv coverage. The redundant arrays on the other hand
+imaging arrays are designed to image localized bright sources and thus favor non-redundant patterns for $uv$ coverage. The redundant arrays, on the other hand,
 focus on high-sensitivity measurements of the same Fourier mode
@@
 \cite{Ali2015} provides the newest upper limit to the power spectrum
-measurements with the 64-element version of PAPER(henceforth referred to as PAPER-64). From some models we roughly expect
+measurements with the 64-element version of PAPER (henceforth referred to as PAPER-64). From some models we roughly expect
 that the sensitivity required for detection is a factor of 10 away.
-This paper explores a new technique to add sensitivity to the analysis
-of redundant arrays, by making use of the earth's rotation. This technique
-can be applied to redundant arrays such as PAPER and HERA to increase
-the sensitivity by up to a factor of 1.3.
 
+This paper explores a new technique to add sensitivity to the power spectrum analysis
+of redundant arrays, by making use of the earth's rotation. 
 The earth's rotation causes the baselines to pick up different modes of the sky with time. 
-This effect is used extensively to improve uv coverage in imaging with minimum redundancy arrays. In
-a maximally redundant array, baselines that are slightly different
-rotate into each other at a time delay. Here we present a method to
+This effect is used extensively to improve $uv$ coverage in imaging with minimum redundancy arrays. In
+a maximally redundant array, we can make use of the fact that baselines that are slightly different
+rotate into each other at a time delay to extract power-spectrum from their respective visibilities. Here we present a method to
 improve the sensitivity of the maximum redundancy arrays by including
-the uv redundancy of different baselines at a time lag.
-Recently the authors of \cite{wterm} proposed an algorithmic search approach to extract information from all near-equivalent baselines.
-We here present a simple and elegant theoretical formalism that in addition allows us to identify good baselines to cross-correlate and predict the weight of sensitivity with speed and accuracy.
+the $uv$ redundancy of different baselines at a time lag.
+Recently, \cite{wterm} proposed an algorithmic approach to extract information from near-equivalent baselines, by carefully treating the $w$-term in the fringe pattern to obtain weights for cross-correlation.
+The algorithm makes use of a time dependent coordinate system ('moving grid') and is thus computationally expensive.
+We here present a formalism that allows us to compute such weights by simply performing a numerical integral.
+This allows us to identify good baselines to cross-correlate and predict the weight of sensitivity with speed and accuracy, thereby making the power spectrum estimation from a large number of baseline pairs computationally tractable.
 
@@
 in the sky. In order to eliminate confusion and ambiguity, we shall
-introduce slightly different terminology. We'll call baselines that
+introduce slightly different terminology. We shall call baselines that
 are the same length and orientation ``equivalent baselines'', inspired
@@
 times. Non-equivalent baselines can also be partially redundant if their
-respected time series is shifted with respect to one another by a
-certain delay. In other words, one baseline can be ``rotated into''
-another. We call these ``near-equivalent baselines''. It is our
+respected time series is shifted with respect to one another by some delay. In other words, one baseline can be ``rotated into''
+another.
+We call these ``near-equivalent baselines''.
+Exactly how ``near" is ``near-enough" depends on the array configuration.
+All non-equivalent baselines contribute non-zero sensitivity to the power spectrum analysis It is our
 goal to a) identify the near-equivalent baselines that give good
-redundancy, b) to find the optimal time offset for a given pair of baselines, and
+redundancy, b) to find the optimal time offset\footnote{By optimal we mean the time offset that corresponds to maximal redundancy. } for a given pair of baselines, and
 c) to quantify the sensitivity improvement associated with cross multiplying
-such a pair of near-equivalent baselines. In section
+such a pair of near-equivalent baselines.
+
+The rest of this paper is organized as follows. In section
 2 we explain the theoretical basis for this cross multiplication.
@@
 an antenna. Top panel shows the antenna positions drawn to scale,
-bottom panel show the antenna labels and distances, excluding the outlier
+bottom panel show the antenna labels and distances, excluding the outrigger
 antennas.
@@
 
-We shall use the 128-element PAPER array to demonstrate, and extend our results to several HERA configurations in Sec \ref{sec:arrconf}. 
+We shall use the 128-element PAPER array to demonstrate our method, and extend our results to several HERA configurations in Sec \ref{sec:arrconf}. 
 The PAPER array is located in the Karoo desert in South Africa (30:43:17.5
@@
 not be counted twice. Antennas in purely north-south baselines
-are close enough to induce cross talks, and hence are not suitable
-for use. The original PAPER-64 analysis (\cite{Ali2015}) used three classes of baselines, here
+are close enough to induce cross-talk, and hence are not suitable
+for use. The original PAPER-64 analysis \citep{Ali2015} used three classes of baselines, here
 equivalent to 
@@
 Given a point source on the sky, each baseline maps the
-source to a point in uv plane. As the earth rotates with respect to
-the source, the point traces out tracks in the uv-plane. 
-We show in Fig. \ref{fig:Tracks} uv tracks of PAPER-128 over 4.8 hours, at 0.15GHz, for a source that passes through zenith.
+source to a point in $uv$ plane. As the earth rotates with respect to
+the source, the point traces out tracks in the $uv$ plane. 
+We show in Fig. \ref{fig:Tracks} $uv$ tracks of PAPER-128 over 4.8 hours, at 0.15GHz, for a source that passes through zenith.
 
@@
 \includegraphics[width=\linewidth]{tracks128}
-\caption{Tracks of PAPER-128 (grid only, excluding outliers) for a hypothetical source that passes through zenith.
+\caption{Tracks of PAPER-128 (grid only, excluding outriggers) for requency $\nu=0.15\text{GHz}$ and a hypothetical source that passes through zenith.
 These tracks are traced out over 0.2 sidereal days, or roughly 4.8
-hours. Color represents different baselines. Frequency is $\nu=0.15\text{GHz}$.
+hours. Color represents different baselines. 
 As the earth rotates, tracks are traced out counterclockwise. }
@@
 redundancy of near-equivalent baselines as crossings
-of the uv tracks. Although this is a valid quick method to identify some baselines to cross-multiply, we find
-that it is not accurate enough for time offset determination.
-The reason is that UV tracks are traces of a point source, which though adequate for traditional narrow beam imaging, are not applicable for wide-beam receivers like that of PAPER.
-Here the uv-imprint of the sky for a given baseline and frequency is really a cloud that's the track convolved with the appropriate point-spread function. To accurately
-constrain redundancy, a more rigorous treatment is thus required. 
-We show sample beams of HERA and PAPER antennas in Fig.\ref{fig:Beam} for reference.
+of the $uv$ tracks.
+As we see in Fig.
+\ref{fig:Tracks}, there are many such crossings.
+Crossing tracks, however does not imply perfect redundancy.
+The reason lies in the finite size of the beams.
+We show sample beams of HERA and PAPER antennas in Fig.
+\ref{fig:Beam} for reference.
+The inaccuracy is expected to be even more significant for the smaller HERA beam, which has a larger corresponding $uv$ point spread function.
+
+Thus although it is a valid quick method to identify some baselines to cross-multiply, track-crossing is
+not accurate enough for time offset determination, nor can give estimate of the degree of redundancy.
+To determine the time-offset corresponding to maximal redundancy, we must develop a more general  formalism that accounts for the point spread function of the finite beams, and estimates the degree of redundancy for general combination of baselines at a general time-offset, as we do in the next section.
 
@@
 for frequency of $\nu=150\text{MHz}$ and Stokes $I$ polarization.
-Notice $A$ in this paper is a ``baseline's beam'', equivalent to squares of the antenna beams shown here.
+Notice $A$ in this paper is a ``baseline's beam'', equivalent to squares of the antenna voltage beams shown here.
 The circles centered around zenith (center of beam) here are
-spaced 10 degrees apart. The finite size of the beams limits the accuracy of the uv track-crossing as a method of redundancy search. \label{fig:Beam}}
+spaced 10 degrees apart. The finite size of the beams limits the accuracy of the $uv$ track-crossing as a method of redundancy search. \label{fig:Beam}}
 \end{figure}
@@
 We take the visibility as commonly defined in the literature (e.g.
-\cite{first-paper}): 
+\citealt{first-paper}): 
 \begin{equation}
@@
 related to $T$, the brightness temperature in the Rayleigh-Jeans
-limit. Note that $A$ here is the baseline beam, i.e. product of two antenna beams, examples of which are shown in Fig.\ref{fig:Beam}. $\phi(\nu)$ is the frequency bandpass profile. In practice,
-power spectrum measurements are typically taken from a few ten MHz centered around the corresponding redshift of interest (e.g. 150 MHz for z=9.5).
+limit.
+$k_B$ is Boltzmann's constant.
+Note that $A$ here is the baseline beam, i.e.
+product of two antenna voltage beams, examples of which are shown in Fig.
+\ref{fig:Beam}.
+$\phi(\nu)$ is the frequency bandpass profile.
+In practice,
+power spectrum measurements are typically taken from a few ten MHz centered around the corresponding redshift of interest (e.g.
+150 MHz for z=9.5).
+We take a moment to note that in Eq.
+\ref{eq:Vis1}, and our subsequent discussion, no flat-sky approximation has been made.
+The angular integral is performed over the dome $d\Omega$.
 
-We define the delay transformed visibility \cite{delay-transform}:
+We define the delay-transformed visibility \citep{delay-transform}:
 \small
@@
 $\boldsymbol{r}$ and $\boldsymbol{k}$. We start by noticing that
-{[}Gerry: change all r and k into bold{]} 
 \[
@@
 \]
-where $\nu_{21}=1420MHz$ is the 21cm transition rest frequency, $\nu_{0}$
+where $\nu_{21}=1420$MHz is the 21cm transition rest frequency, $\nu_{0}$
 a reference central frequency with corresponding redshift $z_{0}$,
@@
 \end{equation}
-Thus $d\nu=-dr/Y$ and $d^{3}r=-X^{2}Yd\Omega d\nu$.
 
@@
 \[
-\begin{aligned}V(\boldsymbol{b},\tau) & =\frac{1}{X^{2}Y}\int_{H}d^{3}rA(\boldsymbol{r})\phi(r)I(\boldsymbol{r})\exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\hat{\boldsymbol{r}}+\tau\right)\nu_{r}\right]\end{aligned}.
+\begin{aligned}V(\boldsymbol{b},\tau) & =\frac{2k_{B}}{\lambda^{2}}\int_{H}\frac{d^{3}r}{X^{2}Y}A(\boldsymbol{r})\phi(r)T(\boldsymbol{r})\exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\hat{\boldsymbol{r}}+\tau\right)\nu_{r}\right]\end{aligned}.
 \]
 \normalsize
-Here $(r_{x},r_{y},r_{z})=(X\hat{\boldsymbol{s}}_{x},X\hat{\boldsymbol{s}}_{y},Y\nu)$, and
+Here $(r_{x},r_{y},r_{z})=(X\hat{\boldsymbol{s}}_{x},X\hat{\boldsymbol{s}}_{y},Y\nu)$, and so $d\nu=-dr/Y$ and $d^{3}r=-X^{2}Yd\Omega d\nu$. 
 $(Xk_{x},Xk_{y},Yk_{z})=\frac{2\pi}{c}(b_{x},b_{y},\tau)$ relate
@@
 Here we choose to fix the sky, and denote the rotated coordinates
-of the beam pattern with the operator $\Gamma$. With implicit bounds
+of the beam pattern with the 3 dimensional rotation operator $\Gamma$. With implicit bounds
 of integrals from $-\infty$ to $\infty$, we have:
@@
  & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\boldsymbol{\kappa}\cdot(\boldsymbol{r}-\boldsymbol{r'})}\right)A^{*}(\boldsymbol{r})A(\Gamma r')\Phi_{b,\tau}(\boldsymbol{r},\Gamma r'),\\
- & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\xi(\boldsymbol{r}-\boldsymbol{r'})A^{*}(\boldsymbol{r})A(\Gamma r')\Phi_{b,\tau}(\boldsymbol{r},\Gamma r'),\\
  & \approx\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}rd^{3}r'\delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})A^{*}(\boldsymbol{r})A(\Gamma r')\Phi_{b,\tau}(\boldsymbol{r},\Gamma r'),\\
  & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}r|A^{*}(\boldsymbol{r})A(\Gamma r)||\phi(\nu_{r})|^{2}\exp\left[-i2\pi\nu_{r}\left(\hat{\boldsymbol{r}}\cdot\frac{\boldsymbol{b}}{c}-\hat{\Gamma r}\cdot\frac{\boldsymbol{b'}}{c}\right)\right],\\
- & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\hat{\Gamma s},\nu)||\phi(\nu)|^{2}\exp\left[-i2\pi\nu\left(\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b}}{c}-\hat{\Gamma s}\cdot\frac{\boldsymbol{b'}}{c}\right)\right],
+ & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\hat{\Gamma s},\nu)||\phi(\nu)|^{2}\exp\left[-i2\pi\nu\left(\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b}}{c}-\Gamma\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b'}}{c}\right)\right],
 \end{aligned}
@@
 sky, and the third to fourth line follows from the assumption that
-the 3D power spectrum varies negligibly over the region of interest
-so that $\hat{P_{21}}(k+k_{2})\approx\hat{P_{21}}(k)$. Since $\Gamma$
-is a sky rotation, it doesn't affect $\nu$, hence we have taken $\nu_{r}$
+the 3D power spectrum varies negligibly over the $k$-space of interest
+so that $\hat{P}_{21}(k+k_{2})\approx\hat{P}_{21}(k)$. Since $\Gamma$
+is a sky rotation, it does not affect $\nu$, hence we have taken $\nu_{r}$
 outside the parenthesis. Notice that the phase factor $\exp\left[-i2\pi\tau\left(\nu-\nu'\right)\right]$
@@
 \begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau')\rangle\\
- & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\hat{\Gamma s},\nu)||\phi(\nu)|^{2}\\
- & \qquad \qquad \qquad \qquad \exp\left[-i2\pi\nu\left(\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b}}{c}-\hat{\Gamma s}\cdot\frac{\boldsymbol{b'}}{c}\right)\right].\end{aligned}
+ & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)||\phi(\nu)|^{2}\\
+ & \qquad \qquad \qquad \qquad \exp\left[-i2\pi\nu\left(\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b}}{c}-\Gamma\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b'}}{c}\right)\right].\end{aligned}
 \label{eq:final}
@@
 \begin{equation}
-\Theta =\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\hat{\Gamma s},\nu)||\phi(\nu)|^{2} e^{-2\pi i\nu\left(\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b}}{c}-\hat{\Gamma s}\cdot\frac{\boldsymbol{b'}}{c}\right)}.
+\Theta =\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)||\phi(\nu)|^{2} e^{-2\pi i\nu\left(\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b}}{c}-\Gamma\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b'}}{c}\right)}.
 \label{eq:Theta}
@@
 So roughly speaking the cross multiplications of visibilities at a time delay
-in uv-space is proportional to the power spectrum times the Fourier
+in $uv$-space is proportional to the power spectrum times the Fourier
 transform of the cross multiplied beam pattern. We can then in principle
@@
 the phase and normalization. As a check, when applied to equivalent baselines,
-$\boldsymbol{b}=\boldsymbol{b'}$, $\hat{\boldsymbol{s}}=\hat{\Gamma s}$, and Eq.(\ref{eq:final}) reduces to Eq.(B9) of \cite{paper32}.
+$\boldsymbol{b}=\boldsymbol{b'}$, $\hat{\boldsymbol{s}}=\Gamma\hat{\boldsymbol{s}}$, and Eq.(\ref{eq:final}) reduces to Eq.(B9) of \cite{paper32}.
 
-If this result is robust, then we can achieve all our goals stated in the introduction, i.e. to identify 
-candidate baseline pairs, to find the optimal time offset, and to quantify the sensibility, by computing the weight $\Theta$ from
-Eq.(\ref{eq:Theta}) for various time offsets. Thus we shall first test its robustness numerically.
+With Eq.
+\ref{eq:opp} and Eq.
+\ref{eq:Theta} we can for any given baseline pair and time delay, estimate the degree of redundancy, here represented by $\Theta$, thereby achieving all our goals stated in the introduction, i.e.
+to identify
+candidate baseline pairs with good redundancy, to find the time offset that maximizes redundancy, and to quantify the degree of such redundancy.
+We can do all the above simply by computing the weight $\Theta$ from
+Eq.(\ref{eq:Theta}) for various time offsets, without having to actually cross-multiply visibilities at different offsets. This makes the task computationally tractable.
 
@@
 
-The physical origin of this frequency dependent phase lies the two visibilities having different phase centers.
+The physical origin of this frequency-dependent phase lies in the two visibilities having different phase centers.
 By default the two visibilities are both phased to zenith at the same time.
-When they are cross-correlated, they must be rephased to the account for the movement of the zenith.
+When they are cross-multiplied with a time lag, they must be rephased to the account for the movement of the zenith.
 Thus in practice we must rephase the visibilities before delay transform.
 
-In Fig.\ref{fig:freqdiff} we compare two channels 0.16 GHz and 0.17 GHz.
+In Fig.
+\ref{fig:freqdiff} we compare the cross-multiplied visibilities of both equivalent and near-equivalent baseline pairs for two channels: 0.16 GHz and 0.17 GHz.
 The top two panels have zero rephasing, and the bottom two are rephased to a time offset of 0.055 sidereal days, the optimal time offset for the given baseline pair.
@@
 
-To test the robustness of Eq.(\ref{eq:final}), we need to compare the amplitude and phase of the integral weight $\Theta$ 
+First we present some numerical tests of our formalism to verify the validity of using $\Theta$ to estimate the degree of redundancy as a function of time lag.
+To do so, we need to compare the amplitude and phase of the integral weight $\Theta$
 for a pair of baselines with products of simulated visibilities of those baselines.
@@
 We may use this simplification without compromising generality because Eq.
-(\ref{eq:final}) is valid for a range of frequency channels if an only if it's valid for each single channels, provided that different channels are combined with rephasing as desscribed in Sec.
+(\ref{eq:final}) is valid for a range of frequency channels if and only if it's valid for each single channels, provided that different channels are combined with rephasing as described in Sec.
 \ref{sec:rephs}.
 
-For the simulation, we take N random realizations of the sky
-on a healpix map (\cite{Heal}, \cite{HealPrimer}) \footnote{We use functionalities in the python package AIPY for healpix mapping
-as well as coordinate transforms. }. For each realization, each pixel is given a Gaussian random value
+For the simulation, we take N=2500 random realizations of the sky
+on a healpix map \citep{Heal, HealPrimer} \footnote{We use functionalities in the python package AIPY for healpix mapping
+as well as coordinate transforms. }, and pick two baselines sep1,0 and sep1,1. For each realization, each pixel is given a Gaussian random value
 of brightness temperature. We then rotate the baseline positions with
-the appropriate rotation matrix, recording the visibilities, for each
-baseline \footnote{As a caveat, there are two obvious ways to achieve the rotation. One
+the appropriate rotation matrix, multiplying the sky by the primary beam to get the visibilities, for each
+baseline\footnote{There are two obvious ways to achieve the rotation. One
 can either fix the sky and rotate the baselines, or the other way
@@
 correlation as a function of time-offset. The peak of the curve then
-corresponds to maximum redundancy. The accuracy of this result is
+corresponds to maximum redundancy.  We do this for both the equivalent (sep1,0 against sep1,0) and near-equivalent (sep1,0 against sep1,1) case. The accuracy of this result is
 limited by (simulated) cosmic variance and finite spacial resolution,
@@
 We generated 2400 instances of Gaussian random sky on healpix maps, computed visibilities and cross correlated them to find the correlation.
-Top panel shows the equivalent baseline pairs sep1,0 against itselt, bottom show set1,0:1:1.
+Top panel shows the equivalent baseline pairs sep1,0 against sep1,0, bottom panel shows set1,0 against sep1,1.
 Evaluations of weight $\Theta$ are shown with solid curves, and the visibility correlations of simulated random sky is given with dashed lines.
@@
 The comparison is show in Fig. \ref{fig:numerics}. 
-Since the sensitivity contribution is inferred from height of the peak of the cross-multiplied visibilities as a function of time. 
-We first compare the simulation to the analytical weight $\Theta$ for a pair of equivalent baselines, then normalize the peaks 
-of the near-equivalent baselines to that of the equivalent ones. 
+Sensitivity contribution is inferred from height of the peak of the cross-multiplied visibilities as a function of time. 
+We first compare the simulation to the analytical weight $\Theta$ for a pair of equivalent baselines (sep1,0, sep1,0), then normalize the peaks 
+of the near-equivalent baselines (sep1,0, sep1,1) to that of the equivalent ones. 
 In Fig. \ref{fig:numerics}  we show on the top comparison of
@@
 
-\begin{widetext}
-\begin{figure}[H]
-\includegraphics[scale=0.5]{fringe_res}
-%\includegraphics[width=\linewidth]{fringe_res}
+\begin{figure*}[h!]
+\includegraphics[width=\linewidth]{fringe_res}
 \caption{Beam fringe pattern of sep2,0(left), sep2,1 at a time delay (middle),
-and their conjugate product. Frequency of $\nu=0.15\text{GHz}$ is
-chosen and only the real components are shown. \label{fig:Beam-fringe-pattern}}
-\end{figure}
-\end{widetext}
+and their conjugate product (right). Frequency of $\nu=0.15\text{GHz}$ is
+chosen and only the real components are shown. The colorbar values are normalized such that the peak of the original beam is 1. \label{fig:Beam-fringe-pattern}}
+\end{figure*}
+
+\begin{figure*}[h!]
+\includegraphics[width=\linewidth]{sensitivity}
+
+\caption{Relative sensitivity contributions of selected baseline combinations in PAPER-128. In the legend m,n:p,q denotes cross
+multiplying PAPER-128 baselines of east-west, north-south separations (m,n) and (p,q) respectively. The top
+panel shows the peak height (degree of correlation) of each baseline
+combination, while the bottom panel multiplies the heights by the
+corresponding multiplicities as in Eq. \eqref{eq:sensul}, and in some cases an extra factor of $\sqrt{2}$ (explained in text). 
+In weighting by the multiplicity (bottom), we have chosen to fix the sensitivity 
+contribution of 1,0:1,0 to unity. The same symbols and colors correspond in the top and bottom panel, with half of the legend shown in each. }
+\label{fig:sensplot}
+\end{figure*}
 
@@
 Having verified Eq. \eqref{eq:final}, we can thus predict the sensitivity contributions of a
-particular baseline pair simply by computing the integral.
-Intuitively we expect the sensitivity to depend on both the uv coverage of the baseline and the patch of sky inside the beam.
-A larger beam like PAPER-beam would tolerate larger time-offsets because more sky area can coincide in the two beams.
-To keep the flow of the discussion, we shall focus on results of PAPER-128, and present the results for HERA in the Appendix.
+particular baseline pair simply by computing the integral Eq. \.
+Intuitively we expect the sensitivity to depend on both the $uv$ coverage of the baseline and the patch of sky inside the beam.
+A larger beam like that of PAPER would tolerate larger time-offsets because more sky area can coincide in the two beams \footnote{Larger beams also imply smaller spread in $uv$ space.
+This could lead to either larger or smaller redundancy, which depends on the overlap of two such point spread functions.}
 Having computed all of the baseline pairs, we find that baseline pairs that are mirror images of each other 
@@
 For example, sep1,0:1,1 is mirror image of sep1,0:1,-1 and these two baseline pairs
-give the same sensibility contribution. Thus in our results we shall only show a subset of representative baseline pairs 
-that give different informations.
+give the same sensibility contribution.
+Thus we shall only show a subset of representative baseline pairs to illustrate the contributions from different classes of baseline pairs.
+For a more complete result see Fig.
+\ref{fig:pairplot}.
 
@@
 
-\begin{widetext}
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{sensitivity}
-
-\caption{Relative sensitivity contributions of selected baseline combinations in PAPER-128. In the legend m,n:p,q denotes cross
-multiplying PAPER-128 baselines of east-west, north-south separations (m,n) and (p,q) respectively. The top
-panel shows the peak height (degree of correlation) of each baseline
-combination, while the bottom panel multiplies the heights by the
-corresponding multiplicities as in Eq. \eqref{eq:sensul}, and in some cases an extra factor of $\sqrt{2}$ (explained in text). 
-In weighting by the multiplicity (bottom), we have chosen to fix the sensitivity 
-contribution of 1,0:1,0 to unity. The same symbols and colors correspond in the top and bottom panel, with half of the legend shown in each. }
-\label{fig:sensplot}
-\end{figure}
-\end{widetext}
-
 To determine that actual relative contribution to sensitivity of these
 baseline pairs, we have to take into account of the multiplicities of
-these baselines, or the size of the equivalency classes in a mathematical
-language. By these we mean how many physical antenna pairs have the
+these baselines. By these we mean how many physical antenna pairs have the
 same length and orientation. Looking at Fig. \ref{fig:sensplot}
@@
 or sep1,1. The latest release of PAPER-64 data uses the 128-equivalent baselines sep2,1,
-sep2,0 and sep2,-1 \cite{Ali2015}, and achieved a $2\sigma$ upper
-limit of $(22.4mK)^{2}$. There, the three sets of equivalent baselines
+sep2,0 and sep2,-1 \citep{Ali2015}, and achieved a $2\sigma$ upper
+limit of $(22.4\text{mK})^{2}$. There, the three sets of equivalent baselines
 are only cross multiplied by itself. Assuming that each baseline delivers
-the same quality of data (i.e. they have the same height of correlation
+the same quality of data (meaning they have the same height of correlation
 peaks, which is in our normalization equal to unity), the relative
@@
 Points that have zero time delay are the equivalent baseline pairs and their weighted correlation values simply reflect the multiplicity factor.
-For clarity of presentation we have "folded over" the negative time delays and combined baseline pairs that are identical modulus parity,
+For clarity of presentation we have ``folded over'' the negative time delays and combined baseline pairs that are identical modulus parity,
 or in other words are mirror images of each other
@@
 
-Finally we estimate the power spectrum by inverse covariance weighting:
+Having defined the modified weight $\widetilde{\Theta}$, we can estimate the power spectrum by inverse covariance weighting:
 \begin{equation}
@@
 \begin{equation}
-\rho \propto \frac{1}{\sigma_P^2} = \sum_{bb'}^N\widetilde{\Theta}^2_{bb'},
+\rho \propto \frac{1}{\sigma_P^2} \propto \frac{1}{\sigma^4_N}\sum_{bb'}^N\widetilde{\Theta}^2_{bb'},
 \end{equation}
+where $\sigma_N$ is a characteristic single-baseline noise level.
 
-So far we have been sloppy in Eq.
-\ref{eq:sensul} for simplicity.
+The scaling in Eq.
+\ref{eq:sensul} was rough for simplicity of motivation.
 As we derive in Appendix \ref{sec:appB}, this weight should be corrected by a factor proportional to the signal to noise ratio of a single-baseline visibility $\rho_0=\sigma_S^2/\sigma_N^2$:
@@
 \subsection{Array Configuration Comparisons \label{sec:arrconf}}
-We run our algorithm over all possible baseline-pairs of  PAPER128, HERA37, HERA128, HERA243 and HERA350. The HERA antenna configurations are shown in Fig. \ref{fig:HeraAntpos} in the Appendix. The  
+We run our algorithm over all possible baseline-pairs of  PAPER128, HERA37, HERA128, HERA243 and HERA350. The HERA antenna configurations are shown in Fig. \ref{fig:HeraAntpos}. The  
 hexagonal design is the densest pattern of antenna-packing.
-The larger arrays are designed with a "gap" dividing the antennas into three different regions.
-The gaps are designed so as to improve UV-coverage and ease calibration without compromising sensitivity, but also produces much more near-equivalent baselines than the versions without the gap.
+The larger arrays are designed with a ``gap'' dividing the antennas into three different regions.
+The gaps are designed so as to improve $uv$ coverage and ease calibration without compromising sensitivity, but also produces many more near-equivalent baselines than the versions without the gap.
 The motivations behind the designs are explained in \cite{HERAconfiguration}.
-The lack of short baselines that are visually close to each other, as well as the shorter beam (Fig.\ref{fig:Beam}) means that we expect to see only longer near-equivalent baselines.
+The lack of short baselines that are visually close to each other, as well as the smaller beam (Fig.
+\ref{fig:Beam}) means that we expect to see only longer near-equivalent baselines.
 The lower multiplicities per class of baselines is made up with the larger number of classes of baseline-pairs, especially given the gap in the larger versions.
 
-Having gotten a sense of the sensitivity from a given pair of baselines, we'd like to study the cumulative sensitivity of the array depending on which baseline pairs we include.
+Having quantified the sensitivity from a given pair of baselines, we study the cumulative sensitivity of the array depending on which baseline pairs we include.
 Evidently we should prefer the pairs with larger $\widetilde{\Theta}$.
@@
 This plot therefore does not compare the absolute sensitivity across the different arrays.
-The stepwise pattern is characteristic of a regular grid; as we step to lower $\widetilde{\Theta}$ large groups of baseline pair classes get included in "batch".
+The stepwise pattern is characteristic of a regular grid; as we step to lower $\widetilde{\Theta}$ large groups of baseline pair classes get included in ``batch''.
 %\begin{centered}
@@
 %\end{centered}
+
+\begin{figure*}[h!]
+\includegraphics[width=0.9\textwidth]{HeraAntpos}
+\label{fig:HeraAntpos}
+
+\caption{Planned Hydrogen Epoch of Reionization Array antenna configurations.
+HERA 37 is expected to complete and start collecting data in summer of 2017, and the other three are planned configurations in the next phases.
+For HERA350, only the 320 elements in the core are shown.
+}
+\end{figure*}
 
@@
 \appendix
-\section{\\Hera Antenna Configurations}
-\begin{widetext}
-\begin{figure}[H]
-\includegraphics[width=0.9\textwidth]{HeraAntpos}
-\label{fig:HeraAntpos}
-\caption{Planned Hydrogen Epoch of Reionization Array antenna configurations.
-HERA 37 is expected to complete and start collecting data in summer of 2017, and the other three are planned configurations in the next phases.
-For HERA350, only the 320 elements in the core are shown.
-}
-\end{figure}
-\end{widetext}
-
-\section{\\Derivation of Noise Covariance \label{sec:appB}}
-
+\section{\label{sec:appB}\\Derivation of Noise Covariance \label{sec:appB}}
+\label{sec:appB}
 In this appendix we give a brief derivation of the effective weight quoted in \ref{sec:sensitivity}.

### 2017-03-30 Rotation paper/draft_working.tex
save

@@
 \begin{equation}
-\rho \propto \frac{1}{\sigma_P^2} \propto \frac{1}{\sigma^4_N}\sum_{bb'}^N\widetilde{\Theta}^2_{bb'},
+\rho \propto \frac{1}{\sigma_P^2} \propto \rho_0^2\sum_{bb'}^N\widetilde{\Theta}^2_{bb'},
 \end{equation}
-where $\sigma_N$ is a characteristic single-baseline noise level.
+where, if $\sigma_S^2$ and $\sigma_N^2$ are the characteristic signal and noise levels of a single-baseline visibility, $\rho_0=\sigma_S^2/\sigma_N^2$ is the signal to noise ratio.
 
-The scaling in Eq.
-\ref{eq:sensul} was rough for simplicity of motivation.
-As we derive in Appendix \ref{sec:appB}, this weight should be corrected by a factor proportional to the signal to noise ratio of a single-baseline visibility $\rho_0=\sigma_S^2/\sigma_N^2$:
+The scaling in Eq. \ref{eq:sensul} was rough for simplicity of motivation. As we derive in Appendix \ref{sec:appB}, this weight should be corrected by a factor proportional $\rho_0$:
 

### 2017-05-02 Rotation paper/draft_working.tex
save

@@
 \documentclass[twocolumn,apj,numberedappendix]{emulateapj}
-\shorttitle{Adding Sensitivity to 21cm Inteferometric Probes of Reionization}
+\shorttitle{Adding Sensitivity to 21\,cm Inteferometric Probes of Reionization}
 \shortauthors{Zhang\&Parsons}
@@
 
-\title{Adding Sensitivity to 21cm Inteferometric Probes of Reionization by Optimizing Choice of Baselines}
+\title{Adding Sensitivity to 21\,cm Inteferometric Probes of Reionization by Optimizing Choice of Baselines}
 
@@
 \begin{abstract}
-The observational effort to measure the primordial 21cm power spectrum with radio interferometers require high sensitivity, or signal to noise ratio.
-We introduce a straightforward and accurate method to extract additional sensitivity from data of redundant radio arrays.
-Our method efficiently finds optimal baselines to cross correlate, including baselines that are slightly different in length and orientation, and quantifies the sensitivity contributions of given baselines.
-Using the configurations and beams of the latest and final 128-element version of Precision Array for Probing the Epoch of Reionization (PAPER-128), as well as the planned 37, 128, 240 and 350 element versions of the Hydrogen Epoch of Reionization Array (HERA) now under development, we illustrate how our method applies to different arrays and predict the sensitivity improvements of including each baseline pairs.
+The observational effort to measure the cosmological 21\,cm power spectrum with radio interferometers requires high sensitivity.
+We introduce a straightforward and accurate method to extract sensitivity from visibility data.
+We develop our method in line with existing visibility-based power spectrum pipelines.
+Our method efficiently finds subsets of baselines to cross correlate, including partially redundant baselines, and quantifies the sensitivity contributions of given baselines.
+Using the 128-element configurations and beams of Precision Array for Probing the Epoch of Reionization (PAPER-128), as well as 4 planned versions of the Hydrogen Epoch of Reionization Array (HERA), we illustrate how our method applies to different arrays and predict the sensitivity improvements of including each baseline pairs.
 We show that inclusion of non-perfectly redundant baselines
-would increase the sensitivity of PAPER-128 and different configurations of HERA by about to $20\%$ to $60\%$, with HERA-350 benefitting the most from the technique. 
+would account for $20\%$ to $60\%$ of the sensitivity of PAPER-128 and different configurations of HERA. 
 \end{abstract}
@@
 
-The epoch of reionization, or cosmic dawn, represents the last key
-stage of our universe's early evolution. Study of this event stands at
+The epoch of reionization represents the last key
+stage of our Universe's early evolution. Study of this event stands at
 the intersection of cosmology and astrophysics. Understanding this
-event is important not only because it is a scientific goal
-of its own, but also because it can provide crucial information
-to fundamental physics of inflation, neutrino mass and phenomenology
+event not only serves as a scientific goal
+of its own, but also as a gateway to crucial information
+regarding fundamental physics of inflation, neutrino mass and phenomenology
 of the first stars and galaxies (e.g. \citealt{LiuOpticalDepth, Liu2016b, Mao2008, DEw21cm, Bull2015, Oyama20131186} etc.).
@@
 comes from measurement of the ``spin-flip'' transition of neutral
-hydrogen of characteristic wavelength 21cm \citep{Furlanetto2006181,PritchardLoeb}.
-While probes such as quasar spectra rely on emission and scattering
-of free electrons, and are thus limited to the lower redshifts towards the end of reionization, the 21cm line directly probes the
+hydrogen of characteristic wavelength 21\,cm \citep{Furlanetto2006181,PritchardLoeb}.
+Other observational studies of reionization include for example Gunn-Peterson measurements of quasi-stellar objects \citep{Fan2006}.
+GP tests rely on Ly$\alpha$ absorption, and thus easily saturate with small degree of ionization and can only probe the end stage of reionization.
+The 21\,cm line can in comparison directly probe the
 abundance and distribution of neutral hydrogen, and thus would potentially
 shed light on all stages of reionization. Radio interferometric efforts
-to measure the 21cm power spectrum has been a top priority in recent years of astronomy.
-Current generation instruments include the Precision Array for Probing
+to measure the 21\,cm power spectrum has been a top priority in recent years of astronomy.
+Current-generation instruments include the Precision Array for Probing
 the Epoch of Reionization (PAPER) \citep{Ali2015,paper32}, Murchison
-Widefield Array (MWA) \citep{Bowman2013, Tingay2013}, Low Frequency Array (LOFAR) \citep{LOFAR}. Next generation instruments such as the Hydrogen Epoch of Reionization
-Array (HERA) (e.g. \citealt{HERA,HERAconfiguration,HERABEAM1,HERADISH2}) currently under construction
-and the Square Kilometer Array Low (SKA-low) (e.g. \citealt{SKA1}) in planning.
+Widefield Array (MWA) \citep{Bowman2013, Tingay2013}, Low Frequency Array (LOFAR) \citep{LOFAR}. Next-generation instruments include the Hydrogen Epoch of Reionization
+Array (HERA) (e.g. \citealt{HERA,HERAconfiguration,HERABEAM1,HERADISH2})  under construction, 
+and the Square Kilometer Array Low (SKA-low) (e.g. \citealt{SKA1}), which is currently in planning.
 
-One of the main challenges of observations of cosmic
-21cm line, interferometric or global, is foreground contamination. Both sources within our
-galaxy, and to a lesser extent distant sources outside of our galaxy emit
-radio contamination (via for example synchrotron processes) up to five orders of magnitude stronger than the
-reionization signal. There are two common methods to deal with the
-foreground contamination in power spectrum measurements. In the first, sometimes called the foreground
-removal technique, individual sources are identified and removed in the image domain.
-The other technique, commonly referred to as ``foreground avoidance'',
-makes use of the fact that most common foreground contaminants have smooth
-spectra, and thus is constrained to a Fourier domain ``foreground
-wedge'' \citep{wedge1,wedge2}. Thus, contamination can be avoided
-by restricting our observation to ``outside the wedge'', where theoretically foreground contamination is limited and noise becomes dominant. The main
-challenge of using the avoidance technique is thus the sensitivity, or
-the signal to noise ratio. This is the motivation for the design of
-the maximum redundancy arrays such as PAPER and HERA. Traditional
-imaging arrays are designed to image localized bright sources and thus favor non-redundant patterns for $uv$ coverage. The redundant arrays, on the other hand,
-focus on high-sensitivity measurements of the same Fourier mode
-with many tightly packed antennas at equal spacing. Since baselines
-of the same length and orientation measure the same Fourier modes
-on the sky, a maximum redundancy array is able to increase the
-signal to noise ratio by averaging over measurements of the same baseline.
+These modern low-frequency radio interferometers aimed at measuring the 21\,cm signal bears designs dramatically different from traditional instruments built for imaging localized bright sources.
+With the goal of detecting a cosmological signal 5 orders of magnitude fainter than foreground contaminations, current efforts are sensitivity starved.
+Maximum-redundancy arrays are thus designed to repeatedly measure the same Fourier signal to increase sensitivity.
+Ranging upwards from 100 elements, these arrays are typically of the drift-scan type to limit cost.
+
+Analysis pipelines of such maximum-redundancy arrays typically fall into two categories.
+In the first, more traditional technique, image is formed in Fourier domain through rotation synthesis, then foreground is either removed or avoided to construct a power spectrum.
+The second, lately popular technique works directly with visibilities from baselines and cross multiply them to form the power spectrum.
+Due to its more natural focus on the instrumental output of visibilities, this technique is shown to fair better in dealing with systematics such as RFIs and instrumental cross-talks, and thus tend to provide good limits on the power spectrum.
 \cite{Ali2015} provides the newest upper limit to the power spectrum
-measurements with the 64-element version of PAPER (henceforth referred to as PAPER-64). From some models we roughly expect
-that the sensitivity required for detection is a factor of 10 away.
+measurements with the 64-element version of PAPER (henceforth referred to as PAPER-64).
+However, there is one caveat with existing visibility-based pipelines.
+While an imaging based power-spectrum pipeline naturally includes all redundancy information, visibility-based pipelines until now only cross multiply fully redundant baselines, in other words baselines of the same length and orientation.
 
-This paper explores a new technique to add sensitivity to the power spectrum analysis
-of redundant arrays, by making use of the earth's rotation. 
-The earth's rotation causes the baselines to pick up different modes of the sky with time. 
-This effect is used extensively to improve $uv$ coverage in imaging with minimum redundancy arrays. In
-a maximally redundant array, we can make use of the fact that baselines that are slightly different
-rotate into each other at a time delay to extract power-spectrum from their respective visibilities. Here we present a method to
-improve the sensitivity of the maximum redundancy arrays by including
-the $uv$ redundancy of different baselines at a time lag.
-Recently, \cite{wterm} proposed an algorithmic approach to extract information from near-equivalent baselines, by carefully treating the $w$-term in the fringe pattern to obtain weights for cross-correlation.
-The algorithm makes use of a time dependent coordinate system ('moving grid') and is thus computationally expensive.
-We here present a formalism that allows us to compute such weights by simply performing a numerical integral.
-This allows us to identify good baselines to cross-correlate and predict the weight of sensitivity with speed and accuracy, thereby making the power spectrum estimation from a large number of baseline pairs computationally tractable.
+Baselines of different length and orientation still contain partially redundant information.
+The use of partially redundant baselines is in principle not a fundamental limitation of visibility-based pipelines.
+In fact, most instrumental reports to date do include partial redundancy.
+Recently, \cite{wterm} proposed an algorithmic approach to extract power-spectrum from partially redundant baselines in tracking measurements, by carefully treating the $w$-term in the fringe pattern to obtain weights for cross-correlation.
+Our work parallels this effort by focusing on arrays of the drift-scan type that do not have tracking capabilities, such as PAPER, HERA and SKA.
+The Earth's rotation causes the baselines in a drift-scan array to pick up different modes of the sky with time.
+Rotation-synthesis makes use of the rotation-induced $uv$ coverage map to form image.
+In a visibility based pipeline, we can extract the same rotation-induced redundancy.
+In this point of view, baselines that are slightly different in length and orientation
+rotate into each other at a time delay.
+We can thus cross-multiply time-shifted visibilities, with the proper weighting, to form power spectrum.
+Due to the large number of elements of modern arrays (upward of 100), the task of cross-multiplying every baseline against every other, scaling as number of array element to the fourth power, can be computationally formidable.
+Our contribution is thus twofold.
+First we introduce a formalism to include pairs of partially-redundant baselines in a visibility based power spectrum pipeline.
+Secondly we show how to use the formalism to automaticaly preselect baseline pairs and time offsets, making the problem computationally efficient.
+More precisely, our formalism allows one to simultaneously a) identify the baselines that give good
+redundancy, b) find the time offset that corresponds to maximal redundancy for a given pair of baselines, and
+c) quantify the sensitivity associated with cross multiplying
+such a pair of partially redundant baselines, which in turns is used as weight to combine measurements in a power spectrum pipeline.
 
-Baselines of the same length and orientation are traditionally called
+[TODO: this paragraph would potentially be promoted to a section] By using pairs of baselines, we also reduce additional mode mixing.
+As discussed in \cite{Hazelton2013}, the varying shapes of overlapping regions in $uv$ space of all contributing baselines lead to chromaticities in the response.
+By considering only pairs of baselines, we fix shapes of overlapping regions, and only let their sizes vary with frequency, just as in the case of fully redundant baselines.
+
+The rest of this paper is organized as follows. In section 2 we introduce some terminology and notations to be used for the rest of the paper. In section
+3 we introduce the formalism for weighting partially redundant baselines.
+In section 4 we present numerical tests of
+this technique as well as the expected sensitivity improvement
+with this method for HERA and PAPER-128 and with section 5 we conclude.
+
+\section{Notation and Terminology}
+
+In order to eliminate confusion and ambiguity, we
+introduce some terminology for the rest of this paper, some of which differs from what is commonly found in the literature.
+
+First, we make the distinction of a \textit{``baseline''}, which refers to two specific antennas, and a \textit{``class of baselines''}, which refers to all baselines of the same length and orientation in a given array.
+
+Baselines of the same class are traditionally called
 ``redundant baselines'', because they measure the same Fourier mode
-in the sky. In order to eliminate confusion and ambiguity, we shall
-introduce slightly different terminology. We shall call baselines that
-are the same length and orientation ``equivalent baselines'', inspired
-by the mathematical notion of equivalency classes. Two equivalent
-baselines will be redundant with each other simultaneously at all
-times. Non-equivalent baselines can also be partially redundant if their
-respected time series is shifted with respect to one another by some delay. In other words, one baseline can be ``rotated into''
-another.
-We call these ``near-equivalent baselines''.
-Exactly how ``near" is ``near-enough" depends on the array configuration.
-All non-equivalent baselines contribute non-zero sensitivity to the power spectrum analysis It is our
-goal to a) identify the near-equivalent baselines that give good
-redundancy, b) to find the optimal time offset\footnote{By optimal we mean the time offset that corresponds to maximal redundancy. } for a given pair of baselines, and
-c) to quantify the sensitivity improvement associated with cross multiplying
-such a pair of near-equivalent baselines.
+in the sky.
+We shall call baselines in the same class \textit{``equivalent baselines''}, and reserve the word \textit{``redundancy''} of two baselines in reference to a variable function of the relative time-offsets of their visibility time series.
+With this terminology, two equivalent
+baselines are fully redundant with each other simultaneously at all
+times. Non-equivalent baselines also have partial redundancy, which is maximized if their
+respected time series are shifted with respect to one another by some delay. In other words, one baseline can be ``rotated into''
+another if they are \textit{``near-equivalent''}.
 
-The rest of this paper is organized as follows. In section
-2 we explain the theoretical basis for this cross multiplication.
-In section 3 we present numerical tests of
-this technique as well as the expected sensitivity improvement
-with this method for HERA and PAPER-128 and with section 4 we conclude.
+\begin{figure}[H]
+\includegraphics[width=\linewidth]{antpos128}
+
+\caption{The PAPER-128 layout. Each dot corresponds to the location of
+an antenna. Top panel shows the antenna positions drawn to scale;
+bottom panel show the antenna labels and distances, excluding the outrigger
+antennas.
+The numbering of the antennas the bottom panel are original labels
+during instrument assembly and does not bear significant
+meaning. In the bottom panel, the two baselines (5,77) and (40,85) are an example of equivalent pair, with separation denoted by \{1, 0\}, for the
+antennas are separated by 1 unit east and 0 unit north. Similarly,
+the baselines (5,69) and (5,85) are examples
+of the classes \{1, 1\} and \{1,-1\}, respectively. Note \{1, 0\} and \{-1, 0\} for example
+are the same baseline classes and should not be counted twice.}
+\label{fig:AntPos}
+\end{figure}
+
+We shall use the 128-element PAPER array (henceforth as PAPER-128) to motivate our formalism and demonstrate our method, and extend our results to several HERA configurations in Sec \ref{sec:arrconf}.
+The PAPER array is located in the Karoo desert in South Africa (30:43:17.5
+S, 21:25:41.8 E). The layout pattern with antenna labels are show
+in Fig. \ref{fig:AntPos}. The array consists of a 112-element grid and 16 "outriggers" used primarily to aid calibration. We see the antenna spacing in North-South
+directions are comparatively close (4m), so that baselines such as
+(0, 44) and (0, 7) are very close to equivalent. In the bottom panel, the two baselines (5, 77) and (40, 85) are an example of equivalent pair, with separation denoted by \{1, 0\}, for the
+antennas are separated by 1 unit east and 0 unit north. Similarly,
+the baselines (5, 69) and (5, 85) are examples
+of \{1, 1\} and \{1, -1\}, respectively.
+Note \{1, 0\} and \{-1, 0\} for example are the same classes and should
+not be counted twice. Antennas in purely north-south baselines
+are close enough to induce cross-talk, and hence are not suitable
+for use. The PAPER-64 analysis of \cite{Ali2015} used three classes of baselines, the PAPER-128
+equivalent of which are 
+\{2, 0\}, \{2, 1\} and \{2, -1\}. There each of these classes
+of baselines are cross multiplied to itself.
 
@@
 
-\begin{figure}[H]
-\includegraphics[width=\linewidth]{antpos128}
-
-\caption{The PAPER-128 layout. Each blue dot corresponds to the location of
-an antenna. Top panel shows the antenna positions drawn to scale,
-bottom panel show the antenna labels and distances, excluding the outrigger
-antennas.
-The numbering of the antennas the bottom panel are original labels
-during instrument assembly and does not bear significant
-meaning. Baselines corresponding to the same separations are called
-equivalent. In the bottom panel, the two baselines (5,77) and (40,85) are an example of equivalent pair, with separation denoted by sep1,0, for the
-antennas are separated by 1 unit east and 0 unit north. Similarly,
-the baselines (5,69) and (5,85) are examples
-of sep1,1 and sep1,-1, respectively. Note sep1,0 and sep-1,0 for example
-are the same baselines and should not be counted twice.}
-\label{fig:AntPos}
-\end{figure}
-
-We shall use the 128-element PAPER array to demonstrate our method, and extend our results to several HERA configurations in Sec \ref{sec:arrconf}. 
-The PAPER array is located in the Karoo desert in South Africa (30:43:17.5
-S, 21:25:41.8 E). The layout pattern with antenna labels are show
-in Fig. \ref{fig:AntPos}. We see the antenna spacing in North-South
-directions are comparatively close (4m), so that baselines such as
-0\_44 and 0\_7 are very close to equivalent. In the bottom panel, the two baselines (5,77) and (40,85) are an example of equivalent pair, with separation denoted by sep1,0, for the
-antennas are separated by 1 unit east and 0 unit north. Similarly,
-the baselines (5,69) and (5,85) are examples
-of sep1,1 and sep1,-1, respectively.
-Note sep1,0 and sep-1,0 for example are the same baselines and should
-not be counted twice. Antennas in purely north-south baselines
-are close enough to induce cross-talk, and hence are not suitable
-for use. The original PAPER-64 analysis \citep{Ali2015} used three classes of baselines, here
-equivalent to 
-sep2,0, sep2,1 and sep2,-1 \cite{Ali2015} of PAPER-128. There each of these classes
-of baselines are cross multiplied to itself. We shall see that in addition these
-baseline classes can be cross multiplied with a time offset.
-
 Given a point source on the sky, each baseline maps the
-source to a point in $uv$ plane. As the earth rotates with respect to
+source to a point in $uv$ plane. As the Earth rotates with respect to
 the source, the point traces out tracks in the $uv$ plane. 
@@
 hours. Color represents different baselines. 
-As the earth rotates, tracks are traced out counterclockwise. }
+As the Earth rotates, tracks are traced out counterclockwise. }
 \label{fig:Tracks}
@@
 We take a moment to note that in Eq.
-\ref{eq:Vis1}, and our subsequent discussion, no flat-sky approximation has been made.
+\ref{eq:Vis1}, no flat-sky approximation has yet been made.
 The angular integral is performed over the dome $d\Omega$.
@@
 \]
-where $\nu_{21}=1420$MHz is the 21cm transition rest frequency, $\nu_{0}$
+where $\nu_{21}=1420$MHz is the 21\,cm transition rest frequency, $\nu_{0}$
 a reference central frequency with corresponding redshift $z_{0}$,
@@
 
-We can thus rewrite the delayed transformed visibility as 
+Further we shall invoke the flat-sky limit
+\begin{equation}
+(r_{x},r_{y})=(X\hat{\boldsymbol{s}}_{x},X\hat{\boldsymbol{s}}_{y}), 
+\end{equation}
+where
+\begin{equation}
+X = \frac{c}{H_0}\int_{0}^{z}\frac{dz'}{E(z')}
+\end{equation}
+is a cosmological scalar. In the same limit $(Xk_{x},Xk_{y},Yk_{z})=\frac{2\pi}{c}(b_{x},b_{y},\tau)$ relates
+the cosmological $k$ to the observation coordinates. For a more rigorous treatment without the flat sky limit, see Appendix [ADRIAN].
+
+We can rewrite the delayed transformed visibility as 
 \small
 \[
-\begin{aligned}V(\boldsymbol{b},\tau) & =\frac{2k_{B}}{\lambda^{2}}\int_{H}\frac{d^{3}r}{X^{2}Y}A(\boldsymbol{r})\phi(r)T(\boldsymbol{r})\exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\hat{\boldsymbol{r}}+\tau\right)\nu_{r}\right]\end{aligned}.
+\begin{aligned}V(\boldsymbol{b},\tau) & =\frac{2k_{B}}{\lambda^{2}}\int_{H}\frac{d^{3}r}{X^{2}Y}A(\boldsymbol{r})\phi(r)T(\boldsymbol{r})\exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\hat{\boldsymbol{r}}+\tau\right)\nu_{r}\right]\end{aligned},
 \]
 \normalsize
-Here $(r_{x},r_{y},r_{z})=(X\hat{\boldsymbol{s}}_{x},X\hat{\boldsymbol{s}}_{y},Y\nu)$, and so $d\nu=-dr/Y$ and $d^{3}r=-X^{2}Yd\Omega d\nu$. 
-$(Xk_{x},Xk_{y},Yk_{z})=\frac{2\pi}{c}(b_{x},b_{y},\tau)$ relate
-the cosmological coordinates $r$ and $k$ to the measured coordinates.
+ where $d\nu=-dr/Y$ and $d^{3}r=-X^{2}Yd\Omega d\nu$.
+
 We have written $\nu_{r}$ to remind us that $\nu$ and $r$ are related
@@
 \begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau')\rangle\\
- & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle\right)A^{*}(\boldsymbol{r})A(\Gamma r')\Phi_{b,\tau}(\boldsymbol{r},\Gamma r'),\\
- & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\boldsymbol{\kappa})T(\boldsymbol{\kappa'})\rangle e^{-i(\boldsymbol{\kappa}\cdot \boldsymbol{r}-\boldsymbol{\kappa'}\cdot\boldsymbol{r'})}\right)A^{*}(\boldsymbol{r})A(\Gamma r')\Phi_{b,\tau}(\boldsymbol{r},\Gamma r'),\\
- & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\boldsymbol{\kappa}\cdot(\boldsymbol{r}-\boldsymbol{r'})}\right)A^{*}(\boldsymbol{r})A(\Gamma r')\Phi_{b,\tau}(\boldsymbol{r},\Gamma r'),\\
- & \approx\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}rd^{3}r'\delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})A^{*}(\boldsymbol{r})A(\Gamma r')\Phi_{b,\tau}(\boldsymbol{r},\Gamma r'),\\
- & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}r|A^{*}(\boldsymbol{r})A(\Gamma r)||\phi(\nu_{r})|^{2}\exp\left[-i2\pi\nu_{r}\left(\hat{\boldsymbol{r}}\cdot\frac{\boldsymbol{b}}{c}-\hat{\Gamma r}\cdot\frac{\boldsymbol{b'}}{c}\right)\right],\\
- & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\hat{\Gamma s},\nu)||\phi(\nu)|^{2}\exp\left[-i2\pi\nu\left(\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b}}{c}-\Gamma\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b'}}{c}\right)\right],
+ & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle\right)A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
+ & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\boldsymbol{\kappa})T(\boldsymbol{\kappa'})\rangle e^{-i(\boldsymbol{\kappa}\cdot \boldsymbol{r}-\boldsymbol{\kappa'}\cdot\boldsymbol{r'})}\right)A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
+ & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\boldsymbol{\kappa}\cdot(\boldsymbol{r}-\boldsymbol{r'})}\right)A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
+ & \approx\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}rd^{3}r'\delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
+ & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}r|A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r})||\phi(\nu_{r})|^{2}\exp\left[-i2\pi\nu_{r}\left(\hat{\boldsymbol{r}}\cdot\frac{\boldsymbol{b}}{c}-\Gamma \hat{\boldsymbol{r}}\cdot\frac{\boldsymbol{b'}}{c}\right)\right],\\
+ & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma \hat{\boldsymbol{s}},\nu)||\phi(\nu)|^{2}\exp\left[-i2\pi\nu\left(\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b}}{c}-\Gamma\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b'}}{c}\right)\right],
 \end{aligned}
@@
 \begin{equation}
-\Phi_{b,\tau}(\boldsymbol{r},\Gamma r')=|{\phi^{*}}(\ensuremath{\nu_{r}})\phi(\nu_{r'})|\exp\left[-i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\Gamma\hat{r'}\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)\right].
+\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'})=|{\phi^{*}}(\ensuremath{\nu_{r}})\phi(\nu_{r'})|\exp\left[-i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\Gamma\hat{r'}\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)\right].
 \end{equation}
 \end{widetext}
-The second to third line of Eq.(\ref{eq:main}) follows from assumption of Gaussian random
-sky, and the third to fourth line follows from the assumption that
+The second to third line of Eq.(\ref{eq:main}) follows from assumption of translational invariance, and the third to fourth line follows from the assumption that
 the 3D power spectrum varies negligibly over the $k$-space of interest
@@
 Before we discuss tests, there is one subtlety when applying Eq.
-\ref{eq:final}.
+\ref{eq:final} to the visibilities from correlators.
 The optimal time offset is given by the $\Gamma$ that maximizes the absolute value of the weight $\Theta$.
@@
 The physical origin of this frequency-dependent phase lies in the two visibilities having different phase centers.
-By default the two visibilities are both phased to zenith at the same time.
+By default the correlators phase the two visibilities  both to zenith at the same time.
 When they are cross-multiplied with a time lag, they must be rephased to the account for the movement of the zenith.
 Thus in practice we must rephase the visibilities before delay transform.
+Doing so is also crucial to reduction of additional mode mixing, as discussed in \cite{Hazelton2013}.
+Shapes of overlapping regions of pairs of baselines are fixed with respect to frequency.
 
@@
 The top two panels have zero rephasing, and the bottom two are rephased to a time offset of 0.055 sidereal days, the optimal time offset for the given baseline pair.
-The first and third panels show the equivalent baseline pairs sep1,0 against itself, and second and fourth panels show sep1,0 against sep1,1.
+The first and third panels show the equivalent baseline pairs \{1, 0\} against itself, and second and fourth panels show \{1, 0\} against \{1, 1\}.
 We see that in the un-rephased case in the second panel, although the magnitude of correlations match up for the two frequencies, the phases do not.
@@
 \begin{figure*}[h!]
-\includegraphics[width=\linewidth]{fringe_res}
+\includegraphics[width=0.9\textwidth]{fringe_res}
 \caption{Beam fringe pattern of sep2,0(left), sep2,1 at a time delay (middle),
@@
 \begin{figure*}[h!]
-\includegraphics[width=\linewidth]{sensitivity}
+\includegraphics[width=0.9\textwidth]{sensitivity}
 
@@
 \label{fig:sensplot}
+\end{figure*}
+
+\begin{figure*}[h!]
+\includegraphics[width=0.9\textwidth]{HeraAntpos}
+\label{fig:HeraAntpos}
+\caption{Planned Hydrogen Epoch of Reionization Array antenna configurations.
+HERA 37 is expected to complete and start collecting data in summer of 2017, and the other three are planned configurations in the next phases.
+For HERA350, only the 320 elements in the core are shown.
+}
 \end{figure*}
@@
 where the sum is over classes of baseline pairs. 
-We define the estimator sensitivity to be the inverse of the power-spectrum noise variance:
+We define the estimator sensitivity to be the inverse of the power spectrum noise variance:
 \begin{equation}
@@
 
-To move forward we must give an estimate of $\rho_0$ in Eq.\ref{eq:tildereal}.
+For a given $\rho_0$ Eq.\ref{eq:tildereal} thus quantifies the relative sensitivity contribution of a baseline pair $bb'$.
 Assuming a reionization signal of $\Delta_{21cm}^2\sim 30mK^2$, observation at 150MHz ($z=8.5$), 120 days of observation with PAPER antennas, we have roughly
@@
 The dashed lines represent the values when only the equivalent baseline-pairs are used.
-We see that in all cases using the near-equivalent baselines lead to more and more significant improvements with lower $\widetilde{\Theta}_{min}$, or in other words when worse baseline pairs are used.
+We see as expected that in all cases using the near-equivalent baselines lead to more and more significant improvements with lower $\widetilde{\Theta}_{min}$, or in other words when worse baseline pairs are used.
 The small HERA37, with no gap (like in HERA350) or short near-equivalent baselines (like in PAPER 128), will not benefit much from the near-equivalent baselines.
@@
 %\end{centered}
-
-\begin{figure*}[h!]
-\includegraphics[width=0.9\textwidth]{HeraAntpos}
-\label{fig:HeraAntpos}
-
-\caption{Planned Hydrogen Epoch of Reionization Array antenna configurations.
-HERA 37 is expected to complete and start collecting data in summer of 2017, and the other three are planned configurations in the next phases.
-For HERA350, only the 320 elements in the core are shown.
-}
-\end{figure*}
-
-In Fig.
-\ref{fig:pairplot} we present a comparison of 4 different properties for baselines that contribute well to the sensitivity ($\widetilde{\Theta}>0.05$, where again the weight for the top class of equivalent pair is normalized to 1).
-The $dT$ vs.
-$\Theta$ and $dT$ vs.
-$\widetilde{\Theta}$ plots are familiar from Fig.
-\ref{fig:sensplot}.
-We only show 3 of the mentioned arrays for visual clarity.
-The other two results are similar barring intuitive differences.
-We make a few observations:
-\begin{itemize}
-
-\item Mapping of equivalent baseline pairs: the equivalent baselines are located in key points in the plots.
-They are the vertical bars in the lower left four plots and the highest peak in the first three histograms, the upper "diagonal edge" in the $\widetilde{\Theta}$ vs.
-$dT$ and the dot of $\Theta=1$.
-
-\item Longer baselines roughly peak at lower time-offset, and lower multiplicity.
-This is intuitive for PAPER since  for the same north-south separation the rotation angle is smaller with longer east-west separation.
-
-\item Interestingly for HERA the near-equivalent baselines form multiple "U" shapes in $dT, \Theta$ plane, somewhat counter-intuitive especially given out last point.
-This is characteristic of the hexagonal layout.
-Baselines of equal length are more redundant at the right time delay than ones that are slightly different in length, though the latter has a shorter optimal time-delay.
-
-\item The top near-equivalent pairs in HERA the same $\Theta$ as in PAPER128, but much lower $\widetilde{\Theta}$.
-This is because they are longer baselines with lower multiplicity.
-In the end these baseline classes still lead to high contributions to total sensitivity (Fig.
-\ref{fig:osens}) because there are a lot more such baseline pairs for HERA.
-
-\end{itemize}
 
@@
 Shift the visibilities in time.
-3) Cross multiply the visibilities of the two baselines  to form the power-spectrum.
+3) Cross multiply the visibilities of the two baselines  to form the power spectrum.
 4) Finally combine the different baseline pairs by appropriate inverse-variance weighting that takes into account the predicted sensitivity contributions of each case.
 We showed that such techniques could lead to $20\%$ to $60\%$ increase in sensitivity for PAPER and HERA.
-\begin{widetext}
-\begin{figure}[H]
-\includegraphics[width=\textwidth]{pairplot}
-
-\caption{Pairplots of the top contributing baseline pairs in three arrays.
-Plotted properties are optimal time delay $dT$, peak height $\Theta$, effective weight $\widetilde{\Theta}$ and baseline length $L$.
-Only those points with $\widetilde{\Theta}>0.05$ are shown (the weight for the top class of equivalent pair is normalized to 1).
-Only one baseline length is shown since all top contributing pairs have very similar lengths as expected.
-The scatter plots are shown with transparency so that darker regions indicate degeneracies.
-Scatter points of the 3 arrays overlap, in order indicated by the legend.
-}
-\label{fig:pairplot}
-\end{figure}
-\end{widetext}
 
@@
 \end{equation}
+
+\section{\label{sec:fullstats}\\Additional Statistics}
+
+\begin{widetext}
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{pairplot}
+
+\caption{Pairplots of the top contributing baseline pairs in three arrays.
+Plotted properties are optimal time delay $dT$, peak height $\Theta$, effective weight $\widetilde{\Theta}$ and baseline length $L$.
+Only those points with $\widetilde{\Theta}>0.05$ are shown (the weight for the top class of equivalent pair is normalized to 1).
+Only one baseline length is shown since all top contributing pairs have very similar lengths as expected.
+The scatter plots are shown with transparency so that darker regions indicate degeneracies.
+Scatter points of the 3 arrays overlap, in order indicated by the legend.
+}
+\label{fig:pairplot}
+\end{figure}
+\end{widetext}
+
+In Fig.
+\ref{fig:pairplot} we present a comparison of 4 different properties for baselines that contribute well to the sensitivity ($\widetilde{\Theta}>0.05$, where again the weight for the top class of equivalent pair is normalized to 1).
+The $dT$ vs.
+$\Theta$ and $dT$ vs.
+$\widetilde{\Theta}$ plots are familiar from Fig.
+\ref{fig:sensplot}.
+We only show 3 of the mentioned arrays for visual clarity.
+The other two results are similar barring intuitive differences.
+We make a few observations:
+\begin{itemize}
+
+\item Mapping of equivalent baseline pairs: the equivalent baselines are located in key points in the plots.
+They are the vertical bars in the lower left four plots and the highest peak in the first three histograms, the upper "diagonal edge" in the $\widetilde{\Theta}$ vs.
+$dT$ and the dot of $\Theta=1$.
+
+\item Longer baselines roughly peak at lower time-offset, and lower multiplicity.
+This is intuitive for PAPER since  for the same north-south separation the rotation angle is smaller with longer east-west separation.
+
+\item Interestingly for HERA the near-equivalent baselines form multiple "U" shapes in $dT, \Theta$ plane, somewhat counter-intuitive especially given out last point.
+This is characteristic of the hexagonal layout.
+Baselines of equal length are more redundant at the right time delay than ones that are slightly different in length, though the latter has a shorter optimal time-delay.
+
+\item The top near-equivalent pairs in HERA the same $\Theta$ as in PAPER128, but much lower $\widetilde{\Theta}$.
+This is because they are longer baselines with lower multiplicity.
+In the end these baseline classes still lead to high contributions to total sensitivity (Fig.
+\ref{fig:osens}) because there are a lot more such baseline pairs for HERA.
+
+\end{itemize}
+
 \bibliographystyle{apj}

### 2017-05-03 Rotation paper/draft_working.tex
save

@@
 The observational effort to measure the cosmological 21\,cm power spectrum with radio interferometers requires high sensitivity.
-We introduce a straightforward and accurate method to extract sensitivity from visibility data.
-We develop our method in line with existing visibility-based power spectrum pipelines.
-Our method efficiently finds subsets of baselines to cross correlate, including partially redundant baselines, and quantifies the sensitivity contributions of given baselines.
-Using the 128-element configurations and beams of Precision Array for Probing the Epoch of Reionization (PAPER-128), as well as 4 planned versions of the Hydrogen Epoch of Reionization Array (HERA), we illustrate how our method applies to different arrays and predict the sensitivity improvements of including each baseline pairs.
-We show that inclusion of non-perfectly redundant baselines
+Current visibility-based power spectrum pipelines, though shown to ease control of systematics, lack the ability to include partial redundancy.
+We introduce a method to include partial redundancy in such power spectrum pipelines of drift-scan arrays.
+Our method efficiently finds pairs of baselines to cross-multiply, and quantifies the sensitivity contributions of given baselines.
+Using the 128-element configurations and beams of the Precision Array for Probing the Epoch of Reionization (PAPER-128), as well as 4 planned versions of the Hydrogen Epoch of Reionization Array (HERA), we illustrate how our method applies to different arrays and predict the sensitivity improvements of including each baseline pairs.
+We show that inclusion of partially redundant baselines
 would account for $20\%$ to $60\%$ of the sensitivity of PAPER-128 and different configurations of HERA. 
@@
 measurements with the 64-element version of PAPER (henceforth referred to as PAPER-64).
-However, there is one caveat with existing visibility-based pipelines.
+However, one disadvantage of existing visibility-based pipelines is the lack of use of partial redundancy.
+Baselines of different length and orientation still contain partially redundant information.
 While an imaging based power-spectrum pipeline naturally includes all redundancy information, visibility-based pipelines until now only cross multiply fully redundant baselines, in other words baselines of the same length and orientation.
 
-Baselines of different length and orientation still contain partially redundant information.
 The use of partially redundant baselines is in principle not a fundamental limitation of visibility-based pipelines.
 In fact, most instrumental reports to date do include partial redundancy.
-Recently, \cite{wterm} proposed an algorithmic approach to extract power-spectrum from partially redundant baselines in tracking measurements, by carefully treating the $w$-term in the fringe pattern to obtain weights for cross-correlation.
+Recently, \cite{wterm} proposed a visibility-based approach to extract power spectrum from partially redundant baselines in tracking measurements.
 Our work parallels this effort by focusing on arrays of the drift-scan type that do not have tracking capabilities, such as PAPER, HERA and SKA.
@@
 In this point of view, baselines that are slightly different in length and orientation
-rotate into each other at a time delay.
+``rotate into'' each other at a time delay.
 We can thus cross-multiply time-shifted visibilities, with the proper weighting, to form power spectrum.
-Due to the large number of elements of modern arrays (upward of 100), the task of cross-multiplying every baseline against every other, scaling as number of array element to the fourth power, can be computationally formidable.
+Due to the large number of elements of modern arrays (upward of 100), the task of cross-multiplying every baseline against every other, scaling as number of array element to the fourth power, can be computationally formidable, and many pairs of baselines provide only negligible redundancy information.
 Our contribution is thus twofold.
 First we introduce a formalism to include pairs of partially-redundant baselines in a visibility based power spectrum pipeline.
-Secondly we show how to use the formalism to automaticaly preselect baseline pairs and time offsets, making the problem computationally efficient.
+Secondly we show how to use the formalism to automatically pre-select baseline pairs and time offsets, making the problem computationally efficient.
 More precisely, our formalism allows one to simultaneously a) identify the baselines that give good
@@
 
-In order to eliminate confusion and ambiguity, we
-introduce some terminology for the rest of this paper, some of which differs from what is commonly found in the literature.
+In order to eliminate confusion and ambiguity for the rest of this paper, we
+introduce some terminology that may differ from what is commonly found in the literature.
 
-First, we make the distinction of a \textit{``baseline''}, which refers to two specific antennas, and a \textit{``class of baselines''}, which refers to all baselines of the same length and orientation in a given array.
-
+We make the distinction of a \textit{baseline}, which corresponds to two specific antennas, and a \textit{class of baselines}, which refers to all baselines of the same length and orientation in a given array.
 Baselines of the same class are traditionally called
@@
 in the sky.
-We shall call baselines in the same class \textit{``equivalent baselines''}, and reserve the word \textit{``redundancy''} of two baselines in reference to a variable function of the relative time-offsets of their visibility time series.
+We shall call baselines in the same class \textit{equivalent baselines}, and reserve the word \textit{redundancy} of two baselines in reference to a variable function of the relative time-offsets of their visibility time series.
 With this terminology, two equivalent
 baselines are fully redundant with each other simultaneously at all
-times. Non-equivalent baselines also have partial redundancy, which is maximized if their
+times. Non-equivalent baselines also have partial redundancy, and the redundancy can be maximized if their
 respected time series are shifted with respect to one another by some delay. In other words, one baseline can be ``rotated into''
-another if they are \textit{``near-equivalent''}.
+another if they are \textit{near-equivalent}.
 
@@
 an antenna. Top panel shows the antenna positions drawn to scale;
-bottom panel show the antenna labels and distances, excluding the outrigger
-antennas.
-The numbering of the antennas the bottom panel are original labels
+bottom panel show the antenna labels and distances of the 112-element grid, in other words excluding the outrigger antennas.
+The numbering of the antennas in the bottom panel are original labels
 during instrument assembly and does not bear significant
-meaning. In the bottom panel, the two baselines (5,77) and (40,85) are an example of equivalent pair, with separation denoted by \{1, 0\}, for the
+meaning. In the bottom panel, the two baselines (5,77) and (40,85) marked with red segments are example of an equivalent pair, both with separation denoted by \{1, 0\}, for the
 antennas are separated by 1 unit east and 0 unit north. Similarly,
-the baselines (5,69) and (5,85) are examples
-of the classes \{1, 1\} and \{1,-1\}, respectively. Note \{1, 0\} and \{-1, 0\} for example
-are the same baseline classes and should not be counted twice.}
+the baselines marked in blue and yellow are examples
+of the classes \{1, 1\} and \{1,-1\}, respectively. Due to the small North-South separation within the grid, \{1, 1\} is expected to be near-equivalent to \{1, 0\}. }
 \label{fig:AntPos}
@@
 S, 21:25:41.8 E). The layout pattern with antenna labels are show
-in Fig. \ref{fig:AntPos}. The array consists of a 112-element grid and 16 "outriggers" used primarily to aid calibration. We see the antenna spacing in North-South
-directions are comparatively close (4m), so that baselines such as
-(0, 44) and (0, 7) are very close to equivalent. In the bottom panel, the two baselines (5, 77) and (40, 85) are an example of equivalent pair, with separation denoted by \{1, 0\}, for the
+in Fig.
+\ref{fig:AntPos}.
+The array consists of a 112-element grid and 16 "outriggers" used primarily to aid calibration.
+In the bottom panel, the two baselines (5, 77) and (40, 85), marked in red, are an example of equivalent pair.
+We denote a equivalency-class of baselines in the PAPER grid by their separations, in this case  \{1, 0\}, for the
 antennas are separated by 1 unit east and 0 unit north. Similarly,
-the baselines (5, 69) and (5, 85) are examples
-of \{1, 1\} and \{1, -1\}, respectively.
-Note \{1, 0\} and \{-1, 0\} for example are the same classes and should
+the baselines marked in blue and yellow are respectively examples
+of \{1, 1\} and \{1, -1\}.
+Note \{1, 0\} and \{-1, 0\} for example are the same class and should
 not be counted twice. Antennas in purely north-south baselines
-are close enough to induce cross-talk, and hence are not suitable
-for use. The PAPER-64 analysis of \cite{Ali2015} used three classes of baselines, the PAPER-128
+are close (4m), and hence these baselines are not suitable
+for cross-multiplication due to cross-talks.
+On the other hand, the close North-South separation means that classes such as \{1, 0\} and  \{1, 1\} are expected to be near-equivalent.
+The PAPER-64 analysis of \cite{Ali2015} used three classes of baselines, the PAPER-128
 equivalent of which are 
 \{2, 0\}, \{2, 1\} and \{2, -1\}. There each of these classes
-of baselines are cross multiplied to itself.
+of baselines are cross multiplied within itself. This paper provides the method for inter-class multiplications.
 

### 2017-05-28 Rotation paper/draft_working.tex
save

@@
 regarding fundamental physics of inflation, neutrino mass and phenomenology
-of the first stars and galaxies (e.g. \citealt{LiuOpticalDepth, Liu2016b, Mao2008, DEw21cm, Bull2015, Oyama20131186} etc.).
+of the first stars and galaxies (e.g. \citealt{LiuOpticalDepth, Liu2016b, Mao2008, DEw21cm, Bull2015, Oyama20131186}).
 
@@
 GP tests rely on Ly$\alpha$ absorption, and thus easily saturate with small degree of ionization and can only probe the end stage of reionization.
-The 21\,cm line can in comparison directly probe the
-abundance and distribution of neutral hydrogen, and thus would potentially
+[other probes include cmb optical depth, cmb ksz, lyman alpha emitters (redshifted as in bubble to no longer be absorbed, big bubbles corresponding to clustered sources), GP, UV luminosity of high redshift galaxy] The 21\,cm brightness temperature distribution is a direct tracer of neutral hydrogen through the epoch of reionization, and thus measurements of the 21\,cm line would potentially
 shed light on all stages of reionization. Radio interferometric efforts
-to measure the 21\,cm power spectrum has been a top priority in recent years of astronomy.
+to measure the 21\,cm power spectrum [spatial pspec of brightness temp fluctuations] has been a top priority in recent years of astronomy.
 Current-generation instruments include the Precision Array for Probing
@@
 
-These modern low-frequency radio interferometers aimed at measuring the 21\,cm signal bears designs dramatically different from traditional instruments built for imaging localized bright sources.
+The highly redshifted 21\,cm signal is faint and diffuse, in contrast to the localized bright sources that have been of interest in many traditional applications of radio astronomy.
+Thus modern low-frequency radio interferometers aimed at measuring the 21\,cm signal bears designs different from traditional instruments.
+Current experiments are sensitivity-starved, with the sensitivity requirement further increased considering the foreground contaminations 5 orders of manitude brighter.
 With the goal of detecting a cosmological signal 5 orders of magnitude fainter than foreground contaminations, current efforts are sensitivity starved.
-Maximum-redundancy arrays are thus designed to repeatedly measure the same Fourier signal to increase sensitivity.
+The arrays are thus designed with multiple copies of the same baselines to repeatedly measure the same Fourier signal to increase sensitivity.
+To satisfy the sensitivity need, modern arrays are large.
 Ranging upwards from 100 elements, these arrays are typically of the drift-scan type to limit cost.
 
-Analysis pipelines of such maximum-redundancy arrays typically fall into two categories.
-In the first, more traditional technique, image is formed in Fourier domain through rotation synthesis, then foreground is either removed or avoided to construct a power spectrum.
-The second, lately popular technique works directly with visibilities from baselines and cross multiply them to form the power spectrum.
-Due to its more natural focus on the instrumental output of visibilities, this technique is shown to fair better in dealing with systematics such as RFIs and instrumental cross-talks, and thus tend to provide good limits on the power spectrum.
-\cite{Ali2015} provides the newest upper limit to the power spectrum
-measurements with the 64-element version of PAPER (henceforth referred to as PAPER-64).
+Analysis pipelines for these new arrays typically fall into two categories.
+In the first, more traditional technique, images are formed in Fourier domain through rotation synthesis, followed by a foreground mitigation step to construct a power spectrum.
+An alternative technique works directly with visibilities from baselines and cross multiplies them to form the power spectrum.
+Tracking closely the instrumental output of visibilities, this technique allows one to deal with systematics such as Radio-Frequency-Interference(RFI) and instrumental cross-talks in a more native context, and thus tend to provide good limits on the power spectrum.
+An example of the visibility-based pipeline was presented in \cite{Ali2015}, which provides the newest upper limit to the power spectrum
+measurements with the 64-element version of PAPER (henceforth as PAPER-64).
 However, one disadvantage of existing visibility-based pipelines is the lack of use of partial redundancy.
@@
 Recently, \cite{wterm} proposed a visibility-based approach to extract power spectrum from partially redundant baselines in tracking measurements.
-Our work parallels this effort by focusing on arrays of the drift-scan type that do not have tracking capabilities, such as PAPER, HERA and SKA.
+Arrays capable of tracking include MWA and LOFAR.
+Our work parallels this effort by focusing on drift-scan arrays that do not have tracking capabilities, such as PAPER, HERA and potentially SKA.
 The Earth's rotation causes the baselines in a drift-scan array to pick up different modes of the sky with time.
@@
 ``rotate into'' each other at a time delay.
-We can thus cross-multiply time-shifted visibilities, with the proper weighting, to form power spectrum.
-Due to the large number of elements of modern arrays (upward of 100), the task of cross-multiplying every baseline against every other, scaling as number of array element to the fourth power, can be computationally formidable, and many pairs of baselines provide only negligible redundancy information.
+We can thus cross-multiply time-shifted visibilities, with the proper weighting, to form power spectra.
+Due to the large number of elements of modern arrays (upward of 100), the task of cross-multiplying every baseline against every other, scaling as number of array elements to the fourth power, can be computationally formidable, and many pairs of baselines provide only negligible redundancy information.
 Our contribution is thus twofold.
@@
 Secondly we show how to use the formalism to automatically pre-select baseline pairs and time offsets, making the problem computationally efficient.
-More precisely, our formalism allows one to simultaneously a) identify the baselines that give good
-redundancy, b) find the time offset that corresponds to maximal redundancy for a given pair of baselines, and
-c) quantify the sensitivity associated with cross multiplying
+More precisely, our formalism allows one to simultaneously identify the baselines that give good
+redundancy, find the time offset that corresponds to maximal redundancy for a given pair of baselines, and quantify the sensitivity associated with cross multiplying
 such a pair of partially redundant baselines, which in turns is used as weight to combine measurements in a power spectrum pipeline.

### 2017-07-02 Rotation paper/draft_preWterm.tex
save

--- 2017-05-28 Rotation paper/draft_working.tex
@@
 \shorttitle{Adding Sensitivity to 21\,cm Inteferometric Probes of Reionization}
-\shortauthors{Zhang\&Parsons}
+\shortauthors{Zhang, Liu \&Parsons}
 \usepackage{float}
@@
 
-\title{Adding Sensitivity to 21\,cm Inteferometric Probes of Reionization by Optimizing Choice of Baselines}
+\title{A Generalized Visbility to Power-Spectrum Framework for 21\,cm Inteferometric Probes of Reionizations}
 
@@
 
-Arguably the most promising observational probe of the epoch of reionization
-comes from measurement of the ``spin-flip'' transition of neutral
+In the past, observational studies of reionization, including Gunn-Peterson measurements of quasi-stellar objects \citep{Fan2006} and cosmic microwave background (CMB) measurements of temperature and anisotropy \citep{Planck2016}, Kinetic Sunyaev-Zeldo'vich effect \citep{kszpatchy} and and Lyman alpha emitter clustering \citep{mcquinnLyA} have given us indications of the rough time-frame of reionization, but only limited constrains on the finer spatial and temporal variations.
+A surge of recent radio-astronomical experiments of reionization focus on measurement of the ``spin-flip'' transition of neutral
 hydrogen of characteristic wavelength 21\,cm \citep{Furlanetto2006181,PritchardLoeb}.
-Other observational studies of reionization include for example Gunn-Peterson measurements of quasi-stellar objects \citep{Fan2006}.
-GP tests rely on Ly$\alpha$ absorption, and thus easily saturate with small degree of ionization and can only probe the end stage of reionization.
-[other probes include cmb optical depth, cmb ksz, lyman alpha emitters (redshifted as in bubble to no longer be absorbed, big bubbles corresponding to clustered sources), GP, UV luminosity of high redshift galaxy] The 21\,cm brightness temperature distribution is a direct tracer of neutral hydrogen through the epoch of reionization, and thus measurements of the 21\,cm line would potentially
-shed light on all stages of reionization. Radio interferometric efforts
-to measure the 21\,cm power spectrum [spatial pspec of brightness temp fluctuations] has been a top priority in recent years of astronomy.
+In complement to the aforementioned probes, the 21\,cm brightness temperature distribution is a direct tracer of neutral hydrogen through the epoch of reionization, and thus tomography of the 21\,cm line is a direct measurement of the full temporal and spatial variations of this event.
+Before possible realization of full scale 21\,cm tomography, many current radio interferometric efforts
+prioritize measuring the spatial power spectrum of 21\,cm brightness temperature fluctuations.
 Current-generation instruments include the Precision Array for Probing
@@
 The highly redshifted 21\,cm signal is faint and diffuse, in contrast to the localized bright sources that have been of interest in many traditional applications of radio astronomy.
+Current experiments are sensitivity-starved, with the sensitivity requirement further increased considering the foreground contaminations 5 orders of magnitude brighter than the cosmological signal of interest.
 Thus modern low-frequency radio interferometers aimed at measuring the 21\,cm signal bears designs different from traditional instruments.
-Current experiments are sensitivity-starved, with the sensitivity requirement further increased considering the foreground contaminations 5 orders of manitude brighter.
-With the goal of detecting a cosmological signal 5 orders of magnitude fainter than foreground contaminations, current efforts are sensitivity starved.
-The arrays are thus designed with multiple copies of the same baselines to repeatedly measure the same Fourier signal to increase sensitivity.
+Many specially designed arrays, such as PAPER and HERA, feature multiple copies of the same baselines to repeatedly measure the same Fourier signal to increase sensitivity.
 To satisfy the sensitivity need, modern arrays are large.
@@
 
-Analysis pipelines for these new arrays typically fall into two categories.
+Analysis pipelines for the 21\,cm power spectrum typically fall into two categories.
 In the first, more traditional technique, images are formed in Fourier domain through rotation synthesis, followed by a foreground mitigation step to construct a power spectrum.
@@
 
-[TODO: this paragraph would potentially be promoted to a section] By using pairs of baselines, we also reduce additional mode mixing.
+%[TODO: this paragraph would potentially be promoted to a section] By using pairs of baselines, we also reduce additional mode mixing.
 As discussed in \cite{Hazelton2013}, the varying shapes of overlapping regions in $uv$ space of all contributing baselines lead to chromaticities in the response.
@@
 during instrument assembly and does not bear significant
-meaning. In the bottom panel, the two baselines (5,77) and (40,85) marked with red segments are example of an equivalent pair, both with separation denoted by \{1, 0\}, for the
+meaning. In the bottom panel, the two baselines (5,77) and (40,85) marked with red segments are example of an equivalent pair, both with separation denoted by \{1,0\}, for the
 antennas are separated by 1 unit east and 0 unit north. Similarly,
 the baselines marked in blue and yellow are examples
-of the classes \{1, 1\} and \{1,-1\}, respectively. Due to the small North-South separation within the grid, \{1, 1\} is expected to be near-equivalent to \{1, 0\}. }
+of the classes \{1,1\} and \{1,-1\}, respectively. Due to the small North-South separation within the grid, \{1,1\} is expected to be near-equivalent to \{1,0\}. }
 \label{fig:AntPos}
@@
 In the bottom panel, the two baselines (5, 77) and (40, 85), marked in red, are an example of equivalent pair.
-We denote a equivalency-class of baselines in the PAPER grid by their separations, in this case  \{1, 0\}, for the
+We denote a equivalency-class of baselines in the PAPER grid by their separations, in this case  \{1,0\}, for the
 antennas are separated by 1 unit east and 0 unit north. Similarly,
 the baselines marked in blue and yellow are respectively examples
-of \{1, 1\} and \{1, -1\}.
-Note \{1, 0\} and \{-1, 0\} for example are the same class and should
+of \{1,1\} and \{1, -1\}.
+Note \{1,0\} and \{-1, 0\} for example are the same class and should
 not be counted twice. Antennas in purely north-south baselines
@@
 for cross-multiplication due to cross-talks.
-On the other hand, the close North-South separation means that classes such as \{1, 0\} and  \{1, 1\} are expected to be near-equivalent.
+On the other hand, the close North-South separation means that classes such as \{1,0\} and  \{1,1\} are expected to be near-equivalent.
 The PAPER-64 analysis of \cite{Ali2015} used three classes of baselines, the PAPER-128
@@
 \{2, 0\}, \{2, 1\} and \{2, -1\}. There each of these classes
-of baselines are cross multiplied within itself. This paper provides the method for inter-class multiplications.
+of baselines are cross multiplied within itself.
+This paper provides the method for inter-class multiplications.
+To do so we will use the short hand notation $\{m,n\}:\{m',n'\}$ to denote a pair of baseline classes to be cross-multiplied.
 
@@
 
-\subsection{Rough Idea}
+\subsection{Rough Idea \label{sec:tracks}}
 
-Given a point source on the sky, each baseline maps the
-source to a point in $uv$ plane. As the Earth rotates with respect to
-the source, the point traces out tracks in the $uv$ plane. 
+Given a point source on the sky, each baseline maps to a point in $uv$ plane. As the Earth rotates with respect to
+the source, the points trace out tracks in the $uv$ plane. 
 We show in Fig. \ref{fig:Tracks} $uv$ tracks of PAPER-128 over 4.8 hours, at 0.15GHz, for a source that passes through zenith.
@@
 \ref{fig:Tracks}, there are many such crossings.
-Crossing tracks, however does not imply perfect redundancy.
-The reason lies in the finite size of the beams.
+However, crossing tracks do not imply perfect redundancy in the case of drift-scan arrays.
+In fact, track-crossing is
+not accurate enough for time offset determination, nor can it give estimate of the degree of redundancy.
+The reason lies in the finite size of the beams and the nature of a drift-scan measurement.
+This shall become evident in the next section, after we develop a more general formalism that accounts for the point spread function of the finite beams, and estimate the degree of redundancy for general combination of baselines at a general time-offset.
 We show sample beams of HERA and PAPER antennas in Fig.
 \ref{fig:Beam} for reference.
-The inaccuracy is expected to be even more significant for the smaller HERA beam, which has a larger corresponding $uv$ point spread function.
-
-Thus although it is a valid quick method to identify some baselines to cross-multiply, track-crossing is
-not accurate enough for time offset determination, nor can give estimate of the degree of redundancy.
-To determine the time-offset corresponding to maximal redundancy, we must develop a more general  formalism that accounts for the point spread function of the finite beams, and estimates the degree of redundancy for general combination of baselines at a general time-offset, as we do in the next section.
 
@@
 
-\caption{Sample beam response of HERA (left) and PAPER (right) antennas, both
-for frequency of $\nu=150\text{MHz}$ and Stokes $I$ polarization.
+\caption{Example beam response of HERA (left) and PAPER (right) antennas.
+Both beams are Stokes $I$ polarization antenna voltage beams, at frequency of $\nu=150\text{MHz}$ and normalized to $1$ at zenith.
 Notice $A$ in this paper is a ``baseline's beam'', equivalent to squares of the antenna voltage beams shown here.
 The circles centered around zenith (center of beam) here are
-spaced 10 degrees apart. The finite size of the beams limits the accuracy of the $uv$ track-crossing as a method of redundancy search. \label{fig:Beam}}
+spaced 10 degrees apart. \label{fig:Beam}}
 \end{figure}
@@
 \subsection{Formalism}
-Below we shall derive theoretical expectations of cross multiplications
-of two near-equivalent baselines. More precisely we shall relate
-the product of visibilities of two different baselines to the power spectrum.
+In this section we formulate the relation between 
+the product of visibilities from two arbitrary baselines and the power spectrum.
 
-We take the visibility as commonly defined in the literature (e.g.
+We begin with the visibility as commonly defined in the literature (e.g.
 \citealt{first-paper}): 
 \begin{equation}
-\begin{aligned}V_{\nu}(\boldsymbol{b}) & =\int d\Omega A_{\nu}(\hat{\boldsymbol{s}})\phi(\nu)I_{\nu}(\hat{\boldsymbol{s}})\exp\left[-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right],\\
- & \approx\frac{2k_{B}}{\lambda^{2}}\int d\Omega A_{\nu}(\hat{\boldsymbol{s}})\phi(\nu)T(\hat{\boldsymbol{s}})\exp\left[-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right],
+\begin{aligned}V_{\nu}(\boldsymbol{b}) & =\int d\Omega A_{\nu}(\hat{\boldsymbol{s}})\phi(\nu)I_{\nu}(\hat{\boldsymbol{s}})\exp\left(-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right),\\
+ & \approx\frac{2k_{B}}{\lambda^{2}}\int d\Omega A_{\nu}(\hat{\boldsymbol{s}})\phi(\nu)T(\hat{\boldsymbol{s}})\exp\left(-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right),
 \end{aligned}
@@
 \end{equation}
-Here $\lambda$ is a mean wavelength, $\boldsymbol{b}$ is the baseline
+Here $k_B$ is Boltzmann's constant, $\lambda$ is the mean wavelength, $\boldsymbol{b}$ is the baseline
 length, $\hat{\boldsymbol{s}}$ and $\Omega$ are a direction in the
-sky and its corresponding solid angle. $A_{\nu}$ is the (frequency
-dependent) beam, and $I$ is the specific intensity, which has been
+sky and its corresponding solid angle. Inside the integral we have $\phi(\nu)$ the frequency bandpass profile, $A_{\nu}$ the (frequency
+dependent) primary beam, and $I$ the specific intensity, which has been
 related to $T$, the brightness temperature in the Rayleigh-Jeans
-limit.
-$k_B$ is Boltzmann's constant.
-Note that $A$ here is the baseline beam, i.e.
-product of two antenna voltage beams, examples of which are shown in Fig.
-\ref{fig:Beam}.
-$\phi(\nu)$ is the frequency bandpass profile.
-In practice,
+limit. The beam reception pattern $A$ is dimensionless,
+normalized to 1 at its peak (zenith), and we assume it to be the same
+for all baselines. In practice,
 power spectrum measurements are typically taken from a few ten MHz centered around the corresponding redshift of interest (e.g.
 150 MHz for z=9.5).
-We take a moment to note that in Eq.
-\ref{eq:Vis1}, no flat-sky approximation has yet been made.
-The angular integral is performed over the dome $d\Omega$.
+Note that in Eq.
+\ref{eq:Vis1}, no flat-sky approximation has yet been made; the angular integral is performed over the dome.
 
-We define the delay-transformed visibility \citep{delay-transform}:
+Delayed-transformed visibility has been widely adopted in recent years since it has been shown that foregrounds are isolated in delay space.
+We define the delay-transformed visibility by Fourier transforming visibility along the frequency axis\citep{delay-transform} as:
 \small
 \begin{equation}
-\begin{aligned}V(\boldsymbol{b},\tau) & =\int d\nu V_{\nu}(\boldsymbol{b})\phi(\nu)\exp\left[-2\pi i\nu\tau\right],\\
- & =\frac{2k_{B}}{\lambda^{2}}\int d\Omega d\nu A(\hat{\boldsymbol{s}},\nu)\phi(\nu)T(\hat{\boldsymbol{s}},\nu)\exp\left[-2\pi i\nu\left(\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}+\tau\right)\right]
+\begin{aligned}V(\boldsymbol{b},\tau) & =\int d\nu V_{\nu}(\boldsymbol{b})\phi(\nu)\exp\left(-2\pi i\nu\tau\right),\\
+ & =\frac{2k_{B}}{\lambda^{2}}\int d\Omega d\nu A(\hat{\boldsymbol{s}},\nu)\phi(\nu)T(\hat{\boldsymbol{s}},\nu)\exp\left[-2\pi i\nu\left(\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}+\tau\right)\right].
 \end{aligned}
-.\label{eq:Vb1}
+\label{eq:Vb1}
 \end{equation}
 \normalsize
-
-Eq. \eqref{eq:Vb1} expresses the delay-transformed visibility as
-an integral over observation coordinates $\hat{\boldsymbol{s}}$ and $\nu$. Ultimately,
+Here the delay $\tau$ is the Fourier variable. Eq. \eqref{eq:Vb1} expresses the delay-transformed visibility as
+an integral over observational coordinates $\hat{\boldsymbol{s}}$ and $\nu$. Ultimately,
 we would like to relate the data, collected with coordinates $\hat{\boldsymbol{s}}$
-and $\nu$, to the power spectrum, written with cosmological coordinates
-$\boldsymbol{r}$ and $\boldsymbol{k}$. We start by noticing that
+and $\nu$, to the power spectrum, written with 
+$\boldsymbol{r}$ and $\boldsymbol{k}$, the cosmological position coordinate and wavenumber. To do so, we start by noticing that
 \[
@@
  & \approx\frac{c}{H_{0}}\int_{0}^{z_{0}}\frac{dz'}{E(z')}-\frac{c(1+z)^{2}}{\nu_{21}H_{0}E(z)}\left(\nu-\nu_{0}\right),\\
- & \equiv D_{c}-Y\Delta\nu,
-\end{aligned}
+ & \equiv X-Y\Delta\nu,
+\end{aligned} \label{eq:r}
 \]
@@
 \begin{equation}
-\nu=\frac{D_{c}-r}{Y}+\nu_{21}.\label{eq:nur}
+\nu=\frac{X-r}{Y}+\nu_{21}.\label{eq:nur}
 \end{equation}
 
-Further we shall invoke the flat-sky limit
+Further we shall invoke the flat-sky limit, with the $z$ direction points to zenith, and $x$ and $y$ point east and north, respectively:
 \begin{equation}
-(r_{x},r_{y})=(X\hat{\boldsymbol{s}}_{x},X\hat{\boldsymbol{s}}_{y}), 
+(r_{x},r_{y})=(X\hat{\boldsymbol{s}}_{x},X\hat{\boldsymbol{s}}_{y}). 
 \end{equation}
-where
-\begin{equation}
-X = \frac{c}{H_0}\int_{0}^{z}\frac{dz'}{E(z')}
-\end{equation}
-is a cosmological scalar. In the same limit $(Xk_{x},Xk_{y},Yk_{z})=\frac{2\pi}{c}(b_{x},b_{y},\tau)$ relates
-the cosmological $k$ to the observation coordinates. For a more rigorous treatment without the flat sky limit, see Appendix [ADRIAN].
+In the same limit $(Xk_{x},Xk_{y},Yk_{z})=\frac{2\pi}{c}(b_{x},b_{y},\tau)$ relates
+the cosmological $k$ to the observational coordinates. For a more rigorous treatment without the flat sky limit, see Appendix [ADRIAN].
 
-We can rewrite the delayed transformed visibility as 
+We can rewrite the delayed-transformed visibility as 
 \small
 \[
-\begin{aligned}V(\boldsymbol{b},\tau) & =\frac{2k_{B}}{\lambda^{2}}\int_{H}\frac{d^{3}r}{X^{2}Y}A(\boldsymbol{r})\phi(r)T(\boldsymbol{r})\exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\hat{\boldsymbol{r}}+\tau\right)\nu_{r}\right]\end{aligned},
+\begin{aligned}V(\boldsymbol{b},\tau) & =\frac{2k_{B}}{\lambda^{2}}\int\frac{d^{3}r}{X^{2}Y}A(\boldsymbol{r})\phi(r)T(\boldsymbol{r})\exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\hat{\boldsymbol{r}}+\tau\right)\nu_{r}\right]\end{aligned},
 \]
@@
  where $d\nu=-dr/Y$ and $d^{3}r=-X^{2}Yd\Omega d\nu$. 
+We have written $\nu_{r}$ as a reminder that $\nu$ and $r$ are related
+by Eq. \eqref{eq:nur}.
 
-We have written $\nu_{r}$ to remind us that $\nu$ and $r$ are related
-by Eq. \eqref{eq:nur}. The beam reception pattern $A$ is dimensionless,
-normalized to 1 at its peak (zenith), and we assume it to be the same
-for all baselines.
-
-With a time offset, the beam pattern has moved relative to the sky.
-Here we choose to fix the sky, and denote the rotated coordinates
-of the beam pattern with the 3 dimensional rotation operator $\Gamma$. With implicit bounds
-of integrals from $-\infty$ to $\infty$, we have:
+Most existing visibility based power-spectrum pipelines for redundant drift-scan  arrays relate the power-spectrum to the conjugate square of the visibilities \citep{first-paper, paper32, Ali2015}.
+We would like to generalize such relations by relating the power-spectrum to the product of two visibilities from two arbitrary baselines and time-offsets.
+The beam pattern of a baseline shifts relative to the sky as the Earth rotates. Here we choose to fix the sky, and denote the rotated coordinates
+of the beam pattern with the 3 dimensional rotation operator $\Gamma$:
+\small
+\[
+\begin{aligned}V(\boldsymbol{b'},\tau') & =\frac{2k_{B}}{\lambda^{2}}\int\frac{d^{3}r}{X^{2}Y}A(\Gamma\boldsymbol{r})\phi(r)T(\boldsymbol{r})\exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\Gamma\hat{\boldsymbol{r}}+\tau'\right)\nu_{r}\right]\end{aligned}.
+\]
+\normalsize
+With implicit bounds of integrals from $-\infty$ to $\infty$, we have:
 \begin{widetext}
@@
  & \approx\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}rd^{3}r'\delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
- & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}r|A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r})||\phi(\nu_{r})|^{2}\exp\left[-i2\pi\nu_{r}\left(\hat{\boldsymbol{r}}\cdot\frac{\boldsymbol{b}}{c}-\Gamma \hat{\boldsymbol{r}}\cdot\frac{\boldsymbol{b'}}{c}\right)\right],\\
- & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma \hat{\boldsymbol{s}},\nu)||\phi(\nu)|^{2}\exp\left[-i2\pi\nu\left(\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b}}{c}-\Gamma\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b'}}{c}\right)\right],
+ & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}r|A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r})||\phi(\nu_{r})|^{2}\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\Gamma \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)\right],\\
+ & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma \hat{\boldsymbol{s}},\nu)||\phi(\nu)|^{2}\exp\left[-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)\right],
 \end{aligned}
@@
 \begin{equation}
-\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'})=|{\phi^{*}}(\ensuremath{\nu_{r}})\phi(\nu_{r'})|\exp\left[-i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\Gamma\hat{r'}\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)\right].
+\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'})=|{\phi^{*}}(\ensuremath{\nu_{r}})\phi(\nu_{r'})|\exp\left[-i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\Gamma\hat{\boldsymbol{r'}}\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)\right].
 \end{equation}
 \end{widetext}
-The second to third line of Eq.(\ref{eq:main}) follows from assumption of translational invariance, and the third to fourth line follows from the assumption that
+The third equality of Eq.(\ref{eq:main}) follows from assumption of translational invariance of sky statistics, and the first inequality follows from the assumption that
 the 3D power spectrum varies negligibly over the $k$-space of interest
-so that $\hat{P}_{21}(k+k_{2})\approx\hat{P}_{21}(k)$. Since $\Gamma$
+so that $\hat{P}_{21}(k+k')\approx\hat{P}_{21}(k)$. Since $\Gamma$
 is a sky rotation, it does not affect $\nu$, hence we have taken $\nu_{r}$
-outside the parenthesis. Notice that the phase factor $\exp\left[-i2\pi\tau\left(\nu-\nu'\right)\right]$
-drops out in the end. This means that correlation is peaked at the same time-offset for all delay channels.
+outside the parenthesis in the phase term $\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\Gamma \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)\right]$.
+Notice that the phase factor $\exp\left[-i2\pi\tau\left(\nu-\nu'\right)\right]$
+drops out in the end. This means that the location and magnitude of the correlation peak does not depend on delay $\tau$, an important point we shall come back to.
 
-Finally, since the beam pattern and bandwidth are given in $\hat{\boldsymbol{s}}$
+Since the beam pattern and bandwidth are given in $\hat{\boldsymbol{s}}$
 and $\nu$, we convert the integral back to these coordinates to get
@@
  & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)||\phi(\nu)|^{2}\\
- & \qquad \qquad \qquad \qquad \exp\left[-i2\pi\nu\left(\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b}}{c}-\Gamma\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b'}}{c}\right)\right].\end{aligned}
+ & \qquad \qquad \qquad \qquad \exp\left[-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)\right].\end{aligned}
 \label{eq:final}
@@
 \begin{equation}
-\Theta =\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)||\phi(\nu)|^{2} e^{-2\pi i\nu\left(\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b}}{c}-\Gamma\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b'}}{c}\right)}.
+\Theta \equiv\int d\nu \Theta_{\nu}, 
 \label{eq:Theta}
 \end{equation}
+and 
+\[
+\Theta_{\nu}\equiv \int \frac{d\Omega}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)||\phi(\nu)|^{2} e^{-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)}e^{i\psi}.
+\label{eq:Thetanu}
+\]
+Notice that $\Theta$ has no dependence on $\tau$. In Eq. \ref{eq:Thetanu} we introduced an extra phase $\psi$, the origin of which we shall explain in \ref{sec:rephs}.
 
-So roughly speaking the cross multiplications of visibilities at a time delay
+Roughly speaking, Eq. \ref{eq:opp} through \ref{eq:Thetanu} tell us that the cross multiplications of visibilities at a time delay
 in $uv$-space is proportional to the power spectrum times the Fourier
-transform of the cross multiplied beam pattern. We can then in principle
-combine information from different baseline pairs if we correct for
-the phase and normalization. As a check, when applied to equivalent baselines,
+transform of the cross multiplied beam pattern. As a check, when applied to equivalent baselines,
 $\boldsymbol{b}=\boldsymbol{b'}$, $\hat{\boldsymbol{s}}=\Gamma\hat{\boldsymbol{s}}$, and Eq.(\ref{eq:final}) reduces to Eq.(B9) of \cite{paper32}. 
-
 With Eq.
 \ref{eq:opp} and Eq.
-\ref{eq:Theta} we can for any given baseline pair and time delay, estimate the degree of redundancy, here represented by $\Theta$, thereby achieving all our goals stated in the introduction, i.e.
+\ref{eq:Theta} we can, for any given pair of baseline classes and time delay, estimate the degree of redundancy, here represented by $\Theta$, thereby achieving all our goals stated in the introduction, i.e.
 to identify
@@
 We can do all the above simply by computing the weight $\Theta$ from
-Eq.(\ref{eq:Theta}) for various time offsets, without having to actually cross-multiply visibilities at different offsets. This makes the task computationally tractable.
+Eq.(\ref{eq:Theta}) for various time offsets, without having to actually cross-multiply visibilities for all baselines and offsets. This makes the task computationally tractable.
+
+\section{Analysis}
+
+\subsection{Track-crossing, beam, and phase}
+We are now in a position to understand the reason that track-crossing, as mentioned section \ref{sec:tracks}, does not correspond exactly to the peak of correlation.
+Typical convention of rotation synthesis has rotation operates on $\boldsymbol{b'}$ instead of $\hat{\boldsymbol{s}}$, in which case Eq.\ref{eq:Theta} becomes
+\begin{equation}
+\Theta \equiv\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)||\phi(\nu)|^{2} e^{-i2\pi\frac{\nu}{c}\hat{\boldsymbol{s}}\cdot\left(\boldsymbol{b}-\widetilde{\Gamma}\boldsymbol{b'}\right)},
+\end{equation}
+where $\widetilde{\Gamma}$ is the inverse of $\Gamma$. Track-crossing corresponds to vanishing of the exponent
+\[
+ \boldsymbol{b}-\widetilde{\Gamma}\boldsymbol{b'} = 0. \label{eq:trackcross}
+ \]
+This condition maximizes $\Theta$ if and only if no other term in the integral depends on $\Gamma$, i.e.
+if and only if $A(\Gamma\hat{\boldsymbol{s}},\nu)=A(\hat{\boldsymbol{s}},\nu)$.
+If the measurement is tracking, $\Gamma$ becomes a rotation around zenith, and the condition is satisfied.
+In a general drift-scan measurement, this is not the case, and track-crossing point of Eq.
+\ref{eq:trackcross} does not maximize $\Theta$.
+
+\begin{figure}[h]
+\includegraphics[width=\linewidth]{opp}
+
+\caption{Numerical comparisons of the visibility correlation peaks to the $\Omega$ factor in Eq.(\ref{eq:final}).
+We generated 2500 instances of Gaussian random sky on healpix maps, computed visibilities and cross correlated them to find the correlation.
+Top panel shows the equivalent baseline pairs \{1,0\} against \{1,0\}, bottom panel shows set1,0 against \{1,1\}.
+Evaluations of weight $\Theta$ are shown with solid curves, and the visibility correlations of simulated random sky is given with dashed lines.
+In both cases blue denotes real part, green the imaginary part, and red the magnitude.
+We see that in both cases the theory and simulation line up in both amplitude and phase.}
+\label{fig:numerics}
+\end{figure}
+
+We shall illustrate with an example from PAPER128.
+In the solid curves of Fig.
+\ref{fig:numerics} we show examples of $\Theta_{\nu}$ as a function of time offset  here with no additional phase.
+The amplitude and real, as well as imaginary parts are shown for equivalent \{1,0\}:\{1,0\} in the top panel, and near-equivalent \{1,0\}:\{1,1\} in the bottom.
+The vertical axis is normalized such that the peak of \{1,0\}:\{1,0\} is unity.
+We see that as expected, \{1,0\}:\{1,0\} peaks at $dT=0$, where $\Theta_{\nu}$ is real.
+\{1,0\}:\{1,1\} on the other hand, peaks with amplitude of approximately 0.4 at $dT\approx0.055$ days and is not real at its peak.
+Since track-crossing corresponds to vanishing exponent, correlation at track-cross points are real.
+This in the bottom plot corresponds to roughly $dT=0.05$ days.
+
+Shown in dashed curves, which almost completely overlap with the solid curves, in Fig.
+\ref{fig:numerics} are cross-multiplications of simulated visibilities.
+For the simulation, we take N=2500 random realizations of the sky
+on a healpix map \citep{Heal, HealPrimer} \footnote{We use functionalities in the python package AIPY for healpix mapping
+as well as coordinate transforms. }, and pick two baselines \{1,0\} and \{1,1\}. For each realization, each pixel is given a Gaussian random value
+of brightness temperature. We then rotate the baseline positions with
+the appropriate rotation matrix, multiplying the sky by the primary beam to get the visibilities, for each
+baseline\footnote{There are two obvious ways to achieve the rotation. One
+can either fix the sky and rotate the baselines, or the other way
+around. We found however, that we must not physically rotate the sky
+map, for the numerical round-offs due to finite resolutions of the
+map turns out to be significant. Thus we let the sky, represented
+by the healpix map, be fixed, and rotate the baselines. }. The resulting visibilities for the two baselines are then convolved
+via the Fourier convolution theorem, to obtain values of the cross
+correlation as a function of time-offset. The peak of the curve then
+corresponds to maximum redundancy.  We do this for both the equivalent (\{1,0\}:\{1,0\}) and near-equivalent (\{1,0\}:\{1,1\}) case.
 
 \subsection{Rephasing \label{sec:rephs}}
-Before we discuss tests, there is one subtlety when applying Eq.
-\ref{eq:final} to the visibilities from correlators.
-The optimal time offset is given by the $\Gamma$ that maximizes the absolute value of the weight $\Theta$.
-However, $\Theta$ is in general complex.
-First integrating over the spatial dimension $\Omega$, we see the phase term at peak time-offset is inevitably frequency dependent.
-This frequency dependence of the phase would lead to destructive interference when we integrate over frequency.
+As we see in Fig.\ref{fig:numerics}, $\Theta$ at the peak of correlation is in general complex, and often far from real.
+Furthermore, this phase of peak correlation is inevitably frequency dependent, as we see in Eq.
+\ref{eq:Thetanu}.
+This frequency dependence of the phase would lead to destructive interference when we integrate over frequency, unless we correct one of the visibilities by an extra phase $\psi_\nu$.
 
-The physical origin of this frequency-dependent phase lies in the two visibilities having different phase centers.
-By default the correlators phase the two visibilities  both to zenith at the same time.
+The physical origin of $\psi_\nu$ lies in the two visibilities having different phase centers.
+By default the correlators of a drift-scan array phase the two visibilities both to zenith at the same time.
 When they are cross-multiplied with a time lag, they must be rephased to the account for the movement of the zenith.
 Thus in practice we must rephase the visibilities before delay transform.
-Doing so is also crucial to reduction of additional mode mixing, as discussed in \cite{Hazelton2013}.
-Shapes of overlapping regions of pairs of baselines are fixed with respect to frequency.
+%Doing so is also crucial to reduction of additional mode mixing, as discussed in \cite{Hazelton2013}. Shapes of overlapping regions of pairs of baselines are fixed with respect to frequency.
 
@@
 The top two panels have zero rephasing, and the bottom two are rephased to a time offset of 0.055 sidereal days, the optimal time offset for the given baseline pair.
-The first and third panels show the equivalent baseline pairs \{1, 0\} against itself, and second and fourth panels show \{1, 0\} against \{1, 1\}.
+The first and third panels show the equivalent baseline pairs \{1,0\}:\{1,0\}, and second and fourth panels show \{1,0\}:\{1,1\}.
 We see that in the un-rephased case in the second panel, although the magnitude of correlations match up for the two frequencies, the phases do not.
@@
 \caption{Comparisons of the peak phases of two different frequencies.
+Specifically shown are the real parts of $\Theta_{\nu}$.
 First and third panels shows equivalent baselines, second and fourth show a pair of near-equivalents.
-For visual simplicity only the real parts are shown.
+The top two panels have zero phase shift ($\psi_{\nu}=0$), and the bottom two are rephased to a time offset of 0.055 sidereal days.
+The first and last panels thus show a coherently rephased series that would add constructively at the respective peaks.
 }
@@
 
-\section{Analysis}
+Once rephased, the overall weight can be constructed by $\Theta=\int \Theta_{\nu}d\nu$.
+We must numerically check that rephasing the weight as in Eq.\ref{eq:Thetanu} is indeed equivalent to rephasing the input visibilities before delay-transform.
+In other words, we want to verify that
+\begin{equation}
+\begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau')\rangle\\
+ & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)||\phi(\nu)|^{2}\\
+ & \qquad \qquad \qquad \qquad \exp\left[-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)\right].\end{aligned}
+\label{eq:final}
+\end{equation}
+
 \subsection{Numerical Test \label{sec:Techniquet}}
@@
 
-For the simulation, we take N=2500 random realizations of the sky
-on a healpix map \citep{Heal, HealPrimer} \footnote{We use functionalities in the python package AIPY for healpix mapping
-as well as coordinate transforms. }, and pick two baselines sep1,0 and sep1,1. For each realization, each pixel is given a Gaussian random value
-of brightness temperature. We then rotate the baseline positions with
-the appropriate rotation matrix, multiplying the sky by the primary beam to get the visibilities, for each
-baseline\footnote{There are two obvious ways to achieve the rotation. One
-can either fix the sky and rotate the baselines, or the other way
-around. We found however, that we must not physically rotate the sky
-map, for the numerical round-offs due to finite resolutions of the
-map turns out to be significant. Thus we let the sky, represented
-by the healpix map, be fixed, and rotate the baselines. }. The resulting visibilities for the two baselines are then convolved
-via the Fourier convolution theorem, to obtain values of the cross
-correlation as a function of time-offset. The peak of the curve then
-corresponds to maximum redundancy.  We do this for both the equivalent (sep1,0 against sep1,0) and near-equivalent (sep1,0 against sep1,1) case. The accuracy of this result is
+ The accuracy of this result is
 limited by (simulated) cosmic variance and finite spacial resolution,
@@
 
-\begin{figure}[h]
-\includegraphics[width=\linewidth]{opp}
-
-\caption{Numerical comparisons of the visibility correlation peaks to the $\Omega$ factor in Eq.(\ref{eq:final}).
-We generated 2400 instances of Gaussian random sky on healpix maps, computed visibilities and cross correlated them to find the correlation.
-Top panel shows the equivalent baseline pairs sep1,0 against sep1,0, bottom panel shows set1,0 against sep1,1.
-Evaluations of weight $\Theta$ are shown with solid curves, and the visibility correlations of simulated random sky is given with dashed lines.
-In both cases blue denotes real part, green the imaginary part, and red the magnitude.
-We see that in both cases the theory and simulation line up in both amplitude and phase.}
-\label{fig:numerics}
-\end{figure}
-
 The comparison is show in Fig. \ref{fig:numerics}. 
 Sensitivity contribution is inferred from height of the peak of the cross-multiplied visibilities as a function of time. 
-We first compare the simulation to the analytical weight $\Theta$ for a pair of equivalent baselines (sep1,0, sep1,0), then normalize the peaks 
-of the near-equivalent baselines (sep1,0, sep1,1) to that of the equivalent ones. 
+We first compare the simulation to the analytical weight $\Theta$ for a pair of equivalent baselines (\{1,0\}, \{1,0\}), then normalize the peaks 
+of the near-equivalent baselines (\{1,0\}, \{1,1\}) to that of the equivalent ones. 
 In Fig. \ref{fig:numerics}  we show on the top comparison of
@@
 peak, whose location is at zero time offset as expected. On the bottom we show
-comparison of a near-equivalent baseline pair, of classes sep1,0:1,1 and sep1,1:1,1. The peaks
+comparison of a near-equivalent baseline pair, of classes \{1,0\}:\{1,1\} and \{1,1\}:\{1,1\}. The peaks
 are normalized with the same factor as the peak height on the top,
@@
 in Fig. \ref{fig:Beam-fringe-pattern}. The left and middle panels show the beam fringe
-patterns for baselines sep2,0 and sep2,1, delayed by 0.0325 sidereal days,
+patterns for baselines \{2,0\} and \{2,1\}, delayed by 0.0325 sidereal days,
 and the right panel shows their cross product. The fringe pattern
@@
 \includegraphics[width=0.9\textwidth]{fringe_res}
-\caption{Beam fringe pattern of sep2,0(left), sep2,1 at a time delay (middle),
+\caption{Beam fringe pattern of \{2,0\}(left), \{2,1\} at a time delay (middle),
 and their conjugate product (right). Frequency of $\nu=0.15\text{GHz}$ is
@@
 \begin{figure*}[h!]
-\includegraphics[width=0.9\textwidth]{sensitivity}
+\includegraphics[width=\textwidth]{sensitivity_1}
 
@@
 combination, while the bottom panel multiplies the heights by the
-corresponding multiplicities as in Eq. \eqref{eq:sensul}, and in some cases an extra factor of $\sqrt{2}$ (explained in text). 
+corresponding multiplicities as in Eq. \eqref{eq:sensul}. 
 In weighting by the multiplicity (bottom), we have chosen to fix the sensitivity 
@@
 give the same amount of redundancies (peak height), with the opposite time offset, as expected from symmetry. 
-For example, sep1,0:1,1 is mirror image of sep1,0:1,-1 and these two baseline pairs
+For example, \{1,0\}:\{1,1\} is mirror image of \{1,0\}:\{1,-1\} and these two baseline pairs
 give the same sensibility contribution.
@@
 to the sky and hence the less overlaps in patch of sky surveyed, b) smaller optimal
-time-offset corresponds to smaller differences in orientation, and hence in length of
-baselines in PAPER.
+time-offset corresponds to smaller differences in orientation and length of the 
+pair of baselines.
 
@@
 same length and orientation. Looking at Fig. \ref{fig:sensplot}
-we see for example sep1,0 will have higher multiplicity than sep2,0,
-or sep1,1. The latest release of PAPER-64 data uses the 128-equivalent baselines sep2,1,
-sep2,0 and sep2,-1 \citep{Ali2015}, and achieved a $2\sigma$ upper
+we see for example \{1,0\} will have higher multiplicity than \{2,0\},
+or \{1,1\}. The latest release of PAPER-64 data uses the 128-equivalent baselines \{2,1\},
+\{2,0\} and \{2,-1\} \citep{Ali2015}, and achieved a $2\sigma$ upper
 limit of $(22.4\text{mK})^{2}$. There, the three sets of equivalent baselines
@@
 First we can average of the visibilities of the equivalent baselines. Since the core of PAPER-128 has 16 by 7 antenna configuration, there
-are $M=(16-|m|)\times(7-|n|)$ copies of the baseline sepm,n. This means
+are $M\equiv(16-|m|)\times(7-|n|)$ copies of the baseline class $\{m,n\}$. This means
 that if we add visibility measurements of all these equivalent baselines,
@@
 level $\sigma_N$ of the visibility.
-The sensitivity contribution of sepm,n, cross multiplied with sepm',n'  thus roughly speaking scales as $\sqrt{\left((16-|m|)(7-|n|)(16-|m'|)(7-|n'|)\right)}=\sqrt{MM'}$.
+The sensitivity contribution of $\{m,n\}$, cross multiplied with $\{m',n'\}$  thus roughly speaking scales as $\sqrt{\left((16-|m|)(7-|n|)(16-|m'|)(7-|n'|)\right)}=\sqrt{MM'}$.
 For cross-multiplications of near-equivalent baselines of
-types sepm,n and sepm',n', we get an effective weight:
+types $\{m,n\}$ and $\{m',n'\}$, we get an effective weight:
 
@@
 Points that have zero time delay are the equivalent baseline pairs and their weighted correlation values simply reflect the multiplicity factor.
-For clarity of presentation we have ``folded over'' the negative time delays and combined baseline pairs that are identical modulus parity,
-or in other words are mirror images of each other
-%\footnote{The parity symmetry is partially broken unless the array is located at the poles.
-But we are ignoring this here for illustrative purposes.
-See Fig.
-\ref{fig:pairplot} for a more accurate result.
-}
-, by summing over them to get an extra factor of $\sqrt{2}$. For example, 
-instead of plotting 1,1:1,1 and 1,-1:1,-1 separately, we plot 1,1:1,1 with twice the multiplicity. Similarly, baseline pairs such as 
-1,0:1,1 also get the factor of 2 because they are identical to 1,0:1,-1. Baseline pairs such as 1,0:1,0, or 1,1:1,-1 do not get
-the factor of $\sqrt{2}$ because their mirror images are themselves.
+For clarity of presentation we have ``folded over'' the negative time delays and combined baseline pairs that are identical modulus parity.
+We point out that the data points shown here are not all the cases of highest correlation.
 
@@
 \end{equation}
-where, if $\sigma_S^2$ and $\sigma_N^2$ are the characteristic signal and noise levels of a single-baseline visibility, $\rho_0=\sigma_S^2/\sigma_N^2$ is the signal to noise ratio.
+where, if $\sigma_S^2$ and $\sigma_N^2$ are the characteristic signal and noise levels of a single-baseline visibility, $\rho_0\equiv\sigma_S^2/\sigma_N^2$ is the signal to noise ratio.
 
@@
 The lower multiplicities per class of baselines is made up with the larger number of classes of baseline-pairs, especially given the gap in the larger versions.
+
+\begin{figure}[H]
+\includegraphics[width=\linewidth]{pairplot_small}
+
+\caption{Pairplots of the top contributing baseline pairs in three arrays.
+Plotted properties are optimal time delay $dT$,  effective weight $\widetilde{\Theta}$ and baseline length $L$.
+Only those points with $\widetilde{\Theta}>0.01$ are shown (the weight for the top class of equivalent pair is normalized to 1).
+Only one baseline length is shown since all top contributing pairs have very similar lengths as expected.
+The scatter plots are shown with transparency so that darker regions indicate degeneracies.
+Scatter points of the 3 arrays overlap, in order indicated by the legend.
+}
+\label{fig:pairplot}
+\end{figure}
+
+In Fig.
+\ref{fig:pairplot} we present a paired distribution of 3 different properties for baselines that contribute well to the sensitivity ($\widetilde{\Theta}>0.01$, where again the weight for the top class of equivalent pair is normalized to 1).
+The $dT$ vs.
+$\widetilde{\Theta}$ plot is familiar from Fig.
+\ref{fig:sensplot}.
+We only show 3 of the mentioned arrays for visual clarity.
+The other two results are similar barring intuitive differences.
+We make a few observations:
+\begin{itemize}
+
+\item Mapping of equivalent baseline pairs: the equivalent baselines are located in key points in the plots.
+They are the vertical bars in the lower left four plots and the highest peak in the first three histograms, the upper "diagonal edge" in the $\widetilde{\Theta}$ vs.
+$dT$ and the linear trend in the $\widetilde{\Theta}$ vs.
+$L$ plot for HERA arrays .
+
+\item Longer baselines roughly peak at lower time-offset, and lower multiplicity.
+This is intuitive for PAPER since  for the same north-south separation the rotation angle is smaller with longer east-west separation.
+
+\item The top near-equivalent pairs in HERA the same $\Theta$ (not shown) as in PAPER128, but much lower $\widetilde{\Theta}$.
+This is because they are longer baselines with lower multiplicity.
+In the end these baseline classes still lead to high contributions to total sensitivity (Fig.
+\ref{fig:osens}) because there are a lot more such baseline pairs for HERA.
+
+\end{itemize}
 
@@
 
-\section{\label{sec:fullstats}\\Additional Statistics}
-
-\begin{widetext}
-\begin{figure}[H]
-\includegraphics[width=\textwidth]{pairplot}
-
-\caption{Pairplots of the top contributing baseline pairs in three arrays.
-Plotted properties are optimal time delay $dT$, peak height $\Theta$, effective weight $\widetilde{\Theta}$ and baseline length $L$.
-Only those points with $\widetilde{\Theta}>0.05$ are shown (the weight for the top class of equivalent pair is normalized to 1).
-Only one baseline length is shown since all top contributing pairs have very similar lengths as expected.
-The scatter plots are shown with transparency so that darker regions indicate degeneracies.
-Scatter points of the 3 arrays overlap, in order indicated by the legend.
-}
-\label{fig:pairplot}
-\end{figure}
-\end{widetext}
-
-In Fig.
-\ref{fig:pairplot} we present a comparison of 4 different properties for baselines that contribute well to the sensitivity ($\widetilde{\Theta}>0.05$, where again the weight for the top class of equivalent pair is normalized to 1).
-The $dT$ vs.
-$\Theta$ and $dT$ vs.
-$\widetilde{\Theta}$ plots are familiar from Fig.
-\ref{fig:sensplot}.
-We only show 3 of the mentioned arrays for visual clarity.
-The other two results are similar barring intuitive differences.
-We make a few observations:
-\begin{itemize}
-
-\item Mapping of equivalent baseline pairs: the equivalent baselines are located in key points in the plots.
-They are the vertical bars in the lower left four plots and the highest peak in the first three histograms, the upper "diagonal edge" in the $\widetilde{\Theta}$ vs.
-$dT$ and the dot of $\Theta=1$.
-
-\item Longer baselines roughly peak at lower time-offset, and lower multiplicity.
-This is intuitive for PAPER since  for the same north-south separation the rotation angle is smaller with longer east-west separation.
-
-\item Interestingly for HERA the near-equivalent baselines form multiple "U" shapes in $dT, \Theta$ plane, somewhat counter-intuitive especially given out last point.
-This is characteristic of the hexagonal layout.
-Baselines of equal length are more redundant at the right time delay than ones that are slightly different in length, though the latter has a shorter optimal time-delay.
-
-\item The top near-equivalent pairs in HERA the same $\Theta$ as in PAPER128, but much lower $\widetilde{\Theta}$.
-This is because they are longer baselines with lower multiplicity.
-In the end these baseline classes still lead to high contributions to total sensitivity (Fig.
-\ref{fig:osens}) because there are a lot more such baseline pairs for HERA.
-
-\end{itemize}
-
 \bibliographystyle{apj}

### 2017-07-02 Rotation paper/draft_2.tex
save

--- 2017-05-02 Rotation paper/draft_working.tex
@@
 \documentclass[twocolumn,apj,numberedappendix]{emulateapj}
-\shorttitle{Adding Sensitivity to 21\,cm Inteferometric Probes of Reionization}
+\shorttitle{Adding Sensitivity to 21cm Inteferometric Probes of Reionization}
 \shortauthors{Zhang\&Parsons}
@@
 
-\title{Adding Sensitivity to 21\,cm Inteferometric Probes of Reionization by Optimizing Choice of Baselines}
+\title{Adding Sensitivity to 21cm Inteferometric Probes of Reionization by Optimizing Choice of Baselines}
 
@@
 \begin{abstract}
-The observational effort to measure the cosmological 21\,cm power spectrum with radio interferometers requires high sensitivity.
+The observational effort to measure the cosmological 21cm power spectrum with radio interferometers requires high sensitivity.
 We introduce a straightforward and accurate method to extract sensitivity from visibility data.
@@
 The epoch of reionization represents the last key
-stage of our Universe's early evolution. Study of this event stands at
+stage of our universe's early evolution. Study of this event stands at
 the intersection of cosmology and astrophysics. Understanding this
@@
 comes from measurement of the ``spin-flip'' transition of neutral
-hydrogen of characteristic wavelength 21\,cm \citep{Furlanetto2006181,PritchardLoeb}.
+hydrogen of characteristic wavelength 21cm \citep{Furlanetto2006181,PritchardLoeb}.
 Other observational studies of reionization include for example Gunn-Peterson measurements of quasi-stellar objects \citep{Fan2006}.
-GP tests rely on Ly$\alpha$ absorption, and thus easily saturate with small degree of ionization and can only probe the end stage of reionization.
-The 21\,cm line can in comparison directly probe the
+GP tests rely on Ly$\alpha$ absorption, and are thus insensitive to the degree of ionization and can only probe the end stage of reionization.
+The 21cm line can in comparison directly probe the
 abundance and distribution of neutral hydrogen, and thus would potentially
 shed light on all stages of reionization. Radio interferometric efforts
-to measure the 21\,cm power spectrum has been a top priority in recent years of astronomy.
+to measure the 21cm power spectrum has been a top priority in recent years of astronomy.
 Current-generation instruments include the Precision Array for Probing
@@
 
-These modern low-frequency radio interferometers aimed at measuring the 21\,cm signal bears designs dramatically different from traditional instruments built for imaging localized bright sources.
-With the goal of detecting a cosmological signal 5 orders of magnitude fainter than foreground contaminations, current efforts are sensitivity starved.
-Maximum-redundancy arrays are thus designed to repeatedly measure the same Fourier signal to increase sensitivity.
-Ranging upwards from 100 elements, these arrays are typically of the drift-scan type to limit cost.
-
-Analysis pipelines of such maximum-redundancy arrays typically fall into two categories.
-In the first, more traditional technique, image is formed in Fourier domain through rotation synthesis, then foreground is either removed or avoided to construct a power spectrum.
-The second, lately popular technique works directly with visibilities from baselines and cross multiply them to form the power spectrum.
-Due to its more natural focus on the instrumental output of visibilities, this technique is shown to fair better in dealing with systematics such as RFIs and instrumental cross-talks, and thus tend to provide good limits on the power spectrum.
+One of the main challenges of observations of the cosmic
+21cm line is foreground contamination. Both sources within our
+galaxy, and to a lesser extent distant sources outside of our galaxy emit
+radio contamination (via for example synchrotron processes) up to five orders of magnitude stronger than the
+reionization signal. There are two common methods to deal with the
+foreground contamination in power spectrum measurements. In the first, sometimes called  ``foreground
+removal" technique, individual sources are identified and removed in the image domain.
+The other technique, commonly referred to as ``foreground avoidance'', form power-spectrum directly from the visibilities, an arguably more natural approach from an instrumental point of view.  
+Earlier studies have shown that the Fourier signature of smooth-spectrum foreground contaminants are constrained to a ``foreground
+wedge'' \citep{wedge1,wedge2}. Contamination can thus be avoided
+by restricting our observation to ``outside the wedge'', where theoretically foreground contamination is limited and noise becomes dominant. The main
+challenge of using the avoidance technique is thus sensitivity. This is the motivation for the design of
+the maximum redundancy arrays such as PAPER and HERA. Traditional
+imaging arrays are designed to image localized bright sources and thus favor non-redundant patterns for $uv$ coverage. The redundant arrays, on the other hand,
+focus on repeated measurements of the same Fourier mode
+with regular grids of antennas. Since baselines
+of the same length and orientation measure the same Fourier modes
+on the sky, a maximum redundancy array increases the
+signal to noise ratio by averaging over measurements of the same baseline.
 \cite{Ali2015} provides the newest upper limit to the power spectrum
 measurements with the 64-element version of PAPER (henceforth referred to as PAPER-64).
-However, there is one caveat with existing visibility-based pipelines.
-While an imaging based power-spectrum pipeline naturally includes all redundancy information, visibility-based pipelines until now only cross multiply fully redundant baselines, in other words baselines of the same length and orientation.
 
-Baselines of different length and orientation still contain partially redundant information.
-The use of partially redundant baselines is in principle not a fundamental limitation of visibility-based pipelines.
-In fact, most instrumental reports to date do include partial redundancy.
-Recently, \cite{wterm} proposed an algorithmic approach to extract power-spectrum from partially redundant baselines in tracking measurements, by carefully treating the $w$-term in the fringe pattern to obtain weights for cross-correlation.
-Our work parallels this effort by focusing on arrays of the drift-scan type that do not have tracking capabilities, such as PAPER, HERA and SKA.
-The Earth's rotation causes the baselines in a drift-scan array to pick up different modes of the sky with time.
-Rotation-synthesis makes use of the rotation-induced $uv$ coverage map to form image.
-In a visibility based pipeline, we can extract the same rotation-induced redundancy.
-In this point of view, baselines that are slightly different in length and orientation
-rotate into each other at a time delay.
-We can thus cross-multiply time-shifted visibilities, with the proper weighting, to form power spectrum.
-Due to the large number of elements of modern arrays (upward of 100), the task of cross-multiplying every baseline against every other, scaling as number of array element to the fourth power, can be computationally formidable.
-Our contribution is thus twofold.
-First we introduce a formalism to include pairs of partially-redundant baselines in a visibility based power spectrum pipeline.
+While an imaging based power-spectrum pipeline naturally includes all redundancy information, visibility-based pipelines until now only cross multiply simultaneously redundant baselines.
+Here we present a method to
+improve the sensitivity output of the visibility-based power-spectrum pipelines by including pairs of partially redundant baselines.
+To extract maximal partial redundancy we make use of the earth's rotation.
+The earth's rotation causes the baselines in a drift-scan array to pick up different modes of the sky with time.
+In a maximally redundant array, we use the fact that baselines that are slightly different
+rotate into each other at a time delay to extract power-spectrum from their respective visibilities.
+Our contribution is twofold.
+First we introduce a formalism to include pairs of partially-redundant baselines.
 Secondly we show how to use the formalism to automaticaly preselect baseline pairs and time offsets, making the problem computationally efficient.
-More precisely, our formalism allows one to simultaneously a) identify the baselines that give good
-redundancy, b) find the time offset that corresponds to maximal redundancy for a given pair of baselines, and
-c) quantify the sensitivity associated with cross multiplying
-such a pair of partially redundant baselines, which in turns is used as weight to combine measurements in a power spectrum pipeline.
-
-[TODO: this paragraph would potentially be promoted to a section] By using pairs of baselines, we also reduce additional mode mixing.
+Recently, \cite{wterm} proposed an algorithmic approach to extract information from partially redundant baselines, by carefully treating the $w$-term in the fringe pattern to obtain weights for cross-correlation.
+The algorithm makes use of a time dependent coordinate system (``moving grid'') and is thus computationally expensive.
+We here present a formalism that allows us to compute such weights by simply performing a numerical integral.
+This allows us to identify good baselines to cross-correlate and predict the weight of sensitivity with speed and accuracy, thereby making the power spectrum estimation from a large number of baseline pairs computationally tractable.
+By using pairs of baselines, we also reduce additional mode mixing.
 As discussed in \cite{Hazelton2013}, the varying shapes of overlapping regions in $uv$ space of all contributing baselines lead to chromaticities in the response.
@@
 
-The rest of this paper is organized as follows. In section 2 we introduce some terminology and notations to be used for the rest of the paper. In section
-3 we introduce the formalism for weighting partially redundant baselines.
-In section 4 we present numerical tests of
+Baselines of the same length and orientation are traditionally called
+``redundant baselines'', because they measure the same Fourier mode
+in the sky. In order to eliminate confusion and ambiguity, we shall
+introduce slightly different terminology. We shall call baselines that
+are the same length and orientation ``equivalent baselines'', inspired
+by the mathematical notion of equivalency classes. Two equivalent
+baselines will be redundant with each other simultaneously at all
+times. Non-equivalent baselines can also be partially redundant, with their partial redundancy maximized if their
+respected time series is shifted with respect to one another by some delay. In other words, one baseline can be ``rotated into''
+another.
+We call these ``near-equivalent baselines''.
+Exactly how ``near" is ``near-enough" depends on the array configuration.
+All non-equivalent baselines contribute non-zero sensitivity to the power spectrum analysis It is our
+goal to a) identify the near-equivalent baselines that give good
+redundancy, b) to find the optimal time offset\footnote{By optimal we mean the time offset that corresponds to maximal redundancy. } for a given pair of baselines, and
+c) to quantify the sensitivity improvement associated with cross multiplying
+such a pair of near-equivalent baselines.
+
+The rest of this paper is organized as follows. In section
+2 we introduce the formalism for weighting partially redundant/near-equivalent baselines.
+In section 3 we present numerical tests of
 this technique as well as the expected sensitivity improvement
-with this method for HERA and PAPER-128 and with section 5 we conclude.
+with this method for HERA and PAPER-128 and with section 4 we conclude.
 
-\section{Notation and Terminology}
+\section{Method}
 
-In order to eliminate confusion and ambiguity, we
-introduce some terminology for the rest of this paper, some of which differs from what is commonly found in the literature.
-
-First, we make the distinction of a \textit{``baseline''}, which refers to two specific antennas, and a \textit{``class of baselines''}, which refers to all baselines of the same length and orientation in a given array.
-
-Baselines of the same class are traditionally called
-``redundant baselines'', because they measure the same Fourier mode
-in the sky.
-We shall call baselines in the same class \textit{``equivalent baselines''}, and reserve the word \textit{``redundancy''} of two baselines in reference to a variable function of the relative time-offsets of their visibility time series.
-With this terminology, two equivalent
-baselines are fully redundant with each other simultaneously at all
-times. Non-equivalent baselines also have partial redundancy, which is maximized if their
-respected time series are shifted with respect to one another by some delay. In other words, one baseline can be ``rotated into''
-another if they are \textit{``near-equivalent''}.
+\subsection{Rough Idea}
 
@@
 
-\caption{The PAPER-128 layout. Each dot corresponds to the location of
-an antenna. Top panel shows the antenna positions drawn to scale;
+\caption{The PAPER-128 layout. Each blue dot corresponds to the location of
+an antenna. Top panel shows the antenna positions drawn to scale,
 bottom panel show the antenna labels and distances, excluding the outrigger
@@
 during instrument assembly and does not bear significant
-meaning. In the bottom panel, the two baselines (5,77) and (40,85) are an example of equivalent pair, with separation denoted by \{1, 0\}, for the
+meaning. Baselines corresponding to the same separations are called
+equivalent. In the bottom panel, the two baselines (5,77) and (40,85) are an example of equivalent pair, with separation denoted by \{1, 0\}, for the
 antennas are separated by 1 unit east and 0 unit north. Similarly,
 the baselines (5,69) and (5,85) are examples
-of the classes \{1, 1\} and \{1,-1\}, respectively. Note \{1, 0\} and \{-1, 0\} for example
-are the same baseline classes and should not be counted twice.}
+of \{1, 1\} and \{1,-1\}, respectively. Note \{1, 0\} and \{-1, 0\} for example
+are the same baselines and should not be counted twice.}
 \label{fig:AntPos}
@@
 
-We shall use the 128-element PAPER array (henceforth as PAPER-128) to motivate our formalism and demonstrate our method, and extend our results to several HERA configurations in Sec \ref{sec:arrconf}.
+We shall use the 128-element PAPER array to demonstrate our method, and extend our results to several HERA configurations in Sec \ref{sec:arrconf}. 
 The PAPER array is located in the Karoo desert in South Africa (30:43:17.5
 S, 21:25:41.8 E). The layout pattern with antenna labels are show
-in Fig. \ref{fig:AntPos}. The array consists of a 112-element grid and 16 "outriggers" used primarily to aid calibration. We see the antenna spacing in North-South
+in Fig. \ref{fig:AntPos}. We see the antenna spacing in North-South
 directions are comparatively close (4m), so that baselines such as
@@
 of \{1, 1\} and \{1, -1\}, respectively.
-Note \{1, 0\} and \{-1, 0\} for example are the same classes and should
+Note \{1, 0\} and \{-1, 0\} for example are the same baselines and should
 not be counted twice. Antennas in purely north-south baselines
 are close enough to induce cross-talk, and hence are not suitable
-for use. The PAPER-64 analysis of \cite{Ali2015} used three classes of baselines, the PAPER-128
-equivalent of which are 
-\{2, 0\}, \{2, 1\} and \{2, -1\}. There each of these classes
-of baselines are cross multiplied to itself.
-
-\section{Method}
-
-\subsection{Rough Idea}
+for use. The original PAPER-64 analysis \citep{Ali2015} used three classes of baselines, here
+equivalent to 
+\{2, 0\}, \{2, 1\} and \{2, -1\} \cite{Ali2015} of PAPER-128. There each of these classes
+of baselines are cross multiplied to itself. We shall see that in addition these
+baseline classes can be cross multiplied with a time offset.
 
 Given a point source on the sky, each baseline maps the
-source to a point in $uv$ plane. As the Earth rotates with respect to
+source to a point in $uv$ plane. As the earth rotates with respect to
 the source, the point traces out tracks in the $uv$ plane. 
@@
 hours. Color represents different baselines. 
-As the Earth rotates, tracks are traced out counterclockwise. }
+As the earth rotates, tracks are traced out counterclockwise. }
 \label{fig:Tracks}
@@
 \]
-where $\nu_{21}=1420$MHz is the 21\,cm transition rest frequency, $\nu_{0}$
+where $\nu_{21}=1420$MHz is the 21cm transition rest frequency, $\nu_{0}$
 a reference central frequency with corresponding redshift $z_{0}$,
@@
 where the sum is over classes of baseline pairs. 
-We define the estimator sensitivity to be the inverse of the power spectrum noise variance:
+We define the estimator sensitivity to be the inverse of the power-spectrum noise variance:
 \begin{equation}
@@
 Shift the visibilities in time.
-3) Cross multiply the visibilities of the two baselines  to form the power spectrum.
+3) Cross multiply the visibilities of the two baselines  to form the power-spectrum.
 4) Finally combine the different baseline pairs by appropriate inverse-variance weighting that takes into account the predicted sensitivity contributions of each case.

### 2017-07-02 Rotation paper/draft_working.tex
save

@@
 \usepackage{natbib}
+\usepackage{gensymb}
 %\usepackage{pdflscape}
@@
 \{2, 0\}, \{2, 1\} and \{2, -1\}. There each of these classes
-of baselines are cross multiplied within itself. This paper provides the method for inter-class multiplications.
+of baselines are cross multiplied within itself.
+This paper provides the method for inter-class multiplications.
+To do so we will use the short hand notation $\{m,n\}:\{m',n'\}$ to denote a pair of baseline classes to be cross-multiplied.
 
-\section{Method}
+\section{Method \label{sec:method}}\label{sec:method}
 
-\subsection{Rough Idea}
+\subsection{$uvw$ tracks \label{sec:tracks}}
 
-Given a point source on the sky, each baseline maps the
-source to a point in $uv$ plane. As the Earth rotates with respect to
-the source, the point traces out tracks in the $uv$ plane. 
-We show in Fig. \ref{fig:Tracks} $uv$ tracks of PAPER-128 over 4.8 hours, at 0.15GHz, for a source that passes through zenith.
+Given a point source on the sky, each baseline maps to a point in $uvw$ space. As the Earth rotates with respect to
+the source, the points trace out tracks in the $uvw$ space. 
+We show in Fig. \ref{fig:Tracks} $uvw$ tracks of PAPER-128 over 12 hours, at 0.15GHz, for a source that passes through zenith.
 
 \begin{figure}[H]
-\includegraphics[width=\linewidth]{tracks128}
-\caption{Tracks of PAPER-128 (grid only, excluding outriggers) for requency $\nu=0.15\text{GHz}$ and a hypothetical source that passes through zenith.
-These tracks are traced out over 0.2 sidereal days, or roughly 4.8
-hours. Color represents different baselines. 
-As the Earth rotates, tracks are traced out counterclockwise. }
+\includegraphics[width=\linewidth]{tracks3D}
+\caption{Tracks of PAPER-128 (only a randomly selected subset of 32 baselines are shown) for frequency $\nu=0.15\text{GHz}$ and a hypothetical source that passes through zenith.
+These tracks are traced out over 0.5 sidereal days. Color represents different baselines. 
+As the Earth rotates, tracks are traced out counterclockwise. Tracks in $uvw$ space do not cross, even though the $uv$ projections do. }
 \label{fig:Tracks}
@@
 
-Roughly speaking, we can identify
+Traditionally, we can identify
 redundancy of near-equivalent baselines as crossings
-of the $uv$ tracks.
+of the $uv$ tracks, a 2 dimensional projection of Fig.
+\ref{fig:Tracks}.
 As we see in Fig.
-\ref{fig:Tracks}, there are many such crossings.
-Crossing tracks, however does not imply perfect redundancy.
-The reason lies in the finite size of the beams.
-We show sample beams of HERA and PAPER antennas in Fig.
-\ref{fig:Beam} for reference.
-The inaccuracy is expected to be even more significant for the smaller HERA beam, which has a larger corresponding $uv$ point spread function.
+\ref{fig:Tracks}, there would be many such crossings.
+However, there are several reasons that $uv$ track-crossing do not imply perfect redundancy.
+In fact, in our case $uv$ track-crossings are
+not accurate enough for time offset determination, nor can it give estimate of the degree of redundancy.
+The most obvious reason is that the 3 dimensional $uvw$ tracks do not actually cross.
+For drift-scan arrays, even a crossing in $uvw$ space does not imply perfect redundancy.
+This shall become evident in the next section, after we develop a more general formalism that accounts for the point spread function of the finite beams, and estimate the degree of redundancy for general combination of baselines at a general time-offset.
+An explicit explanation of track-crossing will be presented in Appendix \ref{sec:appA}.
 
-Thus although it is a valid quick method to identify some baselines to cross-multiply, track-crossing is
-not accurate enough for time offset determination, nor can give estimate of the degree of redundancy.
-To determine the time-offset corresponding to maximal redundancy, we must develop a more general  formalism that accounts for the point spread function of the finite beams, and estimates the degree of redundancy for general combination of baselines at a general time-offset, as we do in the next section.
+\subsection{Formalism}
+In this section we formulate the relation between 
+the product of visibilities from two arbitrary baselines and the power spectrum.
+
+We begin with the visibility as commonly defined in the literature (e.g.
+\citealt{first-paper}): 
+\begin{equation}
+\begin{aligned}V_{\nu}(\boldsymbol{b}) & =\int d\Omega A(\hat{\boldsymbol{s}},\nu)\phi(\nu)I_{\nu}(\hat{\boldsymbol{s}})\exp\left(-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right),\\
+ & \approx\frac{2k_{B}}{\lambda^{2}}\int d\Omega A(\hat{\boldsymbol{s}},\nu)\phi(\nu)T(\hat{\boldsymbol{s}})\exp\left(-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right),
+\end{aligned}
+\label{eq:Vis1}
+\end{equation}
+Here $k_B$ is Boltzmann's constant, $\lambda$ is the mean wavelength, $\boldsymbol{b}$ is the baseline length, $\hat{\boldsymbol{s}}$ and $\Omega$ are a direction in the
+sky and its corresponding solid angle. Inside the integral we have $\phi(\nu)$ the frequency bandpass profile, $A_{\nu}$ the (frequency
+dependent) primary beam, and $I$ the specific intensity, which has been
+related to $T$, the brightness temperature in the Rayleigh-Jeans
+limit. The beam reception pattern $A$ is dimensionless,
+normalized to 1 at its peak (zenith), and we assume it to be the same
+for all baselines. We show sample beams of HERA and PAPER antennas in Fig. \ref{fig:Beam} for reference.
 
@@
 
-\caption{Sample beam response of HERA (left) and PAPER (right) antennas, both
-for frequency of $\nu=150\text{MHz}$ and Stokes $I$ polarization.
+\caption{Example beam response of HERA (left) and PAPER (right) antennas.
+Both beams are Stokes $I$ polarization antenna voltage beams, at frequency of $\nu=150\text{MHz}$ and normalized to $1$ at zenith.
 Notice $A$ in this paper is a ``baseline's beam'', equivalent to squares of the antenna voltage beams shown here.
 The circles centered around zenith (center of beam) here are
-spaced 10 degrees apart. The finite size of the beams limits the accuracy of the $uv$ track-crossing as a method of redundancy search. \label{fig:Beam}}
+spaced 10 degrees apart. \label{fig:Beam}}
 \end{figure}
 
-\subsection{Formalism}
-Below we shall derive theoretical expectations of cross multiplications
-of two near-equivalent baselines. More precisely we shall relate
-the product of visibilities of two different baselines to the power spectrum.
+In practice,
+power spectrum measurements are typically taken from a few ten MHz centered around the corresponding redshift of interest (e.g. 150 MHz for z=9.5).
 
-We take the visibility as commonly defined in the literature (e.g.
-\citealt{first-paper}): 
-\begin{equation}
-\begin{aligned}V_{\nu}(\boldsymbol{b}) & =\int d\Omega A_{\nu}(\hat{\boldsymbol{s}})\phi(\nu)I_{\nu}(\hat{\boldsymbol{s}})\exp\left[-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right],\\
- & \approx\frac{2k_{B}}{\lambda^{2}}\int d\Omega A_{\nu}(\hat{\boldsymbol{s}})\phi(\nu)T(\hat{\boldsymbol{s}})\exp\left[-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right],
-\end{aligned}
-\label{eq:Vis1}
-\end{equation}
-Here $\lambda$ is a mean wavelength, $\boldsymbol{b}$ is the baseline
-length, $\hat{\boldsymbol{s}}$ and $\Omega$ are a direction in the
-sky and its corresponding solid angle. $A_{\nu}$ is the (frequency
-dependent) beam, and $I$ is the specific intensity, which has been
-related to $T$, the brightness temperature in the Rayleigh-Jeans
-limit.
-$k_B$ is Boltzmann's constant.
-Note that $A$ here is the baseline beam, i.e.
-product of two antenna voltage beams, examples of which are shown in Fig.
-\ref{fig:Beam}.
-$\phi(\nu)$ is the frequency bandpass profile.
-In practice,
-power spectrum measurements are typically taken from a few ten MHz centered around the corresponding redshift of interest (e.g.
-150 MHz for z=9.5).
-We take a moment to note that in Eq.
-\ref{eq:Vis1}, no flat-sky approximation has yet been made.
-The angular integral is performed over the dome $d\Omega$.
-
-We define the delay-transformed visibility \citep{delay-transform}:
+Delayed-transformed visibility has been widely adopted in recent years since it has been shown that foregrounds are isolated in delay space.
+We define the delay-transformed visibility by Fourier transforming visibility along the frequency axis\citep{delay-transform} as:
 \small
 \begin{equation}
-\begin{aligned}V(\boldsymbol{b},\tau) & =\int d\nu V_{\nu}(\boldsymbol{b})\phi(\nu)\exp\left[-2\pi i\nu\tau\right],\\
- & =\frac{2k_{B}}{\lambda^{2}}\int d\Omega d\nu A(\hat{\boldsymbol{s}},\nu)\phi(\nu)T(\hat{\boldsymbol{s}},\nu)\exp\left[-2\pi i\nu\left(\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}+\tau\right)\right]
+\begin{aligned}V(\boldsymbol{b},\tau) & =\int d\nu V_{\nu}(\boldsymbol{b})\phi(\nu)\exp\left(-2\pi i\nu\tau\right),\\
+ & =\frac{2k_{B}}{\lambda^{2}}\int d\Omega d\nu A(\hat{\boldsymbol{s}},\nu)T(\hat{\boldsymbol{s}},\nu)\exp\left[-2\pi i\nu\left(\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}+\tau\right)\right]. 
 \end{aligned}
-.\label{eq:Vb1}
+\label{eq:Vb1}
 \end{equation}
 \normalsize
+Here the delay $\tau$ is the Fourier variable. Eq. \eqref{eq:Vb1} expresses the delay-transformed visibility as
+an integral over observational coordinates $\hat{\boldsymbol{s}}$ and $\nu$.
+For notational simplicity here and in the rest of this paper, we shall absorb the bandpass $\phi$ inside the primary beam $A$.
 
-Eq. \eqref{eq:Vb1} expresses the delay-transformed visibility as
-an integral over observation coordinates $\hat{\boldsymbol{s}}$ and $\nu$. Ultimately,
+Ultimately,
 we would like to relate the data, collected with coordinates $\hat{\boldsymbol{s}}$
-and $\nu$, to the power spectrum, written with cosmological coordinates
-$\boldsymbol{r}$ and $\boldsymbol{k}$. We start by noticing that
+and $\nu$, to the power spectrum, written with 
+$\boldsymbol{r}$ and $\boldsymbol{k}$, the cosmological position coordinate and wavenumber. To do so, we start by noticing that
 \[
@@
 \begin{equation}
-\nu=\frac{D_{c}-r}{Y}+\nu_{21}.\label{eq:nur}
+\nu=\frac{X-r}{Y}+\nu_{21}.\label{eq:nur}
 \end{equation}
 
-Further we shall invoke the flat-sky limit
+In the thin-shell limit, we can write:
 \begin{equation}
-(r_{x},r_{y})=(X\hat{\boldsymbol{s}}_{x},X\hat{\boldsymbol{s}}_{y}), 
+d^2r=X^2d\Omega. 
+\label{eq:thinshell}
 \end{equation}
-where
-\begin{equation}
-X = \frac{c}{H_0}\int_{0}^{z}\frac{dz'}{E(z')}
-\end{equation}
-is a cosmological scalar. In the same limit $(Xk_{x},Xk_{y},Yk_{z})=\frac{2\pi}{c}(b_{x},b_{y},\tau)$ relates
-the cosmological $k$ to the observation coordinates. For a more rigorous treatment without the flat sky limit, see Appendix [ADRIAN].
+Note that in Eq.\ref{eq:thinshell} no flat-sky approximation has been made; the angular integral is still performed over the dome.
 
-We can rewrite the delayed transformed visibility as 
+We can rewrite the delayed-transformed visibility as 
 \small
 \[
-\begin{aligned}V(\boldsymbol{b},\tau) & =\frac{2k_{B}}{\lambda^{2}}\int_{H}\frac{d^{3}r}{X^{2}Y}A(\boldsymbol{r})\phi(r)T(\boldsymbol{r})\exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\hat{\boldsymbol{r}}+\tau\right)\nu_{r}\right]\end{aligned},
+\begin{aligned}V(\boldsymbol{b},\tau) & =\frac{2k_{B}}{\lambda^{2}}\int\frac{d^{3}r}{X^{2}Y}A(\boldsymbol{r})T(\boldsymbol{r})\exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\hat{\boldsymbol{r}}+\tau\right)\nu_{r}\right]\end{aligned},
 \]
@@
  where $d\nu=-dr/Y$ and $d^{3}r=-X^{2}Yd\Omega d\nu$. 
+We have written $\nu_{r}$ as a reminder that $\nu$ and $r$ are related
+by Eq. \eqref{eq:nur}.
 
-We have written $\nu_{r}$ to remind us that $\nu$ and $r$ are related
-by Eq. \eqref{eq:nur}. The beam reception pattern $A$ is dimensionless,
-normalized to 1 at its peak (zenith), and we assume it to be the same
-for all baselines.
-
-With a time offset, the beam pattern has moved relative to the sky.
-Here we choose to fix the sky, and denote the rotated coordinates
-of the beam pattern with the 3 dimensional rotation operator $\Gamma$. With implicit bounds
-of integrals from $-\infty$ to $\infty$, we have:
+Most existing visibility based power-spectrum pipelines for redundant drift-scan  arrays relate the power-spectrum to the conjugate square of the visibilities \citep{first-paper, paper32, Ali2015}.
+We would like to generalize such relations by relating the power-spectrum to the product of two visibilities from two arbitrary baselines and time-offsets.
+The beam pattern of a baseline shifts relative to the sky as the Earth rotates. Here we choose to fix the sky, and denote the rotated coordinates
+of the beam pattern with the 3 dimensional rotation operator $\Gamma$:
+\small
+\[
+\begin{aligned}V(\boldsymbol{b'},\tau') & =\frac{2k_{B}}{\lambda^{2}}\int\frac{d^{3}r}{X^{2}Y}A(\Gamma\boldsymbol{r})T(\boldsymbol{r})\exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\Gamma\hat{\boldsymbol{r}}+\tau'\right)\nu_{r}\right]\end{aligned}.
+\]
+\normalsize
+With implicit bounds of integrals from $-\infty$ to $\infty$, we have:
 \begin{widetext}
 \begin{equation}
-\begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau')\rangle\\
+\begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau)\rangle\\
  & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle\right)A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
@@
  & \approx\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}rd^{3}r'\delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
- & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}r|A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r})||\phi(\nu_{r})|^{2}\exp\left[-i2\pi\nu_{r}\left(\hat{\boldsymbol{r}}\cdot\frac{\boldsymbol{b}}{c}-\Gamma \hat{\boldsymbol{r}}\cdot\frac{\boldsymbol{b'}}{c}\right)\right],\\
- & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma \hat{\boldsymbol{s}},\nu)||\phi(\nu)|^{2}\exp\left[-i2\pi\nu\left(\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b}}{c}-\Gamma\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b'}}{c}\right)\right],
+ & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}r|A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r})|\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\Gamma \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)\right],\\
+ & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma \hat{\boldsymbol{s}},\nu)|\exp\left[-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)\right],
 \end{aligned}
@@
 \begin{equation}
-\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'})=|{\phi^{*}}(\ensuremath{\nu_{r}})\phi(\nu_{r'})|\exp\left[-i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\Gamma\hat{r'}\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)\right].
+\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'})=\exp\left[-i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\Gamma\hat{\boldsymbol{r'}}\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)\right].
 \end{equation}
 \end{widetext}
-The second to third line of Eq.(\ref{eq:main}) follows from assumption of translational invariance, and the third to fourth line follows from the assumption that
+The third equality of Eq.(\ref{eq:main}) follows from assumption of translational invariance of sky statistics, and the first inequality follows from the assumption that
 the 3D power spectrum varies negligibly over the $k$-space of interest
-so that $\hat{P}_{21}(k+k_{2})\approx\hat{P}_{21}(k)$. Since $\Gamma$
+so that $\hat{P}_{21}(k+k')\approx\hat{P}_{21}(k)$.
+The careful reader would have noticed that we have also assumed here a definitive relation between $k$ and $b,\tau$.
+In the flat-sky limit for example, such a relation would take the form $(Xk_{x},Xk_{y},Yk_{z})=\frac{2\pi}{c}(b_{x},b_{y},\tau)$.
+Instrumental chromaticity and foreground isolation will be examined in \ref{sec:chromaticity}.
+Since $\Gamma$
 is a sky rotation, it does not affect $\nu$, hence we have taken $\nu_{r}$
-outside the parenthesis. Notice that the phase factor $\exp\left[-i2\pi\tau\left(\nu-\nu'\right)\right]$
-drops out in the end. This means that correlation is peaked at the same time-offset for all delay channels.
+outside the parenthesis in the phase term $\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\Gamma \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)\right]$.
+Notice that the phase factor $\exp\left[-i2\pi\tau\left(\nu-\nu'\right)\right]$
+drops out in the end. This means that the location and magnitude of the correlation peak does not depend on delay $\tau$, an important point we shall come back to.
 
-Finally, since the beam pattern and bandwidth are given in $\hat{\boldsymbol{s}}$
+Since the beam pattern and bandwidth are given in $\hat{\boldsymbol{s}}$
 and $\nu$, we convert the integral back to these coordinates to get
@@
 \begin{equation}
-\begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau')\rangle\\
- & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)||\phi(\nu)|^{2}\\
- & \qquad \qquad \qquad \qquad \exp\left[-i2\pi\nu\left(\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b}}{c}-\Gamma\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b'}}{c}\right)\right].\end{aligned}
+\begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau)\rangle\\
+ & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)|\\
+ & \qquad \qquad \qquad \qquad \exp\left[-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)\right].\end{aligned}
 \label{eq:final}
@@
 \begin{equation}
- P(k_{b,\tau}) = \left(\frac{\lambda^{2}}{2k_{B}}\right)^{2} \frac{\langle V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau')\rangle}{\Theta}, 
+ P(k_{b,\tau}) = \left(\frac{\lambda^{2}}{2k_{B}}\right)^{2} \frac{\langle V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau)\rangle}{\Theta}, 
  \label{eq:opp}
@@
 \begin{equation}
-\Theta =\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)||\phi(\nu)|^{2} e^{-2\pi i\nu\left(\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b}}{c}-\Gamma\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b'}}{c}\right)}.
+\Theta \equiv\int d\nu \Theta_{\nu}, 
 \label{eq:Theta}
 \end{equation}
+and
 
-So roughly speaking the cross multiplications of visibilities at a time delay
+\[
+\Theta_{\nu} \equiv \int \frac{d\Omega}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)| e^{-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)}e^{i\psi_{\nu}}.
+\label{eq:Thetanu}
+\]
+
+Notice that $\Theta$ has no dependence on $\tau$.
+In Eq.
+\ref{eq:Thetanu} we introduced an extra phase $\psi$, the origin of which we shall explain in \ref{sec:rephs}.
+We point out that although all our derivations focused on drift-scan telescopes, we can get the analogous result for tracking measurements simply by noticing that for a tracking primary beam, $\Gamma$ becomes a rotation around zenith, and so:
+\[
+\Theta_{\nu} \equiv \int \frac{d\Omega}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\hat{\boldsymbol{s}},\nu)| e^{-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)}e^{i\psi_{\nu}}.
+\label{eq:Thetanu_tracking}
+\]
+
+Roughly speaking, Eq. \ref{eq:opp} through \ref{eq:Thetanu} tell us that the cross multiplications of visibilities at a time delay
 in $uv$-space is proportional to the power spectrum times the Fourier
-transform of the cross multiplied beam pattern. We can then in principle
-combine information from different baseline pairs if we correct for
-the phase and normalization. As a check, when applied to equivalent baselines,
+transform of the cross multiplied beam pattern. As a check, when applied to equivalent baselines,
 $\boldsymbol{b}=\boldsymbol{b'}$, $\hat{\boldsymbol{s}}=\Gamma\hat{\boldsymbol{s}}$, and Eq.(\ref{eq:final}) reduces to Eq.(B9) of \cite{paper32}. 
-
 With Eq.
 \ref{eq:opp} and Eq.
-\ref{eq:Theta} we can for any given baseline pair and time delay, estimate the degree of redundancy, here represented by $\Theta$, thereby achieving all our goals stated in the introduction, i.e.
+\ref{eq:Theta} we can, for any given pair of baseline classes and time delay, estimate the degree of redundancy, here represented by $\Theta$, thereby achieving all our goals stated in the introduction, i.e.
 to identify
@@
 We can do all the above simply by computing the weight $\Theta$ from
-Eq.(\ref{eq:Theta}) for various time offsets, without having to actually cross-multiply visibilities at different offsets. This makes the task computationally tractable.
+Eq.(\ref{eq:Theta}) for various time offsets, without having to actually cross-multiply visibilities for all baselines and offsets. This makes the task computationally tractable.
 
 \subsection{Rephasing \label{sec:rephs}}
-Before we discuss tests, there is one subtlety when applying Eq.
-\ref{eq:final} to the visibilities from correlators.
-The optimal time offset is given by the $\Gamma$ that maximizes the absolute value of the weight $\Theta$.
-However, $\Theta$ is in general complex.
-First integrating over the spatial dimension $\Omega$, we see the phase term at peak time-offset is inevitably frequency dependent.
-This frequency dependence of the phase would lead to destructive interference when we integrate over frequency.
+As we see in Eq.
+\ref{eq:Thetanu}, $\Theta_{\nu}$ at the peak of correlation is in general complex, and often far from real.
+Furthermore, this phase of peak correlation is inevitably frequency dependent.
+This frequency dependence of the phase would lead to destructive interference when we integrate over frequency, unless we correct $\Theta_{\nu}$ by an extra phase $\psi_\nu$.
 
-The physical origin of this frequency-dependent phase lies in the two visibilities having different phase centers.
-By default the correlators phase the two visibilities  both to zenith at the same time.
-When they are cross-multiplied with a time lag, they must be rephased to the account for the movement of the zenith.
-Thus in practice we must rephase the visibilities before delay transform.
-Doing so is also crucial to reduction of additional mode mixing, as discussed in \cite{Hazelton2013}.
-Shapes of overlapping regions of pairs of baselines are fixed with respect to frequency.
+%Doing so is also crucial to reduction of additional mode mixing, as discussed in \cite{Hazelton2013}. Shapes of overlapping regions of pairs of baselines are fixed with respect to frequency.
 
-In Fig.
-\ref{fig:freqdiff} we compare the cross-multiplied visibilities of both equivalent and near-equivalent baseline pairs for two channels: 0.16 GHz and 0.17 GHz.
+To illustrate the effect of rephasing, we compare in Fig.
+\ref{fig:freqdiff} real parts of $\Theta_{\nu}$ of both equivalent and near-equivalent baseline pairs for two channels: 0.16 GHz and 0.17 GHz.
 The top two panels have zero rephasing, and the bottom two are rephased to a time offset of 0.055 sidereal days, the optimal time offset for the given baseline pair.
-The first and third panels show the equivalent baseline pairs \{1, 0\} against itself, and second and fourth panels show \{1, 0\} against \{1, 1\}.
+The first and third panels show the equivalent baseline pairs \{1,0\}:\{1,0\}, and second and fourth panels show \{1,0\}:\{1,1\}.
 We see that in the un-rephased case in the second panel, although the magnitude of correlations match up for the two frequencies, the phases do not.
@@
 
+The physical origin of $\psi_\nu$ lies in the two visibilities having different phase centers.
+By default the correlators of a drift-scan array phase the two visibilities both to zenith at the same time.
+When they are cross-multiplied with a time lag, the visibilities must be rephased before delay transform to the account for the movement of the zenith.
+One can check that introduction of a phase in Eq.
+\ref{eq:Vis1} does carry through the derivation to Eq \ref{eq:Thetanu}.
+The effect of the phase thus roughly corresponds in a shift of the delay mode measured, and we should expect $\psi_{\nu}$ to be in first order a linear function of $\nu$.
+In Fig.
+\ref{fig:phi_nu} we show the peak phases for for a given baseline pair of PAPER128 (\{2,0\}:\{2,1\}), comparing the drift-scan phase dependence with that of the same baseline with hypothetical tracking-elements (Eq.
+\ref{eq:Thetanu} and Eq.
+\ref{eq:Thetanu_tracking}).
+In the drift-scan case, we indeed see a linear relation corresponding to a delay of $\Delta\tau\approx17$ns.
+In the tracking case, because of minimal zenith movement, only second order effects are observed.
+
+\begin{figure}[H]
+\includegraphics[width=1\linewidth]{phi_nu}
+
+\caption{Frequency dependent peak phases for PAPER-128 baseline pair \{2,0\}:\{2,1\}.
+Both the drift-scan case and a hypothetical tracking baseline are shown.
+We have also fixed the global phase in both cases to 0 at 0.145GHz.
+The drift-scan case exhibits linear behavior and to first order corresponds to a shift in delay space to account for the movement of zenith.
+}
+\label{fig:phi_nu}
+\end{figure}
+
+Once rephased, the overall weight can be constructed by $\Theta=\int \Theta_{\nu}d\nu$, and thus
+\begin{equation}
+\begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle\\
+ & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)|\\
+ & \qquad \qquad \qquad \qquad \exp\left[-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)\right].\end{aligned}
+\label{eq:final_new}
+\end{equation}
+
 \section{Analysis}
@@
 
-First we present some numerical tests of our formalism to verify the validity of using $\Theta$ to estimate the degree of redundancy as a function of time lag.
+Bringing together the discussion from Section \ref{sec:method}, we present a numerical check of Eq.\ref{eq:final_new}, including rephasing.
 To do so, we need to compare the amplitude and phase of the integral weight $\Theta$
 for a pair of baselines with products of simulated visibilities of those baselines.
-For computational simplicity, we shall use a single frequency channel of $150$MHz for the comparison.
-We may use this simplification without compromising generality because Eq.
-(\ref{eq:final}) is valid for a range of frequency channels if and only if it's valid for each single channels, provided that different channels are combined with rephasing as described in Sec.
-\ref{sec:rephs}.
+We use 10 frequency channels evenly spaced from $145$MHz to $155$MHz for the comparison.
+We rephase one of the simulated visibilities, as well as computed $\Theta_{\nu}$ to the same delay as computed from the peak phase of the latter.
 
-For the simulation, we take N=2500 random realizations of the sky
+\begin{figure}[h]
+\includegraphics[width=\linewidth]{numerics}
+
+\caption{Numerical comparisons of the visibility correlation peaks to the $\Omega$ factor in Eq.(\ref{eq:final}).
+We generated a Gaussian random sky on healpix maps, computed visibilities and cross correlated them to find the correlation.
+Top panel shows the equivalent baseline pairs \{2,0\} against \{2,0\}, bottom panel shows \{2,0\} against \{2,1\}.
+Evaluations of weight $\Theta$ are shown green, and the visibility correlations of simulated random sky is given in blue.
+In both cases only the real parts are shown.
+We see that in both cases the theory and simulation line up in both amplitude and phase.}
+\label{fig:numerics}
+\end{figure}
+
+Fig.
+\ref{fig:numerics} shows the real parts of the resulting degree of redundancy as a function of time offset, for equivalent baseline pairs (\{2,0\}:\{2,0\}) in the top panel and near-equivalent (\{2,0\}:\{2,1\}) in the bottom.
+Shown in green for the two cases are the computed values of $\Theta$, both normalized to the peak of the equivalent case.
+The plot on the bottom is rephased with a delay of $\Delta\tau\approx17$ns as outlined in Section \ref{sec:rephs}.
+Since both y-axes are normalized to the same scale, we see the equivalent baseline reach maximal redundancy with 0 time-offset, while the near-equivalent pair reaches maximum redundancy of $\Theta\approx0.5$ at $dT\approx0.325$days.
+
+Almost completely overlapping with the green curves are the cross-multiplied visibilities from a simple simulation, shown in blue.
+For the simulation, we populate random values of brightness temperature
 on a healpix map \citep{Heal, HealPrimer} \footnote{We use functionalities in the python package AIPY for healpix mapping
-as well as coordinate transforms. }, and pick two baselines sep1,0 and sep1,1. For each realization, each pixel is given a Gaussian random value
-of brightness temperature. We then rotate the baseline positions with
+as well as coordinate transforms. }. We rotate the baseline positions with
 the appropriate rotation matrix, multiplying the sky by the primary beam to get the visibilities, for each
@@
 map turns out to be significant. Thus we let the sky, represented
-by the healpix map, be fixed, and rotate the baselines. }. The resulting visibilities for the two baselines are then convolved
+by the healpix map, be fixed, and rotate the baselines.
+}.
+For the near-equivalent case, we rephase the delayed visibility (\{2,0\}) by the same delay $\Delta\tau\approx17$ns, followed by delay-transform.
+The resulting delay-space visibilities for the two baselines are then convolved
 via the Fourier convolution theorem, to obtain values of the cross
-correlation as a function of time-offset. The peak of the curve then
-corresponds to maximum redundancy.  We do this for both the equivalent (sep1,0 against sep1,0) and near-equivalent (sep1,0 against sep1,1) case. The accuracy of this result is
-limited by (simulated) cosmic variance and finite spacial resolution,
-and hence can be beat down by averaging over a large number of universes.
-A numerical estimate of the error of the peak height with $N=1$ is
-$\lesssim20\%$, and thus with $N=2500,$ we achieve an error of peak
-height $<0.5\%$.
+correlation as a function of time-offset.   We do this for both the equivalent (\{2,0\}:\{2,0\}) and near-equivalent (\{2,0\}:\{2,1\}) case.
+Since the blue and green curves overlap, we have verified that in this case Eq.\ref{eq:final_new} is valid and the computed $\Theta$ can be used as weight to obtain the power spectrum.
 
-\begin{figure}[h]
-\includegraphics[width=\linewidth]{opp}
-
-\caption{Numerical comparisons of the visibility correlation peaks to the $\Omega$ factor in Eq.(\ref{eq:final}).
-We generated 2400 instances of Gaussian random sky on healpix maps, computed visibilities and cross correlated them to find the correlation.
-Top panel shows the equivalent baseline pairs sep1,0 against sep1,0, bottom panel shows set1,0 against sep1,1.
-Evaluations of weight $\Theta$ are shown with solid curves, and the visibility correlations of simulated random sky is given with dashed lines.
-In both cases blue denotes real part, green the imaginary part, and red the magnitude.
-We see that in both cases the theory and simulation line up in both amplitude and phase.}
-\label{fig:numerics}
-\end{figure}
-
-The comparison is show in Fig. \ref{fig:numerics}. 
-Sensitivity contribution is inferred from height of the peak of the cross-multiplied visibilities as a function of time. 
-We first compare the simulation to the analytical weight $\Theta$ for a pair of equivalent baselines (sep1,0, sep1,0), then normalize the peaks 
-of the near-equivalent baselines (sep1,0, sep1,1) to that of the equivalent ones. 
-In Fig. \ref{fig:numerics}  we show on the top comparison of
-the convolution of an equivalent baseline of PAPER-128, normalized to 1 at the
-peak, whose location is at zero time offset as expected. On the bottom we show
-comparison of a near-equivalent baseline pair, of classes sep1,0:1,1 and sep1,1:1,1. The peaks
-are normalized with the same factor as the peak height on the top,
-i.e. the two plots have thus the same scale on the vertical axis.
-The height of the plot thus quantifies the added sensitivity. We see at a peak of around 0.055 sidereal days, or 1.32 hours, the two baselines are maximally redundant.
-
-At the optimal time separation, the integral in Eq. \eqref{eq:final}
-is maximized. Thus as another check we expect the two beams to have to same fringe pattern
-(frequency and phase). Due to the time delay, however, the beam center
-would be slightly shifted with respect to each other. This we show
-in Fig. \ref{fig:Beam-fringe-pattern}. The left and middle panels show the beam fringe
-patterns for baselines sep2,0 and sep2,1, delayed by 0.0325 sidereal days,
-and the right panel shows their cross product. The fringe pattern
-indeed cancels out as we expect.
+\subsection{Chromaticity \label{sec:chromaticity}}
+Manageable chromaticity of the instrument is crucial in delay-transform technique of foreground isolation.
+Previous studies have shown that smooth foreground are constrainted in a "wedge" structure in cosmological $k$ space \citep{wedge1, wedge2}.
+We must show that our method does not destroy the ability of delay-transform to isolate foreground.[ .....
+]
 
 \begin{figure*}[h!]
-\includegraphics[width=0.9\textwidth]{fringe_res}
-\caption{Beam fringe pattern of sep2,0(left), sep2,1 at a time delay (middle),
-and their conjugate product (right). Frequency of $\nu=0.15\text{GHz}$ is
-chosen and only the real components are shown. The colorbar values are normalized such that the peak of the original beam is 1. \label{fig:Beam-fringe-pattern}}
-\end{figure*}
-
-\begin{figure*}[h!]
-\includegraphics[width=0.9\textwidth]{sensitivity}
+\includegraphics[width=\textwidth]{sensitivity_1}
 
@@
 hexagonal design is the densest pattern of antenna-packing.
-The larger arrays are designed with a ``gap'' dividing the antennas into three different regions.
+The larger arrays are designed with a ``gap'' dividing the antennas into three different groups.
 The gaps are designed so as to improve $uv$ coverage and ease calibration without compromising sensitivity, but also produces many more near-equivalent baselines than the versions without the gap.
 The motivations behind the designs are explained in \cite{HERAconfiguration}.
-The lack of short baselines that are visually close to each other, as well as the smaller beam (Fig.
+Compared to PAPER128, the hexagonal pattern of HERA lack short baselines that are oriented close to each other, and the smaller beam (Fig.
 \ref{fig:Beam}) means that we expect to see only longer near-equivalent baselines.
 The lower multiplicities per class of baselines is made up with the larger number of classes of baseline-pairs, especially given the gap in the larger versions.
+
+\begin{figure}[H]
+\includegraphics[width=\linewidth]{pairplot}
+
+\caption{Pairplots of the top contributing baseline pairs in three arrays.
+Plotted properties are optimal time delay $dT$,  effective weight $\widetilde{\Theta}$ and baseline length $L$.
+Only those points with $\widetilde{\Theta}>0.01$ are shown (the weight for the top class of equivalent pair is normalized to 1).
+Only one baseline length is shown since all top contributing pairs have very similar lengths as expected.
+The scatter plots are shown with transparency so that darker regions indicate degeneracies.
+Scatter points of the 3 arrays overlap, in order indicated by the legend.
+}
+\label{fig:pairplot}
+\end{figure}
+
+In Fig.
+\ref{fig:pairplot} we present pair-distributions of 3 different properties for baselines that contribute well to the sensitivity ($\widetilde{\Theta}>0.01$, where again the weight for the top class of equivalent pair is normalized to 1).
+The properties shown are effective weight $\widetilde{\Theta}$, optimal time offset $dT$, and average baseline length $L$.
+We only show 3 of the mentioned arrays for visual clarity.
+The other two results are similar barring intuitive differences.
+We discuss each cross-distribution plot:
+\begin{itemize}
+
+\item $dT$ vs. $\widetilde{\Theta}$:
+This relation  is familiar from Fig. \ref{fig:sensplot}. The points at $dT=0$ are the equivalent baselines. Note PAPER128 has more data points with high $dT$.
+
+\item $dT$ vs. $L$:
+PAPER128 shows the trend that longer baselines correspond to lower $dT$.
+HERA arrays do not exhibit this trend here because of a selection effect.
+Only baseline pairs with $\widetilde{\Theta}>0.01$ are shown.
+Shorter HERA baselines require much longer $dT$ to overlap, partly because of the hexagonal structure requires $60\degree$ of rotation to overlap baselines, partly because of the smaller primary beam.
+Thus most short HERA baseline pairs are not shown in this figure.
+
+\item $\widetilde{\Theta}$ vs. $L$:
+The general trend to be seen here is that longer baselines tend to have lower $\widetilde{\Theta}$.
+This is due to the lower multiplicity of longer baselines.
+The trend is particularly obvious in the linear structure for the HERA arrays, which are the equivalent baselines.
+Since they all have the unity $\Theta$ by definition, the linearly decreasing trend of $\widetilde{\Theta}$ is a direct measure of the baseline multiplicity structures of the HERA array configurations.
+
+\item The top near-equivalent pairs in HERA the same $\Theta$ (not shown) as in PAPER128, but much lower $\widetilde{\Theta}$.
+This is because they are longer baselines with lower multiplicity.
+In the end these baseline classes still lead to high contributions to total sensitivity (Fig.
+\ref{fig:osens}) because there are a lot more such baseline pairs for HERA.
+
+\end{itemize}
 
@@
 Current generations of redundant radio arrays, such as those probing the power spectrum of the epoch of reionization could benefit from data analysis techniques that improve the sensitivity.
-We present an intuitive analysis of cross-multiplying baselines that are close in length and orientation to each other.
-Given an antenna array configuration, our method quickly identifies the best baseline pairs to cross-multiply and predict the expected sensitivity contribution.
-With the predicted result one can improve existing power-spectrum pipelines through 3 simple steps.
+We present a visibility and delay-transform based method of cross-multiplying baselines that are close in length and orientation to each other, thereby extracting the sensitivity contained in such partial redundancy.
+Our discussion focuses on drift-scan arrays but also provide the analogous expressions for tracking measurements.
+Our method relies on cross-multiplications of visibilities at a time-offset.
+For drift-scan arrays, the movement of zenith during this time-offset requires a frequency-dependent rephasing prior to delay transform.
+We show that such zenith-rephasing lead to manageable chromatic leakage.
+Given an antenna array configuration, our method identifies the best baseline pairs to cross-multiply and predict the optimal time-offset $\Delta T$, weight $\Theta$ and rephase delay $\Delta\tau$.
+With the predicted results one can incorporate partial redundancy into existing delay-tranform based power-spectrum pipelines through a few steps.
 1).
-Rephase the visibilities prior to delay transforming by the zenith displacement (or more accurately by the phase predicted by our numerical analysis).
+Rephase the visibilities prior to delay transforming by $\Delta\tau$.
 2).
-Shift the visibilities in time.
+Shift the visibilities in time by $\Delta T$.
 3) Cross multiply the visibilities of the two baselines  to form the power spectrum.
-4) Finally combine the different baseline pairs by appropriate inverse-variance weighting that takes into account the predicted sensitivity contributions of each case.
-We showed that such techniques could lead to $20\%$ to $60\%$ increase in sensitivity for PAPER and HERA.
+4) combine the different baseline pairs by appropriate inverse-variance weighting that takes into account the predicted sensitivity contributions of each case.
+We showed that incorporation of partial redundancy accounts for $20\%$ to $60\%$ sensitivity for various configurations of PAPER and HERA.
 
@@
 \appendix
+\section{\label{sec:appA}\\Track-crossing}
+In this section we explain the reason why track-crossing, even in $uvw$ space, do not correspond to peak correlation for drift-scan measurements. 
+Typical convention of rotation synthesis has rotation operates on $\boldsymbol{b'}$ instead of $\hat{\boldsymbol{s}}$, in which case Eq.\ref{eq:Theta} becomes
+\begin{equation}
+\Theta \equiv\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)||\phi(\nu)|^{2} e^{-i2\pi\frac{\nu}{c}\hat{\boldsymbol{s}}\cdot\left(\boldsymbol{b}-\widetilde{\Gamma}\boldsymbol{b'}\right)},
+\label{eq:drift}
+\end{equation}
+where $\widetilde{\Gamma}$ is the inverse of $\Gamma$. 
+For tracking elements, $\Gamma$ becomes a rotation around zenith:
+\begin{equation}
+\Theta \equiv\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\hat{\boldsymbol{s}},\nu)||\phi(\nu)|^{2} e^{-i2\pi\frac{\nu}{c}\hat{\boldsymbol{s}}\cdot\left(\boldsymbol{b}-\widetilde{\Gamma}\boldsymbol{b'}\right)},
+\label{eq:tracking}
+\end{equation}
+We are now in a position to understand the reason that track-crossing, even in $uvw$ space, as mentioned section \ref{sec:tracks}, would not correspond exactly to the peak of correlation.
+Track-crossing corresponds to vanishing of the exponent
+\[
+ \boldsymbol{b}-\widetilde{\Gamma}\boldsymbol{b'} = 0. \label{eq:trackcross}
+ \]
+This condition maximizes $\Theta$ if and only if no other term in the integral depends on $\Gamma$.
+We see from Eq.
+\ref{tracking} that this condition is satisfied.
+In a general drift-scan measurement, however, track-crossing point of Eq.
+\ref{eq:trackcross} does not maximize $\Theta$.
+
 \section{\label{sec:appB}\\Derivation of Noise Covariance \label{sec:appB}}

### 2017-07-02 Rotation paper/draft_working.tex
save

@@
 
-\caption{The PAPER-128 layout. Each dot corresponds to the location of
+\caption{The PAPER128 layout. Each dot corresponds to the location of
 an antenna. Top panel shows the antenna positions drawn to scale;
@@
 \ref{fig:freqdiff} real parts of $\Theta_{\nu}$ of both equivalent and near-equivalent baseline pairs for two channels: 0.16 GHz and 0.17 GHz.
-The top two panels have zero rephasing, and the bottom two are rephased to a time offset of 0.055 sidereal days, the optimal time offset for the given baseline pair.
-The first and third panels show the equivalent baseline pairs \{1,0\}:\{1,0\}, and second and fourth panels show \{1,0\}:\{1,1\}.
+The top two panels have zero rephasing, and the bottom two are rephased to a time offset of 0.0325 sidereal days, the optimal time offset for the given baseline pair.
+The first and third panels show the equivalent baseline pairs \{2,0\}:\{2,0\}, and second and fourth panels show \{2,0\}:\{2,1\}.
 We see that in the un-rephased case in the second panel, although the magnitude of correlations match up for the two frequencies, the phases do not.
@@
 \ref{eq:Vis1} does carry through the derivation to Eq \ref{eq:Thetanu}.
-The effect of the phase thus roughly corresponds in a shift of the delay mode measured, and we should expect $\psi_{\nu}$ to be in first order a linear function of $\nu$.
+The effect of the phase thus roughly corresponds in a shift of the delay mode measured, and we should expect $\psi_{\nu}$ to be in first order a linear function of $\nu$:
+\[
+\psi_{\nu}\sim 2\pi\Delta\tau\nu.
+\]
 In Fig.
@@
 In the tracking case, because of minimal zenith movement, only second order effects are observed.
+The origin of the second order effects can be seen as due to the $w$-term, or more precisely the fact that $uvw$ tracks do not cross when their 2 dimensional projections do.
+We refer the reader to Appendix \ref{sec:appA} for further explanation.
 
@@
 \begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle\\
- & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)|\\
+ & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)|e^{i\psi_{\nu}}\\
  & \qquad \qquad \qquad \qquad \exp\left[-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)\right].\end{aligned}
@@
 
-\caption{Relative sensitivity contributions of selected baseline combinations in PAPER-128. In the legend m,n:p,q denotes cross
-multiplying PAPER-128 baselines of east-west, north-south separations (m,n) and (p,q) respectively. The top
-panel shows the peak height (degree of correlation) of each baseline
+\caption{Relative sensitivity contributions of selected baseline combinations in PAPER-128.
+Each labeled point in the middle and bottom panels correspond to a cross-correlated baseline pair \{m,n\}:\{p,q\}.
+The shape of the symbol encodes the first baseline \{m,n\}, as displayed in the top left legend panel.
+The edge and face colors encode the second baseline \{p,q\}, as displayed in the top right legend panel.
+The middle
+panel shows the peak height ($\Theta$) of each baseline
 combination, while the bottom panel multiplies the heights by the
 corresponding multiplicities as in Eq. \eqref{eq:sensul}. 
-In weighting by the multiplicity (bottom), we have chosen to fix the sensitivity 
-contribution of 1,0:1,0 to unity. The same symbols and colors correspond in the top and bottom panel, with half of the legend shown in each. }
+In both the middle and bottom panels, we have chosen to fix the value  of \{1,0\}:\{1,0\} to unity.  }
 \label{fig:sensplot}
@@
 
-\begin{figure*}[h!]
-\includegraphics[width=0.9\textwidth]{HeraAntpos}
+\begin{figure}[h]
+\includegraphics[width=\linewidth]{HeraAntpos}
 \label{fig:HeraAntpos}
@@
 }
-\end{figure*}
+\end{figure}
 
@@
 \appendix
-\section{\label{sec:appA}\\Track-crossing}
-In this section we explain the reason why track-crossing, even in $uvw$ space, do not correspond to peak correlation for drift-scan measurements. 
+\section{\\Track-crossing and $w$-term \label{sec:appA}}
+\label{sec:appA}
+In this section we link our results from Section \ref{sec:method} to the traditional views of rotation synthesis and in particular $uv$ tracks. 
 Typical convention of rotation synthesis has rotation operates on $\boldsymbol{b'}$ instead of $\hat{\boldsymbol{s}}$, in which case Eq.\ref{eq:Theta} becomes
@@
 \end{equation}
-We are now in a position to understand the reason that track-crossing, even in $uvw$ space, as mentioned section \ref{sec:tracks}, would not correspond exactly to the peak of correlation.
+
 Track-crossing corresponds to vanishing of the exponent
@@
  \]
-This condition maximizes $\Theta$ if and only if no other term in the integral depends on $\Gamma$.
-We see from Eq.
-\ref{tracking} that this condition is satisfied.
-In a general drift-scan measurement, however, track-crossing point of Eq.
-\ref{eq:trackcross} does not maximize $\Theta$.
+If exponent vanishes, $\Theta_{\nu}$ is real, and one would not observe any non-zero phase at the peak.
+However, with inclusion of the $w$-term, we see that track crossings in $uv$-plane do not actually imply crossing in the $uvw$ space, therefore the exponent Eq.
+\ref{eq:trackcross} does not vanish and we observe non-zero, frequency dependent peak phases even in the case of tracking arrays, as shown in Fig.
+\ref{fig:phi_nu}.
+
+Furthermore, in the case of drift-scan arrays, even a crossing in $uvw$ space does not necessarily correspond to the maximum of correlation.
+From Eq.
+\ref{eq:drift}, we see the track-crossing condition Eq.
+\ref{eq:trackcross} maximizes $\Theta$ if and only if no other term in the integral depends on $\Gamma$.
+We see that this is true for the tracking case (Eq.
+\ref{eq:tracking}), but not the drift-scan case (Eq.
+\ref{eq:drift}).
 

### 2017-07-07 Rotation paper/draft_working.tex
save

@@
 Yunfan Gerry Zhang \altaffilmark{1},
-Adrian C. Liu \altaffilmark{1, 2},
+Adrian Liu \altaffilmark{1, 2},
 Aaron R. Parsons\altaffilmark{1, 2}
@@
 \begin{abstract}
-The observational effort to measure the cosmological 21\,cm power spectrum with radio interferometers requires high sensitivity.
+Measuring the cosmological 21\,cm power spectrum with radio interferometers requires high sensitivity.
+Many modern low-frequency interferometers are thus designed with antennas placed on regular grid to maximize redundancy.
 Current visibility-based power spectrum pipelines, though shown to ease control of systematics, lack the ability to include partial redundancy.
 We introduce a method to include partial redundancy in such power spectrum pipelines of drift-scan arrays.
-Our method efficiently finds pairs of baselines to cross-multiply, and quantifies the sensitivity contributions of given baselines.
-Using the 128-element configurations and beams of the Precision Array for Probing the Epoch of Reionization (PAPER-128), as well as 4 planned versions of the Hydrogen Epoch of Reionization Array (HERA), we illustrate how our method applies to different arrays and predict the sensitivity improvements of including each baseline pairs.
+Our method pairs baselines to cross-multiply at a time lag, and quantifies the sensitivity contributions of each pair of baselines.
+Using the 128-element configurations and beams of the Donald C.
+Backer Precision Array for Probing the Epoch of Reionization (PAPER-128), and 4 planned versions of the Hydrogen Epoch of Reionization Array (HERA), we illustrate how our method applies to different arrays and predict the sensitivity improvements of including each baseline pairs.
 We show that inclusion of partially redundant baselines
@@
 
-In the past, observational studies of reionization, including Gunn-Peterson measurements of quasi-stellar objects \citep{Fan2006} and cosmic microwave background (CMB) measurements of temperature and anisotropy \citep{Planck2016}, Kinetic Sunyaev-Zeldo'vich effect \citep{kszpatchy} and and Lyman alpha emitter clustering \citep{mcquinnLyA} have given us indications of the rough time-frame of reionization, but only limited constrains on the finer spatial and temporal variations.
-A surge of recent radio-astronomical experiments of reionization focus on measurement of the ``spin-flip'' transition of neutral
+Observational studies of reionization, including Gunn-Peterson measurements of quasi-stellar objects \citep{Fan2006} and cosmic microwave background temperature and anisotropy measurements (CMB;\citealt{Planck2016}), the Kinetic Sunyaev-Zeldo'vich effect \citep{kszpatchy} and Lyman alpha emitter clustering \citep{mcquinnLyA} have given us indications of the rough time-frame of reionization, but only limited constraints on the finer spatial and temporal structures.
+A surge of recent radio-astronomical experiments of reionization focus on measuring the ``spin-flip'' transition of neutral
 hydrogen of characteristic wavelength 21\,cm \citep{Furlanetto2006181,PritchardLoeb}.
-In complement to the aforementioned probes, the 21\,cm brightness temperature distribution is a direct tracer of neutral hydrogen through the epoch of reionization, and thus tomography of the 21\,cm line is a direct measurement of the full temporal and spatial variations of this event.
-Before possible realization of full scale 21\,cm tomography, many current radio interferometric efforts
-prioritize measuring the spatial power spectrum of 21\,cm brightness temperature fluctuations.
-Current-generation instruments include the Precision Array for Probing
-the Epoch of Reionization (PAPER) \citep{Ali2015,paper32}, Murchison
-Widefield Array (MWA) \citep{Bowman2013, Tingay2013}, Low Frequency Array (LOFAR) \citep{LOFAR}. Next-generation instruments include the Hydrogen Epoch of Reionization
-Array (HERA) (e.g. \citealt{HERA,HERAconfiguration,HERABEAM1,HERADISH2})  under construction, 
-and the Square Kilometer Array Low (SKA-low) (e.g. \citealt{SKA1}), which is currently in planning.
+The 21\,cm brightness temperature is a direct tracer of neutral hydrogen through the epoch of reionization, and thus tomography of the 21\,cm line is a direct measurement of the full temporal and spatial variations of this event.
+Before realizing of full-scale 21\,cm tomography, many current radio interferometric efforts
+aim to measure the spatial power spectrum of 21\,cm brightness temperature fluctuations.
+Current-generation instruments include the Donald C. Backer Precision Array for Probing
+the Epoch of Reionization (PAPER; \citealt{Ali2015,paper32}), Murchison
+Widefield Array (MWA; \citealt{Bowman2013, Tingay2013}), Low Frequency Array (LOFAR; \citealt{LOFAR}). Next-generation instruments include the Hydrogen Epoch of Reionization
+Array (HERA; e.g. \citealt{HERA,HERAconfiguration,HERABEAM1,HERADISH2})  under construction, 
+and the Square Kilometer Array Low (SKA-low; e.g. \citealt{SKA1}) currently in planning stages.
 
-The highly redshifted 21\,cm signal is faint and diffuse, in contrast to the localized bright sources that have been of interest in many traditional applications of radio astronomy.
-Current experiments are sensitivity-starved, with the sensitivity requirement further increased considering the foreground contaminations 5 orders of magnitude brighter than the cosmological signal of interest.
-Thus modern low-frequency radio interferometers aimed at measuring the 21\,cm signal bears designs different from traditional instruments.
-Many specially designed arrays, such as PAPER and HERA, feature multiple copies of the same baselines to repeatedly measure the same Fourier signal to increase sensitivity.
+The highly redshifted 21\,cm signal is faint and diffuse, in contrast to the localized bright sources targeted by many traditional radio telescopes.
+Current 21\,cm experiments are sensitivity-starved, with the sensitivity requirement further increased considering the foreground contaminations 5 orders of magnitude brighter than the cosmological signal of interest.
+Low-frequency radio interferometers aiming to measure the 21\,cm signal are thus designed differently from traditional instruments.
+Specially designed arrays such as PAPER and HERA feature multiple copies of the same baselines to repeatedly measure the same Fourier signal to increase sensitivity \citep{first-paper}.
 To satisfy the sensitivity need, modern arrays are large.
@@
 Analysis pipelines for the 21\,cm power spectrum typically fall into two categories.
-In the first, more traditional technique, images are formed in Fourier domain through rotation synthesis, followed by a foreground mitigation step to construct a power spectrum.
-An alternative technique works directly with visibilities from baselines and cross multiplies them to form the power spectrum.
-Tracking closely the instrumental output of visibilities, this technique allows one to deal with systematics such as Radio-Frequency-Interference(RFI) and instrumental cross-talks in a more native context, and thus tend to provide good limits on the power spectrum.
+In the first, images are formed in Fourier domain through rotation synthesis, followed by a foreground mitigation step to construct a power spectrum.
+An alternative technique works directly with visibilities from baselines, delay-transforming and cross-multiplying them to form the power spectrum.
+This technique avoids many systematics associated with combining data from different baselines and tracks the native sampling of the interferometer.
 An example of the visibility-based pipeline was presented in \cite{Ali2015}, which provides the newest upper limit to the power spectrum
 measurements with the 64-element version of PAPER (henceforth as PAPER-64).
-However, one disadvantage of existing visibility-based pipelines is the lack of use of partial redundancy.
-Baselines of different length and orientation still contain partially redundant information.
-While an imaging based power-spectrum pipeline naturally includes all redundancy information, visibility-based pipelines until now only cross multiply fully redundant baselines, in other words baselines of the same length and orientation.
+However, one disadvantage of existing visibility-based pipelines is their lack of use of partial redundancy.
+Baselines of different lengths and orientations contain partially redundant information.
+While imaging based power-spectrum pipelines naturally includes all redundancy information, visibility-based pipelines so far only cross multiply fully redundant baselines, i.e.
+baselines of the same length and orientation.
 
-The use of partially redundant baselines is in principle not a fundamental limitation of visibility-based pipelines.
-In fact, most instrumental reports to date do include partial redundancy.
-Recently, \cite{wterm} proposed a visibility-based approach to extract power spectrum from partially redundant baselines in tracking measurements.
-Arrays capable of tracking include MWA and LOFAR.
-Our work parallels this effort by focusing on drift-scan arrays that do not have tracking capabilities, such as PAPER, HERA and potentially SKA.
+Visibility-based pipelines are not fundamentally limited from using partially redundant baselines.
+In fact, most sensitivity forecasts to date do include partial redundancy.
+Recently, \cite{wterm} proposed a visibility-based approach to extract power spectrum from partially redundant baselines in sky-tracking measurements.
+Arrays capable of such tracking include MWA and LOFAR.
+Our work parallels this effort by focusing on drift-scan arrays that do not have tracking capabilities, such as PAPER, HERA and potentially SKA-low.
+
 The Earth's rotation causes the baselines in a drift-scan array to pick up different modes of the sky with time.
@@
 
-The rest of this paper is organized as follows. In section 2 we introduce some terminology and notations to be used for the rest of the paper. In section
+The rest of this paper is organized as follows. In section 2 we introduce some terminology and notation used in the rest of the paper. In section
 3 we introduce the formalism for weighting partially redundant baselines.
@@
 this technique as well as the expected sensitivity improvement
-with this method for HERA and PAPER-128 and with section 5 we conclude.
+ this method provides for HERA and PAPER-128 pipelines, and with section 5 we conclude.
 
@@
 
-In order to eliminate confusion and ambiguity for the rest of this paper, we
+In order to avoid confusion and ambiguity for the rest of this paper, we
 introduce some terminology that may differ from what is commonly found in the literature.
 
-We make the distinction of a \textit{baseline}, which corresponds to two specific antennas, and a \textit{class of baselines}, which refers to all baselines of the same length and orientation in a given array.
+We make the distinction between a \textit{baseline}, which corresponds to two specific antennas, and a \textit{class of baselines}, which refers to all baselines of the same length and orientation in a given array.
 Baselines of the same class are traditionally called
@@
 times. Non-equivalent baselines also have partial redundancy, and the redundancy can be maximized if their
-respected time series are shifted with respect to one another by some delay. In other words, one baseline can be ``rotated into''
-another if they are \textit{near-equivalent}.
+respected time series are shifted with respect to one another by some delay.
 
 \begin{figure}[H]
-\includegraphics[width=\linewidth]{antpos128}
+\includegraphics[width=\linewidth]{antconfig}
 
 \caption{The PAPER128 layout. Each dot corresponds to the location of
-an antenna. Top panel shows the antenna positions drawn to scale;
-bottom panel show the antenna labels and distances of the 112-element grid, in other words excluding the outrigger antennas.
+an antenna. The top panel shows the antenna positions drawn to scale;
+the bottom panel show the antenna labels and distances of the 112-element grid, in other words excluding the outrigger antennas.
 The numbering of the antennas in the bottom panel are original labels
 during instrument assembly and does not bear significant
-meaning. In the bottom panel, the two baselines (5,77) and (40,85) marked with red segments are example of an equivalent pair, both with separation denoted by \{1,0\}, for the
-antennas are separated by 1 unit east and 0 unit north. Similarly,
+meaning. In the bottom panel, the two baselines marked with red segments are example of an equivalent pair, each with antennas separated by 1 unit east and 0 unit north. Similarly,
 the baselines marked in blue and yellow are examples
-of the classes \{1,1\} and \{1,-1\}, respectively. Due to the small North-South separation within the grid, \{1,1\} is expected to be near-equivalent to \{1,0\}. }
+of classes that , due to the small North-South separation within the grid (see top panel), are expected to be nearly equivalent to the red baselines. }
 \label{fig:AntPos}
@@
 The PAPER array is located in the Karoo desert in South Africa (30:43:17.5
-S, 21:25:41.8 E). The layout pattern with antenna labels are show
+S, 21:25:41.8 E). The layout pattern with antenna labels is shown
 in Fig.
 \ref{fig:AntPos}.
-The array consists of a 112-element grid and 16 "outriggers" used primarily to aid calibration.
-In the bottom panel, the two baselines (5, 77) and (40, 85), marked in red, are an example of equivalent pair.
+The array consists of a 112-element core is a rectangular grid, and 16 "outriggers" used primarily to aid calibration.
+In the bottom panel, the two baselines marked in red, are an example of equivalent pair.
 We denote a equivalency-class of baselines in the PAPER grid by their separations, in this case  \{1,0\}, for the
@@
 are close (4m), and hence these baselines are not suitable
-for cross-multiplication due to cross-talks.
+for cross-multiplication due to cross-coupling.
 On the other hand, the close North-South separation means that classes such as \{1,0\} and  \{1,1\} are expected to be near-equivalent.
@@
 
+In this section we introduce our method to cross-multiply nearly equivalent baseline classes.
+
 \subsection{$uvw$ tracks \label{sec:tracks}}
 
-Given a point source on the sky, each baseline maps to a point in $uvw$ space. As the Earth rotates with respect to
+Relative to a phase center on the sky, each baseline maps to a point in $uvw$ space. As the Earth rotates with respect to
 the source, the points trace out tracks in the $uvw$ space. 
-We show in Fig. \ref{fig:Tracks} $uvw$ tracks of PAPER-128 over 12 hours, at 0.15GHz, for a source that passes through zenith.
+We show in Fig. \ref{fig:Tracks} $uvw$ tracks of PAPER-128 over 12 sidereal hours, at 0.15GHz, relative to a phase center that passes through zenith.
 
 \begin{figure}[H]
-\includegraphics[width=\linewidth]{tracks3D}
-\caption{Tracks of PAPER-128 (only a randomly selected subset of 32 baselines are shown) for frequency $\nu=0.15\text{GHz}$ and a hypothetical source that passes through zenith.
-These tracks are traced out over 0.5 sidereal days. Color represents different baselines. 
+\includegraphics[width=\linewidth]{rotation}
+\caption{Tracks of PAPER-128 for a selection of 16 baselines and frequency $\nu=0.15\text{GHz}$ and a phase center that passes through zenith.
+Tracks are traced out over half a sidereal day. Color represents different baselines. 
 As the Earth rotates, tracks are traced out counterclockwise. Tracks in $uvw$ space do not cross, even though the $uv$ projections do. }
@@
 
-Traditionally, we can identify
-redundancy of near-equivalent baselines as crossings
+Equivalent baselines follow identical $uvw$ tracks. Traditionally, we can identify
+redundancy of nearly equivalent baselines as crossings
 of the $uv$ tracks, a 2 dimensional projection of Fig.
@@
 The most obvious reason is that the 3 dimensional $uvw$ tracks do not actually cross.
-For drift-scan arrays, even a crossing in $uvw$ space does not imply perfect redundancy.
+For drift-scan arrays, even a hypothetical crossing in $uvw$ space would not imply perfect redundancy.
 This shall become evident in the next section, after we develop a more general formalism that accounts for the point spread function of the finite beams, and estimate the degree of redundancy for general combination of baselines at a general time-offset.
-An explicit explanation of track-crossing will be presented in Appendix \ref{sec:appA}.
+Then, relation of track-crossing and redundancy for both drift-scan and tracking measurements will be explored in Appendix \ref{sec:appA}.
 
@@
 In this section we formulate the relation between 
-the product of visibilities from two arbitrary baselines and the power spectrum.
+the product of delay-transformed visibilities from two arbitrary baselines and the power spectrum.
 
@@
 \end{equation}
-Here $k_B$ is Boltzmann's constant, $\lambda$ is the mean wavelength, $\boldsymbol{b}$ is the baseline length, $\hat{\boldsymbol{s}}$ and $\Omega$ are a direction in the
-sky and its corresponding solid angle. Inside the integral we have $\phi(\nu)$ the frequency bandpass profile, $A_{\nu}$ the (frequency
-dependent) primary beam, and $I$ the specific intensity, which has been
+where $k_B$ is Boltzmann's constant, $\lambda$ is the mean wavelength, $\boldsymbol{b}$ is the baseline vector, $\hat{\boldsymbol{s}}$ and $\Omega$ are a direction in the
+sky and its corresponding solid angle.
+Inside the integral we have $\phi(\nu)$ the frequency bandpass profile, $A_{\nu}$ the (frequency-dependent) primary beam, and $I$ the specific intensity, which has been
 related to $T$, the brightness temperature in the Rayleigh-Jeans
-limit. The beam reception pattern $A$ is dimensionless,
+limit. The beam power pattern $A$ is dimensionless,
 normalized to 1 at its peak (zenith), and we assume it to be the same
@@
 
-In practice,
-power spectrum measurements are typically taken from a few ten MHz centered around the corresponding redshift of interest (e.g. 150 MHz for z=9.5).
+Power-spectrum measurements are typically taken from $\sim10$MHz centered around the corresponding redshift of interest (e.g. 150 MHz for z=9.5).
 
-Delayed-transformed visibility has been widely adopted in recent years since it has been shown that foregrounds are isolated in delay space.
-We define the delay-transformed visibility by Fourier transforming visibility along the frequency axis\citep{delay-transform} as:
+Delayed-transformed visibility has gained popularity in recent years since it has been shown that foregrounds are isolated in delay space.
+We define the delay-transformed visibility by Fourier transforming visibility along the frequency axis \citep{delay-transform} as:
 \small
@@
 \normalsize
-Here the delay $\tau$ is the Fourier variable. Eq. \eqref{eq:Vb1} expresses the delay-transformed visibility as
+Here the delay $\tau$ is the Fourier dual of $\nu$. Eq. \eqref{eq:Vb1} expresses the delay-transformed visibility as
 an integral over observational coordinates $\hat{\boldsymbol{s}}$ and $\nu$.
@@
 \end{equation}
-Note that in Eq.\ref{eq:thinshell} no flat-sky approximation has been made; the angular integral is still performed over the dome.
+Note Eq. \eqref{eq:thinshell} does not require the flat-sky approximation; the angular integral is still performed over the curved sky.
 
@@
 
-Most existing visibility based power-spectrum pipelines for redundant drift-scan  arrays relate the power-spectrum to the conjugate square of the visibilities \citep{first-paper, paper32, Ali2015}.
+Existing visibility based power-spectrum pipelines for redundant drift-scan  arrays relate the power-spectrum to the conjugate square of the visibilities \citep{delay-transform, paper32, Ali2015}.
 We would like to generalize such relations by relating the power-spectrum to the product of two visibilities from two arbitrary baselines and time-offsets.
 The beam pattern of a baseline shifts relative to the sky as the Earth rotates. Here we choose to fix the sky, and denote the rotated coordinates
-of the beam pattern with the 3 dimensional rotation operator $\Gamma$:
+in the topocentric frame with the 3 dimensional rotation operator $\Gamma$:
 \small
@@
 \begin{equation}
-\begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau)\rangle\\
+\begin{aligned} & \langle e^{i\psi_{\nu}} V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau)\rangle\\
  & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle\right)A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
@@
  & \approx\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}rd^{3}r'\delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
- & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}r|A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r})|\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\Gamma \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)\right],\\
- & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma \hat{\boldsymbol{s}},\nu)|\exp\left[-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)\right],
+ & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}rA^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r})\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\Gamma \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)+i\psi_{\nu}\right],\\
+ & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma \hat{\boldsymbol{s}},\nu)\exp\left[-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)+i\psi_{\nu}\right],
 \end{aligned}
@@
 \end{equation}
-
-where in transition from cosmological coordinates back to observing coordinates we have written $\hat{\boldsymbol{r}}\equiv\hat{\boldsymbol{s}}$, and 
+where we introduced a frequency-dependent phase $\psi_{\nu}$, which to first order corresponds to phasing the two visibilities to the same phase center.
+\footnote{Turns out for drift-scan measurements, the first-order interpretation is largely sufficient, but we keep the term for now to reserve the option to determine its exact form numerically.} In transition from cosmological coordinates back to observing coordinates we have written $\hat{\boldsymbol{r}}\equiv\hat{\boldsymbol{s}}$, and
 \begin{equation}
-\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'})=\exp\left[-i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\Gamma\hat{\boldsymbol{r'}}\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)\right].
+\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'})\equiv\exp\left[-i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\Gamma\hat{\boldsymbol{r'}}\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)+i\psi_{\nu}\right].
 \end{equation}
 \end{widetext}
-The third equality of Eq.(\ref{eq:main}) follows from assumption of translational invariance of sky statistics, and the first inequality follows from the assumption that
-the 3D power spectrum varies negligibly over the $k$-space of interest
-so that $\hat{P}_{21}(k+k')\approx\hat{P}_{21}(k)$.
-The careful reader would have noticed that we have also assumed here a definitive relation between $k$ and $b,\tau$.
+The third equality of Eq.(\ref{eq:main}) follows from assumption of translational invariance of statistics of the 21\,cm field, and the first inequality follows from the assumption that
+the 3D power spectrum varies negligibly over the $k$-space of interest.
+The careful reader may have noticed that we have also assumed here a definitive relation between $k$ and $b,\tau$.
 In the flat-sky limit for example, such a relation would take the form $(Xk_{x},Xk_{y},Yk_{z})=\frac{2\pi}{c}(b_{x},b_{y},\tau)$.
@@
 Notice that the phase factor $\exp\left[-i2\pi\tau\left(\nu-\nu'\right)\right]$
-drops out in the end. This means that the location and magnitude of the correlation peak does not depend on delay $\tau$, an important point we shall come back to.
+drops out in the end. This means that the location and magnitude of the correlation peak does not depend on delay $\tau$.
 
@@
 \begin{equation}
-\begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau)\rangle\\
- & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)|\\
- & \qquad \qquad \qquad \qquad \exp\left[-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)\right].\end{aligned}
+\begin{aligned} & \langle e^{i\psi_{\nu}} V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau)\rangle\\
+ & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)\\
+ & \qquad \qquad \qquad \qquad \exp\left[-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)+i\psi_{\nu}\right].
+ \end{aligned}
 \label{eq:final}
@@
 \begin{equation}
- P(k_{b,\tau}) = \left(\frac{\lambda^{2}}{2k_{B}}\right)^{2} \frac{\langle V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau)\rangle}{\Theta}, 
+ P(k_{b,\tau}) = \left(\frac{\lambda^{2}}{2k_{B}}\right)^{2} \frac{\langle e^{i\psi_{\nu}} V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau)\rangle}{\Theta}, 
  \label{eq:opp}
 \end{equation}
-where the weight
+where the weight is defined as 
 \begin{equation}
@@
 \end{equation}
-and
+with
 
 \[
-\Theta_{\nu} \equiv \int \frac{d\Omega}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)| e^{-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)}e^{i\psi_{\nu}}.
+\Theta_{\nu} \equiv \int \frac{d\Omega}{X^{2}Y}A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu) e^{-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)}e^{i\psi_{\nu}}.
 \label{eq:Thetanu}
@@
 In Eq.
-\ref{eq:Thetanu} we introduced an extra phase $\psi$, the origin of which we shall explain in \ref{sec:rephs}.
+\eqref{eq:Thetanu} we introduced an extra phase $\psi$, the origin of which we shall explain in \ref{sec:rephs}.
 We point out that although all our derivations focused on drift-scan telescopes, we can get the analogous result for tracking measurements simply by noticing that for a tracking primary beam, $\Gamma$ becomes a rotation around zenith, and so:
 \[
-\Theta_{\nu} \equiv \int \frac{d\Omega}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\hat{\boldsymbol{s}},\nu)| e^{-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)}e^{i\psi_{\nu}}.
+\Theta_{\nu} \equiv \int \frac{d\Omega}{X^{2}Y}A^{*}(\hat{\boldsymbol{s}},\nu)A(\hat{\boldsymbol{s}},\nu) e^{-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)}e^{i\psi_{\nu}}.
 \label{eq:Thetanu_tracking}
@@
 
-Roughly speaking, Eq. \ref{eq:opp} through \ref{eq:Thetanu} tell us that the cross multiplications of visibilities at a time delay
-in $uv$-space is proportional to the power spectrum times the Fourier
-transform of the cross multiplied beam pattern. As a check, when applied to equivalent baselines,
+Roughly speaking, Eqs. \eqref{eq:opp} through \eqref{eq:Thetanu} tell us that the cross multiplications of visibilities at a time delay is proportional to the power spectrum times the Fourier
+transform of the cross multiplied beam patterns. As a check, when applied to equivalent baselines,
 $\boldsymbol{b}=\boldsymbol{b'}$, $\hat{\boldsymbol{s}}=\Gamma\hat{\boldsymbol{s}}$, and Eq.(\ref{eq:final}) reduces to Eq.(B9) of \cite{paper32}. 
 With Eq.
-\ref{eq:opp} and Eq.
-\ref{eq:Theta} we can, for any given pair of baseline classes and time delay, estimate the degree of redundancy, here represented by $\Theta$, thereby achieving all our goals stated in the introduction, i.e.
-to identify
-candidate baseline pairs with good redundancy, to find the time offset that maximizes redundancy, and to quantify the degree of such redundancy.
+\eqref{eq:opp} and Eq.
+\eqref{eq:Theta} we can, for any given pair of baseline classes and time delay, estimate the degree of redundancy, here represented by $\Theta$.
+This allows us to achieve our goals stated in the introduction: to identify
+candidate baseline pairs with significant redundancy, to find the time offset that maximizes redundancy, and to quantify the degree of such redundancy.
 We can do all the above simply by computing the weight $\Theta$ from
-Eq.(\ref{eq:Theta}) for various time offsets, without having to actually cross-multiply visibilities for all baselines and offsets. This makes the task computationally tractable.
+Eq.(\ref{eq:Theta}) for various time offsets.
 
 \subsection{Rephasing \label{sec:rephs}}
-As we see in Eq.
-\ref{eq:Thetanu}, $\Theta_{\nu}$ at the peak of correlation is in general complex, and often far from real.
+If $\psi_{\nu}=0$ in  Eq.
+\eqref{eq:Thetanu}, $\Theta_{\nu}$ at the peak of correlation is in general complex, and often far from real.
 Furthermore, this phase of peak correlation is inevitably frequency dependent.
@@
 
-%Doing so is also crucial to reduction of additional mode mixing, as discussed in \cite{Hazelton2013}. Shapes of overlapping regions of pairs of baselines are fixed with respect to frequency.
+To first order, the physical origin of $\psi_\nu$ lies in the two visibilities having different phase centers.
+By default the correlators of a drift-scan array phase the two visibilities both to zenith at the same time.
+When they are cross-multiplied with a time lag, the visibilities must be rephased before delay transform to the account for the movement of the zenith.
+The effect of the phase thus roughly corresponds in a shift of the delay mode measured, and we should expect $\psi_{\nu}$ to be in first order a linear function of $\nu$:
+\[
+\psi_{\nu}\sim 2\pi\Delta\tau\nu.
+\]
+In Fig.
+\ref{fig:phi_nu} we show the peak phases for for a given baseline pair of PAPER128 (\{2,0\}:\{2,1\}), comparing the drift-scan phase dependence with that of the same baseline with hypothetical tracking-elements (Eq.
+\eqref{eq:Thetanu} and Eq.
+\eqref{eq:Thetanu_tracking}).
+In the drift-scan case, we indeed see a linear relation corresponding to a delay of $\Delta\tau\approx15.7$ns.
+In the tracking case, because of minimal zenith movement, only second order effects are observed.
+The origin of the second order effects can be seen as due to the $w$-term, or more precisely the fact that $uvw$ tracks do not cross when their 2 dimensional projections do.
+We refer the reader to Appendix \ref{sec:appA} for further explanation.
+In both cases, the full effects are encapsulated in the phase of $\Theta_\nu(\psi_\nu=0)$ and can be thus determined empirically.
+
+\begin{figure}[H]
+\includegraphics[width=1\linewidth]{phi_nu}
+
+\caption{Frequency dependent peak phases for PAPER-128 baseline pair \{2,0\}:\{2,1\}.
+Both the drift-scan case and a hypothetical tracking baseline are shown in solid lines.
+The first-order, linear effects are shown in dashed lines.
+We have also fixed the global phase in both cases to 0 at 0.15GHz.
+The drift-scan case exhibits linear behavior and conforms well to the first order effect of a shift in delay space due to the movement of zenith.
+}
+\label{fig:phi_nu}
+\end{figure}
 
 To illustrate the effect of rephasing, we compare in Fig.
-\ref{fig:freqdiff} real parts of $\Theta_{\nu}$ of both equivalent and near-equivalent baseline pairs for two channels: 0.16 GHz and 0.17 GHz.
-The top two panels have zero rephasing, and the bottom two are rephased to a time offset of 0.0325 sidereal days, the optimal time offset for the given baseline pair.
+\ref{fig:freqdiff} real parts of $\Theta_{\nu}$ of both equivalent and nearly equivalent baseline pairs for two channels: 0.16 GHz and 0.17 GHz.
+The top two panels have zero rephasing, and the bottom two are linearly rephased to a differential delay of $\Delta\tau\approx15.7$ns, as we determined above.
 The first and third panels show the equivalent baseline pairs \{2,0\}:\{2,0\}, and second and fourth panels show \{2,0\}:\{2,1\}.
@@
 The wider the frequency profile, the more destructive the interference would be.
-In the rephased case, the phases of the near-equivalent case match up and can be added without compromising sensitivity.
+In the rephased case, the phases of the nearly equivalent case match up at the peak of correlation and can be added without compromising sensitivity.
 We should thus rephase the data separately for each set of baseline pairs.
@@
 Specifically shown are the real parts of $\Theta_{\nu}$.
-First and third panels shows equivalent baselines, second and fourth show a pair of near-equivalents.
+First and third panels shows equivalent baselines, second and fourth show a pair of nearly equivalents.
 The top two panels have zero phase shift ($\psi_{\nu}=0$), and the bottom two are rephased to a time offset of 0.055 sidereal days.
@@
 \label{fig:freqdiff}
-\end{figure}
-
-The physical origin of $\psi_\nu$ lies in the two visibilities having different phase centers.
-By default the correlators of a drift-scan array phase the two visibilities both to zenith at the same time.
-When they are cross-multiplied with a time lag, the visibilities must be rephased before delay transform to the account for the movement of the zenith.
-One can check that introduction of a phase in Eq.
-\ref{eq:Vis1} does carry through the derivation to Eq \ref{eq:Thetanu}.
-The effect of the phase thus roughly corresponds in a shift of the delay mode measured, and we should expect $\psi_{\nu}$ to be in first order a linear function of $\nu$:
-\[
-\psi_{\nu}\sim 2\pi\Delta\tau\nu.
-\]
-In Fig.
-\ref{fig:phi_nu} we show the peak phases for for a given baseline pair of PAPER128 (\{2,0\}:\{2,1\}), comparing the drift-scan phase dependence with that of the same baseline with hypothetical tracking-elements (Eq.
-\ref{eq:Thetanu} and Eq.
-\ref{eq:Thetanu_tracking}).
-In the drift-scan case, we indeed see a linear relation corresponding to a delay of $\Delta\tau\approx17$ns.
-In the tracking case, because of minimal zenith movement, only second order effects are observed.
-The origin of the second order effects can be seen as due to the $w$-term, or more precisely the fact that $uvw$ tracks do not cross when their 2 dimensional projections do.
-We refer the reader to Appendix \ref{sec:appA} for further explanation.
-
-\begin{figure}[H]
-\includegraphics[width=1\linewidth]{phi_nu}
-
-\caption{Frequency dependent peak phases for PAPER-128 baseline pair \{2,0\}:\{2,1\}.
-Both the drift-scan case and a hypothetical tracking baseline are shown.
-We have also fixed the global phase in both cases to 0 at 0.145GHz.
-The drift-scan case exhibits linear behavior and to first order corresponds to a shift in delay space to account for the movement of zenith.
-}
-\label{fig:phi_nu}
 \end{figure}
@@
 \section{Analysis}
+
+In this section we delve into visualizing and testing our method, and explore sensitivity contributions of various baseline classes for PAPER and HERA array configurations. 
+\subsection{Visualization \label{sec:visual}}
+
+So far for clarity and generality we have avoided the traditional formulation of radio astronomy in terms of the $uv$-plane.
+To gain some visual intuition of the formalism in Section \ref{sec:method}, we show explicit beam fringe patterns of the two baselines and their interactions.
+For this section, we relax into the flat sky approximation and examine a single frequency of $150$MHz.
+
+At the time-offset of maximum redundancy, we expect two beams to have to same fringe pattern in both frequency and phase.
+Due to the time delay, however, the beam centers would be slightly shifted with respect to each other.
+This we show in the top panels of Fig.
+\ref{fig:beamfringe}.
+The top left and middle panels show the beam-fringe
+patterns (real parts) for baselines \{2,0\} and \{2,1\}, delayed by 0.0325 sidereal days.
+The product of the beam fringe patterns in top right shows that the fringes
+ cancel out as we expect.
+On the bottom of Fig.
+\ref{fig:beamfringe} we show the instantaneous $uv$ coverage of two baselines and their product.
+We see from the bottom right panel that the cross-multiplied baselines recover  power concentrated at $u\approx\frac{30\text{m}}{c/1.5\times10^8\text{Hz}}\approx15$.
+Note the bottom right panel is not the Fourier transform of the top right panel.
+
+\begin{figure*}[h!]
+\includegraphics[width=\textwidth]{BeamFringe}
+
+\caption{PAPER128 beam-fringe patterns and their point-spread functions on the $uv$-plane for a frequency of $150$MHz.
+Selected baselines have 30 meter East-West separation, and 0 and 4 meter North-South separation, respectively.The top panels show the real parts of the beam-fringe patterns of the baselines (left and middle), and their conjugate product (right).
+The beam-fringe values are normalized such that the peak of the original beam is unity.
+The bottom left and middle panels show the $uv$ beam spread of the two beam-fringe patterns.
+The product of the beam spreads displayed in the bottom right panel shows that power concentrated at $u\approx15$ is recovered.
+}
+\label{fig:beamfringe}
+\end{figure*}
+
 \subsection{Numerical Test \label{sec:Techniquet}}
 
-Bringing together the discussion from Section \ref{sec:method}, we present a numerical check of Eq.\ref{eq:final_new}, including rephasing.
+Bringing together the discussion from Section \ref{sec:method}, we present a numerical check of Eq.
+\eqref{eq:final_new}, including rephasing.
 To do so, we need to compare the amplitude and phase of the integral weight $\Theta$
@@
 Fig.
-\ref{fig:numerics} shows the real parts of the resulting degree of redundancy as a function of time offset, for equivalent baseline pairs (\{2,0\}:\{2,0\}) in the top panel and near-equivalent (\{2,0\}:\{2,1\}) in the bottom.
+\ref{fig:numerics} shows the real parts of the resulting degree of redundancy as a function of time offset, for equivalent baseline pairs (\{2,0\}:\{2,0\}) in the top panel and nearly equivalent (\{2,0\}:\{2,1\}) in the bottom.
 Shown in green for the two cases are the computed values of $\Theta$, both normalized to the peak of the equivalent case.
 The plot on the bottom is rephased with a delay of $\Delta\tau\approx17$ns as outlined in Section \ref{sec:rephs}.
-Since both y-axes are normalized to the same scale, we see the equivalent baseline reach maximal redundancy with 0 time-offset, while the near-equivalent pair reaches maximum redundancy of $\Theta\approx0.5$ at $dT\approx0.325$days.
+Since both y-axes are normalized to the same scale, we see the equivalent baseline reach maximal redundancy with 0 time-offset, while the nearly equivalent pair reaches maximum redundancy of $\Theta\approx0.5$ at $dT\approx0.325$days.
 
@@
 }.
-For the near-equivalent case, we rephase the delayed visibility (\{2,0\}) by the same delay $\Delta\tau\approx17$ns, followed by delay-transform.
+For the nearly equivalent case, we rephase the delayed visibility (\{2,0\}) by the same delay $\Delta\tau\approx17$ns, followed by delay-transform.
 The resulting delay-space visibilities for the two baselines are then convolved
 via the Fourier convolution theorem, to obtain values of the cross
-correlation as a function of time-offset.   We do this for both the equivalent (\{2,0\}:\{2,0\}) and near-equivalent (\{2,0\}:\{2,1\}) case.
-Since the blue and green curves overlap, we have verified that in this case Eq.\ref{eq:final_new} is valid and the computed $\Theta$ can be used as weight to obtain the power spectrum.
+correlation as a function of time-offset.   We do this for both the equivalent (\{2,0\}:\{2,0\}) and nearly equivalent (\{2,0\}:\{2,1\}) case.
+Since the blue and green curves overlap, we have verified that in this case Eq. \eqref{eq:final_new} is valid and the computed $\Theta$ can be used as weight to obtain the power spectrum.
 
@@
 The sensitivity contribution of $\{m,n\}$, cross multiplied with $\{m',n'\}$  thus roughly speaking scales as $\sqrt{\left((16-|m|)(7-|n|)(16-|m'|)(7-|n'|)\right)}=\sqrt{MM'}$.
-For cross-multiplications of near-equivalent baselines of
+For cross-multiplications of nearly equivalent baselines of
 types $\{m,n\}$ and $\{m',n'\}$, we get an effective weight:
@@
 
-The scaling in Eq. \ref{eq:sensul} was rough for simplicity of motivation. As we derive in Appendix \ref{sec:appB}, this weight should be corrected by a factor proportional $\rho_0$:
+The scaling in Eq. \eqref{eq:sensul} was rough for simplicity of motivation. As we derive in Appendix \ref{sec:appB}, this weight should be corrected by a factor proportional $\rho_0$:
 
@@
 
-For a given $\rho_0$ Eq.\ref{eq:tildereal} thus quantifies the relative sensitivity contribution of a baseline pair $bb'$.
+For a given $\rho_0$ Eq.
+\eqref{eq:tildereal} thus quantifies the relative sensitivity contribution of a baseline pair $bb'$.
 Assuming a reionization signal of $\Delta_{21cm}^2\sim 30mK^2$, observation at 150MHz ($z=8.5$), 120 days of observation with PAPER antennas, we have roughly
@@
 where $L$ is the baseline length.
-We only need a single characteristic $L$ even in the near-equivalent case because only baselines of nearly equal length would have high redundancy.As expected, baseline-pairs that have smaller $\widetilde{\Theta}$ contribute less to the sensitivity.
+We only need a single characteristic $L$ even in the nearly equivalent case because only baselines of nearly equal length would have high redundancy.As expected, baseline-pairs that have smaller $\widetilde{\Theta}$ contribute less to the sensitivity.
 
@@
 The larger arrays are designed with a ``gap'' dividing the antennas into three different groups.
-The gaps are designed so as to improve $uv$ coverage and ease calibration without compromising sensitivity, but also produces many more near-equivalent baselines than the versions without the gap.
+The gaps are designed so as to improve $uv$ coverage and ease calibration without compromising sensitivity, but also produces many more nearly equivalent baselines than the versions without the gap.
 The motivations behind the designs are explained in \cite{HERAconfiguration}.
 Compared to PAPER128, the hexagonal pattern of HERA lack short baselines that are oriented close to each other, and the smaller beam (Fig.
-\ref{fig:Beam}) means that we expect to see only longer near-equivalent baselines.
+\ref{fig:Beam}) means that we expect to see only longer nearly equivalent baselines.
 The lower multiplicities per class of baselines is made up with the larger number of classes of baseline-pairs, especially given the gap in the larger versions.
@@
 
-\item The top near-equivalent pairs in HERA the same $\Theta$ (not shown) as in PAPER128, but much lower $\widetilde{\Theta}$.
+\item The top nearly equivalent pairs in HERA the same $\Theta$ (not shown) as in PAPER128, but much lower $\widetilde{\Theta}$.
 This is because they are longer baselines with lower multiplicity.
@@
 The dashed lines represent the values when only the equivalent baseline-pairs are used.
-We see as expected that in all cases using the near-equivalent baselines lead to more and more significant improvements with lower $\widetilde{\Theta}_{min}$, or in other words when worse baseline pairs are used.
-The small HERA37, with no gap (like in HERA350) or short near-equivalent baselines (like in PAPER 128), will not benefit much from the near-equivalent baselines.
+We see as expected that in all cases using the nearly equivalent baselines lead to more and more significant improvements with lower $\widetilde{\Theta}_{min}$, or in other words when worse baseline pairs are used.
+The small HERA37, with no gap (like in HERA350) or short nearly equivalent baselines (like in PAPER 128), will not benefit much from the nearly equivalent baselines.
 The maximum benefits for other cases are expected to be around $20\%$ to $60\%$.
-PAPER128 is designed with highly redundant near-equivalent baselines, and thus  these baselines start contributing at higher $\widetilde{\Theta}_{min}$, but the gapped HERA configurations will benefit even more from near-equivalent baselines at low $\widetilde{\Theta}_{min}$ due to there being more classes of such pairs.
+PAPER128 is designed with highly redundant nearly equivalent baselines, and thus  these baselines start contributing at higher $\widetilde{\Theta}_{min}$, but the gapped HERA configurations will benefit even more from nearly equivalent baselines at low $\widetilde{\Theta}_{min}$ due to there being more classes of such pairs.
 Note that here we normalized $\rho$ such that the contribution of the top equivalent baseline pair, such as (sep0,1:0,1 in the PAPER128 case) are 1.
@@
 \caption{Sensitivity of redundant arrays as a function of the minimum effective weight.
-Dashed lines represent when only equivalent baseline pairs are used, while solid lines indicate use of both the equivalent and near-equivalent baselines are used.
+Dashed lines represent when only equivalent baseline pairs are used, while solid lines indicate use of both the equivalent and nearly equivalent baselines are used.
 The y-axis is normalized independently for each array such that the contribution of the top single equivalent pair class is unity, and thus does not indicate a comparison of absolute sensitivity across the different arrays shown.
@@
 In this section we link our results from Section \ref{sec:method} to the traditional views of rotation synthesis and in particular $uv$ tracks. 
-Typical convention of rotation synthesis has rotation operates on $\boldsymbol{b'}$ instead of $\hat{\boldsymbol{s}}$, in which case Eq.\ref{eq:Theta} becomes
+Typical convention of rotation synthesis has rotation operates on $\boldsymbol{b'}$ instead of $\hat{\boldsymbol{s}}$, in which case Eq. \eqref{eq:Theta} becomes
 \begin{equation}
@@
 However, with inclusion of the $w$-term, we see that track crossings in $uv$-plane do not actually imply crossing in the $uvw$ space, therefore the exponent Eq.
-\ref{eq:trackcross} does not vanish and we observe non-zero, frequency dependent peak phases even in the case of tracking arrays, as shown in Fig.
+\eqref{eq:trackcross} does not vanish and we observe non-zero, frequency dependent peak phases even in the case of tracking arrays, as shown in Fig.
 \ref{fig:phi_nu}.
@@
 From Eq.
-\ref{eq:drift}, we see the track-crossing condition Eq.
-\ref{eq:trackcross} maximizes $\Theta$ if and only if no other term in the integral depends on $\Gamma$.
+\eqref{eq:drift}, we see the track-crossing condition Eq.
+\eqref{eq:trackcross} maximizes $\Theta$ if and only if no other term in the integral depends on $\Gamma$.
 We see that this is true for the tracking case (Eq.
-\ref{eq:tracking}), but not the drift-scan case (Eq.
-\ref{eq:drift}).
+\eqref{eq:tracking}), but not the drift-scan case (Eq.
+\eqref{eq:drift}).
+Only in the special case of a tracking measurement, where the $w$-term happens to be negligible, do $uv$-track crossings correspond exactly to maxima of correlations.
+Incidentally, at the crossing points the exponent in Eq.
+\eqref{eq:trackcross} vanishes and $\psi_\nu$ is zero for all $\nu$.
+In general, second order effects due to the presence of $w$ is observed for the tracking measurements as in Fig.
+\ref{fig:phi_nu}.
 
@@
 \end{equation}
-where we have defined a slightly modified version of the effective weight (compare with Eq. \ref{eq:sensul}):
+where we have defined a slightly modified version of the effective weight (compare with Eq. \eqref{eq:sensul}):
 \begin{equation}

### 2017-07-07 Rotation paper/draft_working.tex
save

@@
 during instrument assembly and does not bear significant
-meaning. In the bottom panel, the two baselines marked with red segments are example of an equivalent pair, each with antennas separated by 1 unit east and 0 unit north. Similarly,
+meaning. In the bottom panel, the two baselines marked with red segments are example of an equivalent pair, each with antennas separated by 2 units east and 0 unit north. Similarly,
 the baselines marked in blue and yellow are examples
@@
 In the bottom panel, the two baselines marked in red, are an example of equivalent pair.
-We denote a equivalency-class of baselines in the PAPER grid by their separations, in this case  \{1,0\}, for the
+We denote a equivalency-class of baselines in the PAPER grid by their separations, in this case  \{2,0\}, for the
 antennas are separated by 1 unit east and 0 unit north. Similarly,
 the baselines marked in blue and yellow are respectively examples
-of \{1,1\} and \{1, -1\}.
-Note \{1,0\} and \{-1, 0\} for example are the same class and should
+of \{2,1\} and \{2, -1\}.
+Note \{2,0\} and \{-2, 0\} for example are the same class and should
 not be counted twice. Antennas in purely north-south baselines
@@
 for cross-multiplication due to cross-coupling.
-On the other hand, the close North-South separation means that classes such as \{1,0\} and  \{1,1\} are expected to be near-equivalent.
+On the other hand, the close North-South separation means that classes such as \{2,0\} and  \{2,1\} are expected to be near-equivalent.
 The PAPER-64 analysis of \cite{Ali2015} used three classes of baselines, the PAPER-128
@@
 The first and third panels show the equivalent baseline pairs \{2,0\}:\{2,0\}, and second and fourth panels show \{2,0\}:\{2,1\}.
-We see that in the un-rephased case in the second panel, although the magnitude of correlations match up for the two frequencies, the phases do not.
-This means summing over frequency leads to destructive interference and signal loss.
+We see that although the amplitude of correlations match up for all time offsets, the phase would only locally match.
+The need for rephasing can thus be understood as a spatial decoherence effect.
+Summing over frequency without proper rephasing leads to destructive interference and signal loss.
 The wider the frequency profile, the more destructive the interference would be.
-In the rephased case, the phases of the nearly equivalent case match up at the peak of correlation and can be added without compromising sensitivity.
-We should thus rephase the data separately for each set of baseline pairs.
+Only after rephasing to the correct time-offset for each baseline pairs can we constructively combine the frequency channels.
 

### 2017-07-07 Rotation paper/draft_working.tex
save

@@
 Visibility-based pipelines are not fundamentally limited from using partially redundant baselines.
-In fact, most sensitivity forecasts to date do include partial redundancy.
+In fact, most sensitivity forecasts to date do include partial redundancy \citep{Pobersens, HERA, JoshAntPos}.
 Recently, \cite{wterm} proposed a visibility-based approach to extract power spectrum from partially redundant baselines in sky-tracking measurements.
@@
 
-Relative to a phase center on the sky, each baseline maps to a point in $uvw$ space. As the Earth rotates with respect to
-the source, the points trace out tracks in the $uvw$ space. 
+Radio interferometric observations are often described in the coordinates $uvw$ define as:
+\[
+(u, v, w) = \frac{\nu}{c}\boldsymbol{b}, 
+\]
+where $\boldsymbol{b}$ is the baseline vector in Cartesian coordinates, with first and second coordinates pointing East and North, respectively.
+$\nu$ is the frequency of observation, and $c$ is the speed of light.
+Relative to a phase center on the sky, each baseline maps to a point in $uvw$ space. As the Earth rotates, the points trace out tracks in the $uvw$ space. 
 We show in Fig. \ref{fig:Tracks} $uvw$ tracks of PAPER-128 over 12 sidereal hours, at 0.15GHz, relative to a phase center that passes through zenith.
@@
 We begin with the visibility as commonly defined in the literature (e.g.
-\citealt{first-paper}): 
+\citealt{TMS, first-paper}): 
 \begin{equation}
@@
 \end{equation}
-where $k_B$ is Boltzmann's constant, $\lambda$ is the mean wavelength, $\boldsymbol{b}$ is the baseline vector, $\hat{\boldsymbol{s}}$ and $\Omega$ are a direction in the
+where $k_B$ is Boltzmann's constant, $\lambda$ is the mean wavelength, $\hat{\boldsymbol{s}}$ and $\Omega$ are a direction in the
 sky and its corresponding solid angle.
-Inside the integral we have $\phi(\nu)$ the frequency bandpass profile, $A_{\nu}$ the (frequency-dependent) primary beam, and $I$ the specific intensity, which has been
+Inside the integral we have $\phi(\nu)$ the frequency bandpass profile, $A(\hat{\boldsymbol{s}},\nu)$ the (frequency-dependent) primary beam, and $I$ the specific intensity, which has been
 related to $T$, the brightness temperature in the Rayleigh-Jeans
@@
 Both beams are Stokes $I$ polarization antenna voltage beams, at frequency of $\nu=150\text{MHz}$ and normalized to $1$ at zenith.
-Notice $A$ in this paper is a ``baseline's beam'', equivalent to squares of the antenna voltage beams shown here.
 The circles centered around zenith (center of beam) here are
@@
 \begin{equation}
-\begin{aligned}V(\boldsymbol{b},\tau) & =\int d\nu V_{\nu}(\boldsymbol{b})\phi(\nu)\exp\left(-2\pi i\nu\tau\right),\\
- & =\frac{2k_{B}}{\lambda^{2}}\int d\Omega d\nu A(\hat{\boldsymbol{s}},\nu)T(\hat{\boldsymbol{s}},\nu)\exp\left[-2\pi i\nu\left(\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}+\tau\right)\right]. 
+\begin{aligned}V(\boldsymbol{b},\tau) & =\int d\nu V_{\nu}(\boldsymbol{b})\phi(\nu)\exp\left(2\pi i\nu\tau\right),\\
+ & =\frac{2k_{B}}{\lambda^{2}}\int d\Omega d\nu A(\hat{\boldsymbol{s}},\nu)T(\hat{\boldsymbol{s}},\nu)\exp\left[-2\pi i\nu\left(\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}-\tau\right)\right]. 
 \end{aligned}
@@
 \[
-\begin{aligned}V(\boldsymbol{b},\tau) & =\frac{2k_{B}}{\lambda^{2}}\int\frac{d^{3}r}{X^{2}Y}A(\boldsymbol{r})T(\boldsymbol{r})\exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\hat{\boldsymbol{r}}+\tau\right)\nu_{r}\right]\end{aligned},
+\begin{aligned}V(\boldsymbol{b},\tau) & =\frac{2k_{B}}{\lambda^{2}}\int\frac{d^{3}r}{X^{2}Y}A(\boldsymbol{r})T(\boldsymbol{r})\exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\hat{\boldsymbol{r}}-\tau\right)\nu_{r}\right],
+\end{aligned}
 \]
@@
 in the topocentric frame with the 3 dimensional rotation operator $\Gamma$:
-\small
+
 \[
-\begin{aligned}V(\boldsymbol{b'},\tau') & =\frac{2k_{B}}{\lambda^{2}}\int\frac{d^{3}r}{X^{2}Y}A(\Gamma\boldsymbol{r})T(\boldsymbol{r})\exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\Gamma\hat{\boldsymbol{r}}+\tau'\right)\nu_{r}\right]\end{aligned}.
+\begin{aligned}V_{\psi}(\boldsymbol{b'},\tau) & =\frac{2k_{B}}{\lambda^{2}}\int\frac{d^{3}r}{X^{2}Y}A(\Gamma\boldsymbol{r})T(\boldsymbol{r})\\
+& \qquad \qquad  \exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\Gamma\hat{\boldsymbol{r}}-\tau\right)\nu_{r}-i\psi_\nu\right],\end{aligned}
 \]
-\normalsize
+where we introduced a frequency-dependent phase $\psi_{\nu}$, which to first order corresponds to phasing the two visibilities to the same phase center.
+\footnote{Turns out for drift-scan measurements, the first-order interpretation is largely sufficient, but we keep the term for now to reserve the option to determine its exact form numerically.}
+
 With implicit bounds of integrals from $-\infty$ to $\infty$, we have:
@@
 \begin{equation}
-\begin{aligned} & \langle e^{i\psi_{\nu}} V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau)\rangle\\
+\begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle\\
  & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle\right)A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
@@
 \end{equation}
-where we introduced a frequency-dependent phase $\psi_{\nu}$, which to first order corresponds to phasing the two visibilities to the same phase center.
-\footnote{Turns out for drift-scan measurements, the first-order interpretation is largely sufficient, but we keep the term for now to reserve the option to determine its exact form numerically.} In transition from cosmological coordinates back to observing coordinates we have written $\hat{\boldsymbol{r}}\equiv\hat{\boldsymbol{s}}$, and
+In transition from cosmological coordinates back to observing coordinates we have written $\hat{\boldsymbol{r}}\equiv\hat{\boldsymbol{s}}$, and 
 \begin{equation}
-\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'})\equiv\exp\left[-i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\Gamma\hat{\boldsymbol{r'}}\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)+i\psi_{\nu}\right].
+\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'})\equiv\exp\left[i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\Gamma\hat{\boldsymbol{r'}}\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)+i\psi_{\nu}\right].
 \end{equation}
@@
 \begin{equation}
-\begin{aligned} & \langle e^{i\psi_{\nu}} V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau)\rangle\\
+\begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle\\
  & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)\\
- & \qquad \qquad \qquad \qquad \exp\left[-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)+i\psi_{\nu}\right].
+ & \qquad \qquad \qquad \qquad \exp\left[i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right].
  \end{aligned}
@@
 \begin{equation}
- P(k_{b,\tau}) = \left(\frac{\lambda^{2}}{2k_{B}}\right)^{2} \frac{\langle e^{i\psi_{\nu}} V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau)\rangle}{\Theta}, 
+ \hat{P}(k_{b,\tau}) \equiv \left(\frac{\lambda^{2}}{2k_{B}}\right)^{2} \frac{V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)}{\Theta}, 
  \label{eq:opp}
@@
 \[
-\Theta_{\nu} \equiv \int \frac{d\Omega}{X^{2}Y}A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu) e^{-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)}e^{i\psi_{\nu}}.
+\Theta_{\nu} \equiv \int \frac{d\Omega}{X^{2}Y}A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu) e^{i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)}e^{-i\psi_{\nu}}.
 \label{eq:Thetanu}
@@
 \[
-\Theta_{\nu} \equiv \int \frac{d\Omega}{X^{2}Y}A^{*}(\hat{\boldsymbol{s}},\nu)A(\hat{\boldsymbol{s}},\nu) e^{-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)}e^{i\psi_{\nu}}.
+\Theta_{\nu} \equiv \int \frac{d\Omega}{X^{2}Y}A^{*}(\hat{\boldsymbol{s}},\nu)A(\hat{\boldsymbol{s}},\nu) e^{i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)}e^{-i\psi_{\nu}}.
 \label{eq:Thetanu_tracking}
@@
 We refer the reader to Appendix \ref{sec:appA} for further explanation.
-In both cases, the full effects are encapsulated in the phase of $\Theta_\nu(\psi_\nu=0)$ and can be thus determined empirically.
+In both cases, the full effects are encapsulated in the phase of $\Theta_\nu$ and can be thus determined empirically without any additional computation.
 
@@
 To illustrate the effect of rephasing, we compare in Fig.
-\ref{fig:freqdiff} real parts of $\Theta_{\nu}$ of both equivalent and nearly equivalent baseline pairs for two channels: 0.16 GHz and 0.17 GHz.
+\ref{fig:freqdiff} real parts of $\Theta_{\nu}$ of both equivalent and nearly equivalent baseline pairs for two channels: 0.145 GHz and 0.155 GHz.
 The top two panels have zero rephasing, and the bottom two are linearly rephased to a differential delay of $\Delta\tau\approx15.7$ns, as we determined above.
 The first and third panels show the equivalent baseline pairs \{2,0\}:\{2,0\}, and second and fourth panels show \{2,0\}:\{2,1\}.
-We see that although the amplitude of correlations match up for all time offsets, the phase would only locally match.
-The need for rephasing can thus be understood as a spatial decoherence effect.
 Summing over frequency without proper rephasing leads to destructive interference and signal loss.
@@
 Only after rephasing to the correct time-offset for each baseline pairs can we constructively combine the frequency channels.
+
+We see from Fig.
+\ref{fig:freqdiff} that although the amplitude of correlations match up for all time offsets, the phase would only locally match.
+The need for rephasing can thus be understood as a symptom of the underlying spatial decoherence; coherence at the phase center would not extend to the entire beam pattern (Fig.
+\ref{fig:beamfringe}).
+Thus while integrating over frequencies and sky-direction in Eq.
+\eqref{eq:Theta} there are necessarily sky-directions that do not add coherently.
+This decoherence across the beam pattern cannot be removed and plays a fundamental role in determining how much sensitivity one can recover from nearly equivalent baselines.
 
@@
 First and third panels shows equivalent baselines, second and fourth show a pair of nearly equivalents.
-The top two panels have zero phase shift ($\psi_{\nu}=0$), and the bottom two are rephased to a time offset of 0.055 sidereal days.
+The top two panels have zero phase shift ($\psi_{\nu}=0$), and the bottom two are rephased to a time offset of 0.0325 sidereal days.
 The first and last panels thus show a coherently rephased series that would add constructively at the respective peaks.
@@
 \end{figure}
-
-Once rephased, the overall weight can be constructed by $\Theta=\int \Theta_{\nu}d\nu$, and thus
-\begin{equation}
-\begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle\\
- & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)|e^{i\psi_{\nu}}\\
- & \qquad \qquad \qquad \qquad \exp\left[-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)\right].\end{aligned}
-\label{eq:final_new}
-\end{equation}
 
@@
 To gain some visual intuition of the formalism in Section \ref{sec:method}, we show explicit beam fringe patterns of the two baselines and their interactions.
-For this section, we relax into the flat sky approximation and examine a single frequency of $150$MHz.
+For this section, we relax into the flat sky approximation so that $(l,m,\sqrt{1-l^2-m^2})\equiv\hat{s}$ and examine a single frequency of $150$MHz.
 
@@
 Bringing together the discussion from Section \ref{sec:method}, we present a numerical check of Eq.
-\eqref{eq:final_new}, including rephasing.
+\eqref{eq:final}, including rephasing.
 To do so, we need to compare the amplitude and phase of the integral weight $\Theta$
@@
 Shown in green for the two cases are the computed values of $\Theta$, both normalized to the peak of the equivalent case.
-The plot on the bottom is rephased with a delay of $\Delta\tau\approx17$ns as outlined in Section \ref{sec:rephs}.
+The plot on the bottom is rephased with a delay of $\Delta\tau\approx15.7$ns as outlined in Section \ref{sec:rephs}.
 Since both y-axes are normalized to the same scale, we see the equivalent baseline reach maximal redundancy with 0 time-offset, while the nearly equivalent pair reaches maximum redundancy of $\Theta\approx0.5$ at $dT\approx0.325$days.
@@
 correlation as a function of time-offset.   We do this for both the equivalent (\{2,0\}:\{2,0\}) and nearly equivalent (\{2,0\}:\{2,1\}) case.
-Since the blue and green curves overlap, we have verified that in this case Eq. \eqref{eq:final_new} is valid and the computed $\Theta$ can be used as weight to obtain the power spectrum.
+Since the blue and green curves overlap, we have verified that in this case Eq. \eqref{eq:final} is valid and the computed $\Theta$ can be used as weight to obtain the power spectrum.
 
@@
 \begin{figure*}[h!]
-\includegraphics[width=\textwidth]{sensitivity_1}
+\includegraphics[width=0.95\textwidth]{sensitivity_1}
 
@@
 \subsection{ Sensitivity \label{sec:sensitivity}}
-
-Having verified Eq. \eqref{eq:final}, we can thus predict the sensitivity contributions of a
-particular baseline pair simply by computing the integral Eq. \.
-Intuitively we expect the sensitivity to depend on both the $uv$ coverage of the baseline and the patch of sky inside the beam.
+In this section we discuss the sensitivity contributions of a variety of baseline pairs from PAPER128.
+Recall the sensitivity contributions of a particular baseline pair is given by $\Theta$ (Eq. \eqref{eq:Theta}). 
+%Intuitively we expect the sensitivity to depend on both the $uv$ coverage of the baseline and the patch of sky inside the beam.
 A larger beam like that of PAPER would tolerate larger time-offsets because more sky area can coincide in the two beams \footnote{Larger beams also imply smaller spread in $uv$ space.
 This could lead to either larger or smaller redundancy, which depends on the overlap of two such point spread functions.}
-Having computed all of the baseline pairs, we find that baseline pairs that are mirror images of each other 
-give the same amount of redundancies (peak height), with the opposite time offset, as expected from symmetry. 
-For example, \{1,0\}:\{1,1\} is mirror image of \{1,0\}:\{1,-1\} and these two baseline pairs
-give the same sensibility contribution.
-Thus we shall only show a subset of representative baseline pairs to illustrate the contributions from different classes of baseline pairs.
-For a more complete result see Fig.
-\ref{fig:pairplot}.
 
-In the top panel of Fig. \ref{fig:sensplot}
-we show the peak heights and locations for a variety of baseline combinations.
-We see that baseline pairs that have crossings at a smaller time delay
+In the middle panel of Fig. \ref{fig:sensplot}
+we show the peak heights $\Theta$ and locations $\Delta T$ for a variety of baseline combinations. Baseline pairs that are mirror images of each other 
+give the same amount of redundancies ($\Theta$), with the opposite time offset, as expected from symmetry. 
+For example, \{1,0\}:\{1,1\} is mirror image of \{1,0\}:\{1,-1\} and these two pairs of baselines
+give the same sensibility contribution. Thus we only show a subset of representative baseline pairs to illustrate the contributions. For a more complete result see Fig. \ref{fig:pairplot}.
+
+Baseline pairs that have smaller optimal time delays
 tend to have higher correlations. In other words, correlation peaks
-that are closer to zero time lag are higher. This is expected since
-a) the longer the time delay, the more the antennas have moved with respect
-to the sky and hence the less overlaps in patch of sky surveyed, b) smaller optimal
-time-offset corresponds to smaller differences in orientation and length of the 
-pair of baselines.
+that are closer to zero time offset are higher. This is expected for two reasons;
+number one is that the longer the time delay that produces redundancy, the more the sky has moved with respect
+to the beams and hence the less overlaps in patch of sky surveyed. The other reason is that smaller optimal
+time-offset corresponds to smaller differences in orientation and length of the pair of baselines and hence better redundancy.
 
@@
 baseline pairs, we have to take into account of the multiplicities of
-these baselines. By these we mean how many physical antenna pairs have the
-same length and orientation. Looking at Fig. \ref{fig:sensplot}
-we see for example \{1,0\} will have higher multiplicity than \{2,0\},
-or \{1,1\}. The latest release of PAPER-64 data uses the 128-equivalent baselines \{2,1\},
+these baseline classes, in other words the number of antenna pairs with the
+same length and orientation. From Fig. \ref{fig:sensplot}
+we see for example \{1,0\} has higher multiplicity than \{2,0\},
+or \{1,1\}. The latest release of PAPER64 data uses the 128-equivalent baselines \{2,1\},
 \{2,0\} and \{2,-1\} \citep{Ali2015}, and achieved a $2\sigma$ upper

### 2017-07-07 save/draft_working.tex
save

\documentclass[preprint2,numberedappendix,tighten,twocolappendix]{aastex6}  % USE THIS TO MAKE BIB, THEN FORMAT USING EMULATEAPJ
%\documentclass[twocolumn,apj,numberedappendix]{emulateapj}
\shorttitle{Adding Sensitivity to 21cm Inteferometric Probes of Reionization}
\shortauthors{Zhang\&Parsons}
\usepackage{float}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[figuresright]{rotating}
%\usepackage{rotating}
\usepackage{natbib}
%\usepackage{pdflscape}
%\usepackage{lscape}
\usepackage{ctable}
\citestyle{aa}
\renewcommand\[{\begin{equation}}
\renewcommand\]{\end{equation}}

\begin{document}

\title{Adding Sensitivity to 21cm Inteferometric Probes of Reionization by Optimizing Choice of Baselines}

%\maketitle
\author{
Yunfan Gerry Zhang \altaffilmark{1},
Aaron R. Parsons\altaffilmark{1}\altaffilmark{2}
}

\altaffiltext{1}{Astronomy Dept., U. California, Berkeley, CA}
\altaffiltext{2}{Radio Astronomy Lab., U. California, Berkeley, CA}
%\altaffiltext{3}{Berkeley Center for Cosmological Physics, Berkeley, CA}
%\altaffiltext{3}{Dept. of Physics and Astronomy, U. Pennsylvania, Philadelphia, PA}
%\altaffiltext{8}{School of Earth and Space Exploration, Arizona State U., Tempe, AZ}

\begin{abstract}
We introduce an accurate method to extract additional sensitivity from data of redundant radio arrays. 
Using the 64 element version of Precision Array for Probing the Epoch of Reionization (PAPER-64), we illustrate 
how our method finds optimal baselines to cross correlate, including baselines that are slightly different in length and orientation. 
We predict the sensitivity improvements of including each baseline pairs. The extremely natural and simple implementation of this method
would increase the sensitivity of PAPER-64 by up to [~50] percent. 
\end{abstract}

\section{Introduction}

The epoch of reionization, or cosmic dawn represents the last key
stage of the universe's early evolution. Study of this event stands at
the intersection of cosmology and astrophysics. Understanding this
event is important not only as a scientific goal
of its own, but also because it potentially provides crucial information
to fundamental nature of inflation, neutrino mass and phenomenology
of the first stars and galaxies alike (e.g. \cite{LiuOpticalDepth}).

Arguably the most promising observational probe of the epoch of reionization
comes from measurement of the ``spin-flip'' transition of neutral
hydrogen of characteristic wavelength 21cm \cite{Furlanetto2006181,PritchardLoeb}.
While probes such as quasar spectra rely on emission and scattering
of free electrons, and are thus limited to the lower redshifts after
the completion of reionization, the 21cm line directly probes the
abundance and distribution of neutral hydrogen, and thus would potentially
shed light on all stages of reionization. Radio interferometry efforts
to measure the 21cm power spectrum has been a top priority in recent years of astronomy.
Current generation instruments include the Precision Array for Probing
the Epoch of Reionization (PAPER) \cite{Ali2015,paper32}, Murchison
Widefield Array (MWA) \cite{MWA}, Low Frequency Array (LOFAR) \cite{LOFAR},
and next generation instruments such as the Hydrogen Epoch of Reionization
Array (HERA) (e.g. \cite{HERA,HERAconfiguration,HERABEAM1,HERADISH2})
and the Square Kilometer Array Low (SKA low) (e.g. \cite{SKA1}) are
under construction.

One of the main challenges of interferometric observations of cosmic
21cm line is the foreground contamination. Both sources within our
galaxy, and to a lesser distant sources outside of our galaxy emit
radio contaminations (via for example synchrotron processes) up to five orders of magnitude stronger than the
reionization signal. There are two common methods to deal with the
foreground contamination. In the first, sometimes called the foreground
removal technique, individual sources are identified and removed in the image domain.
The other technique, commonly referred to as 'foreground avoidance',
makes use of the fact that most common foreground contaminations have a smooth
spectrum, and thus is constrained to a Fourier domain ``foreground
wedge'' \cite{wedge1,wedge2}. Thus contamination can be avoided
by restricting our observation to ``outside the wedge'', where the data is currently noise dominated. The main
challenge of using the avoidance technique is thus the sensitivity, or
the signal to noise ratio. This is the motivation for the design of
the maximum redundancy arrays such as PAPER and HERA. While traditional
imaging arrays focus on Fourier mode uv coverage, the redundant arrays
have many tightly packed antennas at equal spacing. Since baselines
of the same length and orientation measure the same Fourier modes
on the sky, the maximum redundancy arrays is able to increase the
signal to noise ratio by averaging over measurements of the same baseline.
\cite{Ali2015} provides the newest upper limit to the power spectrum
measurements with the PAPER-64 data. From some models we roughly expect
that the sensitivity required for detection is a factor of 10 away.
This paper explores a new technique to add sensitivity to the analysis
of redundant arrays, by making use of the earth's rotation. This technique
can be applied to redundant arrays such as PAPER and HERA to increase
the sensitivity by up to a factor of [1.5].

The earth's rotation causes the baselines to pick up different modes of the sky with time. 
This effect is used extensively to improve uv coverage in imaging with minimum redundancy arrays. In
a maximally redundant array, baselines that are slightly different
rotate into each other at a time delay. Here we present a method to
improve the sensitivity of the maximum redundancy arrays by considering
the uv redundancy of different baselines at a time lag.

Baselines of the same length and orientation are traditionally called
``redundant baselines'', because they measure the same Fourier mode
in the sky. In order to eliminate confusion and ambiguity, we shall
introduce slightly different terminology. We'll call baselines that
are the same length and orientation ``equivalent baselines'', inspired
by the mathematical notion of equivalency classes. Two equivalent
baselines will be redundant with each other simultaneously at all
times. Some non-equivalent baselines can also be redundant if their
respected time series is shifted with respect to one another by a
certain delay. In other words, one baseline can be ``rotated into''
another. We call these ``near-equivalent baselines''. It is our
goal to a) identify the near-equivalent baselines that give good
redundancy, b) to find the optimal time offset for a given pair of baselines, and
c) to quantify the sensitivity improvement associated with cross multiplying
such a pair of near-equivalent baselines. In section
2 we explain the theoretical basis for this cross multiplication.
In section 3 we present details of the numerical implementation of
this technique. In section 4 we discuss the expected sensitivity improvement
with this method and with section 5 we conclude.

\section{Method}

\subsection{Rough Idea}
\begin{figure}[H]
\includegraphics[width=\linewidth]{64layout}

\caption{The PAPER-64 layout. Each blue dot corresponds to the location of
an antenna. Top panel shows the antenna positions drawn to scale,
bottom panel show the antenna labels and distances.\label{fig:The-PAPER-64}
The numbering of the antennas the bottom panel are original labels
during instrument assembly and does not bear significant
meaning. Baselines corresponding to the same separations are called
equivalent. In the bottom panel, the two baselines labeled with the
red arrow are equivalent, with separation denoted by sep1,0, for the
antennas are separated by 1 unit east and 0 unit north. Similarly,
the baselines labeled with the yellow and green arrows are examples
of sep1,1 and sep-1,1, respectively. Note sep1,0 and sep-1,0 for example
are the same baselines and should not be counted twice. }
\end{figure}

We shall use the 64-antenna PAPER array to demonstrate. 
The PAPER-64 array is located in the Karoo desert in South Africa (30:43:17.5
S, 21:25:41.8 E). The layout pattern with antenna labels are show
in Fig. \ref{fig:The-PAPER-64}. We see the antenna spacing in North-South
directions are comparatively close (4m), so that baselines such as
0\_44 and 0\_7 are very close to equivalent. In the bottom panel,
the two baselines labeled with the red arrow are equivalent, with
separation denoted by sep1,0, for the antennas are separated by 1
unit east and 0 unit north. Similarly, the baselines labeled with
the yellow and green arrows are examples of sep1,1 and sep-1,1, respectively.
Note sep1,0 and sep-1,0 for example are the same baselines and should
not be counted twice. Turns out antennas in purely north-south baselines
are close enough to induce cross talks, and hence are not suitable
for use. The original PAPER-64 analysis thus used three classes of baselines,
sep1,0, sep1,1 and sep-1,1 \cite{Ali2015}. There each of these classes
of baselines are cross multiplied to itself. We shall see that in addition these
baseline classes can be cross multiplied with a time offset.

Given a point source on the sky, each baseline maps the
source to a point in uv plane. As the earth rotates with respect to
the source, the point traces out tracks in the uv-plane. 
The uv tracks of PAPER-64 over 7.2 hours, at 0.15GHz, for a source that passes through zenith, are shown in Fig. \ref{fig:Tracks-of-PAPER-64}.

\begin{figure}[H]
\includegraphics[width=\linewidth]{TRACKSpt3days}
\caption{Tracks of PAPER-64 for a hypothetical source that passes through zenith.
These tracks are traced out over 0.3 sidereal days, or roughly 7.2
hours. Color represents different baselines. Frequency is $\nu=0.15\text{GHz}$.
As the earth rotates, tracks are traced out counterclockwise. \label{fig:Tracks-of-PAPER-64}}
\end{figure}

Roughly speaking, we can identify
redundancy of near-equivalent baselines as crossings
of the uv tracks. Although this is a valid quick method to identify some baselines to cross-multiply, we find
that it is not accurate enough for time offset determination. The reason is, for wide beam antennas such as PAPER-64, the sources are not point-like. 
In other words, when tracks of
one point of the sky are crossing in uv-plane, those corresponding to another is not. To accurately
constrain redundancy we must take into account of finite beam width. A more rigorous treatment is thus required.  
We show the beam of PAPER-64 in Fig.\ref{fig:Beam-of-PAPER-64.} for reference:

\begin{figure}[H]
\includegraphics[width=\linewidth]{BEAM}

\caption{Beam of PAPER-64. Shown here is the values of $A$ (Eq. \eqref{eq:Vis1})
for frequency of $\nu=150\text{MHz}$. Notice $A$ is a ``baseline's
beam''. Since all antennas in PAPER-64 have the same beam, $A$ here
is just the square of an antenna beam, and is normalized to 1 at the
center. The circles centered around zenith (center of beam) here are
spaced 10 degrees apart. Such a wide beam limits the accuracy of the uv track-crossing as a method of redundancy search. \label{fig:Beam-of-PAPER-64.}}
\end{figure}

\subsection{Formalism}
Below we shall derive theoretical expectations of cross multiplications
of two near-equivalent baselines. More precisely we shall relate
the product of visibilities of two different baselines to the power spectrum.

We take the visibility as commonly defined in the literature (e.g.
\cite{first-paper}): 
\begin{equation}
\begin{aligned}V_{\nu}(b) & =\int d\Omega A_{\nu}(\hat{s})\phi(\nu)I_{\nu}(\hat{s})\exp\left[-2\pi i\frac{\nu}{c}b\cdot\hat{s}\right],\\
 & \approx\frac{2k_{B}}{\lambda^{2}}\int d\Omega A_{\nu}(\hat{s})\phi(\nu)T(\hat{s})\exp\left[-2\pi i\frac{\nu}{c}b\cdot\hat{s}\right],
\end{aligned}
\label{eq:Vis1}
\end{equation}
Here $\lambda$ is a mean wavelength, $\boldsymbol{b}$ is the baseline
length, $\hat{\boldsymbol{s}}$ and $\Omega$ are a direction in the
sky and its corresponding solid angle. $A_{\nu}$ is the (frequency
dependent) beam, and $I$ is the specific intensity, which has been
related to $T$, the brightness temperature in the Rayleigh-Jeans
limit. $\phi(\nu)$ is the frequency band profile. For PAPER-64, for example,
the bands studied are roughly a top-hat from 144MHz to 154 MHz.

We define the delay transformed visibility \cite{delay-transform}:
\small
\begin{equation}
\begin{aligned}V(b,\tau) & =\int d\nu V_{\nu}(b)\phi(\nu)\exp\left[-2\pi i\nu\tau\right],\\
 & =\frac{2k_{B}}{\lambda^{2}}\int d\Omega d\nu A(\hat{s},\nu)\phi(\nu)T(\hat{s},\nu)\exp\left[-2\pi i\nu\left(\frac{b\cdot\hat{s}}{c}+\tau\right)\right]
\end{aligned}
.\label{eq:Vb1}
\end{equation}
\normalsize

Eq. \eqref{eq:Vb1} expresses the delay-transformed visibility as
an integral over observation coordinates $\hat{s}$ and $\nu$. Ultimately,
we would like to relate the data, collected with coordinates $\hat{s}$
and $\nu$, to the power spectrum, written with cosmological coordinates
$\boldsymbol{r}$ and $\boldsymbol{k}$. We start by noticing that
{[}Gerry: change all r and k into bold{]} 
\[
\begin{aligned}r & =\frac{c}{H_{0}}\int_{0}^{z}\frac{dz'}{E(z')},\\
 & \approx\frac{c}{H_{0}}\int_{0}^{z_{0}}\frac{dz'}{E(z')}-\frac{c(1+z)^{2}}{\nu_{21}H_{0}E(z)}\left(\nu-\nu_{0}\right),\\
 & \equiv D_{c}-Y\Delta\nu,
\end{aligned}
\]
where $\nu_{21}=1420MHz$ is the 21cm transition rest frequency, $\nu_{0}$
a reference central frequency with corresponding redshift $z_{0}$,
and 
\[
E(z)=\sqrt{\Omega_{m}(1+z)^{3}+\Omega_{\Lambda}}.
\]
Inverting for $\nu$:
\begin{equation}
\nu=\frac{D_{c}-r}{Y}+\nu_{21}.\label{eq:nur}
\end{equation}
Thus $d\nu=-dr/Y$ and $d^{3}r=-X^{2}Yd\Omega d\nu$.

We can thus rewrite the delayed transformed visibility as 
\small
\[
\boxed{\begin{aligned}V(b,\tau) & =\frac{1}{X^{2}Y}\int_{H}d^{3}rA(\vec{r})\phi(r)I(\vec{r})\exp\left[-2\pi i\left(\frac{b}{c}\cdot\hat{r}+\tau\right)\nu_{r}\right]\end{aligned}
}.
\]
\normalsize
Here $(r_{x},r_{y},r_{z})=(X\hat{s}_{x},X\hat{s}_{y},Y\nu)$, and
$(Xk_{x},Xk_{y},Yk_{z})=\frac{2\pi}{c}(b_{x},b_{y},\eta)$ relate
the cosmological coordinates $r$ and $k$ to the measured coordinates.
We have written $\nu_{r}$ to remind us that $\nu$ and $r$ are related
by Eq. \eqref{eq:nur}. The beam reception pattern $A$ is dimensionless,
normalized to 1 at its peak (zenith), and we assume it to be the same
for all baselines.

With a time offset, the beam pattern has moved relative to the sky.
Here we choose to fix the sky, and denote the rotated coordinates
of the beam pattern with the operator $\Gamma$. With implicit bounds
of integrals from $-\infty$ to $\infty$, we have:
\begin{widetext}
\begin{equation}
\begin{aligned} & \langle V^{*}(b,\tau)V(b',\tau')\rangle\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\langle T^{*}(r)T(r')\rangle\right)A^{*}(r)A(\Gamma r')\Phi_{b,\tau}(r,\Gamma r'),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\kappa)T(\kappa')\rangle e^{-i(\kappa r-\kappa'r')}\right)A^{*}(r)A(\Gamma r')\Phi_{b,\tau}(r,\Gamma r'),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\kappa(r-r')}\right)A^{*}(r)A(\Gamma r')\Phi_{b,\tau}(r,\Gamma r'),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\xi(r-r')A^{*}(r)A(\Gamma r')\Phi_{b,\tau}(r,\Gamma r'),\\
 & \approx\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}rd^{3}r'\delta_{D}^{(3)}(r-r')A^{*}(r)A(\Gamma r')\Phi_{b,\tau}(r,\Gamma r'),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}r|A^{*}(r)A(\Gamma r)||\phi(\nu_{r})|^{2}\exp\left[-i2\pi\nu_{r}\left(\hat{r}\cdot\frac{b}{c}-\hat{\Gamma r}\cdot\frac{b'}{c}\right)\right],\\
 & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{s},\nu)A(\hat{\Gamma s},\nu)||\phi(\nu)|^{2}\exp\left[-i2\pi\nu\left(\hat{s}\cdot\frac{b}{c}-\hat{\Gamma s}\cdot\frac{b'}{c}\right)\right],
\end{aligned}
\label{eq:main}
\end{equation}

where in transition from cosmological coordinates back to observing coordinates we have written $\hat{r}\equiv\hat{s}$, and 
\begin{equation}
\Phi_{b,\tau}(r,\Gamma r')=|{\phi^{*}}(\ensuremath{\nu_{r}})\phi(\nu_{r'})|\exp\left[-i\frac{2\pi}{c}\left(b\cdot\nu_{r}\hat{r}-b'\cdot\nu_{r'}\Gamma\hat{r'}\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)\right].
\end{equation}
\end{widetext}
The second to third line of Eq.(\ref{eq:main}) follows from assumption of Gaussian random
sky, and the third to fourth line follows from the assumption that
the 3D power spectrum varies negligibly over the region of interest
so that $\hat{P_{21}}(k+k_{2})\approx\hat{P_{21}}(k)$. Since $\Gamma$
is a sky rotation, it doesn't affect $\nu$, hence we have taken $\nu_{r}$
outside the parenthesis. Notice that the phase factor $\exp\left[-i2\pi\eta\left(\nu-\nu'\right)\right]$
drops out in the end.

Finally, since the beam pattern and bandwidth are given in $\hat{s}$
and $\nu$, we convert the integral back to these coordinates to get
the general relation between the delay-transformed visibilities and
the power spectrum:

\begin{equation}
\boxed{\begin{aligned} & \langle V^{*}(b,\tau)V(b',\tau')\rangle\\
 & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{s},\nu)A(\hat{\Gamma s},\nu)||\phi(\nu)|^{2}\\
 & \qquad \qquad \qquad \qquad \exp\left[-i2\pi\nu\left(\hat{s}\cdot\frac{b}{c}-\hat{\Gamma s}\cdot\frac{b'}{c}\right)\right].\end{aligned}}
\label{eq:final}
\end{equation}
So roughly speaking the cross multiplications of visibilities at a time delay
in uv-space is proportional to the power spectrum times the Fourier
transform of the cross multiplied beam pattern. We can then in principle
combine information from different baseline pairs if we correct for
the phase and normalization. As a check, when applied to equivalent baselines,
$b=b'$, $\hat{s}=\hat{\Gamma s}$, and Eq.(\ref{eq:final}) reduces to Eq.(B9) of \cite{paper32}.

If this result is robust, then we can achieve all our goals stated in the introduction, i.e. to identify 
candidate baseline pairs, to find the optimal time offset, and to quantify the sensibility, by computing the integral
in Eq.(\ref{eq:final}) for various time offsets. Thus we shall first test its robustness numerically.

\section{Analysis}
\subsection{Numerical Test \label{sec:Techniquet}}

As the earth rotates, a point source traces out a track in UV plane
for a given baseline. Shown in Fig. \ref{fig:Tracks-of-PAPER-64}
are the rotation tracks for a source that passes through zenith for
the baselines of PAPER-64. As we mentioned earlier, crossings of these tracks are approximate, 
and only approximate points of maximum redundancy of near-equivalent baselines, because
of the wide field of view of the PAPER beam, shown in Fig.\ref{fig:Beam-of-PAPER-64.}).

In practice, we first implement a search algorithm that grids the
uv plane and search for intersections these tracks, then fine tune
the time delay by computing a) cross correlated visibilities from
simulated random sky, and b) theoretical proportion factor from Eq.
\eqref{eq:main} for a range of time delays. It is also worth pointing
out that the near-redundant pairs is not only useful at a particular
time corresponding to the crossing. The search algorithm uses a fiducial
point source in the sky to find the offset, and all time-stamps separated
by this offset corresponds to UV-track crossing of a point in the
sky, and thus all data of the two baselines can be utilized.

At the optimal time separation, the integral in Eq. \eqref{eq:final}
is maximized. We expect the two beams to have to same fringe pattern
(frequency and phase). Due to the time delay, however, the beam center
would be slightly shifted with respect to each other. This we show
in Fig. {[}insert{]}. The left and middle panels show the beam fringe
patterns for baselines sep0,1 and sep1,1, delayed by 0.0325 days,
and the right panel shows their cross product. The fringe pattern
indeed cancels out as we expect.

\begin{widetext}
\begin{figure}[H]
\includegraphics[scale=0.5]{fringe_res}

\caption{Beam fringe pattern of sep1.0(left), sep1,1 at a time delay (middle),
and their conjugate product. Frequency of $\nu=0.15\text{GHz}$ is
chosen. \label{fig:Beam-fringe-pattern}}
\end{figure}
\end{widetext}

Below we go into a little more details on the numerical techniques
we used to obtain the results.

\subsection{Track crossing search {[}skip all together?{]}}

The crossing algorithm sets up grids on the UV-plane, creates a point
source that passes through zenith, and searches for intersections
of the UV tracks. The search compares each track to each other track
to locate the closest approaches. To optimize performance, the algorithm
first sets up coarse grids, locates the cells that contain data points
from more than one tracks, then fine-grains the cell to determine
the points of closest approach. Fig {[}insert{]} illustrates this
procedure. After all the crossings have been identified, we rank the
crossings by computing the proportion factor in Eq. \eqref{eq:main}.
This algorithm is mainly used to determine which baselines to cross
correlate, and give a rough estimate of the time offset. We indeed
find that baselines sep0,1 and sep1,1 give the closest approach with
a time offset of roughly 0.03 sidereal days. In reality, a wide beam
samples a patch of sky, not just a point source. And thus the real
UV representation of a baseline is the track convolved with an appropriate
point spread function. Accounting for these factors accurately in
this context is difficult. Thus for accurate determination of the
time offset, we resort to two other methods.

\subsection{Gaussian sky simulation \label{sub:Gaussian-sky-simulation}}

Having found the pairs of baselines to cross-correlate, and a rough
estimate of the time offset, we make a more accurate determination
of the time offset. Here we take N random realizations of the sky
on a healpix map\footnote{We use functionalities in the python package AIPY for healpix mapping
as well as coordinate transforms. }. For each realization, each pixel is given a Gaussian random value
of brightness temperature. We then rotate the baseline positions with
the appropriate rotation matrix, recording the visibilities, for each
baseline, as given in Eq. 1 \footnote{As a caveat, there are two obvious ways to achieve the rotation. One
can either fix the sky and rotate the baselines, or the other way
around. We found however, that we must not physically rotate the sky
map, for the numerical roundoffs due to finite resolutions of the
map turns out to be significant. Thus we let the sky, represented
by the healpix map, be fixed, and rotate the baselines. }. The resulting visibilities for the two baselines are then convolved
via the Fourier convolution theorem, to obtain values of the cross
correlation as a function of time-offset. The peak of the curve then
corresponds to maximum redundancy. The accuracy of this result is
limited by (simulated) cosmic variance and finite spacial resolution,
and hence can be beat down by averaging over a large number of universes.
A numerical estimate of the error of the peak height with $N=1$ is
$\lesssim20\%$, and thus with $N=2500,$ we achieve an error of peak
height $<0.5\%$. An example is shown in Fig. {[}insert{]}

\subsection{Theoretical determination \label{sub:Theoretical-determination}}

To test our understanding of the time-offset, and to verify our derivation
in Eq. \eqref{eq:main} makes sense, we must check the curve obtained
in Fig. {[}insert{]} against one from the multiplicative factor in
Eq. \eqref{eq:main}. In particular, we would really like to compare
the peak location as well as height. For the latter, the most useful
information comes in its comparison with that of an equivalent baseline.
Thus in Fig. {[}insert{]} below we show on the left comparison of
the convolution of an equivalent baseline, normalized to 1 at the
peak, whose location is at 0 offset as expected. On the right we show
comparison of the baseline pair found in the grid search. The peaks
are normalized with the same factor as the peak height on the left,
i.e. the two plots have thus the same scale on the vertical axis.
The height of the plot on the right thus quantifies the added sensitivity.

\begin{figure}[H]
\includegraphics[width=\linewidth]{redun_agree}

\includegraphics[width=\linewidth]{simucorroff}

\caption{The PAPER-64 layout. Top panel shows the antenna positions drawn to
scale, bottom panel show the antenna labels and distances. }
\end{figure}

\subsection{Expected Sensitivity Contributions}

In section \eqref{sub:Gaussian-sky-simulation} and \eqref{sub:Theoretical-determination}
we showed the delays and amount of overlap for the baselines with
separation sep10 vs sep11. In the top panel of Fig. \eqref{fig:sensplot} below
we show the peak heights and locations for a variety of baseline combinations.
We see that baseline pairs that have crossings at a smaller time delay
tend to have higher correlations. In other words, correlation peaks
that are closer to zero time lag are higher. This is expected since
a) the longer the time delay, the more the antennas have moved with respect
to the sky and hence the less overlaps in patch of sky surveyed, b) smaller optimal
time-offset corresponds to smaller differences in orientation, and hence in length of
baselines in PAPER-64.

\begin{figure}[H]
\includegraphics[width=\linewidth]{sensitivity}
\label{fig:sensplot}
\caption{Relative sensitivity contributions of selected baseline combinations in PAPER-64. In the legend m,n:p,q denotes cross
multiplying PAPER-64 baselines of east-west, north-south separations (m,n) and (p,q) respectively. The top
panel shows the peak height (degree of correlation) of each baseline
combination, while the bottom panel multiplies the heights by the
corresponding multiplicities as in Eq. \eqref{eq:sensul}, and in some cases an extra factor of 2 (explained in text). 
In weighting by the multiplicity (bottom), we have chosen to fix the sensitivity 
contribution of 1,0:1,0 to unity. }

\end{figure}

To determine that actual relative contribution to sensitivity of these
baseline pairs, we have to take into account of the multiplicities of
these baselines, or the size of the equivalency classes in a mathematical
language. By these we mean how many physical antenna pairs have the
same length and orientation. Looking at Fig. \ref{fig:The-PAPER-64}
we see for example sep1,0 will have higher redundancy than sep2,0,
or sep1,1. The latest release of PAPER-64 data uses baselines sep-1,1,
sep0,1 and sep1,1 \cite{Ali2015}, and achieved a $2\sigma$ upper
limit of $22.4mK^{2}$. There, the three sets of equivalent baselines
are only cross multiplied by itself. Assuming that each baseline delivers
the same quality of data (i.e. they have the same height of correlation
peaks, which is in our normalization equal to unity), the relative
contribution to sensitivity can be estimated.

First we can average of the visibilities of the baselines of the same
separations. Since PAPER-64 has 8 by 8 antenna configuration, there
are $(8-|m|)\times(8-|n|)$ copies of the baseline sepm,n. This means
that if we add visibility measurements of all these equivalent baselines,
we get a factor of $\sqrt{(8-|m|)\times(8-|n|)}$ reduction in noise
level. The sensitivity contribution of sepm,n, being the signal to
noise ratio of the power spectrum, is thus proportional to $(8-|m|)\times(8-|n|)$.
Next we can average over the power-spectrum taken from baselines of
the different separations. Thus the total cumulative sensitivity is
given by 
\begin{equation}
\frac{P_{S}}{P_{N}}\propto\sqrt{\sum_{(n,m)}\left((8-|m|)\times(8-|n|)\right)^{2}},\label{eq:sensequiv}
\end{equation}
where the sum is only over pairs of m and n. If we now include power
spectrum from cross-multiplications of near-equivalent baselines of
types sepm,n and sepm',n', we get a general cumulative sensitivity:
\small
\begin{equation}
\frac{P_{S}}{P_{N}}\propto\sqrt{\sum_{(n,m,n',m')}\Omega_{mnm'n'}\left((8-|m|)(8-|n|)(8-|m'|)(8-|n'|)\right)},\label{eq:sensul}
\end{equation}
\normalsize
where $\Omega$ is the correlation peak height show in top panel of Fig. \eqref{fig:sensplot},
and the sum is again only over tuples of m,n,m',n'. This form now
includes the equivalent baselines in Eq. \eqref{eq:sensequiv}, with
$\Omega_{m=m',n=n'}=1$. 
Shown in the bottom panel of Fig. \eqref{fig:sensplot} is the peak heights weighted
by this multiplicity factor (before taking the square root), with one difference. The difference is, 
since we have "folded over" the negative time delays, we have combined baseline pairs that are identical modulus parity, 
or in other words are mirror images of each other, by summing over them to get an extra factor of two. For example, 
instead of plotting 1,1:1,1 and 1,-1:1,-1 separately, we plot 1,1:1,1 with twice the multiplicity. Similarly, baseline pairs such as 
1,0:1,1 also get the factor of 2 because they are identical to 1,0:1,-1. Baseline pairs such as 1,0:1,0, or 1,1:1,-1 do not get
the factor of 2 because their mirror images are themselves.

\section{Conclusion}

\bibliographystyle{apj}
%\nocite{*}
\bibliography{draft_working}

\end{document}

### 2017-07-08 Rotation paper/draft_working.tex
save

@@
  & \approx\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}rd^{3}r'\delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
- & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}rA^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r})\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\Gamma \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)+i\psi_{\nu}\right],\\
- & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma \hat{\boldsymbol{s}},\nu)\exp\left[-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)+i\psi_{\nu}\right],
+ & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}rA^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r})\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\Gamma \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right],\\
+ & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma \hat{\boldsymbol{s}},\nu)\exp\left[-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right],
 \end{aligned}
@@
 \begin{equation}
-\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'})\equiv\exp\left[i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\Gamma\hat{\boldsymbol{r'}}\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)+i\psi_{\nu}\right].
+\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'})\equiv\exp\left[i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\Gamma\hat{\boldsymbol{r'}}\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)-i\psi_{\nu}\right].
 \end{equation}
@@
 To determine that actual relative contribution to sensitivity of these
-baseline pairs, we have to take into account of the multiplicities of
+baseline pairs, we have to account for the multiplicities of
 these baseline classes, in other words the number of antenna pairs with the
-same length and orientation. From Fig. \ref{fig:sensplot}
+same length and orientation.
+We would like to estimate an effective weight $\widetilde{\Theta}$ that accounts for both the peak height $\Theta$ and the multiplicities of the baseline classes.
+From Fig.
+\ref{fig:sensplot}
 we see for example \{1,0\} has higher multiplicity than \{2,0\},
-or \{1,1\}. The latest release of PAPER64 data uses the 128-equivalent baselines \{2,1\},
-\{2,0\} and \{2,-1\} \citep{Ali2015}, and achieved a $2\sigma$ upper
-limit of $(22.4\text{mK})^{2}$. There, the three sets of equivalent baselines
-are only cross multiplied by itself. Assuming that each baseline delivers
-the same quality of data (meaning they have the same height of correlation
-peaks, which is in our normalization equal to unity), the relative
-contribution to sensitivity can be estimated.
+or \{1,1\}. The latest release of PAPER64 data uses the PAPER128-equivalent baselines \{2,1\},
+\{2,0\} and \{2,-1\} \citep{Ali2015}. There, the three sets of equivalent baseline classes
+are only cross multiplied with itself. Assuming that each equivalent baseline delivers
+the same $\Theta$, the relative contribution to sensitivity can be estimated.
 
-First we can average of the visibilities of the equivalent baselines. Since the core of PAPER-128 has 16 by 7 antenna configuration, there
+First we can average the visibilities of the equivalent baselines. Since the core of PAPER-128 has 16 by 7 antenna configuration, there
 are $M\equiv(16-|m|)\times(7-|n|)$ copies of the baseline class $\{m,n\}$. This means
-that if we add visibility measurements of all these equivalent baselines,
+that if we add visibility measurements of all the equivalent baselines,
 we get a factor of $\sqrt{M}$ reduction in noise
 level $\sigma_N$ of the visibility.
-The sensitivity contribution of $\{m,n\}$, cross multiplied with $\{m',n'\}$  thus roughly speaking scales as $\sqrt{\left((16-|m|)(7-|n|)(16-|m'|)(7-|n'|)\right)}=\sqrt{MM'}$.
+The sensitivity contribution of $\{m,n\}$, cross multiplied with $\{m',n'\}$  thus roughly scales as $\sqrt{\left((16-|m|)(7-|n|)(16-|m'|)(7-|n'|)\right)}=\sqrt{MM'}$.
 For cross-multiplications of nearly equivalent baselines of
@@
 
-Shown in the bottom panel of Fig. \ref{fig:sensplot} is the peak heights weighted
+Shown in the bottom panel of Fig. \ref{fig:sensplot} are the peak heights weighted
 by the multiplicity factor.
@@
 
-Having defined the modified weight $\widetilde{\Theta}$, we can estimate the power spectrum by inverse covariance weighting:
+With $\widetilde{\Theta}$, we can estimate the power spectrum by inverse covariance weighting:
 \begin{equation}
@@
 For a given $\rho_0$ Eq.
-\eqref{eq:tildereal} thus quantifies the relative sensitivity contribution of a baseline pair $bb'$.
-Assuming a reionization signal of $\Delta_{21cm}^2\sim 30mK^2$, observation at 150MHz ($z=8.5$), 120 days of observation with PAPER antennas, we have roughly
+\eqref{eq:tildereal} quantifies the relative sensitivity contribution of a baseline pair $bb'$.
+Assuming a reionization signal of $\Delta_{21cm}^2\sim 30mK^2$, observation centered at 150MHz ($z=8.5$), and 120 days of integration with PAPER antennas, we have roughly
 (See Eq.(20) in \cite{first-paper})
@@
 \end{equation}
-where $L$ is the baseline length.
-We only need a single characteristic $L$ even in the nearly equivalent case because only baselines of nearly equal length would have high redundancy.As expected, baseline-pairs that have smaller $\widetilde{\Theta}$ contribute less to the sensitivity.
+where $L$ is the average baseline length between the pair. As expected, baseline-pairs that have smaller $\widetilde{\Theta}$ contribute less to the sensitivity.
 
@@
 The larger arrays are designed with a ``gap'' dividing the antennas into three different groups.
-The gaps are designed so as to improve $uv$ coverage and ease calibration without compromising sensitivity, but also produces many more nearly equivalent baselines than the versions without the gap.
+The gaps are designed so as to improve $uv$ coverage and ease calibration without compromising sensitivity, but also produces many more nearly equivalent baselines than a pure hexagonal layout.
 The motivations behind the designs are explained in \cite{HERAconfiguration}.
-Compared to PAPER128, the hexagonal pattern of HERA lack short baselines that are oriented close to each other, and the smaller beam (Fig.
+Compared to PAPER128, the hexagonal pattern of HERA lack short baselines oriented close to each other, and the smaller beam (Fig.
 \ref{fig:Beam}) means that we expect to see only longer nearly equivalent baselines.
-The lower multiplicities per class of baselines is made up with the larger number of classes of baseline-pairs, especially given the gap in the larger versions.
+The lower multiplicities per class of baselines is compensated by the larger number of classes of baseline-pairs, especially given the gap in the larger versions.
 
@@
 The dashed lines represent the values when only the equivalent baseline-pairs are used.
-We see as expected that in all cases using the nearly equivalent baselines lead to more and more significant improvements with lower $\widetilde{\Theta}_{min}$, or in other words when worse baseline pairs are used.
+We see as expected that in all cases using the nearly equivalent baselines lead to more and more significant improvements with lower $\widetilde{\Theta}_{min}$, or in other words when more baseline pairs are used.
 The small HERA37, with no gap (like in HERA350) or short nearly equivalent baselines (like in PAPER 128), will not benefit much from the nearly equivalent baselines.
 The maximum benefits for other cases are expected to be around $20\%$ to $60\%$.
-PAPER128 is designed with highly redundant nearly equivalent baselines, and thus  these baselines start contributing at higher $\widetilde{\Theta}_{min}$, but the gapped HERA configurations will benefit even more from nearly equivalent baselines at low $\widetilde{\Theta}_{min}$ due to there being more classes of such pairs.
-Note that here we normalized $\rho$ such that the contribution of the top equivalent baseline pair, such as (sep0,1:0,1 in the PAPER128 case) are 1.
+PAPER128 is designed with highly redundant nearly equivalent baselines, and thus  these baselines start contributing at higher $\widetilde{\Theta}_{min}$, but the gapped HERA configurations will benefit even more from nearly equivalent baselines at low $\widetilde{\Theta}_{min}$ due to the presence of more classes of such pairs.
+Note that here we normalized $\rho$ such that the contribution of the top equivalent baseline pair, such as are 1.
 This plot therefore does not compare the absolute sensitivity across the different arrays.
@@
 In this appendix we give a brief derivation of the effective weight quoted in \ref{sec:sensitivity}.
-We combine the different power spectrum measurements by inverse variance weighting \footnote{In practice techniques such as bootstrapping is often used, see for example \citep{Ali2015}}.
-We shall separate the visibility and power spectrum into signal and noise contributions:
+We combine the power spectrum measurements from distinct baseline classese by inverse variance weighting.
+We separate the visibility and power spectrum into signal and noise contributions:
 \begin{equation}
@@
 \end{equation}
-We shall denote the noise variance of power spectrum and visibility
+We write the noise variance of power spectrum and visibility as:
 \begin{equation}
@@
 \end{equation}
-One may notice that we have used a single covariance for the complex quantity visibility.
-It's simple to show that the same result holds if we use a separate real and imaginary components, as long as they are independent of each other.
+One may notice that we have used a single covariance for the complex visibility.
+It is straightforward to show that the same result holds if a separate real and imaginary components are used, as long as they are independent of each other.
 In fact, for simplicity and without loss of generality we shall treat the visibility as a real quantity in the rest of this derivation.
@@
 
-Then the variance of $P$ constructed with visibilities $V_1$ and $V_2$ from two baseline classes can be estimated\footnote{We assume all noise terms to be independent for simplicity, in practice the correlation of different measurements ifrom equivalent baselines are alleviated by grouping the baselines in the class and the days of observation, as in \cite{Ali2015}}:
+The variance of $P$ constructed with visibilities $V_1$ and $V_2$ from two baseline classes can be estimated\footnote{We assume all noise terms to be independent for simplicity, in practice the correlation of different measurements ifrom equivalent baselines are alleviated by grouping the baselines in the class and the days of observation, as in \cite{Ali2015}}:
 \begin{equation}
@@
 \end{equation}
-
-where in the second last line we have substituted visibility noise variance. In the final line we used Wick's theorem and the fact that the signal from two visibilities are equal.
+where in the second last line we have substituted in the visibility noise variance. In the final line we used Wick's theorem and the fact that the signal from the two visibilities are equal.
 
@@
 \end{equation}
-where $\sigma_0$ is some single-baseline noise level. Letting $\rho_0=V_S^2/\sigma_0^2$ be the signal to noise ratio for a single baseline, we can write
+where $\sigma_0$ is the single-baseline noise level. Letting $\rho_0=V_S^2/\sigma_0^2$ be the signal to noise ratio for a single baseline, we can write
 
@@
 \sigma_P^2 & \propto  \frac{\sigma_0^4}{\Theta^2} \left[ \rho_0 \left(\frac{1}{M_1}+\frac{1}{M_2} \right) + \frac{1}{M_1 M_2}\right], \\
-&\propto \frac{1}{\widetilde{\Theta}_{12}^2},
+&\propto \frac{1}{\widetilde{\Theta}_{12}^2}.
 \end{aligned}
 \end{equation}
-where we have defined a slightly modified version of the effective weight (compare with Eq. \eqref{eq:sensul}):
+Thus we have defined a slightly modified version of the effective weight in Eq.  \eqref{eq:tildereal}:
 \begin{equation}

### 2017-07-11 Rotation paper/draft_working.tex
saving rotation figures

@@
 Relative to a phase center on the sky, each baseline maps to a point in $uvw$ space. As the Earth rotates, the points trace out tracks in the $uvw$ space. 
-We show in Fig. \ref{fig:Tracks} $uvw$ tracks of PAPER-128 over 12 sidereal hours, at 0.15GHz, relative to a phase center that passes through zenith.
+We show in Fig.
+\ref{fig:Tracks} $uvw$ tracks of the 3 PAPER-128 baselines colored in Fig.
+\ref{fig:AntPos}, projected onto $uv$ and $vw$ planes.
+The tracks are traced over 12 sidereal hours, at 0.15GHz, relative to a phase center that passes through zenith.
 
-\begin{figure}[H]
-\includegraphics[width=\linewidth]{rotation}
-\caption{Tracks of PAPER-128 for a selection of 16 baselines and frequency $\nu=0.15\text{GHz}$ and a phase center that passes through zenith.
-Tracks are traced out over half a sidereal day. Color represents different baselines. 
-As the Earth rotates, tracks are traced out counterclockwise. Tracks in $uvw$ space do not cross, even though the $uv$ projections do. }
+\begin{figure}[h]
+\includegraphics[width=\linewidth]{rotation_new}
+\caption{Tracks of 3 PAPER-128 baseline classes shown in Fig.
+\ref{fig:AntPos}, here with the same respective colors.
+Tracks are traced out over half a sidereal day for frequency $\nu=0.15\text{GHz}$ and a phase center that passes through zenith.
+The top panel shows $uv$ tracks with crossings among each pair of baseline classes.
+Projection to the $vw$ plane in the bottom panel shows that tracks in $uvw$ space do not cross.
+}
 \label{fig:Tracks}
@@
 \ref{fig:Tracks}.
-As we see in Fig.
+As we see in the top panel of Fig.
 \ref{fig:Tracks}, there would be many such crossings.
@@
 not accurate enough for time offset determination, nor can it give estimate of the degree of redundancy.
-The most obvious reason is that the 3 dimensional $uvw$ tracks do not actually cross.
+The most obvious reason is that the 3 dimensional $uvw$ tracks do not actually cross, as is evident from the $vw$ projection in the bottom panel of Fig.
+\ref{fig:Tracks}.
 For drift-scan arrays, even a hypothetical crossing in $uvw$ space would not imply perfect redundancy.
@@
 normalized to 1 at its peak (zenith), and we assume it to be the same
-for all baselines. We show sample beams of HERA and PAPER antennas in Fig. \ref{fig:Beam} for reference.
+for all baselines. %We show sample beams of HERA and PAPER antennas in Fig. \ref{fig:Beam} for reference.
 
-\begin{figure}[H]
-\includegraphics[width=1.2\linewidth]{Beams}
+%\begin{figure}[H]
+%\includegraphics[width=1.2\linewidth]{Beams}
 
-\caption{Example beam response of HERA (left) and PAPER (right) antennas.
+%\caption{Example beam response of HERA (left) and PAPER (right) antennas.
 Both beams are Stokes $I$ polarization antenna voltage beams, at frequency of $\nu=150\text{MHz}$ and normalized to $1$ at zenith.
 The circles centered around zenith (center of beam) here are
-spaced 10 degrees apart. \label{fig:Beam}}
-\end{figure}
+%spaced 10 degrees apart. \label{fig:Beam}}
+%\end{figure}
 
@@
 where we introduced a frequency-dependent phase $\psi_{\nu}$, which to first order corresponds to phasing the two visibilities to the same phase center.
-\footnote{Turns out for drift-scan measurements, the first-order interpretation is largely sufficient, but we keep the term for now to reserve the option to determine its exact form numerically.}
+\footnote{Turns out for many cases, the first-order interpretation is largely sufficient, but we still keep the term to reserve the option to determine its exact form numerically.
+}
 
@@
 \caption{PAPER128 beam-fringe patterns and their point-spread functions on the $uv$-plane for a frequency of $150$MHz.
-Selected baselines have 30 meter East-West separation, and 0 and 4 meter North-South separation, respectively.The top panels show the real parts of the beam-fringe patterns of the baselines (left and middle), and their conjugate product (right).
+Selected baselines have 30 meter East-West separation, and 0 and 4 meter North-South separation, respectively.
+The top panels show the real parts of the beam-fringe patterns of the baselines (left and middle), and their conjugate product (right).
 The beam-fringe values are normalized such that the peak of the original beam is unity.
-The bottom left and middle panels show the $uv$ beam spread of the two beam-fringe patterns.
+The bottom left and middle panels show the peak-normalized $uv$ beam power of the two beam-fringe patterns.
 The product of the beam spreads displayed in the bottom right panel shows that power concentrated at $u\approx15$ is recovered.
@@
 The motivations behind the designs are explained in \cite{HERAconfiguration}.
-Compared to PAPER128, the hexagonal pattern of HERA lack short baselines oriented close to each other, and the smaller beam (Fig.
-\ref{fig:Beam}) means that we expect to see only longer nearly equivalent baselines.
+Compared to PAPER128, the hexagonal pattern of HERA lack short baselines oriented close to each other, and the smaller beam means that we expect to see only longer nearly equivalent baselines.
 The lower multiplicities per class of baselines is compensated by the larger number of classes of baseline-pairs, especially given the gap in the larger versions.

### 2017-07-11 Rotation paper/draft_working.tex
Added chromaticity section

--- 2017-07-08 Rotation paper/draft_working.tex
@@
 \shorttitle{Adding Sensitivity to 21\,cm Inteferometric Probes of Reionization}
-\shortauthors{Zhang, Liu \&Parsons}
+\shortauthors{Zhang, Liu \& Parsons}
 \usepackage{float}
@@
 \subsection{Formalism}
+\label{sec:formalism}
 In this section we formulate the relation between 
@@
 \subsection{Chromaticity \label{sec:chromaticity}}
-Manageable chromaticity of the instrument is crucial in delay-transform technique of foreground isolation.
+%Manageable chromaticity of the instrument is crucial in delay-transform technique of foreground isolation.
 Previous studies have shown that smooth foreground are constrainted in a "wedge" structure in cosmological $k$ space \citep{wedge1, wedge2}.
@@
 ]
+
+A crucial concern when measuring the $21\,\textrm{cm}$ power spectrum is the possibility of astrophysical foreground contamination.
+Being orders of magnitude larger than the cosmological signal, even small residuals from an imperfect isolation of these foregrounds can masquerade as a false detection.
+Recent literature treatments of this problem have shown that if one combines the mathematical properties of interferometric measurement with the empirical fact that foregrounds are spectral smooth, a clean separation of foregrounds and cosmological signal can in principle be achieved in harmonic space.
+These studies suggest that foregrounds preferentially appear in harmonic modes of the sky corresponding to fluctuations that are angularly fine but spectrally smooth, a region in harmonic space that is colloquially known as ``the wedge".
+One's foreground mitigation strategy can then consist mainly of avoiding those modes by working in the complementary region known as ``the EoR window".
+However, for this strategy to be viable, it is necessary to show that one's data analysis pipeline does not corrupt the separation between the wedge and the EoR window.
+We will now do so for our proposed multi-baseline power spectrum estimator.
+
+%To evaluate a power spectrum estimator's propensity to cause leakage of foregrounds into the EoR window, one may compute the window function $W(\mathbf{k},\mathbf{k}^\prime)$, which relates the expectation value of the power spectrum estimate $\hat{P}$ to the true power spectrum $P$ via
+%\begin{equation}
+%\langle \hat{P}(\mathbf{k}) \rangle = \int dk^\prime P(\mathbf{k}^\prime) W(\mathbf{k},\mathbf{k}^\prime).
+%\end{equation}
+%The window function thus captures the linear combination of modes that source the power spectrum estimate at a particular $\mathbf{k}$.
+If the window function for a $\mathbf{k}$ mode within the EoR window is significant at $\mathbf{k}^\prime$ modes where foregrounds are strong, the EoR window will not be clean of foreground contamination.
+
+As discussed in previous sections, the combination of multi-baseline information for wide-field telescopes is inherently a three-dimensional problem.
+Thus, to fully capture possible curved sky effects, we employ a set of basis functions where angular fluctuations are encoded by spherical harmonics, and radial fluctuations by spherical Bessel functions.
+Modes in the sky are indexed by the magnitude $k$ of the wavevector $\mathbf{k}$ and spherical harmonic indices $(\ell, m)$.
+In other words, one defines
+\begin{equation}
+\overline{T}_{\ell m} (k) \equiv \sqrt{\frac{2}{\pi}} \int d\Omega dr r^2 j_\ell (kr) Y_{\ell m}^* (\hat{\mathbf{r}}) T (\hat{\mathbf{r}}),
+\end{equation}
+where $Y_{\ell m}$ is a spherical harmonic function and $j_\ell$ is the $\ell$th-order spherical Bessel function of the first kind.
+The power spectrum is then related to these spherical harmonic Bessel modes via the relation
+\begin{equation}
+\label{eq:SHB}
+\langle \overline{T}_{\ell m} (k) \overline{T}_{\ell^\prime m^\prime}^* (k^\prime) \rangle = \frac{\delta^D (k - k^\prime)}{k^2} \delta_{\ell \ell^\prime} \delta_{m m^\prime} P(k).
+\end{equation}
+
+A delay-transformed visibility $V(\mathbf{b}, \tau) $ is related to the spherical Fourier-Bessel modes by
+\begin{equation}
+V(\mathbf{b}, \tau) = \frac{2 k_B}{\lambda^2} \sqrt{\frac{2}{\pi}} \sum_{\ell m} \int \! dk\, k^2 g_{\ell m} (k; \mathbf{b}, \tau) \overline{T}_{\ell m} (k),
+\end{equation}
+where
+\begin{equation}
+g_{\ell m} (k ; \mathbf{b}, \tau) = \int \! d\Omega d\nu A(\hat{\mathbf{r}},\nu) Y_{\ell m} (\hat{\mathbf{r}}) j_\ell (k r) e^{i 2 \pi \nu (\tau - \mathbf{b}\cdot \hat{\mathbf{r}} / c)}
+\end{equation}
+and the analogous quantity for $V_\psi$ is given by
+\begin{eqnarray}
+\label{eq:glmpsi}
+g_{\ell m}^\psi (k ; \mathbf{b}, \tau) = \int && d\Omega d\nu A(\boldsymbol \Gamma \hat{\mathbf{r}},\nu) Y_{\ell m} (\hat{\mathbf{r}}) j_\ell (k r) \nonumber \\
+&& \times  e^{i 2 \pi \nu (\tau - \mathbf{b}\cdot \boldsymbol \Gamma \hat{\mathbf{r}} / c) + i \psi_\nu}.
+\end{eqnarray}
+Note that unlike the primary beam and the fringe pattern, the spherical harmonic function does \emph{not} rotate, since it originated from a spherical harmonic expansion of the sky temperature, which is fixed.
+With these expressions and a little algebra combining Eqs.
+\eqref{eq:SHB} through \eqref{eq:glmpsi}, the expectation value of our two-baseline estimator from Section \ref{sec:formalism} is then given by
+\begin{eqnarray}
+\hat{P}(k) &=& \left(\frac{\lambda^{2}}{2k_{B}}\right)^{2} \frac{ \langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau) \rangle}{\Theta} \nonumber \\
+&=& \sum_\ell \int \!dk\, W_{\ell}(k; \mathbf{b}, \mathbf{b}^\prime , \tau) P(k),
+\end{eqnarray}
+where we have defined the \emph{window function}
+\begin{equation}
+W_{\ell}(k; \mathbf{b}, \mathbf{b}^\prime , \tau) = \frac{2 k^2}{\pi \Theta} \sum_m g_{\ell m} (k ; \mathbf{b}, \tau) g_{\ell m}^{\psi*} (k ; \mathbf{b}^\prime, \tau).
+\end{equation}
+By construction, the window function sums to unity when integrated over all $k$ and summed over all $\ell$.
+We may therefore interpret our estimator $\hat{P}$ of the power spectrum to be a weighted average over all possible $\ell$ and $k$ modes, with the window function providing the weights of this average.
+This means that window functions can serve as indicators of foreground leakage into the EoR window.
+One writes down the estimator for a hypothetical measurement of a power spectrum mode within the EoR window, and additionally computes the window function for that estimator.
+Ideally, the window function for $\hat{P}(k)$ will be sharply peaked around $k$.
+In general, however, window functions will have wings that encroach upon other modes.
+If the window function is substantially non-zero at low-$k$ modes---which is where foreground emission typically resides, given its spectral smoothness---then the EoR window will be contaminated by foregrounds.
+
+Figure \ref{fig:multiblwindows} shows four example window functions.
+The two window functions in the top plot are for single-baseline power spectrum estimators, i.e., where one squares the delay spectrum from a single baseline and normalizes the result.
+The window function localized at low $\ell$ is for a $30\,\textrm{m}$ baseline, while the window function localized at high $\ell$ is for a $180\,\textrm{m}$ baseline.
+The delay spectra from these baselines can alternatively be multiplied by those from different---but partially redundant---baselines of similar length to form power spectrum estimates using Eq.
+\eqref{eq:opp}.
+The window functions for these estimates are shown in the bottom plot.
+In all cases, the window functions that we show are for $\tau = 703\,\textrm{ns}$, and were generated using primary beam models for PAPER.
+For numerical convenience we assumed that the bandpass (which, recall from the conventions established in Section \ref{sec:formalism} was absorbed into the definition of the primary beam) takes the form of a half-period sine curve that peaks at $150\,\textrm{MHz}$ and goes to zero $5\,\textrm{MHz}$ on either side of the peak.
+
+One sees that all four window functions are fairly localized around specific values of $k$, which suggests that both the single- and multi-baseline delay-spectrum estimators are good estimators of the power spectrum.
+At higher $\ell$, the window functions for both estimators become elongated in $k$, with stronger tails towards the low $k$ modes where foregrounds reside.
+This is the phenomenology of the wedge, where at fine angular scales the foreground leakage to higher $k$ modes are more pronounced.
+Importantly, however, we note that the window functions for the multi-baseline estimator are no more elongated than for the single-baseline estimator.
+One may thus conclude that the our proposed multi-baseline estimator does not result in extra foreground contamination of the EoR window.
+
+%This suggests that the 
+%Note that the right hand side has no explicit dependence on $\ell$ or $m$, aside from the Kronecker Delta functions to enforce that the different spherical harmonic modes are statistically independent.
+This is a consequence of the statistically translation-invariant nature of the cosmological signal.
+For foregrounds, this invariance is broken because the signal in question is a broadband signal at $z \sim 0$, rather than a signal from a spectral line over a range of redshifts.
+The nature of its angular fluctuations are therefore inherently different from its radial
+
+\begin{figure*}[h!]
+\includegraphics[width=1.0\textwidth]{multiblwindows.png}
+\caption{Sample window functions for delay-spectrum estimates of the power spectrum explored in this paper.
+The top plot shows window functions for a traditional delay-spectrum estimate, where a single baseline's delay spectrum is squared.
+The bottom plot shows the window functions resulting from cross-multiplying baselines of similar length in the estimator given by Equation \eqref{eq:opp}, with the time offset between the time series of these baselines determined by the methods described in Section \ref{sec:method}.
+In each plot, the window function on the right corresponds to the window function for a baseline of length $b \sim 180\,\textrm{m}$, while the window function on the left is for $b\sim 30\,\textrm{m}$.
+The general extent of the window functions for the multi-baseline estimator are seen to be roughly the same as for the single-baseline estimator, suggesting that the former is just as good as the latter in keeping the EoR window free of foregrounds.}
+\label{fig:multiblwindows}
+\end{figure*}
 

### 2017-07-12 Rotation paper/draft_working.tex
incorporated chromaticity section

--- 2017-07-11 Rotation paper/draft_working.tex
@@
 \documentclass[twocolumn,apj,numberedappendix]{emulateapj}
-\shorttitle{Adding Sensitivity to 21\,cm Inteferometric Probes of Reionization}
-\shortauthors{Zhang, Liu \&Parsons}
+\shorttitle{A Two-Baseline 21\,cm Power Spectrum Estimator}
+\shortauthors{Zhang, Liu \& Parsons}
 \usepackage{float}
@@
 
-\title{A Generalized Visbility to Power-Spectrum Framework for 21\,cm Inteferometric Probes of Reionizations}
+\title{Unlocking Sensitivity with A Two-baseline Power Spectrum Estimator for 21\,cm Inteferometric Probes of Reionization }
 
@@
 Arrays capable of such tracking include MWA and LOFAR.
-Our work parallels this effort by focusing on drift-scan arrays that do not have tracking capabilities, such as PAPER, HERA and potentially SKA-low.
+Our work introduces a two-baseline power spectrum estimator that can be applied to both tracking capable arrays and drift-scan only telescopes, such as PAPER, HERA and potentially SKA-low.
 
@@
 We can thus cross-multiply time-shifted visibilities, with the proper weighting, to form power spectra.
-Due to the large number of elements of modern arrays (upward of 100), the task of cross-multiplying every baseline against every other, scaling as number of array elements to the fourth power, can be computationally formidable, and many pairs of baselines provide only negligible redundancy information.
+Due to the large number of elements of modern arrays, the task of cross-multiplying every baseline against every other, scaling as number of array elements to the fourth power, can be computationally formidable, and many pairs of baselines provide only negligible redundancy information.
 Our contribution is thus twofold.
-First we introduce a formalism to include pairs of partially-redundant baselines in a visibility based power spectrum pipeline.
-Secondly we show how to use the formalism to automatically pre-select baseline pairs and time offsets, making the problem computationally efficient.
+First we introduce a formalism to estimate the power spectrum from pairs of partially-redundant baselines in a visibility pipeline.
+Secondly we show how to use the two-baseline estimator to automatically pre-select baseline pairs and time offsets, making the problem computationally efficient.
 More precisely, our formalism allows one to simultaneously identify the baselines that give good
@@
 
-\subsection{Formalism}
-In this section we formulate the relation between 
-the product of delay-transformed visibilities from two arbitrary baselines and the power spectrum.
+\subsection{Formalism \label{sec:formalism}}
+\label{sec:formalism}
+In this section we formulate a power spectrum estimator from
+the product of delay-transformed visibilities from two arbitrary baselines.
 
@@
 \begin{aligned}V(\boldsymbol{b},\tau) & =\int d\nu V_{\nu}(\boldsymbol{b})\phi(\nu)\exp\left(2\pi i\nu\tau\right),\\
- & =\frac{2k_{B}}{\lambda^{2}}\int d\Omega d\nu A(\hat{\boldsymbol{s}},\nu)T(\hat{\boldsymbol{s}},\nu)\exp\left[-2\pi i\nu\left(\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}-\tau\right)\right]. 
+ & =\int d\Omega d\nu B(\hat{\boldsymbol{s}},\nu)T(\hat{\boldsymbol{s}},\nu)\exp\left[-2\pi i\nu\left(\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}-\tau\right)\right]. 
 \end{aligned}
@@
 Here the delay $\tau$ is the Fourier dual of $\nu$. Eq. \eqref{eq:Vb1} expresses the delay-transformed visibility as
-an integral over observational coordinates $\hat{\boldsymbol{s}}$ and $\nu$.
-For notational simplicity here and in the rest of this paper, we shall absorb the bandpass $\phi$ inside the primary beam $A$.
+an integral over observational coordinates $\hat{\boldsymbol{s}}$ and $\nu$. For notational simplicity here and in the rest of this paper, we have defined the quantity
+\[
+B(\hat{\boldsymbol{s}},\nu) \equiv \frac{2k_B}{\lambda^2}\phi(\nu)A(\hat{\boldsymbol{s}},\nu). 
+\label{eq:B}
+\]
 
@@
 \[
-\begin{aligned}V(\boldsymbol{b},\tau) & =\frac{2k_{B}}{\lambda^{2}}\int\frac{d^{3}r}{X^{2}Y}A(\boldsymbol{r})T(\boldsymbol{r})\exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\hat{\boldsymbol{r}}-\tau\right)\nu_{r}\right],
+\begin{aligned}V(\boldsymbol{b},\tau) & =\int\frac{d^{3}r}{X^{2}Y}B(\boldsymbol{r})T(\boldsymbol{r})\exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\hat{\boldsymbol{r}}-\tau\right)\nu_{r}\right],
 \end{aligned}
@@
 The beam pattern of a baseline shifts relative to the sky as the Earth rotates. Here we choose to fix the sky, and denote the rotated coordinates
-in the topocentric frame with the 3 dimensional rotation operator $\Gamma$:
+in the topocentric frame with the 3 dimensional rotation operator $\boldsymbol{\Gamma}$:
 
 \[
-\begin{aligned}V_{\psi}(\boldsymbol{b'},\tau) & =\frac{2k_{B}}{\lambda^{2}}\int\frac{d^{3}r}{X^{2}Y}A(\Gamma\boldsymbol{r})T(\boldsymbol{r})\\
-& \qquad \qquad  \exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\Gamma\hat{\boldsymbol{r}}-\tau\right)\nu_{r}-i\psi_\nu\right],\end{aligned}
+\begin{aligned}V_{\psi}(\boldsymbol{b'},\tau) & = \int\frac{d^{3}r}{X^{2}Y}B(\boldsymbol{\Gamma}\boldsymbol{r})T(\boldsymbol{r})\\
+& \qquad \qquad  \exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\boldsymbol{\Gamma}\hat{\boldsymbol{r}}-\tau\right)\nu_{r}-i\psi_\nu\right],\end{aligned}
 \]
@@
 \begin{equation}
-\begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle\\
- & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle\right)A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
- & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\boldsymbol{\kappa})T(\boldsymbol{\kappa'})\rangle e^{-i(\boldsymbol{\kappa}\cdot \boldsymbol{r}-\boldsymbol{\kappa'}\cdot\boldsymbol{r'})}\right)A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
- & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\boldsymbol{\kappa}\cdot(\boldsymbol{r}-\boldsymbol{r'})}\right)A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
- & \approx\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}rd^{3}r'\delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
- & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}rA^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r})\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\Gamma \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right],\\
- & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma \hat{\boldsymbol{s}},\nu)\exp\left[-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right],
+\begin{aligned} \langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle
+ & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle\right)B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}),\\
+ & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\boldsymbol{\kappa})T(\boldsymbol{\kappa'})\rangle e^{-i(\boldsymbol{\kappa}\cdot \boldsymbol{r}-\boldsymbol{\kappa'}\cdot\boldsymbol{r'})}\right)B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}),\\
+ & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\boldsymbol{\kappa}\cdot(\boldsymbol{r}-\boldsymbol{r'})}\right)B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}),\\
+ & \approx P(k_{b,\tau})\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2} \delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}),\\
+ & = P(k_{b,\tau})\int \frac{d^{3}r}{(X^2Y)^2}B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r})\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma} \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right],\\
+ & = P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}B^{*}(\hat{\boldsymbol{s}},\nu)B(\boldsymbol{\Gamma} \hat{\boldsymbol{s}},\nu)\exp\left[-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma}\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right],
 \end{aligned}
@@
 \end{equation}
-In transition from cosmological coordinates back to observing coordinates we have written $\hat{\boldsymbol{r}}\equiv\hat{\boldsymbol{s}}$, and 
+
+with
 \begin{equation}
-\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'})\equiv\exp\left[i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\Gamma\hat{\boldsymbol{r'}}\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)-i\psi_{\nu}\right].
+\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\equiv\exp\left[i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\boldsymbol{\Gamma}\hat{\boldsymbol{r'}}\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)-i\psi_{\nu}\right].
 \end{equation}
+
 \end{widetext}
-The third equality of Eq.(\ref{eq:main}) follows from assumption of translational invariance of statistics of the 21\,cm field, and the first inequality follows from the assumption that
-the 3D power spectrum varies negligibly over the $k$-space of interest.
-The careful reader may have noticed that we have also assumed here a definitive relation between $k$ and $b,\tau$.
-In the flat-sky limit for example, such a relation would take the form $(Xk_{x},Xk_{y},Yk_{z})=\frac{2\pi}{c}(b_{x},b_{y},\tau)$.
-Instrumental chromaticity and foreground isolation will be examined in \ref{sec:chromaticity}.
-Since $\Gamma$
+
+Above $\delta_D$ is the Dirac delta-function and in transition from cosmological coordinates back to observing coordinates we have written $\hat{\boldsymbol{r}}\equiv\hat{\boldsymbol{s}}$.
+The third equality of Eq.(\ref{eq:main}) follows from assumption of translational invariance of statistics of the 21\,cm field.
+For the inequality in the fourth line we have assumed that the  3D power spectrum varies negligibly over the $k$-space of interest and thus pulled out the integral a power spectrum centered at 
+\[
+k_{\boldsymbol{b}, \tau} \equiv 2\pi \sqrt{\left(\frac{\tau}{Y}\right)^2 + \left(\frac{b}{\lambda X}\right)^2}. 
+\label{eq:kbtau}
+\]
+In Section \ref{sec:chromaticity} we will examine the exact form of the window function probed as well as its effect on foreground isolation.  Since $\boldsymbol{\Gamma}$
 is a sky rotation, it does not affect $\nu$, hence we have taken $\nu_{r}$
-outside the parenthesis in the phase term $\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\Gamma \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)\right]$.
+outside the parenthesis in the phase term $\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma} \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)\right]$.
 Notice that the phase factor $\exp\left[-i2\pi\tau\left(\nu-\nu'\right)\right]$
@@
 
-Since the beam pattern and bandwidth are given in $\hat{\boldsymbol{s}}$
+Since the beam pattern and bandpass are given in $\hat{\boldsymbol{s}}$
 and $\nu$, we convert the integral back to these coordinates to get
@@
 \begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle\\
- & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)\\
- & \qquad \qquad \qquad \qquad \exp\left[i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right].
+ & \qquad \approx P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}B^{*}(\hat{\boldsymbol{s}},\nu)B(\boldsymbol{\Gamma}\hat{\boldsymbol{s}},\nu)\\
+ & \qquad \qquad \qquad \qquad \exp\left[i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma}\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right].
  \end{aligned}
@@
 
-In other words the power spectrum estimate from visibilities of a baseline pair is given by 
+We can therefore form the power spectrum estimator for the baseline pair $\{\boldsymbol b, \boldsymbol  b'\}$:
+
 \begin{equation}
- \hat{P}(k_{b,\tau}) \equiv \left(\frac{\lambda^{2}}{2k_{B}}\right)^{2} \frac{V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)}{\Theta}, 
+ \hat{P}(k_{b,\tau}) \equiv \frac{V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)}{\Theta}, 
  \label{eq:opp}
@@
 \[
-\Theta_{\nu} \equiv \int \frac{d\Omega}{X^{2}Y}A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu) e^{i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)}e^{-i\psi_{\nu}}.
+\Theta_{\nu} \equiv e^{-i\psi_{\nu}} \int \frac{d\Omega}{X^{2}Y}B^{*}(\hat{\boldsymbol{s}},\nu)B(\boldsymbol{\Gamma}\hat{\boldsymbol{s}},\nu) e^{i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma}\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)}.
 \label{eq:Thetanu}
@@
 \eqref{eq:Thetanu} we introduced an extra phase $\psi$, the origin of which we shall explain in \ref{sec:rephs}.
-We point out that although all our derivations focused on drift-scan telescopes, we can get the analogous result for tracking measurements simply by noticing that for a tracking primary beam, $\Gamma$ becomes a rotation around zenith, and so:
+We point out that although all our derivations focused on drift-scan telescopes, we can get the analogous result for tracking measurements simply by noticing that for a tracking primary beam, $\boldsymbol{\Gamma}$ becomes a rotation around zenith, and so:
 \[
-\Theta_{\nu} \equiv \int \frac{d\Omega}{X^{2}Y}A^{*}(\hat{\boldsymbol{s}},\nu)A(\hat{\boldsymbol{s}},\nu) e^{i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)}e^{-i\psi_{\nu}}.
+\Theta_{\nu} \equiv e^{-i\psi_{\nu}} \int \frac{d\Omega}{X^{2}Y}B^{*}(\hat{\boldsymbol{s}},\nu)B(\hat{\boldsymbol{s}},\nu) e^{i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma}\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)}.
 \label{eq:Thetanu_tracking}
@@
 transform of the cross multiplied beam patterns. As a check, when applied to equivalent baselines,
-$\boldsymbol{b}=\boldsymbol{b'}$, $\hat{\boldsymbol{s}}=\Gamma\hat{\boldsymbol{s}}$, and Eq.(\ref{eq:final}) reduces to Eq.(B9) of \cite{paper32}. 
+$\boldsymbol{b}=\boldsymbol{b'}$, $\hat{\boldsymbol{s}}=\boldsymbol{\Gamma}\hat{\boldsymbol{s}}$, and Eq.(\ref{eq:final}) reduces to Eq.(B9) of \cite{paper32}. 
 With Eq.
@@
 ]
+
+A crucial concern when measuring the $21\,\textrm{cm}$ power spectrum is the possibility of astrophysical foreground contamination.
+Being orders of magnitude larger than the cosmological signal, even small residuals from an imperfect isolation of these foregrounds can masquerade as a false detection.
+Recent literature treatments of this problem have shown that if one combines the mathematical properties of interferometric measurement with the empirical fact that foregrounds are spectrally smooth, a clean separation of foregrounds and cosmological signal can in principle be achieved in harmonic space \citep{wedge1, wedge2}.
+These studies suggest that foregrounds preferentially appear in harmonic modes of the sky corresponding to fluctuations that are angularly fine but spectrally smooth, a region in harmonic space colloquially known as ``the wedge".
+One's foreground mitigation strategy can then consist mainly of avoiding those modes by working in the complementary region known as ``the EoR window".
+However, for this strategy to be viable, it is necessary to show that one's data analysis pipeline does not corrupt the separation between the wedge and the EoR window.
+We will now do so for our proposed two-baseline power spectrum estimator.
+
+%To evaluate a power spectrum estimator's propensity to cause leakage of foregrounds into the EoR window, one may compute the window function $W(\mathbf{k},\mathbf{k}^\prime)$, which relates the expectation value of the power spectrum estimate $\hat{P}$ to the true power spectrum $P$ via
+%\begin{equation}
+%\langle \hat{P}(\mathbf{k}) \rangle = \int dk^\prime P(\mathbf{k}^\prime) W(\mathbf{k},\mathbf{k}^\prime).
+%\end{equation}
+%The window function thus captures the linear combination of modes that source the power spectrum estimate at a particular $\mathbf{k}$.
+If the window function for a $\mathbf{k}$ mode within the EoR window is significant at $\mathbf{k}^\prime$ modes where foregrounds are strong, the EoR window will not be clean of foreground contamination.
+
+Traditionally, spectral and angular dimensions are cylindrically binned with the coordinates $k_\bot$ and  $k_\parallel$, where the angular direction $k_\bot$ is probed by the beam and line-of-sight direction $k_\parallel$ is probed by the frequency spectrum.
+Despite their usefulness, however, $(k_\bot, k_\parallel)$ are local coordinates that are well defined only in the narrow field-of-view limit.
+As discussed in previous sections, the combination of information from nearly equivalent baselines for wide-field telescopes is inherently a three-dimensional problem.
+Although the estimator of Section \ref{sec:method} does not have such limitations, to discuss foreground signature with possible curved sky effects, we must employ a set of basis functions where angular fluctuations are encoded by spherical harmonics, and radial fluctuations by spherical Bessel functions.
+
+\cite{sphere} discusses in detail power spectrum analyses using spherical harmonics.
+In such as basis, modes in the sky are indexed by the magnitude $k$ of the wavevector $\mathbf{k}$ and spherical harmonic indices $(\ell, m)$.
+In other words, one defines
+\begin{equation}
+\overline{T}_{\ell m} (k) \equiv \sqrt{\frac{2}{\pi}} \int d\Omega dr r^2 j_\ell (kr) Y_{\ell m}^* (\hat{\mathbf{r}}) T (\hat{\mathbf{r}}),
+\end{equation}
+where $Y_{\ell m}$ is a spherical harmonic function and $j_\ell$ is the $\ell$th-order spherical Bessel function of the first kind.
+The power spectrum is then related to these spherical harmonic Bessel modes via the relation
+\begin{equation}
+\label{eq:SHB}
+\langle \overline{T}_{\ell m} (k) \overline{T}_{\ell^\prime m^\prime}^* (k^\prime) \rangle = \frac{\delta^D (k - k^\prime)}{k^2} \delta_{\ell \ell^\prime} \delta_{m m^\prime} P(k).
+\end{equation}
+
+A delay-transformed visibility $V(\mathbf{b}, \tau) $ is related to the spherical Fourier-Bessel modes by
+\begin{equation}
+V(\mathbf{b}, \tau) = \sqrt{\frac{2}{\pi}} \sum_{\ell m} \int \! dk\, k^2 g_{\ell m} (k; \mathbf{b}, \tau) \overline{T}_{\ell m} (k),
+\end{equation}
+where
+\begin{equation}
+g_{\ell m} (k ; \mathbf{b}, \tau) = \int \! d\Omega d\nu B(\hat{\mathbf{r}},\nu) Y_{\ell m} (\hat{\mathbf{r}}) j_\ell (k r) e^{i 2 \pi \nu (\tau - \mathbf{b}\cdot \hat{\mathbf{r}} / c)}
+\end{equation}
+and the analogous quantity for $V_\psi$ is given by
+\begin{eqnarray}
+\label{eq:glmpsi}
+g_{\ell m}^\psi (k ; \mathbf{b}, \tau) = \int && d\Omega d\nu B(\boldsymbol \Gamma \hat{\mathbf{r}},\nu) Y_{\ell m} (\hat{\mathbf{r}}) j_\ell (k r) \nonumber \\
+&& \times  e^{i 2 \pi \nu (\tau - \mathbf{b}\cdot \boldsymbol \Gamma \hat{\mathbf{r}} / c) + i \psi_\nu}.
+\end{eqnarray}
+Note that unlike the primary beam and the fringe pattern, the spherical harmonic function does \emph{not} rotate, since it originated from a spherical harmonic expansion of the sky temperature, which is fixed.
+With these expressions and a little algebra combining Eqs.
+\eqref{eq:SHB} through \eqref{eq:glmpsi}, we arrive at a form of the two-baseline estimator (compare with Eq.
+\eqref{eq:final}):
+\begin{eqnarray}
+\hat{P}(k) &=& \frac{ \langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau) \rangle}{\Theta} \nonumber \\
+&=& \sum_\ell \int \!dk\, W_{\ell}(k; \mathbf{b}, \mathbf{b}^\prime , \tau) P(k),
+\end{eqnarray}
+where we have defined the \emph{window function}
+\begin{equation}
+W_{\ell}(k; \mathbf{b}, \mathbf{b}^\prime , \tau) = \frac{2 k^2}{\pi \Theta} \sum_m g_{\ell m} (k ; \mathbf{b}, \tau) g_{\ell m}^{\psi*} (k ; \mathbf{b}^\prime, \tau).
+\end{equation}
+By construction, the window function sums to unity when integrated over all $k$ and summed over all $\ell$.
+We may therefore interpret our estimator $\hat{P}$ of the power spectrum to be a weighted average over all possible $\ell$ and $k$ modes, with the window function providing the weights of this average.
+This means that window functions can serve as indicators of foreground leakage into the EoR window.
+One writes down the estimator for a hypothetical measurement of a power spectrum mode within the EoR window, and additionally computes the window function for that estimator.
+Ideally, the window function for $\hat{P}(k)$ will be sharply peaked around $k$.
+In general, however, window functions will have wings that encroach upon other modes.
+If the window function is substantially non-zero at low-$k$ modes---which is where foreground emission typically resides, given its spectral smoothness---then the EoR window will be contaminated by foregrounds.
+
+Figure \ref{fig:multiblwindows} shows four example window functions.
+The two window functions in the top plot are for equivalent baseline power spectrum estimators, i.e., where one squares the delay spectrum from a single baseline and normalizes the result.
+The window function localized at low $\ell$ is for the baseline \{2,0\} of around $30\,\textrm{m}$ in length, while the window function localized at high $\ell$ is for \{12,0\}, roughly $180\,\textrm{m}$ in length.
+The window functions for near-equivalent baseline estimates are shown in the bottom plot, for \{2,0\}:\{2,1\} and \{12,0\}:\{12,1\} at low- and high-$l$, respectively.
+In all cases, the window functions that we show are for $\tau = 703\,\textrm{ns}$, and were generated using primary beam models for PAPER.
+For numerical convenience we assumed that the bandpass $\phi_\nu$ (see Eq.
+\eqref{eq:B}) takes the form of a half-period sine curve that peaks at $150\,\textrm{MHz}$ and goes to zero $5\,\textrm{MHz}$ on either side of the peak.
+
+One sees that all four window functions are fairly localized around specific values of $k$, which to a good approximation are described by Eq.
+\eqref{eq:kbtau}.
+This suggests that both equivalent- and near-equivalent-baseline delay-spectrum estimators are good estimators of the power spectrum.
+At higher $\ell$, the window functions for both estimators become elongated in $k$, with stronger tails towards the low $k$ modes where foregrounds reside.
+This is the phenomenology of the wedge, where at fine angular scales the foreground leakage to higher $k$ modes are more pronounced.
+Importantly, however, we note that the window functions for the near-equivalent-baseline estimator are no more elongated than for the equivalent-baseline estimator.
+One may thus conclude that the our proposed near-equivalent-baseline estimator does not result in extra foreground contamination of the EoR window.
+
+%This suggests that the 
+%Note that the right hand side has no explicit dependence on $\ell$ or $m$, aside from the Kronecker Delta functions to enforce that the different spherical harmonic modes are statistically independent.
+This is a consequence of the statistically translation-invariant nature of the cosmological signal.
+For foregrounds, this invariance is broken because the signal in question is a broadband signal at $z \sim 0$, rather than a signal from a spectral line over a range of redshifts.
+The nature of its angular fluctuations are therefore inherently different from its radial
+
+\begin{figure}[h]
+\includegraphics[width=1.1\linewidth]{multiblwindows}
+\caption{Sample window functions for delay-spectrum estimates of the power spectrum explored in this paper.
+The top plot shows window functions for a traditional equivalent-baseline delay-spectrum estimate, where a single baseline's delay spectrum is squared.
+The bottom plot shows the window functions resulting from nearly equivalent baselines.
+In each plot, the window function on the right corresponds to the window function for baselines of length $b \sim 180\,\textrm{m}$ in the East West direction, while the window function on the left is for $b\sim 30\,\textrm{m}$.
+The general extent of the window functions for the two-baseline estimator are seen to be roughly the same as for the single-baseline estimator, suggesting that the former is just as good as the latter in keeping the EoR window free of foregrounds.}
+\label{fig:multiblwindows}
+\end{figure}
 
@@
 \begin{equation}
-\Theta \equiv\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)||\phi(\nu)|^{2} e^{-i2\pi\frac{\nu}{c}\hat{\boldsymbol{s}}\cdot\left(\boldsymbol{b}-\widetilde{\Gamma}\boldsymbol{b'}\right)},
+\Theta \equiv\int\frac{d\Omega d\nu}{X^{2}Y}B^{*}(\hat{\boldsymbol{s}},\nu)B(\boldsymbol{\Gamma}\hat{\boldsymbol{s}},\nu) e^{-i2\pi\frac{\nu}{c}\hat{\boldsymbol{s}}\cdot\left(\boldsymbol{b}-\widetilde{\boldsymbol{\Gamma}}\boldsymbol{b'}\right)},
 \label{eq:drift}
 \end{equation}
-where $\widetilde{\Gamma}$ is the inverse of $\Gamma$. 
-For tracking elements, $\Gamma$ becomes a rotation around zenith:
+where $\widetilde{\boldsymbol{\Gamma}}$ is the inverse of $\boldsymbol{\Gamma}$. 
+For tracking elements, $\boldsymbol{\Gamma}$ becomes a rotation around zenith:
 \begin{equation}
-\Theta \equiv\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\hat{\boldsymbol{s}},\nu)||\phi(\nu)|^{2} e^{-i2\pi\frac{\nu}{c}\hat{\boldsymbol{s}}\cdot\left(\boldsymbol{b}-\widetilde{\Gamma}\boldsymbol{b'}\right)},
+\Theta \equiv\int\frac{d\Omega d\nu}{X^{2}Y}B^{*}(\hat{\boldsymbol{s}},\nu)B(\hat{\boldsymbol{s}},\nu) e^{-i2\pi\frac{\nu}{c}\hat{\boldsymbol{s}}\cdot\left(\boldsymbol{b}-\widetilde{\boldsymbol{\Gamma}}\boldsymbol{b'}\right)},
 \label{eq:tracking}
@@
 \[
- \boldsymbol{b}-\widetilde{\Gamma}\boldsymbol{b'} = 0. \label{eq:trackcross}
+ \boldsymbol{b}-\widetilde{\boldsymbol{\Gamma}}\boldsymbol{b'} = 0. \label{eq:trackcross}
  \]
@@
 \eqref{eq:drift}, we see the track-crossing condition Eq.
-\eqref{eq:trackcross} maximizes $\Theta$ if and only if no other term in the integral depends on $\Gamma$.
+\eqref{eq:trackcross} maximizes $\Theta$ if and only if no other term in the integral depends on $\boldsymbol{\Gamma}$.
 We see that this is true for the tracking case (Eq.

### 2017-07-14 Rotation paper/draft_working.tex
saving progress

@@
 in the sky.
-We shall call baselines in the same class \textit{equivalent baselines}, and reserve the word \textit{redundancy} of two baselines in reference to a variable function of the relative time-offsets of their visibility time series.
+We shall call baselines in the same class \textit{equivalent baselines}, and reserve the word \textit{redundancy} of two baselines in reference to a variable function of the relative time offsets of their visibility time series.
 With this terminology, two equivalent
@@
 This paper provides the method for inter-class multiplications.
-To do so we will use the short hand notation $\{m,n\}:\{m',n'\}$ to denote a pair of baseline classes to be cross-multiplied.
+To do so we will use the short hand notation $\{m,n:m',n'\}$ to denote a pair of baseline classes to be cross-multiplied.
 
@@
 \ref{fig:AntPos}, projected onto $uv$ and $vw$ planes.
-The tracks are traced over 12 sidereal hours, at 0.15GHz, relative to a phase center that passes through zenith.
+The tracks are traced over 12 sidereal hours, at 0.15\,GHz, relative to a phase center that passes through zenith.
 
@@
 \ref{fig:AntPos}, here with the same respective colors.
-Tracks are traced out over half a sidereal day for frequency $\nu=0.15\text{GHz}$ and a phase center that passes through zenith.
+Tracks are traced out over half a sidereal day for frequency $\nu=0.15\text{\,GHz}$ and a phase center that passes through zenith.
 The top panel shows $uv$ tracks with crossings among each pair of baseline classes.
@@
 For drift-scan arrays, even a hypothetical crossing in $uvw$ space would not imply perfect redundancy.
-This shall become evident in the next section, after we develop a more general formalism that accounts for the point spread function of the finite beams, and estimate the degree of redundancy for general combination of baselines at a general time-offset.
+This shall become evident in the next section, after we develop a more general formalism that accounts for the point spread function of the finite beams, and estimate the degree of redundancy for general combination of baselines at a general time offset.
 Then, relation of track-crossing and redundancy for both drift-scan and tracking measurements will be explored in Appendix \ref{sec:appA}.
@@
 %\caption{Example beam response of HERA (left) and PAPER (right) antennas.
-Both beams are Stokes $I$ polarization antenna voltage beams, at frequency of $\nu=150\text{MHz}$ and normalized to $1$ at zenith.
+Both beams are Stokes $I$ polarization antenna voltage beams, at frequency of $\nu=150\text{\,MHz}$ and normalized to $1$ at zenith.
 The circles centered around zenith (center of beam) here are
@@
 
-Power-spectrum measurements are typically taken from $\sim10$MHz centered around the corresponding redshift of interest (e.g. 150 MHz for z=9.5).
+Power-spectrum measurements are typically taken from $\sim10$\,MHz centered around the corresponding redshift of interest (e.g. 150 \,MHz for z=9.5).
 
@@
 \]
-where $\nu_{21}=1420$MHz is the 21\,cm transition rest frequency, $\nu_{0}$
+where $\nu_{21}=1420$\,MHz is the 21\,cm transition rest frequency, $\nu_{0}$
 a reference central frequency with corresponding redshift $z_{0}$,
@@
 Existing visibility based power-spectrum pipelines for redundant drift-scan  arrays relate the power-spectrum to the conjugate square of the visibilities \citep{delay-transform, paper32, Ali2015}.
-We would like to generalize such relations by relating the power-spectrum to the product of two visibilities from two arbitrary baselines and time-offsets.
+We would like to generalize such relations by relating the power-spectrum to the product of two visibilities from two arbitrary baselines and time offsets.
 The beam pattern of a baseline shifts relative to the sky as the Earth rotates. Here we choose to fix the sky, and denote the rotated coordinates
@@
  & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\boldsymbol{\kappa}\cdot(\boldsymbol{r}-\boldsymbol{r'})}\right)B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}),\\
- & \approx P(k_{b,\tau})\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2} \delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}),\\
- & = P(k_{b,\tau})\int \frac{d^{3}r}{(X^2Y)^2}B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r})\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma} \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right],\\
- & = P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}B^{*}(\hat{\boldsymbol{s}},\nu)B(\boldsymbol{\Gamma} \hat{\boldsymbol{s}},\nu)\exp\left[-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma}\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right],
+ & \approx P(k_{\bar{b}, \tau})\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2} \delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}),\\
+ & = P(k_{\bar{b}, \tau})\int \frac{d^{3}r}{(X^2Y)^2}B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r})\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma} \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right],\\
+ & = P(k_{\bar{b}, \tau})\int\frac{d\Omega d\nu}{X^{2}Y}B^{*}(\hat{\boldsymbol{s}},\nu)B(\boldsymbol{\Gamma} \hat{\boldsymbol{s}},\nu)\exp\left[-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma}\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right],
 \end{aligned}
@@
 \begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle\\
- & \qquad \approx P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}B^{*}(\hat{\boldsymbol{s}},\nu)B(\boldsymbol{\Gamma}\hat{\boldsymbol{s}},\nu)\\
+ & \qquad \approx P(k_{\bar{b}, \tau})\int\frac{d\Omega d\nu}{X^{2}Y}B^{*}(\hat{\boldsymbol{s}},\nu)B(\boldsymbol{\Gamma}\hat{\boldsymbol{s}},\nu)\\
  & \qquad \qquad \qquad \qquad \exp\left[i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma}\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right].
@@
 \begin{equation}
- \hat{P}(k_{b,\tau}) \equiv \frac{V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)}{\Theta}, 
+ \hat{P}(k_{\bar{b}, \tau}) \equiv \frac{V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)}{\Theta}, 
  \label{eq:opp}
@@
 In Fig.
-\ref{fig:phi_nu} we show the peak phases for for a given baseline pair of PAPER128 (\{2,0\}:\{2,1\}), comparing the drift-scan phase dependence with that of the same baseline with hypothetical tracking-elements (Eq.
+\ref{fig:phi_nu} we show the peak phases for for a given baseline pair of PAPER128 (\{2,0:2,1\}), comparing the drift-scan phase dependence with that of the same baseline with hypothetical tracking-elements (Eq.
 \eqref{eq:Thetanu} and Eq.
@@
 
-\caption{Frequency dependent peak phases for PAPER-128 baseline pair \{2,0\}:\{2,1\}.
+\caption{Frequency dependent peak phases for PAPER-128 baseline pair \{2,0:2,1\}.
 Both the drift-scan case and a hypothetical tracking baseline are shown in solid lines.
 The first-order, linear effects are shown in dashed lines.
-We have also fixed the global phase in both cases to 0 at 0.15GHz.
+We have also fixed the global phase in both cases to 0 at 0.15\,GHz.
 The drift-scan case exhibits linear behavior and conforms well to the first order effect of a shift in delay space due to the movement of zenith.
@@
 To illustrate the effect of rephasing, we compare in Fig.
-\ref{fig:freqdiff} real parts of $\Theta_{\nu}$ of both equivalent and nearly equivalent baseline pairs for two channels: 0.145 GHz and 0.155 GHz.
+\ref{fig:freqdiff} real parts of $\Theta_{\nu}$ of both equivalent and nearly equivalent baseline pairs for two channels: 0.145\,GHz and 0.155\,GHz.
 The top two panels have zero rephasing, and the bottom two are linearly rephased to a differential delay of $\Delta\tau\approx15.7$ns, as we determined above.
-The first and third panels show the equivalent baseline pairs \{2,0\}:\{2,0\}, and second and fourth panels show \{2,0\}:\{2,1\}.
+The first and third panels show the equivalent baseline pairs \{2,0:2,0\}, and second and fourth panels show \{2,0:2,1\}.
 Summing over frequency without proper rephasing leads to destructive interference and signal loss.
 The wider the frequency profile, the more destructive the interference would be.
-Only after rephasing to the correct time-offset for each baseline pairs can we constructively combine the frequency channels.
+Only after rephasing to the correct time offset for each baseline pairs can we constructively combine the frequency channels.
 
@@
 To gain some visual intuition of the formalism in Section \ref{sec:method}, we show explicit beam fringe patterns of the two baselines and their interactions.
-For this section, we relax into the flat sky approximation so that $(l,m,\sqrt{1-l^2-m^2})\equiv\hat{s}$ and examine a single frequency of $150$MHz.
+For this section, we relax into the flat sky approximation so that $(l,m,\sqrt{1-l^2-m^2})\equiv\hat{s}$ and examine a single frequency of $150$\,MHz.
 
-At the time-offset of maximum redundancy, we expect two beams to have to same fringe pattern in both frequency and phase.
+At the time offset of maximum redundancy, we expect two beams to have to same fringe pattern in both frequency and phase.
 Due to the time delay, however, the beam centers would be slightly shifted with respect to each other.
@@
 
-\caption{PAPER128 beam-fringe patterns and their point-spread functions on the $uv$-plane for a frequency of $150$MHz.
+\caption{PAPER128 beam-fringe patterns and their point-spread functions on the $uv$-plane for a frequency of $150$\,MHz.
 Selected baselines have 30 meter East-West separation, and 0 and 4 meter North-South separation, respectively.
@@
 for a pair of baselines with products of simulated visibilities of those baselines.
-We use 10 frequency channels evenly spaced from $145$MHz to $155$MHz for the comparison.
+We use 10 frequency channels evenly spaced from $145$\,MHz to $155$\,MHz for the comparison.
 We rephase one of the simulated visibilities, as well as computed $\Theta_{\nu}$ to the same delay as computed from the peak phase of the latter.
@@
 Fig.
-\ref{fig:numerics} shows the real parts of the resulting degree of redundancy as a function of time offset, for equivalent baseline pairs (\{2,0\}:\{2,0\}) in the top panel and nearly equivalent (\{2,0\}:\{2,1\}) in the bottom.
+\ref{fig:numerics} shows the real parts of the resulting degree of redundancy as a function of time offset, for equivalent baseline pairs (\{2,0:2,0\}) in the top panel and nearly equivalent (\{2,0:2,1\}) in the bottom.
 Shown in green for the two cases are the computed values of $\Theta$, both normalized to the peak of the equivalent case.
 The plot on the bottom is rephased with a delay of $\Delta\tau\approx15.7$ns as outlined in Section \ref{sec:rephs}.
-Since both y-axes are normalized to the same scale, we see the equivalent baseline reach maximal redundancy with 0 time-offset, while the nearly equivalent pair reaches maximum redundancy of $\Theta\approx0.5$ at $dT\approx0.325$days.
+Since both y-axes are normalized to the same scale, we see the equivalent baseline reach maximal redundancy with 0 time offset, while the nearly equivalent pair reaches maximum redundancy of $\Theta\approx0.5$ at $dT\approx0.325$days.
 
@@
 via the Fourier convolution theorem, to obtain values of the cross
-correlation as a function of time-offset.   We do this for both the equivalent (\{2,0\}:\{2,0\}) and nearly equivalent (\{2,0\}:\{2,1\}) case.
+correlation as a function of time offset.   We do this for both the equivalent (\{2,0:2,0\}) and nearly equivalent (\{2,0:2,1\}) case.
 Since the blue and green curves overlap, we have verified that in this case Eq. \eqref{eq:final} is valid and the computed $\Theta$ can be used as weight to obtain the power spectrum.
@@
 The window function localized at low $\ell$ is for the baseline \{2,0\} of around $30\,\textrm{m}$ in length, while the window function localized at high $\ell$ is for \{12,0\}, roughly $180\,\textrm{m}$ in length.
-The window functions for near-equivalent baseline estimates are shown in the bottom plot, for \{2,0\}:\{2,1\} and \{12,0\}:\{12,1\} at low- and high-$l$, respectively.
+The window functions for near-equivalent baseline estimates are shown in the bottom plot, for \{2,0:2,1\} and \{12,0:12,1\} at low- and high-$\bar{b}$, respectively.
 In all cases, the window functions that we show are for $\tau = 703\,\textrm{ns}$, and were generated using primary beam models for PAPER.
 For numerical convenience we assumed that the bandpass $\phi_\nu$ (see Eq.
-\eqref{eq:B}) takes the form of a half-period sine curve that peaks at $150\,\textrm{MHz}$ and goes to zero $5\,\textrm{MHz}$ on either side of the peak.
+\eqref{eq:B}) takes the form of a half-period sine curve that peaks at $150\,\textrm{\,MHz}$ and goes to zero $5\,\textrm{\,MHz}$ on either side of the peak.
 
@@
 \caption{Relative sensitivity contributions of selected baseline combinations in PAPER-128.
-Each labeled point in the middle and bottom panels correspond to a cross-correlated baseline pair \{m,n\}:\{p,q\}.
+Each labeled point in the middle and bottom panels correspond to a cross-correlated baseline pair \{m,n:p,q\}.
 The shape of the symbol encodes the first baseline \{m,n\}, as displayed in the top left legend panel.
@@
 corresponding multiplicities as in Eq. \eqref{eq:sensul}. 
-In both the middle and bottom panels, we have chosen to fix the value  of \{1,0\}:\{1,0\} to unity.  }
+In both the middle and bottom panels, we have chosen to fix the value  of \{1,0:1,0\} to unity.  }
 \label{fig:sensplot}
@@
 In this section we discuss the sensitivity contributions of a variety of baseline pairs from PAPER128.
-Recall the sensitivity contributions of a particular baseline pair is given by $\Theta$ (Eq. \eqref{eq:Theta}). 
+Recall the sensitivity contribution of a particular baseline pair is given by $\Theta$ (Eq.
+\eqref{eq:Theta}).
+From now one we use $\Theta_{\boldsymbol{bb'}}$ and $\Delta T_{\boldsymbol{bb'}}$ to denote the peak value and location of $\Theta$, or in other words the maximum of the redundancy weight $|\Theta(\Delta T)|$ and the corresponding offset for a given baseline pair $\boldsymbol{b}:\boldsymbol{b'}$.
 %Intuitively we expect the sensitivity to depend on both the $uv$ coverage of the baseline and the patch of sky inside the beam.
-A larger beam like that of PAPER would tolerate larger time-offsets because more sky area can coincide in the two beams \footnote{Larger beams also imply smaller spread in $uv$ space.
+A larger beam like that of PAPER would tolerate larger time offsets because more sky area can coincide in the two beams \footnote{Larger beams also imply smaller spread in $uv$ space.
 This could lead to either larger or smaller redundancy, which depends on the overlap of two such point spread functions.}
@@
 In the middle panel of Fig. \ref{fig:sensplot}
-we show the peak heights $\Theta$ and locations $\Delta T$ for a variety of baseline combinations. Baseline pairs that are mirror images of each other 
-give the same amount of redundancies ($\Theta$), with the opposite time offset, as expected from symmetry. 
-For example, \{1,0\}:\{1,1\} is mirror image of \{1,0\}:\{1,-1\} and these two pairs of baselines
-give the same sensibility contribution. Thus we only show a subset of representative baseline pairs to illustrate the contributions. For a more complete result see Fig. \ref{fig:pairplot}.
+we show $\Theta_{\boldsymbol{bb'}}$ and $\Delta T_{\boldsymbol{bb'}}$ for a variety of baseline combinations. Baseline pairs that are mirror images of each other 
+give the same amount of redundancies ($\Theta_{\boldsymbol{bb'}}$), with the opposite time offset, as expected from symmetry. 
+For example, \{1,0:1,1\} is mirror image of \{1,0:1,-1\} and these two pairs of baselines
+give the same sensibility contribution. Thus we only show a subset of representative baseline pairs to illustrate the contributions. For a more complete result, see Fig. \ref{fig:pairplot}.
 
-Baseline pairs that have smaller optimal time delays
+As is immediately clear from the figure, baseline pairs that have smaller optimal time offsets
 tend to have higher correlations. In other words, correlation peaks
 that are closer to zero time offset are higher. This is expected for two reasons;
-number one is that the longer the time delay that produces redundancy, the more the sky has moved with respect
-to the beams and hence the less overlaps in patch of sky surveyed. The other reason is that smaller optimal
-time-offset corresponds to smaller differences in orientation and length of the pair of baselines and hence better redundancy.
+one is that the longer the time delay that produces redundancy, the more the sky has moved with respect
+to the beams and hence the smaller the overlapping patch of sky surveyed. The other reason is that smaller optimal
+time offset corresponds to smaller differences in orientation and length of the pair of baselines and hence better redundancy.
 
@@
 baseline pairs, we have to account for the multiplicities of
-these baseline classes, in other words the number of antenna pairs with the
+these baseline classes or, in other words, the number of antenna pairs with the
 same length and orientation.
-We would like to estimate an effective weight $\widetilde{\Theta}$ that accounts for both the peak height $\Theta$ and the multiplicities of the baseline classes.
+We would like to estimate an effective weight $\widetilde{\Theta}_{\boldsymbol{bb'}}$ that accounts for both the peak height $\Theta_{\boldsymbol{bb'}}$ and the multiplicities of the baseline classes.
 From Fig.
@@
 we see for example \{1,0\} has higher multiplicity than \{2,0\},
-or \{1,1\}. The latest release of PAPER64 data uses the PAPER128-equivalent baselines \{2,1\},
-\{2,0\} and \{2,-1\} \citep{Ali2015}. There, the three sets of equivalent baseline classes
-are only cross multiplied with itself. Assuming that each equivalent baseline delivers
-the same $\Theta$, the relative contribution to sensitivity can be estimated.
+or \{1,1\}. Assuming that each equivalent baseline delivers
+the same $\Theta_{\boldsymbol{bb}}$, the relative contribution to sensitivity can be estimated.
 
@@
 \begin{equation}
-\widetilde{\Theta}_{bb'} \propto \Theta_{bb'}\times\sqrt{MM'}.\label{eq:sensul}
+\widetilde{\Theta_{\boldsymbol{bb'}}} \propto \Theta_{\boldsymbol{bb'}}\sqrt{MM'}.\label{eq:sensul}
 \end{equation}
@@
 Points that have zero time delay are the equivalent baseline pairs and their weighted correlation values simply reflect the multiplicity factor.
-For clarity of presentation we have ``folded over'' the negative time delays and combined baseline pairs that are identical modulus parity.
-We point out that the data points shown here are not all the cases of highest correlation.
+For clarity of presentation we have ``folded over'' the negative time delays and combined baseline pairs that are mirror images.
 
-With $\widetilde{\Theta}$, we can estimate the power spectrum by inverse covariance weighting:
+With $\widetilde{\Theta}_{\boldsymbol{bb'}}$, we can estimate the power spectrum by inverse covariance weighting.
+We again assume that the power spectrum $P$ varies negligibly over the $k$-space of interest, so that measurements at different $k_{\bar{b}, \tau}$ can be combined to obtain an estimator at some $k_\tau$:
 \begin{equation}
 \begin{aligned}
- P(k_{\tau}) &= \frac{\sum_{bb'}P(k_{b,\tau})/\sigma_P^2(bb')}{\sum_{bb'}\sigma_P^2(bb')}, \\
- &= \frac{\sum_{bb'}P(k_{b,\tau})\widetilde{\Theta}_{bb'}^2}{\sum_{bb'}\widetilde{\Theta}_{bb'}^2},
+ \hat{P}(k_{\tau}) &= \frac{\sum_{\boldsymbol{bb'}}\hat{P}(k_{\bar{b}, \tau})/\sigma_P^2(\boldsymbol{bb'})}{\sum_{\boldsymbol{bb'}}1/\sigma_P^2(\boldsymbol{bb'})}, \\
+ &= \frac{\sum_{\boldsymbol{bb'}}\hat{P}(k_{\bar{b}, \tau})\widetilde{\Theta}_{\boldsymbol{bb'}}^2}{\sum_{\boldsymbol{bb'}}\widetilde{\Theta}_{\boldsymbol{bb'}}^2},
  \end{aligned}
 \end{equation}
-where the sum is over classes of baseline pairs. 
-We define the estimator sensitivity to be the inverse of the power spectrum noise variance:
+where the sum is over classes of baseline pairs, and the power spectrum noise covariance $\sigma_P^2$ is proportional to the effective weight $\widetilde{\Theta}$.
+
+We define the estimator sensitivity to be the inverse of the total power spectrum noise variance $\Sigma_P^2$:
 \begin{equation}
-\rho \propto \frac{1}{\sigma_P^2} \propto \rho_0^2\sum_{bb'}^N\widetilde{\Theta}^2_{bb'},
+\rho \propto 1/\Sigma_P^2 \propto \rho_0^2\sum_{\boldsymbol{bb'}}^N\widetilde{\Theta}^2_{\boldsymbol{bb'}},
 \end{equation}
@@
 \label{eq:tildereal}
-\widetilde{\Theta}_{bb'}=\frac{\Theta_{bb'}\sqrt{M_bM_{b'}}}{\sqrt{1 + \rho_0 \left(M_b+M_{b'} \right)}}.
+\widetilde{\Theta}_{\boldsymbol{bb'}} =\frac{\Theta_{\boldsymbol{bb'}}\sqrt{M_bM_{b'}}}{\sqrt{1 + \rho_0 \left(M_b+M_{b'} \right)}}.
 \end{equation}
@@
 \eqref{eq:tildereal} quantifies the relative sensitivity contribution of a baseline pair $bb'$.
-Assuming a reionization signal of $\Delta_{21cm}^2\sim 30mK^2$, observation centered at 150MHz ($z=8.5$), and 120 days of integration with PAPER antennas, we have roughly
+Assuming a reionization signal of $\sim 30\,\text{mK}$, observation centered at 150\,MHz ($z=8.5$), and 120 days of integration with PAPER antennas, we have roughly
 (See Eq.(20) in \cite{first-paper})
 \begin{equation}
-\rho_0 \sim 0.001\left[\frac{L}{40m}\right] \left[\frac{0.1hMpc^{-1}}{k}\right]^3, 
+\rho_0 \sim 0.001\left[\frac{L}{40\,\text{m}}\right] \left[\frac{0.1h\,\text{Mpc}^{-1}}{k}\right]^3, 
 \end{equation}
-where $L$ is the average baseline length between the pair. As expected, baseline-pairs that have smaller $\widetilde{\Theta}$ contribute less to the sensitivity.
+where $\bar{b}$ is the average baseline length between the pair.
 
@@
 The motivations behind the designs are explained in \cite{HERAconfiguration}.
-Compared to PAPER128, the hexagonal pattern of HERA lack short baselines oriented close to each other, and the smaller beam means that we expect to see only longer nearly equivalent baselines.
+Compared to PAPER128, the hexagonal pattern of HERA lack short baselines oriented close to each other, and thus we expect to see only longer nearly equivalent baselines with high correlations.
 The lower multiplicities per class of baselines is compensated by the larger number of classes of baseline-pairs, especially given the gap in the larger versions.
@@
 \caption{Pairplots of the top contributing baseline pairs in three arrays.
-Plotted properties are optimal time delay $dT$,  effective weight $\widetilde{\Theta}$ and baseline length $L$.
-Only those points with $\widetilde{\Theta}>0.01$ are shown (the weight for the top class of equivalent pair is normalized to 1).
+Plotted properties are optimal time delay $\Delta T_{\boldsymbol{bb'}}$,  effective weight $\widetilde{\Theta}_{\boldsymbol{bb'}}$ and baseline length $\bar{b}$.
+Only those points with $\widetilde{\Theta}_{\boldsymbol{bb'}}>0.01$ are shown (the weight for the top class of equivalent pair is normalized to 1).
 Only one baseline length is shown since all top contributing pairs have very similar lengths as expected.
@@
 In Fig.
-\ref{fig:pairplot} we present pair-distributions of 3 different properties for baselines that contribute well to the sensitivity ($\widetilde{\Theta}>0.01$, where again the weight for the top class of equivalent pair is normalized to 1).
-The properties shown are effective weight $\widetilde{\Theta}$, optimal time offset $dT$, and average baseline length $L$.
-We only show 3 of the mentioned arrays for visual clarity.
-The other two results are similar barring intuitive differences.
-We discuss each cross-distribution plot:
+\ref{fig:pairplot} we present pair-distributions of three different properties for baselines that contribute well to the sensitivity ($\widetilde{\Theta}_{\boldsymbol{bb'}}>0.01$, where again the weight for the top class of equivalent pair is normalized to 1).
+The properties shown are effective weight $\widetilde{\Theta}_{\boldsymbol{bb'}}$, optimal time offset $\Delta T_{\boldsymbol{bb'}}$, and average baseline length $\bar{b}$.
+We only show three of the mentioned arrays for visual clarity.
+The other two results are qualitatively similar.
+We point out some key features:
 \begin{itemize}
 
-\item $dT$ vs. $\widetilde{\Theta}$:
-This relation  is familiar from Fig. \ref{fig:sensplot}. The points at $dT=0$ are the equivalent baselines. Note PAPER128 has more data points with high $dT$.
+\item $\Delta T_{\boldsymbol{bb'}}$ vs. $\widetilde{\Theta}_{\boldsymbol{bb'}}$:
+This relation  is familiar from Fig.
+\ref{fig:sensplot}.
+The points at $\Delta T=0$ are the equivalent baselines.
+Note HERA arrays have fewer data points with high $\Delta T_{\boldsymbol{bb'}}$, due to their comparatively narrow beam.
 
-\item $dT$ vs. $L$:
-PAPER128 shows the trend that longer baselines correspond to lower $dT$.
+\item $\Delta T_{\boldsymbol{bb'}}$ vs. $\bar{b}$:
+PAPER128 shows the trend that longer baselines correspond to lower $\Delta T_{\boldsymbol{bb'}}$.
 HERA arrays do not exhibit this trend here because of a selection effect.
-Only baseline pairs with $\widetilde{\Theta}>0.01$ are shown.
-Shorter HERA baselines require much longer $dT$ to overlap, partly because of the hexagonal structure requires $60\degree$ of rotation to overlap baselines, partly because of the smaller primary beam.
+Only baseline pairs with $\widetilde{\Theta}_{\boldsymbol{bb'}}>0.01$ are shown.
+Shorter HERA baselines require much longer $\Delta T$ to overlap, partly because of the hexagonal structure requires $60\degree$ of rotation to overlap baselines, partly because of the smaller primary beam.
 Thus most short HERA baseline pairs are not shown in this figure.
 
-\item $\widetilde{\Theta}$ vs. $L$:
-The general trend to be seen here is that longer baselines tend to have lower $\widetilde{\Theta}$.
+\item $\widetilde{\Theta}_{\boldsymbol{bb'}}$ vs. $\bar{b}$:
+The general trend to be seen here is that longer baselines tend to have lower $\widetilde{\Theta}_{\boldsymbol{bb'}}$.
 This is due to the lower multiplicity of longer baselines.
 The trend is particularly obvious in the linear structure for the HERA arrays, which are the equivalent baselines.
-Since they all have the unity $\Theta$ by definition, the linearly decreasing trend of $\widetilde{\Theta}$ is a direct measure of the baseline multiplicity structures of the HERA array configurations.
+Since they all have the unity $\Theta_{\boldsymbol{bb'}}$ by definition, the linearly decreasing trend of $\widetilde{\Theta}_{\boldsymbol{bb'}}$ is a direct measure of the baseline multiplicity structures of the HERA array configurations.
 
-\item The top nearly equivalent pairs in HERA the same $\Theta$ (not shown) as in PAPER128, but much lower $\widetilde{\Theta}$.
+\item The top sensitivity-contributing near-equivalent pair in each array is normalized to the same $\Theta_{\boldsymbol{bb'}}$ (not shown), but those of HERA arrays have much lower $\widetilde{\Theta}_{\boldsymbol{bb'}}$ than the top pair of PAPER128.
 This is because they are longer baselines with lower multiplicity.
@@
 
+Clearly not all baseline pairs give significant contribution to sensitivity.
+Due to the large number of baseline pairs, it may not be computationally feasible to include all pairs.
 Having quantified the sensitivity from a given pair of baselines, we study the cumulative sensitivity of the array depending on which baseline pairs we include.
-Evidently we should prefer the pairs with larger $\widetilde{\Theta}$.
+Evidently we should prefer the pairs with larger $\widetilde{\Theta}_{\boldsymbol{bb'}}$.
 In Fig.
-\ref{fig:osens} we plot $\rho$ against the minimum $\widetilde{\Theta}$.
-$\rho(\widetilde{\Theta}_{min})$ is the sensitivity of the array when baseline-pairs that have $\widetilde{\Theta}>\widetilde{\Theta}_{min}$ are included.
+\ref{fig:osens} we plot $\rho$ against the minimum $\widetilde{\Theta}_{\boldsymbol{bb'}}$.
+$\rho(\widetilde{\Theta}_{min})$ is the sensitivity of the array when baseline-pairs that have $\widetilde{\Theta}_{\boldsymbol{bb'}}>\widetilde{\Theta}_{min}$ are included.
 The dashed lines represent the values when only the equivalent baseline-pairs are used.
@@
 This plot therefore does not compare the absolute sensitivity across the different arrays.
-The stepwise pattern is characteristic of a regular grid; as we step to lower $\widetilde{\Theta}$ large groups of baseline pair classes get included in ``batch''.
+The stepwise pattern is characteristic of a regular grid; as we step to lower $\widetilde{\Theta}_{\boldsymbol{bb'}}$ large groups of baseline pair classes get included in ``batch''.
 %\begin{centered}
@@
 Dashed lines represent when only equivalent baseline pairs are used, while solid lines indicate use of both the equivalent and nearly equivalent baselines are used.
-The y-axis is normalized independently for each array such that the contribution of the top single equivalent pair class is unity, and thus does not indicate a comparison of absolute sensitivity across the different arrays shown.
+The vertical axis is normalized independently for each array such that the contribution of the top single equivalent pair class is unity, and thus does not indicate a comparison of absolute sensitivity across the different arrays shown.
 }
@@
 \ref{fig:numerics}.
-Foreground exhibits long-range %correlations and would lead to high responses at wide range of time-offsets.
+Foreground exhibits long-range %correlations and would lead to high responses at wide range of time offsets.
 %Instrumental noise for different baselines/antenna-pairs, on the other hand, are uncorrelated and thus exhibit responses consistent with zero at all time offsets.
@@
 Our discussion focuses on drift-scan arrays but also provide the analogous expressions for tracking measurements.
-Our method relies on cross-multiplications of visibilities at a time-offset.
-For drift-scan arrays, the movement of zenith during this time-offset requires a frequency-dependent rephasing prior to delay transform.
+Our method relies on cross-multiplications of visibilities at a time offset.
+For drift-scan arrays, the movement of zenith during this time offset requires a frequency-dependent rephasing prior to delay transform.
 We show that such zenith-rephasing lead to manageable chromatic leakage.
-Given an antenna array configuration, our method identifies the best baseline pairs to cross-multiply and predict the optimal time-offset $\Delta T$, weight $\Theta$ and rephase delay $\Delta\tau$.
+Given an antenna array configuration, our method identifies the best baseline pairs to cross-multiply and predict the optimal time offset $\Delta T_{\boldsymbol{bb'}}$, weight $\Theta_{\boldsymbol{bb'}}$ and rephase delay $\Delta\tau_{\boldsymbol{bb'}}$.
 With the predicted results one can incorporate partial redundancy into existing delay-tranform based power-spectrum pipelines through a few steps.
 1).
-Rephase the visibilities prior to delay transforming by $\Delta\tau$.
+Rephase the visibilities prior to delay transforming by $\Delta\tau_{\boldsymbol{bb'}}$.
 2).
@@
 In this section we link our results from Section \ref{sec:method} to the traditional views of rotation synthesis and in particular $uv$ tracks.
-Typical convention of rotation synthesis has rotation operates on $\boldsymbol{b'}$ instead of $\hat{\boldsymbol{s}}$, in which case Eq. \eqref{eq:Theta} becomes
+in describing the relative motion of the baseline relative to the sky, we have in Section \ref{sec:method} rotated the direction coordinate $\hat{s}$ with $\boldsymbol{\Gamma}$.
+The typical convention of rotation synthesis has rotation operates on $\boldsymbol{b'}$ instead of $\hat{\boldsymbol{s}}$, in which case Eq. \eqref{eq:Theta} takes the equivalent form
 \begin{equation}
-\Theta \equiv\int\frac{d\Omega d\nu}{X^{2}Y}B^{*}(\hat{\boldsymbol{s}},\nu)B(\boldsymbol{\Gamma}\hat{\boldsymbol{s}},\nu) e^{-i2\pi\frac{\nu}{c}\hat{\boldsymbol{s}}\cdot\left(\boldsymbol{b}-\widetilde{\boldsymbol{\Gamma}}\boldsymbol{b'}\right)},
+\Theta \equiv\int\frac{d\Omega d\nu}{X^{2}Y}B^{*}(\hat{\boldsymbol{s}},\nu)B(\boldsymbol{\Gamma}\hat{\boldsymbol{s}},\nu) e^{-i2\pi\frac{\nu}{c}\hat{\boldsymbol{s}}\cdot\left(\boldsymbol{b}-\boldsymbol{\Gamma}^{-1}\boldsymbol{b'}\right)},
 \label{eq:drift}
 \end{equation}
-where $\widetilde{\boldsymbol{\Gamma}}$ is the inverse of $\boldsymbol{\Gamma}$. 
-For tracking elements, $\boldsymbol{\Gamma}$ becomes a rotation around zenith:
+where $\boldsymbol{\Gamma}^{-1}$ is the inverse of $\boldsymbol{\Gamma}$. 
+And for tracking elements we have analogously:
 \begin{equation}
-\Theta \equiv\int\frac{d\Omega d\nu}{X^{2}Y}B^{*}(\hat{\boldsymbol{s}},\nu)B(\hat{\boldsymbol{s}},\nu) e^{-i2\pi\frac{\nu}{c}\hat{\boldsymbol{s}}\cdot\left(\boldsymbol{b}-\widetilde{\boldsymbol{\Gamma}}\boldsymbol{b'}\right)},
+\Theta \equiv\int\frac{d\Omega d\nu}{X^{2}Y}B^{*}(\hat{\boldsymbol{s}},\nu)B(\hat{\boldsymbol{s}},\nu) e^{-i2\pi\frac{\nu}{c}\hat{\boldsymbol{s}}\cdot\left(\boldsymbol{b}-\boldsymbol{\Gamma}^{-1}\boldsymbol{b'}\right)},
 \label{eq:tracking}
@@
 \[
- \boldsymbol{b}-\widetilde{\boldsymbol{\Gamma}}\boldsymbol{b'} = 0. \label{eq:trackcross}
+ \boldsymbol{b}-\boldsymbol{\Gamma}^{-1}\boldsymbol{b'} = 0. \label{eq:trackcross}
  \]
-If exponent vanishes, $\Theta_{\nu}$ is real, and one would not observe any non-zero phase at the peak.
-However, with inclusion of the $w$-term, we see that track crossings in $uv$-plane do not actually imply crossing in the $uvw$ space, therefore the exponent Eq.
+If exponent vanishes, $\Theta_{\nu}(\Delta T)$ is real, and one would not observe any non-zero phase at its peak value.
+However, with the inclusion of the $w$-term, we see that track crossings in $uv$-plane do not actually imply crossing in the $uvw$ space, and therefore the exponent or, equivalently, Eq.
 \eqref{eq:trackcross} does not vanish and we observe non-zero, frequency dependent peak phases even in the case of tracking arrays, as shown in Fig.
@@
 Only in the special case of a tracking measurement, where the $w$-term happens to be negligible, do $uv$-track crossings correspond exactly to maxima of correlations.
-Incidentally, at the crossing points the exponent in Eq.
+In this case, at the crossing points the exponent in Eq.
 \eqref{eq:trackcross} vanishes and $\psi_\nu$ is zero for all $\nu$.
-In general, second order effects due to the presence of $w$ is observed for the tracking measurements as in Fig.
+In general, second-order effects due to the presence of $w$ is observed for the tracking measurements as in Fig.
 \ref{fig:phi_nu}.
@@
 \label{sec:appB}
-In this appendix we give a brief derivation of the effective weight quoted in \ref{sec:sensitivity}.
+In this appendix we give a brief derivation of the effective weight $\widetilde{\Theta}$ quoted in Section \ref{sec:sensitivity}.
 We combine the power spectrum measurements from distinct baseline classese by inverse variance weighting.
@@
 In fact, for simplicity and without loss of generality we shall treat the visibility as a real quantity in the rest of this derivation.
-Note that though we can assume $\langle V_N^{odd-power}\rangle=0$, the same is not true for $P_N$.
+Note that though we can assume $\langle V_N^{m}\rangle=0$, where $m$ is odd. The same is not true for $P_N$.
 
@@
 &\propto \langle \frac{(V_{1S}+V_{1N})^2 (V_{2S}+V_{2N})^2}{\Theta^2} \rangle - \langle \frac{(V_{1S}+V_{1N}) (V_{2S}+V_{2N})}{\Theta} \rangle ^2,\\
-&= \frac{1}{\Theta^2} \left( V_{1S}^2\sigma_{V2}^2+V_{2S}^2\sigma_{V1}^2+\langle V_{1N}^2 V_{2N}^2\rangle\right), \\
-&= \frac{1}{\Theta^2} \left[ V_{S}^2(\sigma_{V2}^2+\sigma_{V1}^2) + \sigma_{V1}^2 \sigma_{V2}^2\right], 
+&= \frac{1}{\Theta_{12}^2} \left( V_{1S}^2\sigma_{V2}^2+V_{2S}^2\sigma_{V1}^2+\langle V_{1N}^2 V_{2N}^2\rangle\right), \\
+&= \frac{1}{\Theta_{12}^2} \left[ V_{S}^2(\sigma_{V2}^2+\sigma_{V1}^2) + \sigma_{V1}^2 \sigma_{V2}^2\right], 
 \end{aligned}
@@
 \begin{aligned}
-\sigma_P^2 & \propto  \frac{\sigma_0^4}{\Theta^2} \left[ \rho_0 \left(\frac{1}{M_1}+\frac{1}{M_2} \right) + \frac{1}{M_1 M_2}\right], \\
+\sigma_P^2 & \propto  \frac{\sigma_0^4}{\Theta_{12}^2} \left[ \rho_0 \left(\frac{1}{M_1}+\frac{1}{M_2} \right) + \frac{1}{M_1 M_2}\right], \\
 &\propto \frac{1}{\widetilde{\Theta}_{12}^2}.

### 2017-07-15 Rotation paper/draft_working.tex
saving progress

@@
 \documentclass[twocolumn,apj,numberedappendix]{emulateapj}
-\shorttitle{A Two-Baseline 21\,cm Power Spectrum Estimator}
+\shorttitle{Unlocking Sensitivity for 21\,cm Power Spectrum Estimators}
 \shortauthors{Zhang, Liu \& Parsons}
@@
 
-\title{Unlocking Sensitivity with A Two-baseline Power Spectrum Estimator for 21\,cm Inteferometric Probes of Reionization }
+\title{Unlocking Sensitivity for Visibility-based Power Spectrum Estimators of the 21\,cm Reionization Signal}
 
@@
 \begin{abstract}
-Measuring the cosmological 21\,cm power spectrum with radio interferometers requires high sensitivity.
-Many modern low-frequency interferometers are thus designed with antennas placed on regular grid to maximize redundancy.
-Current visibility-based power spectrum pipelines, though shown to ease control of systematics, lack the ability to include partial redundancy.
-We introduce a method to include partial redundancy in such power spectrum pipelines of drift-scan arrays.
-Our method pairs baselines to cross-multiply at a time lag, and quantifies the sensitivity contributions of each pair of baselines.
-Using the 128-element configurations and beams of the Donald C.
-Backer Precision Array for Probing the Epoch of Reionization (PAPER-128), and 4 planned versions of the Hydrogen Epoch of Reionization Array (HERA), we illustrate how our method applies to different arrays and predict the sensitivity improvements of including each baseline pairs.
-We show that inclusion of partially redundant baselines
-would account for $20\%$ to $60\%$ of the sensitivity of PAPER-128 and different configurations of HERA. 
+Radio interferometers seeking to measure the cosmological 21\,cm power spectrum require high sensitivity.
+Several modern low-frequency interferometers feature antennas placed on regular grid to maximize redundancy.
+Current visibility-based power spectrum pipelines, though shown to ease control of systematics, lack the ability to make use of partial redundancy.
+We introduce a method to leverage partial redundancy in such power spectrum pipelines for drift-scan arrays.
+Our method cross-multiplies baseline pairs at a time lag, and quantifies the sensitivity contributions of each pair of baselines.
+Using the configurations and beams of the 128-element Donald C.
+Backer Precision Array for Probing the Epoch of Reionization (PAPER-128), and four staged deployments of the Hydrogen Epoch of Reionization Array (HERA), we illustrate how our method applies to different arrays and predict the sensitivity improvements associated with each baseline pairs.
+We show that for upcoming arrays with increasing number of antennas, partial redundancy will be of increasing importance in unlocking full sensitivities of the arrays.
 \end{abstract}
@@
 of its own, but also as a gateway to crucial information
-regarding fundamental physics of inflation, neutrino mass and phenomenology
+regarding fundamental physics of inflation, neutrino mass, and the phenomenology
 of the first stars and galaxies (e.g. \citealt{LiuOpticalDepth, Liu2016b, Mao2008, DEw21cm, Bull2015, Oyama20131186}).
 
-Observational studies of reionization, including Gunn-Peterson measurements of quasi-stellar objects \citep{Fan2006} and cosmic microwave background temperature and anisotropy measurements (CMB;\citealt{Planck2016}), the Kinetic Sunyaev-Zeldo'vich effect \citep{kszpatchy} and Lyman alpha emitter clustering \citep{mcquinnLyA} have given us indications of the rough time-frame of reionization, but only limited constraints on the finer spatial and temporal structures.
+Observational studies of reionization, including Gunn-Peterson measurements of quasi-stellar objects \citep{Fan2006} and Cosmic Microwave Background temperature and anisotropy measurements (CMB; \citealt{Planck2016}), the Kinetic Sunyaev-Zeldo'vich effect \citep{kszpatchy} and Lyman alpha galaxy observations \citep{mcquinnLyA} have given us indications of the rough time-frame of reionization, but only limited constraints on the finer spatial and temporal structures.
 A surge of recent radio-astronomical experiments of reionization focus on measuring the ``spin-flip'' transition of neutral
-hydrogen of characteristic wavelength 21\,cm \citep{Furlanetto2006181,PritchardLoeb}.
-The 21\,cm brightness temperature is a direct tracer of neutral hydrogen through the epoch of reionization, and thus tomography of the 21\,cm line is a direct measurement of the full temporal and spatial variations of this event.
-Before realizing of full-scale 21\,cm tomography, many current radio interferometric efforts
+hydrogen with a characteristic wavelength of 21\,cm \citep{Furlanetto2006181,PritchardLoeb}.
+The 21\,cm brightness temperature is a direct tracer of neutral hydrogen through the epoch of reionization, and thus by measuring the three dimensional distribution of the 21\,cm signal we measure the full temporal and spatial variations of this event.
+However, before realizing full-scale 21\,cm tomography, many current radio interferometric efforts
 aim to measure the spatial power spectrum of 21\,cm brightness temperature fluctuations.
@@
 The highly redshifted 21\,cm signal is faint and diffuse, in contrast to the localized bright sources targeted by many traditional radio telescopes.
-Current 21\,cm experiments are sensitivity-starved, with the sensitivity requirement further increased considering the foreground contaminations 5 orders of magnitude brighter than the cosmological signal of interest.
+Current 21\,cm experiments are sensitivity-starved, with the sensitivity requirement further increased considering the foreground contaminations five orders of magnitude brighter than the cosmological signal of interest.
 Low-frequency radio interferometers aiming to measure the 21\,cm signal are thus designed differently from traditional instruments.
-Specially designed arrays such as PAPER and HERA feature multiple copies of the same baselines to repeatedly measure the same Fourier signal to increase sensitivity \citep{first-paper}.
-To satisfy the sensitivity need, modern arrays are large.
+To satisfy the sensitivity needs, modern arrays are large.
 Ranging upwards from 100 elements, these arrays are typically of the drift-scan type to limit cost.
+To further improve sensitivity, specially designed arrays such as PAPER and HERA feature multiple copies of the same baselines to repeatedly measure the same Fourier signal\citep{first-paper}.
 
@@
 Baselines of different lengths and orientations contain partially redundant information.
-While imaging based power-spectrum pipelines naturally includes all redundancy information, visibility-based pipelines so far only cross multiply fully redundant baselines, i.e.
+While imaging based power-spectrum pipelines naturally include all redundancy information, visibility-based pipelines so far only cross multiply fully redundant baselines, i.e.
 baselines of the same length and orientation.
 
-Visibility-based pipelines are not fundamentally limited from using partially redundant baselines.
+Visibility-based pipelines are not fundamentally limited to using only fully redundant baselines.
 In fact, most sensitivity forecasts to date do include partial redundancy \citep{Pobersens, HERA, JoshAntPos}.
@@
 Arrays capable of such tracking include MWA and LOFAR.
-Our work introduces a two-baseline power spectrum estimator that can be applied to both tracking capable arrays and drift-scan only telescopes, such as PAPER, HERA and potentially SKA-low.
+Our work introduces a two-baseline power spectrum estimator that can be applied to both tracking capable arrays and drift-scan only telescopes, such as PAPER, HERA, and potentially SKA-low.
 
+The basic idea of our proposed visibility-based, two-baseline power spectrum estimator is as follows.
 The Earth's rotation causes the baselines in a drift-scan array to pick up different modes of the sky with time.
@@
 
-The rest of this paper is organized as follows. In section 2 we introduce some terminology and notation used in the rest of the paper. In section
-3 we introduce the formalism for weighting partially redundant baselines.
-In section 4 we present numerical tests of
+The rest of this paper is organized as follows. In Section 2 we introduce some terminology and notation used in the rest of the paper. In Section
+3 we introduce the formalism for weighting partially redundant baselines and derive the power spectrum estimator.
+In Section 4 we present numerical tests of
 this technique as well as the expected sensitivity improvement
- this method provides for HERA and PAPER-128 pipelines, and with section 5 we conclude.
+ this method provides for HERA and PAPER-128 pipelines. With Section 5 we conclude.
 
@@
 times. Non-equivalent baselines also have partial redundancy, and the redundancy can be maximized if their
-respected time series are shifted with respect to one another by some delay.
+respected time series are shifted with respect to one another by some time offset.
 
@@
 
-We shall use the 128-element PAPER array (henceforth as PAPER-128) to motivate our formalism and demonstrate our method, and extend our results to several HERA configurations in Sec \ref{sec:arrconf}.
+We shall use the 128-element PAPER array (henceforth referred to as PAPER-128) to motivate our formalism and demonstrate our method, and extend our results to several HERA configurations in Section \ref{sec:arrconf}.
 The PAPER array is located in the Karoo desert in South Africa (30:43:17.5
-S, 21:25:41.8 E). The layout pattern with antenna labels is shown
+S, 21:25:41.8 E). The layout pattern is shown
 in Fig.
 \ref{fig:AntPos}.
-The array consists of a 112-element core is a rectangular grid, and 16 "outriggers" used primarily to aid calibration.
+The array consists of a 112-element core in a rectangular grid, and 16 ``outriggers'' used primarily to aid calibration.
 In the bottom panel, the two baselines marked in red, are an example of equivalent pair.
 We denote a equivalency-class of baselines in the PAPER grid by their separations, in this case  \{2,0\}, for the
-antennas are separated by 1 unit east and 0 unit north. Similarly,
+antennas are separated by 2 units east and 0 units north. Similarly,
 the baselines marked in blue and yellow are respectively examples
 of \{2,1\} and \{2, -1\}.
-Note \{2,0\} and \{-2, 0\} for example are the same class and should
-not be counted twice. Antennas in purely north-south baselines
+Note that \{2,0\} and \{-2, 0\} for example are the same class of baselines and should not be counted twice. Antennas in purely north-south baselines
 are close (4m), and hence these baselines are not suitable
-for cross-multiplication due to cross-coupling.
-On the other hand, the close North-South separation means that classes such as \{2,0\} and  \{2,1\} are expected to be near-equivalent.
+for use due to cross-coupling.
+On the other hand, the small North-South separation means that classes such as \{2,0\} and  \{2,1\} are expected to be near-equivalent.
 The PAPER-64 analysis of \cite{Ali2015} used three classes of baselines, the PAPER-128
 equivalent of which are 
-\{2, 0\}, \{2, 1\} and \{2, -1\}. There each of these classes
+\{2, 0\}, \{2, 1\} and \{2, -1\}. There, each of these classes
 of baselines are cross multiplied within itself.
 This paper provides the method for inter-class multiplications.
-To do so we will use the short hand notation $\{m,n:m',n'\}$ to denote a pair of baseline classes to be cross-multiplied.
+We will use the short hand notation $\{m,n:m',n'\}$ to denote a pair of baseline classes to be cross-multiplied.
 
@@
 
-Radio interferometric observations are often described in the coordinates $uvw$ define as:
+Radio interferometric observations are often described in the coordinates $uvw$, defined as:
 \[
@@
 \]
-where $\boldsymbol{b}$ is the baseline vector in Cartesian coordinates, with first and second coordinates pointing East and North, respectively.
-$\nu$ is the frequency of observation, and $c$ is the speed of light.
+where $\boldsymbol{b}$ is the baseline vector in Cartesian coordinates, with first and second coordinates pointing East and North, respectively, and  $\nu$ is the frequency of observation, and $c$ is the speed of light.
 Relative to a phase center on the sky, each baseline maps to a point in $uvw$ space. As the Earth rotates, the points trace out tracks in the $uvw$ space. 
 We show in Fig.
-\ref{fig:Tracks} $uvw$ tracks of the 3 PAPER-128 baselines colored in Fig.
-\ref{fig:AntPos}, projected onto $uv$ and $vw$ planes.
-The tracks are traced over 12 sidereal hours, at 0.15\,GHz, relative to a phase center that passes through zenith.
+\ref{fig:Tracks} $uvw$ tracks of the three PAPER128 baselines colored in Fig.
+\ref{fig:AntPos}, projected onto the $uv$ and $vw$ planes.
+The tracks are traced over 12 sidereal hours, at 0.15\,GHz, relative to a phase center that pdrifts from zenith.
 
@@
 However, there are several reasons that $uv$ track-crossing do not imply perfect redundancy.
-In fact, in our case $uv$ track-crossings are
-not accurate enough for time offset determination, nor can it give estimate of the degree of redundancy.
 The most obvious reason is that the 3 dimensional $uvw$ tracks do not actually cross, as is evident from the $vw$ projection in the bottom panel of Fig.
 \ref{fig:Tracks}.
+Therefore, in our case $uv$ track-crossings are
+not accurate enough for time offset determination, nor can it give estimate of the degree of redundancy.
 For drift-scan arrays, even a hypothetical crossing in $uvw$ space would not imply perfect redundancy.
 This shall become evident in the next section, after we develop a more general formalism that accounts for the point spread function of the finite beams, and estimate the degree of redundancy for general combination of baselines at a general time offset.
-Then, relation of track-crossing and redundancy for both drift-scan and tracking measurements will be explored in Appendix \ref{sec:appA}.
+Relation between track-crossing and redundancy for both drift-scan and tracking measurements is further explored in Appendix \ref{sec:appA}.
 
@@
 In this section we formulate a power spectrum estimator from
-the product of delay-transformed visibilities from two arbitrary baselines.
+the product of delay-transformed visibilities from two arbitrary baselines. We shall work in the case of drift scan telescopes and extent the results to tracking measurements.
 
@@
 \end{equation}
-where $k_B$ is Boltzmann's constant, $\lambda$ is the mean wavelength, $\hat{\boldsymbol{s}}$ and $\Omega$ are a direction in the
+where $k_B$ is Boltzmann's constant, $\lambda$ is the mean wavelength, and $\hat{\boldsymbol{s}}$ and $d \Omega$ denote a direction in the
 sky and its corresponding solid angle.
-Inside the integral we have $\phi(\nu)$ the frequency bandpass profile, $A(\hat{\boldsymbol{s}},\nu)$ the (frequency-dependent) primary beam, and $I$ the specific intensity, which has been
+Inside the integral we have $\phi(\nu)$ as the frequency bandpass profile, $A(\hat{\boldsymbol{s}},\nu)$ as the (frequency-dependent) primary beam, and $I$ as the specific intensity, which has been
 related to $T$, the brightness temperature in the Rayleigh-Jeans
@@
 normalized to 1 at its peak (zenith), and we assume it to be the same
-for all baselines. %We show sample beams of HERA and PAPER antennas in Fig. \ref{fig:Beam} for reference.
+for all baselines.
+Power-spectrum measurements are typically taken from $\sim10$\,MHz centered around the corresponding redshift of interest (e.g.
+150\,MHz for redshift $z\approx9.5$), thus imposing a sharp drop-off on $\phi(\nu)$.
 
+ %We show sample beams of HERA and PAPER antennas in Fig. \ref{fig:Beam} for reference.
 %\begin{figure}[H]
@@
 
-Power-spectrum measurements are typically taken from $\sim10$\,MHz centered around the corresponding redshift of interest (e.g. 150 \,MHz for z=9.5).
-
-Delayed-transformed visibility has gained popularity in recent years since it has been shown that foregrounds are isolated in delay space.
-We define the delay-transformed visibility by Fourier transforming visibility along the frequency axis \citep{delay-transform} as:
+We work with delayed-transformed visibilities since it has been shown that foregrounds are isolated in delay space.
+We define the delay-transformed visibility by Fourier transforming visibility along the frequency axis \citep{delay-transform}, giving:
 \small
@@
 \begin{aligned}r & =\frac{c}{H_{0}}\int_{0}^{z}\frac{dz'}{E(z')},\\
- & \approx\frac{c}{H_{0}}\int_{0}^{z_{0}}\frac{dz'}{E(z')}-\frac{c(1+z)^{2}}{\nu_{21}H_{0}E(z)}\left(\nu-\nu_{0}\right),\\
+ & \approx\frac{c}{H_{0}}\int_{0}^{z_{0}}\frac{dz'}{E(z')}-\frac{c(1+z_0)^{2}}{\nu_{21}H_{0}E(z_0)}\left(\nu-\nu_{0}\right),\\
  & \equiv X-Y\Delta\nu,
@@
 \]
-where $\nu_{21}=1420$\,MHz is the 21\,cm transition rest frequency, $\nu_{0}$
-a reference central frequency with corresponding redshift $z_{0}$,
+where $r$ is the magnitude of the vector $\bold{r}$, $\nu_{21}=1420$\,MHz is the 21\,cm transition rest frequency, $\nu_{0}$ is
+a reference central frequency with corresponding redshift $z_{0}$, $H_0$ is the current-day Hubble constant,  
 and 
@@
 \begin{equation}
-\nu=\frac{X-r}{Y}+\nu_{21}.\label{eq:nur}
+\nu=\frac{X-r}{Y}+\nu_{0}.\label{eq:nur}
 \end{equation}
@@
 The beam pattern of a baseline shifts relative to the sky as the Earth rotates. Here we choose to fix the sky, and denote the rotated coordinates
-in the topocentric frame with the 3 dimensional rotation operator $\boldsymbol{\Gamma}$:
+in the topocentric frame with the three-dimensional rotation operator $\boldsymbol{\Gamma}$:
 
@@
 \]
-where we introduced a frequency-dependent phase $\psi_{\nu}$, which to first order corresponds to phasing the two visibilities to the same phase center.
-\footnote{Turns out for many cases, the first-order interpretation is largely sufficient, but we still keep the term to reserve the option to determine its exact form numerically.
+where we introduced a frequency-dependent phase $\psi_{\nu}$, which to linear-order in $\nu$ corresponds to phasing the two visibilities to the same phase center.
+\footnote{Turns out for many cases, the linear-order interpretation is largely sufficient, but we still keep the term to reserve the option to determine its exact form numerically.
 }
@@
 \begin{aligned} \langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle
- & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle\right)B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}),\\
+ & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}),\\
  & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\boldsymbol{\kappa})T(\boldsymbol{\kappa'})\rangle e^{-i(\boldsymbol{\kappa}\cdot \boldsymbol{r}-\boldsymbol{\kappa'}\cdot\boldsymbol{r'})}\right)B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}),\\
  & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\boldsymbol{\kappa}\cdot(\boldsymbol{r}-\boldsymbol{r'})}\right)B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}),\\
- & \approx P(k_{\bar{b}, \tau})\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2} \delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}),\\
- & = P(k_{\bar{b}, \tau})\int \frac{d^{3}r}{(X^2Y)^2}B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r})\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma} \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right],\\
- & = P(k_{\bar{b}, \tau})\int\frac{d\Omega d\nu}{X^{2}Y}B^{*}(\hat{\boldsymbol{s}},\nu)B(\boldsymbol{\Gamma} \hat{\boldsymbol{s}},\nu)\exp\left[-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma}\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right],
+ & \approx P(k_{\bar{b}, \tau})\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2} \delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'}) \Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}),\\
+ & = P(k_{\bar{b}, \tau})\int \frac{d^{3}r}{(X^2Y)^2}B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r})\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma} \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right],
 \end{aligned}
@@
 
-with
+where $\delta_D$ is the Dirac delta-function and
 \begin{equation}
-\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\equiv\exp\left[i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\boldsymbol{\Gamma}\hat{\boldsymbol{r'}}\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)-i\psi_{\nu}\right].
+\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\equiv\exp\left[i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\boldsymbol{\Gamma}\hat{\boldsymbol{r}}'\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)-i\psi_{\nu}\right].
 \end{equation}
@@
 
-Above $\delta_D$ is the Dirac delta-function and in transition from cosmological coordinates back to observing coordinates we have written $\hat{\boldsymbol{r}}\equiv\hat{\boldsymbol{s}}$.
 The third equality of Eq.(\ref{eq:main}) follows from assumption of translational invariance of statistics of the 21\,cm field. 
@@
 \[
-k_{\boldsymbol{b}, \tau} \equiv 2\pi \sqrt{\left(\frac{\tau}{Y}\right)^2 + \left(\frac{b}{\lambda X}\right)^2}. 
+k_{\boldsymbol{b}, \tau} \equiv 2\pi \sqrt{\left(\frac{\tau}{Y}\right)^2 + \left(\frac{\bar{b}}{\lambda X}\right)^2},
 \label{eq:kbtau}
 \]
-In Section \ref{sec:chromaticity} we will examine the exact form of the window function probed as well as its effect on foreground isolation.  Since $\boldsymbol{\Gamma}$
-is a sky rotation, it does not affect $\nu$, hence we have taken $\nu_{r}$
-outside the parenthesis in the phase term $\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma} \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)\right]$.
+where $\bar{b}$ is approximately the mean of $b$ and $b'$. 
+In Section \ref{sec:chromaticity} we will examine the exact form of the Fourier-space footprint probed as well as its effect on foreground isolation.  
+%Since $\boldsymbol{\Gamma}$ is a sky rotation, it does not affect $\nu$, hence we have taken $\nu_{r}$ outside the parenthesis in the phase term $\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma} \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)\right]$.
 Notice that the phase factor $\exp\left[-i2\pi\tau\left(\nu-\nu'\right)\right]$
-drops out in the end. This means that the location and magnitude of the correlation peak does not depend on delay $\tau$.
+drops out in the end.
 
 Since the beam pattern and bandpass are given in $\hat{\boldsymbol{s}}$
-and $\nu$, we convert the integral back to these coordinates to get
+and $\nu$, we convert the last line of Eq. \eqref{eq:main} back to these coordinates to get
 the general relation between the delay-transformed visibilities and
@@
 
-We can therefore form the power spectrum estimator for the baseline pair $\{\boldsymbol b, \boldsymbol  b'\}$:
+We can therefore form the power spectrum estimator for the baseline pair $\{\boldsymbol b: \boldsymbol  b'\}$:
 
@@
 Notice that $\Theta$ has no dependence on $\tau$.
-In Eq.
-\eqref{eq:Thetanu} we introduced an extra phase $\psi$, the origin of which we shall explain in \ref{sec:rephs}.
-We point out that although all our derivations focused on drift-scan telescopes, we can get the analogous result for tracking measurements simply by noticing that for a tracking primary beam, $\boldsymbol{\Gamma}$ becomes a rotation around zenith, and so:
+We point out that although all our derivations focused on drift-scan telescopes, we can get the analogous result for tracking measurements simply by noticing that for a tracking primary beam, and so by the radial-symmetry of the primary beam, we have:
 \[
@@
 transform of the cross multiplied beam patterns. As a check, when applied to equivalent baselines,
-$\boldsymbol{b}=\boldsymbol{b'}$, $\hat{\boldsymbol{s}}=\boldsymbol{\Gamma}\hat{\boldsymbol{s}}$, and Eq.(\ref{eq:final}) reduces to Eq.(B9) of \cite{paper32}. 
+$\boldsymbol{b}=\boldsymbol{b'}$, the peak correlation occurs at $\Delta T=0$ and $\hat{\boldsymbol{s}}=\boldsymbol{\Gamma}\hat{\boldsymbol{s}}$, in which case Eq.
+\eqref{eq:final} reduces to Eq.
+(B9) of \cite{paper32}.
 With Eq.
@@
 We can do all the above simply by computing the weight $\Theta$ from
-Eq.(\ref{eq:Theta}) for various time offsets.
+Eq. \eqref{eq:Theta} for various time offsets.
 
@@
 
-To first order, the physical origin of $\psi_\nu$ lies in the two visibilities having different phase centers.
+To linear order in $\nu$, $\psi_\nu$ cancels the decoherence due to the two visibilities having different phase centers.
 By default the correlators of a drift-scan array phase the two visibilities both to zenith at the same time.
-When they are cross-multiplied with a time lag, the visibilities must be rephased before delay transform to the account for the movement of the zenith.
+When they are cross-multiplied with a time lag, the visibilities must be rephased before delay transform to account for the movement of the phase center.
 The effect of the phase thus roughly corresponds in a shift of the delay mode measured, and we should expect $\psi_{\nu}$ to be in first order a linear function of $\nu$:
 \[
-\psi_{\nu}\sim 2\pi\Delta\tau\nu.
+\psi_{\nu}\sim 2\pi\Delta\tau\nu,
 \]
+where $\Delta\tau$ is a shift in delay corresponding to the phase center movement.
+
 In Fig.
-\ref{fig:phi_nu} we show the peak phases for for a given baseline pair of PAPER128 (\{2,0:2,1\}), comparing the drift-scan phase dependence with that of the same baseline with hypothetical tracking-elements (Eq.
+\ref{fig:phi_nu} we show the phases of $\Theta_\nu$ at the optimal time offset or, in other words, when $|\Theta_\nu|$ is maximized.
+The phases are shown for the baseline pair \{2,0:2,1\} of PAPER128.
+The drift-scan phase dependence is compared with that of the same baseline with hypothetical tracking-elements (Eq.
 \eqref{eq:Thetanu} and Eq.
 \eqref{eq:Thetanu_tracking}).
-In the drift-scan case, we indeed see a linear relation corresponding to a delay of $\Delta\tau\approx15.7$ns.
-In the tracking case, because of minimal zenith movement, only second order effects are observed.
-The origin of the second order effects can be seen as due to the $w$-term, or more precisely the fact that $uvw$ tracks do not cross when their 2 dimensional projections do.
+In the drift-scan case, we indeed see a linear relation corresponding to a delay of $\Delta\tau\approx15.7$\,ns.
+In the tracking case, since the phase center is fixed to the sky, only second order effects are observed.
+The origin of the second order effects can be seen as due to the $w$-term, or more precisely the fact that $uvw$ tracks do not cross when their two-dimensional projections do.
 We refer the reader to Appendix \ref{sec:appA} for further explanation.
@@
 \ref{fig:freqdiff} real parts of $\Theta_{\nu}$ of both equivalent and nearly equivalent baseline pairs for two channels: 0.145\,GHz and 0.155\,GHz.
-The top two panels have zero rephasing, and the bottom two are linearly rephased to a differential delay of $\Delta\tau\approx15.7$ns, as we determined above.
+The top two panels have zero rephasing, and the bottom two are linearly rephased to a differential delay of $\Delta\tau\approx15.7$\,ns, as we determined above.
 The first and third panels show the equivalent baseline pairs \{2,0:2,0\}, and second and fourth panels show \{2,0:2,1\}.
-Summing over frequency without proper rephasing leads to destructive interference and signal loss.
+Summing over frequency without proper rephasing leads to destructive interference and sensitivity loss.
 The wider the frequency profile, the more destructive the interference would be.
@@
 \ref{fig:freqdiff} that although the amplitude of correlations match up for all time offsets, the phase would only locally match.
-The need for rephasing can thus be understood as a symptom of the underlying spatial decoherence; coherence at the phase center would not extend to the entire beam pattern (Fig.
-\ref{fig:beamfringe}).
+The need for rephasing can thus be understood as a symptom of the underlying spatial decoherence; coherence at the phase center would not extend to the entire beam pattern (see Section \ref{sec:visual}).
 Thus while integrating over frequencies and sky-direction in Eq.
@@
 
-\caption{PAPER128 beam-fringe patterns and their point-spread functions on the $uv$-plane for a frequency of $150$\,MHz.
+\caption{PAPER-128 beam-fringe patterns and their point-spread functions on the $uv$-plane for a frequency of $150$\,MHz.
 Selected baselines have 30 meter East-West separation, and 0 and 4 meter North-South separation, respectively.
@@
 \subsection{ Sensitivity \label{sec:sensitivity}}
-In this section we discuss the sensitivity contributions of a variety of baseline pairs from PAPER128.
+In this section we discuss the sensitivity contributions of a variety of baseline pairs from PAPER-128.
 Recall the sensitivity contribution of a particular baseline pair is given by $\Theta$ (Eq.
 \eqref{eq:Theta}).
-From now one we use $\Theta_{\boldsymbol{bb'}}$ and $\Delta T_{\boldsymbol{bb'}}$ to denote the peak value and location of $\Theta$, or in other words the maximum of the redundancy weight $|\Theta(\Delta T)|$ and the corresponding offset for a given baseline pair $\boldsymbol{b}:\boldsymbol{b'}$.
+From now one we use $\Theta_{\boldsymbol{bb'}}$ and $\Delta T_{\boldsymbol{bb'}}$ to denote the peak value and location of $\Theta$, or in other words the maximum of the redundancy weight $|\Theta(\Delta T)|$ and the corresponding offset for a given baseline pair $\{\boldsymbol{b}:\boldsymbol{b'}\}$.
 %Intuitively we expect the sensitivity to depend on both the $uv$ coverage of the baseline and the patch of sky inside the beam.
@@
 \label{eq:tildereal}
-\widetilde{\Theta}_{\boldsymbol{bb'}} =\frac{\Theta_{\boldsymbol{bb'}}\sqrt{M_bM_{b'}}}{\sqrt{1 + \rho_0 \left(M_b+M_{b'} \right)}}.
+\widetilde{\Theta}_{\boldsymbol{bb'}} =\frac{\Theta_{\boldsymbol{bb'}}\sqrt{MM'}}{\sqrt{1 + \rho_0 \left(M+M' \right)}}.
 \end{equation}
@@
 For a given $\rho_0$ Eq.
-\eqref{eq:tildereal} quantifies the relative sensitivity contribution of a baseline pair $bb'$.
+\eqref{eq:tildereal} quantifies the relative sensitivity contribution of a baseline pair $\{\boldsymbol{b}:\boldsymbol{b'}\}$.
 Assuming a reionization signal of $\sim 30\,\text{mK}$, observation centered at 150\,MHz ($z=8.5$), and 120 days of integration with PAPER antennas, we have roughly
@@
 \subsection{Array Configuration Comparisons \label{sec:arrconf}}
-We run our algorithm over all possible baseline-pairs of  PAPER128, HERA37, HERA128, HERA243 and HERA350. The HERA antenna configurations are shown in Fig. \ref{fig:HeraAntpos}. The  
+We run our algorithm over all possible baseline-pairs of  PAPER-128, HERA37, HERA128, HERA243 and HERA350. The HERA antenna configurations are shown in Fig. \ref{fig:HeraAntpos}. The  
 hexagonal design is the densest pattern of antenna-packing.
@@
 The motivations behind the designs are explained in \cite{HERAconfiguration}.
-Compared to PAPER128, the hexagonal pattern of HERA lack short baselines oriented close to each other, and thus we expect to see only longer nearly equivalent baselines with high correlations.
+Compared to PAPER-128, the hexagonal pattern of HERA lack short baselines oriented close to each other, and thus we expect to see only longer nearly equivalent baselines with high correlations.
 The lower multiplicities per class of baselines is compensated by the larger number of classes of baseline-pairs, especially given the gap in the larger versions.
@@
 \item $\Delta T_{\boldsymbol{bb'}}$ vs. $\bar{b}$:
-PAPER128 shows the trend that longer baselines correspond to lower $\Delta T_{\boldsymbol{bb'}}$.
+PAPER-128 shows the trend that longer baselines correspond to lower $\Delta T_{\boldsymbol{bb'}}$.
 HERA arrays do not exhibit this trend here because of a selection effect.
@@
 
-\item The top sensitivity-contributing near-equivalent pair in each array is normalized to the same $\Theta_{\boldsymbol{bb'}}$ (not shown), but those of HERA arrays have much lower $\widetilde{\Theta}_{\boldsymbol{bb'}}$ than the top pair of PAPER128.
+\item The top sensitivity-contributing near-equivalent pair in each array is normalized to the same $\Theta_{\boldsymbol{bb'}}$ (not shown), but those of HERA arrays have much lower $\widetilde{\Theta}_{\boldsymbol{bb'}}$ than the top pair of PAPER-128.
 This is because they are longer baselines with lower multiplicity.
@@
 The maximum benefits for other cases are expected to be around $20\%$ to $60\%$.
-PAPER128 is designed with highly redundant nearly equivalent baselines, and thus  these baselines start contributing at higher $\widetilde{\Theta}_{min}$, but the gapped HERA configurations will benefit even more from nearly equivalent baselines at low $\widetilde{\Theta}_{min}$ due to the presence of more classes of such pairs.
+PAPER-128 is designed with highly redundant nearly equivalent baselines, and thus  these baselines start contributing at higher $\widetilde{\Theta}_{min}$, but the gapped HERA configurations will benefit even more from nearly equivalent baselines at low $\widetilde{\Theta}_{min}$ due to the presence of more classes of such pairs.
 Note that here we normalized $\rho$ such that the contribution of the top equivalent baseline pair, such as are 1.
@@
 \section{Conclusion}
-Redundant arrays are designed to maximize sensitivity.
-Current generations of redundant radio arrays, such as those probing the power spectrum of the epoch of reionization could benefit from data analysis techniques that improve the sensitivity.
-We present a visibility and delay-transform based method of cross-multiplying baselines that are close in length and orientation to each other, thereby extracting the sensitivity contained in such partial redundancy.
-Our discussion focuses on drift-scan arrays but also provide the analogous expressions for tracking measurements.
-Our method relies on cross-multiplications of visibilities at a time offset.
-For drift-scan arrays, the movement of zenith during this time offset requires a frequency-dependent rephasing prior to delay transform.
-We show that such zenith-rephasing lead to manageable chromatic leakage.
-Given an antenna array configuration, our method identifies the best baseline pairs to cross-multiply and predict the optimal time offset $\Delta T_{\boldsymbol{bb'}}$, weight $\Theta_{\boldsymbol{bb'}}$ and rephase delay $\Delta\tau_{\boldsymbol{bb'}}$.
-With the predicted results one can incorporate partial redundancy into existing delay-tranform based power-spectrum pipelines through a few steps.
-1).
-Rephase the visibilities prior to delay transforming by $\Delta\tau_{\boldsymbol{bb'}}$.
-2).
-Shift the visibilities in time by $\Delta T$.
-3) Cross multiply the visibilities of the two baselines  to form the power spectrum.
-4) combine the different baseline pairs by appropriate inverse-variance weighting that takes into account the predicted sensitivity contributions of each case.
-We showed that incorporation of partial redundancy accounts for $20\%$ to $60\%$ sensitivity for various configurations of PAPER and HERA.
+The 21\,cm brightness temperature is a direct tracer of reionization.
+Recently a surge of efforts in radio astronomy aims to measure the 21\,cm brightness temperature power spectrum of reionization.
+Measurement of the faint and diffuse signal requires high sensitivity, thus arrays are designed with multiple copies of the same baselines to enable repeated measurements of the same Fourier modes.
+
+Current visibility-based power spectrum pipelines have the advantage of easy-to-manage systematics, but have yet to unlock the full sensitivity of the arrays as they do not incorporate partially redundant baselines.
+We present a visibility and delay-transform based method to overcome this limitation by extracting the sensitivity contained in such partial redundancy.
+Applicable to both tracking and drift scan measurements, our method relies on cross-multiplications of visibilities that are offset in time.
+Given an antenna array configuration, our method identifies the best baseline pairs to cross-multiply and predicts the optimal time offset $\Delta T_{\boldsymbol{bb'}}$, weight $\Theta_{\boldsymbol{bb'}}$, and the necessary rephasing factor $\psi_\nu$.
+With the predicted results one can incorporate partial redundancy into existing delay-transform based power-spectrum pipelines.
+
+One benefit of delay-transform based power spectrum estimation is the ability to isolation foreground in Fourier space.
+We show that our proposed estimator does not lead to extra foreground leakage into the EoR window.
+We then apply our estimator to a variety array configurations.
+We show that for different configurations of PAPER and HERA, the incorporation of partial redundancy accounts for $20\%$ to $60\%$ of the array's sensitivity.
+The proposed power spectrum estimator retains benefits of delay-transform techniques for isolating foregrounds and avoiding systematics and enables one to unlock the full sensitivity of a redundant radio interferometer while simultaneously enjoying the benefits of delay-transform techniques for isolating foregrounds and avoiding systematics.
 

### 2017-07-15 Rotation paper/draft_working.tex
saving progress

@@
 On the other hand, the small North-South separation means that classes such as \{2,0\} and  \{2,1\} are expected to be near-equivalent.
-The PAPER-64 analysis of \cite{Ali2015} used three classes of baselines, the PAPER-128
+The PAPER-64 analysis of \cite{Ali2015} uses three classes of baselines, the PAPER-128
 equivalent of which are 
@@
 
-We work with delayed-transformed visibilities since it has been shown that foregrounds are isolated in delay space.
-We define the delay-transformed visibility by Fourier transforming visibility along the frequency axis \citep{delay-transform}, giving:
-\small
-\begin{equation}
-\begin{aligned}V(\boldsymbol{b},\tau) & =\int d\nu V_{\nu}(\boldsymbol{b})\phi(\nu)\exp\left(2\pi i\nu\tau\right),\\
- & =\int d\Omega d\nu B(\hat{\boldsymbol{s}},\nu)T(\hat{\boldsymbol{s}},\nu)\exp\left[-2\pi i\nu\left(\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}-\tau\right)\right]. 
-\end{aligned}
-\label{eq:Vb1}
-\end{equation}
-\normalsize
-Here the delay $\tau$ is the Fourier dual of $\nu$. Eq. \eqref{eq:Vb1} expresses the delay-transformed visibility as
-an integral over observational coordinates $\hat{\boldsymbol{s}}$ and $\nu$. For notational simplicity here and in the rest of this paper, we have defined the quantity
+Ultimately we would like to relate observed visibility to the power spectrum $P$, defined such that
 \[
-B(\hat{\boldsymbol{s}},\nu) \equiv \frac{2k_B}{\lambda^2}\phi(\nu)A(\hat{\boldsymbol{s}},\nu). 
-\label{eq:B}
+\langle \widetilde{T}^{*}(\boldsymbol{k})\widetilde{T}(\boldsymbol{k'})\rangle = \left(2\pi\right)^3\delta_D^3(\bold{k}-\bold{k'})P(k),
+\label{eq:Pofk}
 \]
-
-Ultimately,
-we would like to relate the data, collected with coordinates $\hat{\boldsymbol{s}}$
-and $\nu$, to the power spectrum, written with 
-$\boldsymbol{r}$ and $\boldsymbol{k}$, the cosmological position coordinate and wavenumber. To do so, we start by noticing that
+where $\bold{k}$ is the cosmological wavenumber, $\delta_D^3$ is the three-dimensional Dirac delta function, and $\widetilde{T}(\boldsymbol{k})$ is the three-dimensional Fourier transform of the brightness temperature field $T(\bold{r})$, with $\bold r$ as the cosmological position coordinate.
+We begin by relating the observational coordinates $\hat{\boldsymbol{s}}$
+and $\nu$, to
+$\boldsymbol{r}$ and $\boldsymbol{k}$: 
 \[
@@
 
-We can rewrite the delayed-transformed visibility as 
+Eqs.
+\eqref{eq:nur} and \eqref{eq:thinshell} show that observational angular and spectral dimensions together correspond to the three dimensional volume of cosmological interest.
+Moreover, recall that the power spectrum in Eq.
+\eqref{eq:Pofk} is proportional to the product of Fourier transforms of the temperature field.
+Since Eq.
+\eqref{eq:Vis1} resembles a Fourier transform along the angular directions, it is natural to perform a further Fourier transform along the radial, or equivalently, frequency axis.
+The transform, called delay-transform \citep{delay-transform}, has gained popularity recently since it has been shown that foregrounds are isolated in delay space \footnote{We discuss foreground isolation in detail in Section \ref{sec:chromaticity}.}.
+We define the delay-transformed visibility as:
+\small
+\begin{equation}
+\begin{aligned}V(\boldsymbol{b},\tau) & =\int d\nu V_{\nu}(\boldsymbol{b})\phi(\nu)\exp\left(2\pi i\nu\tau\right),\\
+ & =\int d\Omega d\nu B(\hat{\boldsymbol{s}},\nu)T(\hat{\boldsymbol{s}},\nu)\exp\left[-2\pi i\nu\left(\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}-\tau\right)\right]. 
+\end{aligned}
+\label{eq:Vb1}
+\end{equation}
+\normalsize
+Here the delay $\tau$ is the Fourier dual of $\nu$. Eq. \eqref{eq:Vb1} expresses the delay-transformed visibility as
+an integral over observational coordinates $\hat{\boldsymbol{s}}$ and $\nu$. For notational simplicity here and in the rest of this paper, we have defined the quantity
+\[
+B(\hat{\boldsymbol{s}},\nu) \equiv \frac{2k_B}{\lambda^2}\phi(\nu)A(\hat{\boldsymbol{s}},\nu). 
+\label{eq:B}
+\]
+
+With Eqs. \eqref{eq:nur} and \eqref{eq:thinshell}, we can rewrite the delayed-transformed visibility in cosmological coordinates as 
 \small
@@
 \end{figure*}
-
-\subsection{Numerical Test \label{sec:Techniquet}}
-
-Bringing together the discussion from Section \ref{sec:method}, we present a numerical check of Eq.
-\eqref{eq:final}, including rephasing.
-To do so, we need to compare the amplitude and phase of the integral weight $\Theta$
-for a pair of baselines with products of simulated visibilities of those baselines.
-We use 10 frequency channels evenly spaced from $145$\,MHz to $155$\,MHz for the comparison.
-We rephase one of the simulated visibilities, as well as computed $\Theta_{\nu}$ to the same delay as computed from the peak phase of the latter.
-
-\begin{figure}[h]
-\includegraphics[width=\linewidth]{numerics}
-
-\caption{Numerical comparisons of the visibility correlation peaks to the $\Omega$ factor in Eq.(\ref{eq:final}).
-We generated a Gaussian random sky on healpix maps, computed visibilities and cross correlated them to find the correlation.
-Top panel shows the equivalent baseline pairs \{2,0\} against \{2,0\}, bottom panel shows \{2,0\} against \{2,1\}.
-Evaluations of weight $\Theta$ are shown green, and the visibility correlations of simulated random sky is given in blue.
-In both cases only the real parts are shown.
-We see that in both cases the theory and simulation line up in both amplitude and phase.}
-\label{fig:numerics}
-\end{figure}
-
-Fig.
-\ref{fig:numerics} shows the real parts of the resulting degree of redundancy as a function of time offset, for equivalent baseline pairs (\{2,0:2,0\}) in the top panel and nearly equivalent (\{2,0:2,1\}) in the bottom.
-Shown in green for the two cases are the computed values of $\Theta$, both normalized to the peak of the equivalent case.
-The plot on the bottom is rephased with a delay of $\Delta\tau\approx15.7$ns as outlined in Section \ref{sec:rephs}.
-Since both y-axes are normalized to the same scale, we see the equivalent baseline reach maximal redundancy with 0 time offset, while the nearly equivalent pair reaches maximum redundancy of $\Theta\approx0.5$ at $dT\approx0.325$days.
-
-Almost completely overlapping with the green curves are the cross-multiplied visibilities from a simple simulation, shown in blue.
-For the simulation, we populate random values of brightness temperature
-on a healpix map \citep{Heal, HealPrimer} \footnote{We use functionalities in the python package AIPY for healpix mapping
-as well as coordinate transforms. }. We rotate the baseline positions with
-the appropriate rotation matrix, multiplying the sky by the primary beam to get the visibilities, for each
-baseline\footnote{There are two obvious ways to achieve the rotation. One
-can either fix the sky and rotate the baselines, or the other way
-around. We found however, that we must not physically rotate the sky
-map, for the numerical round-offs due to finite resolutions of the
-map turns out to be significant. Thus we let the sky, represented
-by the healpix map, be fixed, and rotate the baselines.
-}.
-For the nearly equivalent case, we rephase the delayed visibility (\{2,0\}) by the same delay $\Delta\tau\approx17$ns, followed by delay-transform.
-The resulting delay-space visibilities for the two baselines are then convolved
-via the Fourier convolution theorem, to obtain values of the cross
-correlation as a function of time offset.   We do this for both the equivalent (\{2,0:2,0\}) and nearly equivalent (\{2,0:2,1\}) case.
-Since the blue and green curves overlap, we have verified that in this case Eq. \eqref{eq:final} is valid and the computed $\Theta$ can be used as weight to obtain the power spectrum.
 
@@
 \end{equation}
-One may notice that we have used a single covariance for the complex visibility.
+One may notice that we use a single covariance for the complex visibility.
 It is straightforward to show that the same result holds if a separate real and imaginary components are used, as long as they are independent of each other.

### 2017-07-18 Rotation paper/draft_working.tex
saving progress

@@
 This technique avoids many systematics associated with combining data from different baselines and tracks the native sampling of the interferometer.
-An example of the visibility-based pipeline was presented in \cite{Ali2015}, which provides the newest upper limit to the power spectrum
+An example of the visibility-based pipeline is presented in \cite{Ali2015}, which provides the newest upper limit to the power spectrum
 measurements with the 64-element version of PAPER (henceforth as PAPER-64).
@@
 transform of the cross multiplied beam patterns. As a check, when applied to equivalent baselines,
-$\boldsymbol{b}=\boldsymbol{b'}$, the peak correlation occurs at $\Delta T=0$ and $\hat{\boldsymbol{s}}=\boldsymbol{\Gamma}\hat{\boldsymbol{s}}$, in which case Eq.
+$\boldsymbol{b}=\boldsymbol{b'}$, the peak correlation occurs at $\Delta t=0$ and $\hat{\boldsymbol{s}}=\boldsymbol{\Gamma}\hat{\boldsymbol{s}}$, in which case Eq.
 \eqref{eq:final} reduces to Eq.
@@
 
-At the time offset of maximum redundancy, we expect two beams to have to same fringe pattern in both frequency and phase.
+At the time offset of maximum redundancy, we expect two baselines to have to same fringe pattern in both frequency and phase.
 Due to the time delay, however, the beam centers would be slightly shifted with respect to each other.
-This we show in the top panels of Fig.
-\ref{fig:beamfringe}.
+In the top panels of Fig.
+\ref{fig:beamfringe}, we show the real parts of peak-normalized beam-fringe patterns.
 The top left and middle panels show the beam-fringe
-patterns (real parts) for baselines \{2,0\} and \{2,1\}, delayed by 0.0325 sidereal days.
+patterns for the baseline classes \{2,0\} and \{2,1\}, offset by 0.0325 sidereal days.
+More specifically, displayed are the real parts of $B(\hat{\boldsymbol{s}},\nu)\exp\left[-i2\pi\frac{\nu}{c}\hat{\boldsymbol{s}}\cdot\boldsymbol{b}\right]$ and $B(\boldsymbol{\Gamma}\hat{\boldsymbol{s}},\nu)\exp\left[-i2\pi\frac{\nu}{c}\boldsymbol{\Gamma}\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right]$, respectively.
+The top right panel show their conjugate product $B^{*}(\hat{\boldsymbol{s}},\nu)B(\boldsymbol{\Gamma}\hat{\boldsymbol{s}},\nu)\exp\left[i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma}\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right]$, where we have chosen global phase $\psi_{\nu}$ such that the peak of the conjugate product is real.
 The product of the beam fringe patterns in top right shows that the fringes
  cancel out as we expect.
-On the bottom of Fig.
-\ref{fig:beamfringe} we show the instantaneous $uv$ coverage of two baselines and their product.
-We see from the bottom right panel that the cross-multiplied baselines recover  power concentrated at $u\approx\frac{30\text{m}}{c/1.5\times10^8\text{Hz}}\approx15$.
-Note the bottom right panel is not the Fourier transform of the top right panel.
+Comparing with Eq.
+\eqref{eq:Thetanu}, we see that $\Theta_\nu$ is the sum over the top right plot.
+Thus we show that at the optimal time offset the fringe patterns of the two baselines cancel out, leading to a substantial weight $\Theta_\nu$.
+The frequency-summed weight $\Theta$ is constructed with rephasing as described in Section \ref{sec:rephs} so that the conjugate product (top right panel) for different frequencies align in phase.
+
+Similarly, we show the correlation of the two baselines in terms of their $uv$ beam spread overlap.
+The bottom left and middle panels are (peak-normalized) instantaneous $uv$ coverage of two baselines while the bottom right panel is their product.
+We see from the bottom right panel that the cross-multiplied baselines recover power concentrated at $u\approx15$.
+The product of the $uv$ beams spread shown in the bottom right panel should not be confused with the Fourier transform of the beam-fringe conjugate product shown in the top right panel.
 
@@
 The window functions for near-equivalent baseline estimates are shown in the bottom plot, for \{2,0:2,1\} and \{12,0:12,1\} at low- and high-$\bar{b}$, respectively.
-In all cases, the window functions that we show are for $\tau = 703\,\textrm{ns}$, and were generated using primary beam models for PAPER.
+In all cases, the window functions that we show are for $\tau = 703\,\textrm{ns}$, and are generated using primary beam models for PAPER.
 For numerical convenience we assumed that the bandpass $\phi_\nu$ (see Eq.
@@
 \eqref{eq:Theta}).
-From now one we use $\Theta_{\boldsymbol{bb'}}$ and $\Delta T_{\boldsymbol{bb'}}$ to denote the peak value and location of $\Theta$, or in other words the maximum of the redundancy weight $|\Theta(\Delta T)|$ and the corresponding offset for a given baseline pair $\{\boldsymbol{b}:\boldsymbol{b'}\}$.
+From now one we use $\Theta_{\boldsymbol{bb'}}$ and $\Delta t_{\boldsymbol{bb'}}$ to denote the peak value and location of $\Theta$, or in other words the maximum of the redundancy weight $|\Theta(\Delta t)|$ and the corresponding offset for a given baseline pair $\{\boldsymbol{b}:\boldsymbol{b'}\}$.
 %Intuitively we expect the sensitivity to depend on both the $uv$ coverage of the baseline and the patch of sky inside the beam.
@@
 In the middle panel of Fig. \ref{fig:sensplot}
-we show $\Theta_{\boldsymbol{bb'}}$ and $\Delta T_{\boldsymbol{bb'}}$ for a variety of baseline combinations. Baseline pairs that are mirror images of each other 
+we show $\Theta_{\boldsymbol{bb'}}$ and $\Delta t_{\boldsymbol{bb'}}$ for a variety of baseline combinations. Baseline pairs that are mirror images of each other 
 give the same amount of redundancies ($\Theta_{\boldsymbol{bb'}}$), with the opposite time offset, as expected from symmetry. 
@@
 
-The scaling in Eq. \eqref{eq:sensul} was rough for simplicity of motivation. As we derive in Appendix \ref{sec:appB}, this weight should be corrected by a factor proportional $\rho_0$:
+The scaling in Eq. \eqref{eq:sensul} is rough for simplicity of motivation. As we derive in Appendix \ref{sec:appB}, this weight should be corrected by a factor proportional $\rho_0$:
 
@@
 \caption{Pairplots of the top contributing baseline pairs in three arrays.
-Plotted properties are optimal time delay $\Delta T_{\boldsymbol{bb'}}$,  effective weight $\widetilde{\Theta}_{\boldsymbol{bb'}}$ and baseline length $\bar{b}$.
+Plotted properties are optimal time delay $\Delta t_{\boldsymbol{bb'}}$,  effective weight $\widetilde{\Theta}_{\boldsymbol{bb'}}$ and baseline length $\bar{b}$.
 Only those points with $\widetilde{\Theta}_{\boldsymbol{bb'}}>0.01$ are shown (the weight for the top class of equivalent pair is normalized to 1).
@@
 \ref{fig:pairplot} we present pair-distributions of three different properties for baselines that contribute well to the sensitivity ($\widetilde{\Theta}_{\boldsymbol{bb'}}>0.01$, where again the weight for the top class of equivalent pair is normalized to 1).
-The properties shown are effective weight $\widetilde{\Theta}_{\boldsymbol{bb'}}$, optimal time offset $\Delta T_{\boldsymbol{bb'}}$, and average baseline length $\bar{b}$.
+The properties shown are effective weight $\widetilde{\Theta}_{\boldsymbol{bb'}}$, optimal time offset $\Delta t_{\boldsymbol{bb'}}$, and average baseline length $\bar{b}$.
 We only show three of the mentioned arrays for visual clarity.
@@
 
-\item $\Delta T_{\boldsymbol{bb'}}$ vs. $\widetilde{\Theta}_{\boldsymbol{bb'}}$:
+\item $\Delta t_{\boldsymbol{bb'}}$ vs. $\widetilde{\Theta}_{\boldsymbol{bb'}}$:
 This relation  is familiar from Fig.
 \ref{fig:sensplot}.
-The points at $\Delta T=0$ are the equivalent baselines.
-Note HERA arrays have fewer data points with high $\Delta T_{\boldsymbol{bb'}}$, due to their comparatively narrow beam.
+The points at $\Delta t=0$ are the equivalent baselines.
+Note HERA arrays have fewer data points with high $\Delta t_{\boldsymbol{bb'}}$, due to their comparatively narrow beam.
 
-\item $\Delta T_{\boldsymbol{bb'}}$ vs. $\bar{b}$:
-PAPER-128 shows the trend that longer baselines correspond to lower $\Delta T_{\boldsymbol{bb'}}$.
+\item $\Delta t_{\boldsymbol{bb'}}$ vs. $\bar{b}$:
+PAPER-128 shows the trend that longer baselines correspond to lower $\Delta t_{\boldsymbol{bb'}}$.
 HERA arrays do not exhibit this trend here because of a selection effect.
 Only baseline pairs with $\widetilde{\Theta}_{\boldsymbol{bb'}}>0.01$ are shown.
-Shorter HERA baselines require much longer $\Delta T$ to overlap, partly because of the hexagonal structure requires $60\degree$ of rotation to overlap baselines, partly because of the smaller primary beam.
+Shorter HERA baselines require much longer $\Delta t$ to overlap, partly because of the hexagonal structure requires $60\degree$ of rotation to overlap baselines, partly because of the smaller primary beam.
 Thus most short HERA baseline pairs are not shown in this figure.
@@
 Applicable to both tracking and drift scan measurements, our method relies on cross-multiplications of visibilities that are offset in time.
-Given an antenna array configuration, our method identifies the best baseline pairs to cross-multiply and predicts the optimal time offset $\Delta T_{\boldsymbol{bb'}}$, weight $\Theta_{\boldsymbol{bb'}}$, and the necessary rephasing factor $\psi_\nu$.
+Given an antenna array configuration, our method identifies the best baseline pairs to cross-multiply and predicts the optimal time offset $\Delta t_{\boldsymbol{bb'}}$, weight $\Theta_{\boldsymbol{bb'}}$, and the necessary rephasing factor $\psi_\nu$.
 With the predicted results one can incorporate partial redundancy into existing delay-transform based power-spectrum pipelines.
@@
  \]
-If exponent vanishes, $\Theta_{\nu}(\Delta T)$ is real, and one would not observe any non-zero phase at its peak value.
+If exponent vanishes, $\Theta_{\nu}(\Delta t)$ is real, and one would not observe any non-zero phase at its peak value.
 However, with the inclusion of the $w$-term, we see that track crossings in $uv$-plane do not actually imply crossing in the $uvw$ space, and therefore the exponent or, equivalently, Eq.

### 2017-07-19 Rotation paper/draft_working.tex
new draft

@@
 
-Observational studies of reionization, including Gunn-Peterson measurements of quasi-stellar objects \citep{Fan2006} and Cosmic Microwave Background temperature and anisotropy measurements (CMB; \citealt{Planck2016}), the Kinetic Sunyaev-Zeldo'vich effect \citep{kszpatchy} and Lyman alpha galaxy observations \citep{mcquinnLyA} have given us indications of the rough time-frame of reionization, but only limited constraints on the finer spatial and temporal structures.
+Observational studies of reionization, including Gunn-Peterson measurements of quasi-stellar objects \citep{Fan2006} and Cosmic Microwave Background temperature and anisotropy measurements (CMB; \citealt{Planck2016}), the Kinetic Sunyaev-Zeldo'vich effect \citep{ZahnKSZ, GeorgeKSZ, kszpatchy} and Lyman alpha galaxy observations \citep{Rhoads2017, Rhoads2012, mcquinnLyA} have given us indications of the rough time-frame of reionization, but only limited constraints on the finer spatial and temporal structures.
 A surge of recent radio-astronomical experiments of reionization focus on measuring the ``spin-flip'' transition of neutral
@@
 Array (HERA; e.g. \citealt{HERA,HERAconfiguration,HERABEAM1,HERADISH2}) under construction, 
-and the Square Kilometer Array Low (SKA-low; e.g. \citealt{SKA1}) currently in planning stages.
+and the Square Kilometer Array Low (SKA-low; e.g. \citealt{skalow2015} and references therein) currently in planning stages.
 
@@
 an integral over observational coordinates $\hat{\boldsymbol{s}}$ and $\nu$. For notational simplicity here and in the rest of this paper, we have defined the quantity
+
 \[
@@
 \begin{figure}[H]
-\includegraphics[width=1.1\linewidth]{rephs}
+\includegraphics[width=1.1\linewidth]{rephs_cm}
 
@@
 patterns for the baseline classes \{2,0\} and \{2,1\}, offset by 0.0325 sidereal days.
-More specifically, displayed are the real parts of $B(\hat{\boldsymbol{s}},\nu)\exp\left[-i2\pi\frac{\nu}{c}\hat{\boldsymbol{s}}\cdot\boldsymbol{b}\right]$ and $B(\boldsymbol{\Gamma}\hat{\boldsymbol{s}},\nu)\exp\left[-i2\pi\frac{\nu}{c}\boldsymbol{\Gamma}\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right]$, respectively.
-The top right panel show their conjugate product $B^{*}(\hat{\boldsymbol{s}},\nu)B(\boldsymbol{\Gamma}\hat{\boldsymbol{s}},\nu)\exp\left[i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma}\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right]$, where we have chosen global phase $\psi_{\nu}$ such that the peak of the conjugate product is real.
+More specifically, displayed are the real parts of $g_b\equiv B(\hat{\boldsymbol{s}},\nu)\exp\left[-i2\pi\frac{\nu}{c}\hat{\boldsymbol{s}}\cdot\boldsymbol{b}\right]$ and $g_{b'}\equiv B(\boldsymbol{\Gamma}\hat{\boldsymbol{s}},\nu)\exp\left[-i2\pi\frac{\nu}{c}\boldsymbol{\Gamma}\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right]$, respectively.
+The top right panel show their conjugate product $g_{b}^{*}g_{b'} = B^{*}(\hat{\boldsymbol{s}},\nu)B(\boldsymbol{\Gamma}\hat{\boldsymbol{s}},\nu)\exp\left[i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma}\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right]$, where we have chosen global phase $\psi_{\nu}$ such that the peak of the conjugate product is real.
 The product of the beam fringe patterns in top right shows that the fringes
@@
 Comparing with Eq.
-\eqref{eq:Thetanu}, we see that $\Theta_\nu$ is the sum over the top right plot.
+\eqref{eq:Thetanu}, we see that $\Theta_\nu=\int d\Omega g_{b}^{*}g_{b'}$ is the sum over the image shown in the top right panel.
 Thus we show that at the optimal time offset the fringe patterns of the two baselines cancel out, leading to a substantial weight $\Theta_\nu$.
@@
 Similarly, we show the correlation of the two baselines in terms of their $uv$ beam spread overlap.
-The bottom left and middle panels are (peak-normalized) instantaneous $uv$ coverage of two baselines while the bottom right panel is their product.
+The bottom left and middle panels are (peak-normalized) instantaneous $uv$ coverage of two baselines ($\widetilde{g}_{b}$ and $\widetilde{g}_{b'}$) while the bottom right panel is their product.
 We see from the bottom right panel that the cross-multiplied baselines recover power concentrated at $u\approx15$.
@@
 
-We define the estimator sensitivity to be the inverse of the total power spectrum noise variance $\Sigma_P^2$:
+We define the estimator sensitivity to be the square root of the inverse of the total power spectrum noise variance $\Sigma_P^2$:
 \begin{equation}
-\rho \propto 1/\Sigma_P^2 \propto \rho_0^2\sum_{\boldsymbol{bb'}}^N\widetilde{\Theta}^2_{\boldsymbol{bb'}},
+\rho \propto 1/\Sigma_P \propto \rho_0\sqrt{\sum_{\boldsymbol{bb'}}^N\widetilde{\Theta}^2_{\boldsymbol{bb'}}},
 \end{equation}
@@
 Note HERA arrays have fewer data points with high $\Delta t_{\boldsymbol{bb'}}$, due to their comparatively narrow beam.
+Also note that all near-equivalent pairs for HERA arrays appear at $\widetilde{\Theta}_{\boldsymbol{bb'}}\lesssim 0.4$.
 
@@
 \item $\widetilde{\Theta}_{\boldsymbol{bb'}}$ vs. $\bar{b}$:
-The general trend to be seen here is that longer baselines tend to have lower $\widetilde{\Theta}_{\boldsymbol{bb'}}$.
+In this plot the HERA arrays each show two superimposed structures.
+The equivalent baselines, all having the same individual weight $\Theta_{\boldsymbol{bb'}}$, appear as elongated structures towards high $\widetilde{\Theta}_{\boldsymbol{bb'}}$.
+The narrowness of these structures is an indication of the approximately linear and one-to-one correspondence between baseline length and multiplicity.
+The wide structures spanning the whole range of $\bar{b}$ are the near-equivalent baseline pairs.
+We see again for HERA arrays all near-equivalent pairs appear at $\widetilde{\Theta}_{\boldsymbol{bb'}}\lesssim 0.4$.
+For PAPER-128, the existence of short near-equivalent pairs with high multiplicity partially fills the high $\widetilde{\Theta}_{\boldsymbol{bb'}}$ region.
+Note the general trend that longer baselines tend to have lower $\widetilde{\Theta}_{\boldsymbol{bb'}}$.
 This is due to the lower multiplicity of longer baselines.
@@
 The small HERA37, with no gap (like in HERA350) or short nearly equivalent baselines (like in PAPER 128), will not benefit much from the nearly equivalent baselines.
-The maximum benefits for other cases are expected to be around $20\%$ to $60\%$.
+Near-equivalent baselines make an increasingly significant contribution to the total sensitivity as the number of array elements are increased.
 PAPER-128 is designed with highly redundant nearly equivalent baselines, and thus  these baselines start contributing at higher $\widetilde{\Theta}_{min}$, but the gapped HERA configurations will benefit even more from nearly equivalent baselines at low $\widetilde{\Theta}_{min}$ due to the presence of more classes of such pairs.
-Note that here we normalized $\rho$ such that the contribution of the top equivalent baseline pair, such as are 1.
-This plot therefore does not compare the absolute sensitivity across the different arrays.
-The stepwise pattern is characteristic of a regular grid; as we step to lower $\widetilde{\Theta}_{\boldsymbol{bb'}}$ large groups of baseline pair classes get included in ``batch''.
+Note that here we normalized sensitivity for each array as the fraction of the total sensitivity when all baseline pairs are used.
+The plot therefore does not compare absolute sensitivity across different arrays.
+The stepwise pattern seen in almost all cases are characteristic of a regular grid; as we step to lower $\widetilde{\Theta}_{\boldsymbol{bb'}}$ large groups of baseline pair classes get included in ``batch''.
+Thus the ``gapless'' HERA37, with more regularity in its antenna configurations than the larger counterparts, exhibit more step patterns in fractional sensitivity.
 %\begin{centered}
@@
 \includegraphics[width=\linewidth]{osens}
-\caption{Sensitivity of redundant arrays as a function of the minimum effective weight.
+\caption{Fractional sensitivity of redundant arrays as a function of the minimum effective weight.
 Dashed lines represent when only equivalent baseline pairs are used, while solid lines indicate use of both the equivalent and nearly equivalent baselines are used.
-The vertical axis is normalized independently for each array such that the contribution of the top single equivalent pair class is unity, and thus does not indicate a comparison of absolute sensitivity across the different arrays shown.
-}
+The vertical axis is normalized independently for each array to the total sensitivity when all baseline pairs, equivalent and near-equivalent, are used.}
 \label{fig:osens}

### 2017-07-21 Rotation paper/draft_working.tex
save

@@
 
-\title{Unlocking Sensitivity for Visibility-based Power Spectrum Estimators of the 21\,cm Reionization Signal}
+\title{Unlocking Sensitivity for Visibility-based Estimators of the 21\,cm Reionization Power Spectrum}
 
@@
 \author{
-Yunfan Gerry Zhang \altaffilmark{1},
-Adrian Liu \altaffilmark{1, 2},
-Aaron R. Parsons\altaffilmark{1, 2}
+Yunfan Gerry Zhang\altaffilmark{1},
+Adrian Liu\altaffilmark{1,2,3},
+Aaron R. Parsons\altaffilmark{1,2}
 }
-
+\email{yunfanz@berkeley.edu}
 \altaffiltext{1}{Astronomy Dept., U. California, Berkeley, CA}
 \altaffiltext{2}{Radio Astronomy Lab., U. California, Berkeley, CA}
+\altaffiltext{3}{Hubble Fellow}
 
 \begin{abstract}
-Radio interferometers seeking to measure the cosmological 21\,cm power spectrum require high sensitivity.
-Several modern low-frequency interferometers feature antennas placed on regular grid to maximize redundancy.
+Radio interferometers designed to measure the cosmological 21\,cm power spectrum require high sensitivity.
+Several modern low-frequency interferometers feature drift-scan antennas placed on regular grid to maximize the number of instantaneously coherent (redundant) measurements.
+However, even for such maximum-redundancy arrays, significant sensitivity comes through partial coherence between baselines.
 Current visibility-based power spectrum pipelines, though shown to ease control of systematics, lack the ability to make use of partial redundancy.
@@
 The epoch of reionization represents the last key
-stage of our Universe's early evolution. Study of this event stands at
+stage of our Universe's early evolution. The study of this event stands at
 the intersection of cosmology and astrophysics. Understanding this
 event not only serves as a scientific goal
-of its own, but also as a gateway to crucial information
+of its own, but also as a gateway to information
 regarding fundamental physics of inflation, neutrino mass, and the phenomenology
@@
 
-\caption{The PAPER128 layout. Each dot corresponds to the location of
+\caption{The PAPER-128 layout. Each dot corresponds to the location of
 an antenna. The top panel shows the antenna positions drawn to scale;
@@
 We show in Fig.
-\ref{fig:Tracks} $uvw$ tracks of the three PAPER128 baselines colored in Fig.
+\ref{fig:Tracks} $uvw$ tracks of the three PAPER-128 baselines colored in Fig.
 \ref{fig:AntPos}, projected onto the $uv$ and $vw$ planes.
@@
 \ref{fig:phi_nu} we show the phases of $\Theta_\nu$ at the optimal time offset or, in other words, when $|\Theta_\nu|$ is maximized.
-The phases are shown for the baseline pair \{2,0:2,1\} of PAPER128.
+The phases are shown for the baseline pair \{2,0:2,1\} of PAPER-128.
 The drift-scan phase dependence is compared with that of the same baseline with hypothetical tracking-elements (Eq.
@@
 \begin{figure}[H]
-\includegraphics[width=1.1\linewidth]{rephs_cm}
+\includegraphics[width=1.1\linewidth]{rephs_rb}
 
@@
 First and third panels shows equivalent baselines, second and fourth show a pair of nearly equivalents.
-The top two panels have zero phase shift ($\psi_{\nu}=0$), and the bottom two are rephased to a time offset of 0.0325 sidereal days.
+The top two panels have zero phase shift ($\psi_{\nu}=0$), and the bottom two are rephased to a time offset of 0.78 hours.
 The first and last panels thus show a coherently rephased series that would add constructively at the respective peaks.
@@
 The top left and middle panels show the beam-fringe
-patterns for the baseline classes \{2,0\} and \{2,1\}, offset by 0.0325 sidereal days.
+patterns for the baseline classes \{2,0\} and \{2,1\}, offset by 0.78 hours.
 More specifically, displayed are the real parts of $g_b\equiv B(\hat{\boldsymbol{s}},\nu)\exp\left[-i2\pi\frac{\nu}{c}\hat{\boldsymbol{s}}\cdot\boldsymbol{b}\right]$ and $g_{b'}\equiv B(\boldsymbol{\Gamma}\hat{\boldsymbol{s}},\nu)\exp\left[-i2\pi\frac{\nu}{c}\boldsymbol{\Gamma}\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right]$, respectively.
-The top right panel show their conjugate product $g_{b}^{*}g_{b'} = B^{*}(\hat{\boldsymbol{s}},\nu)B(\boldsymbol{\Gamma}\hat{\boldsymbol{s}},\nu)\exp\left[i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma}\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right]$, where we have chosen global phase $\psi_{\nu}$ such that the peak of the conjugate product is real.
+The top right panel show their conjugate product $g_{b}^{*}g_{b'}e^{-i\psi_\nu} = B^{*}(\hat{\boldsymbol{s}},\nu)B(\boldsymbol{\Gamma}\hat{\boldsymbol{s}},\nu)\exp\left[i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma}\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right]$, where we have chosen global phase $\psi_{\nu}$ such that the peak of the conjugate product is real.
 The product of the beam fringe patterns in top right shows that the fringes
@@
 
-\pagebreak
+\acknowledgments
+
+The authors thank Zaki Ali, Carina Cheng, Dave DeBoer, and Miguel Morales for helpful discussions.
+This research was completed as part of the University of California Cosmic Dawn Initiative, with support from the University of California Office of the President Multicampus Research Programs and Initiative through award MR-15-328388.
+Additional support was provided by NSF AST grants 1129258 and 1636646.
+AL acknowledges support for this work by NASA through Hubble Fellowship grant \#HST-HF2-51363.001-A awarded by the Space Telescope Science Institute, which is operated by the Association of Universities for Research in Astronomy, Inc., for NASA, under contract NAS5-26555.
 

### 2017-07-21 Rotation paper/draft_working.tex
save

@@
 Analysis pipelines for the 21\,cm power spectrum typically fall into two categories.
-In the first, images are formed in Fourier domain through rotation synthesis, followed by a foreground mitigation step to construct a power spectrum.
+In the first, images are formed in Fourier domain through rotation synthesis, followed by a foreground mitigation step to construct a power spectrum \citep{LOFARresult, MWAresult1, MWAresult0}.
 An alternative technique works directly with visibilities from baselines, delay-transforming and cross-multiplying them to form the power spectrum.
@@
 the bottom panel show the antenna labels and distances of the 112-element grid, in other words excluding the outrigger antennas.
-The numbering of the antennas in the bottom panel are original labels
-during instrument assembly and does not bear significant
-meaning. In the bottom panel, the two baselines marked with red segments are example of an equivalent pair, each with antennas separated by 2 units east and 0 unit north. Similarly,
+In the bottom panel, the two baselines marked with red segments are example of an equivalent pair, each with antennas separated by 2 units east and 0 unit north. Similarly,
 the baselines marked in blue and yellow are examples
@@
 \begin{equation}
-\begin{aligned}V_{\nu}(\boldsymbol{b}) & =\int d\Omega A(\hat{\boldsymbol{s}},\nu)\phi(\nu)I_{\nu}(\hat{\boldsymbol{s}})\exp\left(-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right),\\
+\begin{aligned}V_{\nu}(\boldsymbol{b}) & =\int d\Omega A(\hat{\boldsymbol{s}},\nu)\phi(\nu)I_{\nu}(\hat{\boldsymbol{s}})\exp\left(-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right)\\
  & \approx\frac{2k_{B}}{\lambda^{2}}\int d\Omega A(\hat{\boldsymbol{s}},\nu)\phi(\nu)T(\hat{\boldsymbol{s}})\exp\left(-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right),
@@
 \[
-\begin{aligned}r & =\frac{c}{H_{0}}\int_{0}^{z}\frac{dz'}{E(z')},\\
- & \approx\frac{c}{H_{0}}\int_{0}^{z_{0}}\frac{dz'}{E(z')}-\frac{c(1+z_0)^{2}}{\nu_{21}H_{0}E(z_0)}\left(\nu-\nu_{0}\right),\\
+\begin{aligned}r & =\frac{c}{H_{0}}\int_{0}^{z}\frac{dz'}{E(z')}\\
+ & \approx\frac{c}{H_{0}}\int_{0}^{z_{0}}\frac{dz'}{E(z')}-\frac{c(1+z_0)^{2}}{\nu_{21}H_{0}E(z_0)}\left(\nu-\nu_{0}\right)\\
  & \equiv X-Y\Delta\nu,
@@
 \begin{equation}
-\begin{aligned}V(\boldsymbol{b},\tau) & =\int d\nu V_{\nu}(\boldsymbol{b})\phi(\nu)\exp\left(2\pi i\nu\tau\right),\\
+\begin{aligned}V(\boldsymbol{b},\tau) & =\int d\nu V_{\nu}(\boldsymbol{b})\phi(\nu)\exp\left(2\pi i\nu\tau\right)\\
  & =\int d\Omega d\nu B(\hat{\boldsymbol{s}},\nu)T(\hat{\boldsymbol{s}},\nu)\exp\left[-2\pi i\nu\left(\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}-\tau\right)\right]. 
@@
 \begin{aligned} \langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle
- & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}),\\
- & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\boldsymbol{\kappa})T(\boldsymbol{\kappa'})\rangle e^{-i(\boldsymbol{\kappa}\cdot \boldsymbol{r}-\boldsymbol{\kappa'}\cdot\boldsymbol{r'})}\right)B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}),\\
- & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\boldsymbol{\kappa}\cdot(\boldsymbol{r}-\boldsymbol{r'})}\right)B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}),\\
- & \approx P(k_{\bar{b}, \tau})\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2} \delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'}) \Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}),\\
+ & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\\
+ & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\boldsymbol{\kappa})T(\boldsymbol{\kappa'})\rangle e^{-i(\boldsymbol{\kappa}\cdot \boldsymbol{r}-\boldsymbol{\kappa'}\cdot\boldsymbol{r'})}\right)B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\\
+ & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\boldsymbol{\kappa}\cdot(\boldsymbol{r}-\boldsymbol{r'})}\right)B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\\
+ & \approx P(k_{\bar{b}, \tau})\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2} \delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'}) \Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\\
  & = P(k_{\bar{b}, \tau})\int \frac{d^{3}r}{(X^2Y)^2}B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r})\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma} \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right],

### 2017-07-21 Rotation paper/draft_working.tex
Aaron edits for first four pages

@@
 However, even for such maximum-redundancy arrays, significant sensitivity comes through partial coherence between baselines.
-Current visibility-based power spectrum pipelines, though shown to ease control of systematics, lack the ability to make use of partial redundancy.
+Current visibility-based power spectrum pipelines, though shown to ease control of systematics, lack the ability to make use of this partial redundancy.
 We introduce a method to leverage partial redundancy in such power spectrum pipelines for drift-scan arrays.
-Our method cross-multiplies baseline pairs at a time lag, and quantifies the sensitivity contributions of each pair of baselines.
+Our method cross-multiplies baseline pairs at a time lag and quantifies the sensitivity contributions of each pair of baselines.
 Using the configurations and beams of the 128-element Donald C.
-Backer Precision Array for Probing the Epoch of Reionization (PAPER-128), and four staged deployments of the Hydrogen Epoch of Reionization Array (HERA), we illustrate how our method applies to different arrays and predict the sensitivity improvements associated with each baseline pairs.
-We show that for upcoming arrays with increasing number of antennas, partial redundancy will be of increasing importance in unlocking full sensitivities of the arrays.
+Backer Precision Array for Probing the Epoch of Reionization (PAPER-128) and staged deployments of the Hydrogen Epoch of Reionization Array (HERA), we illustrate how our method applies to different arrays and predict the sensitivity improvements associated with pairing partially coherent baselines.
+As the number of antennas increases, we find partial redundancy to be of increasing importance in unlocking the full sensitivity of upcoming arrays.
 \end{abstract}
@@
 hydrogen with a characteristic wavelength of 21\,cm \citep{Furlanetto2006181,PritchardLoeb}.
-The 21\,cm brightness temperature is a direct tracer of neutral hydrogen through the epoch of reionization, and thus by measuring the three dimensional distribution of the 21\,cm signal we measure the full temporal and spatial variations of this event.
+The 21\,cm brightness temperature is a direct tracer of neutral hydrogen through the epoch of reionization.
+Therefore, by measuring the three dimensional distribution of the 21\,cm signal we measure the full temporal and spatial variations of this event.
 However, before realizing full-scale 21\,cm tomography, many current radio interferometric efforts
@@
 Widefield Array (MWA; \citealt{Bowman2013, Tingay2013}), Low Frequency Array (LOFAR; \citealt{LOFAR}). Next-generation instruments include the Hydrogen Epoch of Reionization
-Array (HERA; e.g. \citealt{HERA,HERAconfiguration,HERABEAM1,HERADISH2}) under construction, 
-and the Square Kilometer Array Low (SKA-low; e.g. \citealt{skalow2015} and references therein) currently in planning stages.
+Array (HERA; e.g. \citealt{HERA,HERAconfiguration,HERABEAM1,HERADISH2}), which is under construction, 
+and the Square Kilometer Array Low (SKA-low; e.g. \citealt{skalow2015} and references therein), which is currently in planning stages.
 
 The highly redshifted 21\,cm signal is faint and diffuse, in contrast to the localized bright sources targeted by many traditional radio telescopes.
-Current 21\,cm experiments are sensitivity-starved, with the sensitivity requirement further increased considering the foreground contaminations five orders of magnitude brighter than the cosmological signal of interest.
+Current 21\,cm experiments are sensitivity-starved, with the experimental challenge further increased when considering foreground contamination five orders of magnitude brighter than the cosmological signal of interest.
 Low-frequency radio interferometers aiming to measure the 21\,cm signal are thus designed differently from traditional instruments.
-To satisfy the sensitivity needs, modern arrays are large.
-Ranging upwards from 100 elements, these arrays are typically of the drift-scan type to limit cost.
-To further improve sensitivity, specially designed arrays such as PAPER and HERA feature multiple copies of the same baselines to repeatedly measure the same Fourier signal\citep{first-paper}.
+To satisfy the sensitivity needs, modern arrays are large (ranging upwards from 100 elements) and are typically of the drift-scan (static-pointing) type to limit cost.
+To further improve sensitivity, experiments such as PAPER and HERA feature multiple copies of the same baselines to repeatedly measure the same Fourier signal\citep{first-paper}.
 
 Analysis pipelines for the 21\,cm power spectrum typically fall into two categories.
-In the first, images are formed in Fourier domain through rotation synthesis, followed by a foreground mitigation step to construct a power spectrum \citep{LOFARresult, MWAresult1, MWAresult0}.
-An alternative technique works directly with visibilities from baselines, delay-transforming and cross-multiplying them to form the power spectrum.
+In the first, images are formed through rotation synthesis, and after a foreground mitigation step, are Fourier transformed to construct a power spectrum \citep{MWAresult0, MWAresult1, LOFARresult}.
+An alternative technique works directly with visibilities from baselines, delay-transforming and cross-multiplying them to form the power spectrum \citep{delay-transform}.
 This technique avoids many systematics associated with combining data from different baselines and tracks the native sampling of the interferometer.
-An example of the visibility-based pipeline is presented in \cite{Ali2015}, which provides the newest upper limit to the power spectrum
+An example of the visibility-based pipeline is presented in \cite{Ali2015}, which provides power spectrum
 measurements with the 64-element version of PAPER (henceforth as PAPER-64).
-However, one disadvantage of existing visibility-based pipelines is their lack of use of partial redundancy.
-Baselines of different lengths and orientations contain partially redundant information.
+However, one disadvantage of existing visibility-based pipelines is their lack of treatment of partial redundancy.
+Baselines that are slightly different in length and orientation
+``rotate into'' each other at a time delay.
+Baselines of different lengths and orientations therefore contain partially coherent information.
+This is the basis of earth-rotation synthesis.
 While imaging based power-spectrum pipelines naturally include all redundancy information, visibility-based pipelines so far only cross multiply fully redundant baselines, i.e.
@@
 In fact, most sensitivity forecasts to date do include partial redundancy \citep{Pobersens, HERA, JoshAntPos}.
-Recently, \cite{wterm} proposed a visibility-based approach to extract power spectrum from partially redundant baselines in sky-tracking measurements.
+Recently, \cite{wterm} proposed a visibility-based approach to produce power spectrum measurements using partially redundant baselines in sky-tracking measurements.
 Arrays capable of such tracking include MWA and LOFAR.
-Our work introduces a two-baseline power spectrum estimator that can be applied to both tracking capable arrays and drift-scan only telescopes, such as PAPER, HERA, and potentially SKA-low.
+Our work introduces a two-baseline power spectrum estimator that can be applied to both tracking-capable arrays and drift-scan telescopes such as PAPER, HERA, and potentially SKA-low.
 
-The basic idea of our proposed visibility-based, two-baseline power spectrum estimator is as follows.
+The basic idea of our proposed visibility-based, two-baseline power-spectrum estimator is as follows.
 The Earth's rotation causes the baselines in a drift-scan array to pick up different modes of the sky with time.
-Rotation-synthesis makes use of the rotation-induced $uv$ coverage map to form image.
-In a visibility based pipeline, we can extract the same rotation-induced redundancy.
-In this point of view, baselines that are slightly different in length and orientation
-``rotate into'' each other at a time delay.
-We can thus cross-multiply time-shifted visibilities, with the proper weighting, to form power spectra.
+Rotation-synthesis makes use of the rotation-induced $uv$ coverage map to form an image.
+With visibility-based pipelines, the same information can be extracted.
+To do so, we cross-multiply time-shifted visibilities, with the proper weighting, to form power spectra.
 Due to the large number of elements of modern arrays, the task of cross-multiplying every baseline against every other, scaling as number of array elements to the fourth power, can be computationally formidable, and many pairs of baselines provide only negligible redundancy information.
 Our contribution is thus twofold.
-First we introduce a formalism to estimate the power spectrum from pairs of partially-redundant baselines in a visibility pipeline.
+First, we introduce a formalism to estimate the power spectrum from pairs of partially-redundant baselines in a visibility pipeline.
 Secondly we show how to use the two-baseline estimator to automatically pre-select baseline pairs and time offsets, making the problem computationally efficient.
-More precisely, our formalism allows one to simultaneously identify the baselines that give good
+More precisely, our formalism allows one to simultaneously identify the baselines that have strong
 redundancy, find the time offset that corresponds to maximal redundancy for a given pair of baselines, and quantify the sensitivity associated with cross multiplying
-such a pair of partially redundant baselines, which in turns is used as weight to combine measurements in a power spectrum pipeline.
+such a pair of partially redundant baselines, which in turn is used as weight to combine measurements in a power spectrum pipeline.
 
@@
 in the sky.
-We shall call baselines in the same class \textit{equivalent baselines}, and reserve the word \textit{redundancy} of two baselines in reference to a variable function of the relative time offsets of their visibility time series.
-With this terminology, two equivalent
+We shall call baselines in the same class \textit{equivalent baselines}, and reserve the \textit{redundancy} of two baselines to mean a variable function of the relative time offsets of their visibility time series.
+With this terminology, equivalent
 baselines are fully redundant with each other simultaneously at all
-times. Non-equivalent baselines also have partial redundancy, and the redundancy can be maximized if their
-respected time series are shifted with respect to one another by some time offset.
+times. Non-equivalent baselines also have partial redundancy, and the redundancy can be maximized 
+by shifting their time series by a relative time offset.
 
@@
 an antenna. The top panel shows the antenna positions drawn to scale;
-the bottom panel show the antenna labels and distances of the 112-element grid, in other words excluding the outrigger antennas.
+the bottom panel show the antenna labels and distances of the 112-element grid, excluding the outrigger antennas.
 In the bottom panel, the two baselines marked with red segments are example of an equivalent pair, each with antennas separated by 2 units east and 0 unit north. Similarly,
 the baselines marked in blue and yellow are examples
-of classes that , due to the small North-South separation within the grid (see top panel), are expected to be nearly equivalent to the red baselines. }
+of classes that, due to the small North-South separation within the grid, are expected to be nearly equivalent to the red baselines. }
 \label{fig:AntPos}
@@
 The array consists of a 112-element core in a rectangular grid, and 16 ``outriggers'' used primarily to aid calibration.
-In the bottom panel, the two baselines marked in red, are an example of equivalent pair.
+In the bottom panel, the two baselines marked in red are an example of an equivalent pair.
 We denote a equivalency-class of baselines in the PAPER grid by their separations, in this case  \{2,0\}, for the
@@
 of \{2,1\} and \{2, -1\}.
-Note that \{2,0\} and \{-2, 0\} for example are the same class of baselines and should not be counted twice. Antennas in purely north-south baselines
+Note that \{2,0\} and \{-2, 0\} are the same class of baselines and should not be counted twice. Antennas in purely north-south baselines
 are close (4m), and hence these baselines are not suitable
-for use due to cross-coupling.
+for sensitive measurements due to cross-coupling.
 On the other hand, the small North-South separation means that classes such as \{2,0\} and  \{2,1\} are expected to be near-equivalent.
@@
 equivalent of which are 
-\{2, 0\}, \{2, 1\} and \{2, -1\}. There, each of these classes
-of baselines are cross multiplied within itself.
+\{2, 0\}, \{2, 1\} and \{2, -1\}.
+There, baselines are cross-multiplied within each class.
 This paper provides the method for inter-class multiplications.
@@
 
-In this section we introduce our method to cross-multiply nearly equivalent baseline classes.
+In this section we introduce our method to cross-multiply non-equivalent baseline classes.
 
@@
 \ref{fig:AntPos}, projected onto the $uv$ and $vw$ planes.
-The tracks are traced over 12 sidereal hours, at 0.15\,GHz, relative to a phase center that pdrifts from zenith.
+The tracks are traced over 12 sidereal hours, at 0.15\,GHz, relative to a phase center that drifts from zenith.
 
@@
 \includegraphics[width=\linewidth]{rotation_new}
-\caption{Tracks of 3 PAPER-128 baseline classes shown in Fig.
+\caption{Tracks of the three PAPER-128 baseline classes shown in Fig.
 \ref{fig:AntPos}, here with the same respective colors.
@@
 The top panel shows $uv$ tracks with crossings among each pair of baseline classes.
-Projection to the $vw$ plane in the bottom panel shows that tracks in $uvw$ space do not cross.
+Projection to the $vw$ plane in the bottom panel shows that although tracks appear to overlap in the $uv$ plane, tracks in $uvw$ space do not cross.
 }
@@
 redundancy of nearly equivalent baselines as crossings
-of the $uv$ tracks, a 2 dimensional projection of Fig.
+of the $uv$ tracks, a two-dimensional projection of $uvw$, as shown in the top panel of Fig.
 \ref{fig:Tracks}.
-As we see in the top panel of Fig.
-\ref{fig:Tracks}, there would be many such crossings.
-However, there are several reasons that $uv$ track-crossing do not imply perfect redundancy.
-The most obvious reason is that the 3 dimensional $uvw$ tracks do not actually cross, as is evident from the $vw$ projection in the bottom panel of Fig.
+However, there are several reasons that $uv$ track-crossing does not imply perfect redundancy.
+The most obvious reason is that the three-dimensional $uvw$ tracks do not actually cross, as is evident from the $vw$ projection in the bottom panel of Fig.
 \ref{fig:Tracks}.
-Therefore, in our case $uv$ track-crossings are
+In general, track-crossings are
 not accurate enough for time offset determination, nor can it give estimate of the degree of redundancy.
 For drift-scan arrays, even a hypothetical crossing in $uvw$ space would not imply perfect redundancy.
-This shall become evident in the next section, after we develop a more general formalism that accounts for the point spread function of the finite beams, and estimate the degree of redundancy for general combination of baselines at a general time offset.
-Relation between track-crossing and redundancy for both drift-scan and tracking measurements is further explored in Appendix \ref{sec:appA}.
+This will become evident in the next section, after we develop a more general formalism that accounts for the point spread function of primary.
+The relation between track-crossing and redundancy for both drift-scan and tracking measurements is further explored in Appendix \ref{sec:appA}.
 
@@
 In this section we formulate a power spectrum estimator from
-the product of delay-transformed visibilities from two arbitrary baselines. We shall work in the case of drift scan telescopes and extent the results to tracking measurements.
+the product of delay-transformed visibilities from two arbitrary baselines. We shall work in the case of drift-scan telescopes and extend the results to tracking measurements.
 
@@
 
-Ultimately we would like to relate observed visibility to the power spectrum $P$, defined such that
+Ultimately we would like to relate the observed visibility to the power spectrum, $P$, defined such that
 \[
@@
 \]
-where $\bold{k}$ is the cosmological wavenumber, $\delta_D^3$ is the three-dimensional Dirac delta function, and $\widetilde{T}(\boldsymbol{k})$ is the three-dimensional Fourier transform of the brightness temperature field $T(\bold{r})$, with $\bold r$ as the cosmological position coordinate.
+where $\bold{k}$ is the cosmological wavenumber, $\delta_D^3$ is the three-dimensional Dirac delta function, and $\widetilde{T}(\boldsymbol{k})$ is the three-dimensional Fourier transform of the brightness temperature field $T(\bold{r})$, with $\bold r$ being the cosmological position coordinate.
 We begin by relating the observational coordinates $\hat{\boldsymbol{s}}$
@@
 \end{equation}
-Note Eq. \eqref{eq:thinshell} does not require the flat-sky approximation; the angular integral is still performed over the curved sky.
+Note that Eq. \eqref{eq:thinshell} does not require the flat-sky approximation; the angular integral is still performed over the curved sky.
 
@@
 \eqref{eq:Vis1} resembles a Fourier transform along the angular directions, it is natural to perform a further Fourier transform along the radial, or equivalently, frequency axis.
-The transform, called delay-transform \citep{delay-transform}, has gained popularity recently since it has been shown that foregrounds are isolated in delay space \footnote{We discuss foreground isolation in detail in Section \ref{sec:chromaticity}.}.
+The transform, called delay-transform \citep{PB2009}, has gained popularity recently since it has been shown that foregrounds are isolated in delay space\footnote{We discuss foreground isolation in detail in Section \ref{sec:chromaticity}.}.
 We define the delay-transformed visibility as:
@@
 Here the delay $\tau$ is the Fourier dual of $\nu$. Eq. \eqref{eq:Vb1} expresses the delay-transformed visibility as
-an integral over observational coordinates $\hat{\boldsymbol{s}}$ and $\nu$. For notational simplicity here and in the rest of this paper, we have defined the quantity
+an integral over observational coordinates $\hat{\boldsymbol{s}}$ and $\nu$. For notational simplicity we have defined the quantity
 
@@
 
-Existing visibility based power-spectrum pipelines for redundant drift-scan  arrays relate the power-spectrum to the conjugate square of the visibilities \citep{delay-transform, paper32, Ali2015}.
+Existing visibility-based power-spectrum pipelines for redundant drift-scan  arrays relate the power-spectrum to the conjugate square of the visibilities \citep{delay-transform, paper32, Ali2015}.
 We would like to generalize such relations by relating the power-spectrum to the product of two visibilities from two arbitrary baselines and time offsets.
-The beam pattern of a baseline shifts relative to the sky as the Earth rotates. Here we choose to fix the sky, and denote the rotated coordinates
+To do this, we must account for the fact that the beam pattern of a baseline shifts relative to the sky as the Earth rotates. Here we choose to fix the sky, and denote the rotated coordinates
 in the topocentric frame with the three-dimensional rotation operator $\boldsymbol{\Gamma}$:
@@
 where we introduced a frequency-dependent phase $\psi_{\nu}$, which to linear-order in $\nu$ corresponds to phasing the two visibilities to the same phase center.
-\footnote{Turns out for many cases, the linear-order interpretation is largely sufficient, but we still keep the term to reserve the option to determine its exact form numerically.
+\footnote{For many cases, the linear-order interpretation is sufficient.
+We keep the general term to reserve the option to determine its exact form numerically.
 }
@@
 
-The third equality of Eq.(\ref{eq:main}) follows from assumption of translational invariance of statistics of the 21\,cm field. 
-For the inequality in the fourth line we have assumed that the  3D power spectrum varies negligibly over the $k$-space of interest and thus pulled out the integral a power spectrum centered at 
+The third equality of Eq.(\ref{eq:main}) follows from assuming the translational invariance of statistics of the 21\,cm field. 
+For the inequality in the fourth line, we have assumed that the  3D power spectrum varies negligibly over the $k$-space of interest and have thus pulled out of the integral the power spectrum centered at
 \[
@@
 Notice that $\Theta$ has no dependence on $\tau$.
-We point out that although all our derivations focused on drift-scan telescopes, we can get the analogous result for tracking measurements simply by noticing that for a tracking primary beam, and so by the radial-symmetry of the primary beam, we have:
+We point out that although all our derivations focused on drift-scan telescopes, we can get the analogous result for tracking measurements simply by noticing that for a tracking primary beam with radial symmetry we have:
 \[
@@
 
-Roughly speaking, Eqs. \eqref{eq:opp} through \eqref{eq:Thetanu} tell us that the cross multiplications of visibilities at a time delay is proportional to the power spectrum times the Fourier
+Roughly speaking, Eqs. \eqref{eq:opp} through \eqref{eq:Thetanu} tell us that the product of visibilities at a time delay is proportional to the power spectrum times the Fourier
 transform of the cross multiplied beam patterns. As a check, when applied to equivalent baselines,

### 2017-07-21 Rotation paper/draft_working.tex
merge with adrian

--- 2017-07-21 Rotation paper/draft_working.tex
@@
 \eqref{eq:Theta}).
-From now one we use $\Theta_{\boldsymbol{bb'}}$ and $\Delta t_{\boldsymbol{bb'}}$ to denote the peak value and location of $\Theta$, or in other words the maximum of the redundancy weight $|\Theta(\Delta t)|$ and the corresponding offset for a given baseline pair $\{\boldsymbol{b}:\boldsymbol{b'}\}$.
+From now on we use $\Theta_{\boldsymbol{bb'}}$ and $\Delta t_{\boldsymbol{bb'}}$ to denote the peak value and location of $\Theta$, i.e.
+the maximum of the redundancy weight $|\Theta(\Delta t)|$ and the corresponding offset for a given baseline pair $\{\boldsymbol{b}:\boldsymbol{b'}\}$.
 %Intuitively we expect the sensitivity to depend on both the $uv$ coverage of the baseline and the patch of sky inside the beam.
@@
 we show $\Theta_{\boldsymbol{bb'}}$ and $\Delta t_{\boldsymbol{bb'}}$ for a variety of baseline combinations. Baseline pairs that are mirror images of each other 
-give the same amount of redundancies ($\Theta_{\boldsymbol{bb'}}$), with the opposite time offset, as expected from symmetry. 
+give the same amount of redundancy ($\Theta_{\boldsymbol{bb'}}$), with the opposite time offset, as expected from symmetry. 
 For example, \{1,0:1,1\} is mirror image of \{1,0:1,-1\} and these two pairs of baselines
@@
 that are closer to zero time offset are higher. This is expected for two reasons;
-one is that the longer the time delay that produces redundancy, the more the sky has moved with respect
-to the beams and hence the smaller the overlapping patch of sky surveyed. The other reason is that smaller optimal
-time offset corresponds to smaller differences in orientation and length of the pair of baselines and hence better redundancy.
+one is that the longer the time offset of maximum redundancy, the more the sky has moved with respect
+to the primary beams and hence the smaller the overlapping patch of sky surveyed. The other reason is that smaller optimal
+time offset corresponds to smaller $w$-term-induced decoherence and hence better redundancy.
 
 To determine that actual relative contribution to sensitivity of these
-baseline pairs, we have to account for the multiplicities of
-these baseline classes or, in other words, the number of antenna pairs with the
+baseline pairs, we also have to account for the multiplicities of
+these baseline classes, i.e. the number of antenna pairs with the
 same length and orientation.
@@
 or \{1,1\}. Assuming that each equivalent baseline delivers
-the same $\Theta_{\boldsymbol{bb}}$, the relative contribution to sensitivity can be estimated.
+the same $\Theta_{\boldsymbol{bb}}$, the relative contribution to sensitivity can be estimated as follows.
 
-First we can average the visibilities of the equivalent baselines. Since the core of PAPER-128 has 16 by 7 antenna configuration, there
+First we can average the visibilities of the equivalent baselines. Since the core of PAPER-128 is a 16 by 7 antenna configuration, there
 are $M\equiv(16-|m|)\times(7-|n|)$ copies of the baseline class $\{m,n\}$. This means
@@
 \begin{equation}
-\widetilde{\Theta_{\boldsymbol{bb'}}} \propto \Theta_{\boldsymbol{bb'}}\sqrt{MM'}.\label{eq:sensul}
+\widetilde{\Theta}_{\boldsymbol{bb'}} \propto \Theta_{\boldsymbol{bb'}}\sqrt{MM'}.\label{eq:sensul}
 \end{equation}
@@
 \begin{aligned}
- \hat{P}(k_{\tau}) &= \frac{\sum_{\boldsymbol{bb'}}\hat{P}(k_{\bar{b}, \tau})/\sigma_P^2(\boldsymbol{bb'})}{\sum_{\boldsymbol{bb'}}1/\sigma_P^2(\boldsymbol{bb'})}, \\
+ \hat{P}(k_{\tau}) &= \frac{\sum_{\boldsymbol{bb'}}\hat{P}(k_{\bar{b}, \tau})/\sigma_P^2(\boldsymbol{bb'})}{\sum_{\boldsymbol{bb'}}1/\sigma_P^2(\boldsymbol{bb'})}\\
  &= \frac{\sum_{\boldsymbol{bb'}}\hat{P}(k_{\bar{b}, \tau})\widetilde{\Theta}_{\boldsymbol{bb'}}^2}{\sum_{\boldsymbol{bb'}}\widetilde{\Theta}_{\boldsymbol{bb'}}^2},
@@
 \end{equation}
-where the sum is over classes of baseline pairs, and the power spectrum noise covariance $\sigma_P^2$ is proportional to the effective weight $\widetilde{\Theta}$.
+where the sum is over classes of baseline pairs, and the power spectrum noise covariance, $\sigma_P^2$, is proportional to the effective weight, $\widetilde{\Theta}$.
 
@@
 
-The scaling in Eq. \eqref{eq:sensul} is rough for simplicity of motivation. As we derive in Appendix \ref{sec:appB}, this weight should be corrected by a factor proportional $\rho_0$:
+The scaling in Eq. \eqref{eq:sensul} is rough for simplicity of motivation. As we derive in Appendix \ref{sec:appB}, this weight should be corrected by a factor proportional to $\rho_0$:
 
@@
 
-For a given $\rho_0$ Eq.
+For a given $\rho_0$, Eq.
 \eqref{eq:tildereal} quantifies the relative sensitivity contribution of a baseline pair $\{\boldsymbol{b}:\boldsymbol{b'}\}$.
-Assuming a reionization signal of $\sim 30\,\text{mK}$, observation centered at 150\,MHz ($z=8.5$), and 120 days of integration with PAPER antennas, we have roughly
-(See Eq.(20) in \cite{first-paper})
+Assuming a reionization signal of $\sim 30\,\text{mK}$, observation centered at 0.15\,GHz ($z=8.5$), and 120 days of integration with PAPER antennas, we have roughly
+(see Eq.(20) in \cite{first-paper})
 \begin{equation}
-\rho_0 \sim 0.001\left[\frac{L}{40\,\text{m}}\right] \left[\frac{0.1h\,\text{Mpc}^{-1}}{k}\right]^3, 
+\rho_0 \sim 0.001\left[\frac{\bar{b}}{40\,\text{m}}\right] \left[\frac{0.1h\,\text{Mpc}^{-1}}{k}\right]^3, 
 \end{equation}
@@
 \subsection{Array Configuration Comparisons \label{sec:arrconf}}
-We run our algorithm over all possible baseline-pairs of  PAPER-128, HERA37, HERA128, HERA243 and HERA350. The HERA antenna configurations are shown in Fig. \ref{fig:HeraAntpos}. The  
+We run our algorithm over all possible baseline-pairs of  PAPER-128, HERA-37, HERA-128, HERA-243 and HERA-350. The HERA antenna configurations are shown in Fig. \ref{fig:HeraAntpos}. The  
 hexagonal design is the densest pattern of antenna-packing.
-The larger arrays are designed with a ``gap'' dividing the antennas into three different groups.
+The larger arrays are fractured with a ``gap'' dividing the antennas into three different groups.
 The gaps are designed so as to improve $uv$ coverage and ease calibration without compromising sensitivity, but also produces many more nearly equivalent baselines than a pure hexagonal layout.
@@
 \caption{Pairplots of the top contributing baseline pairs in three arrays.
-Plotted properties are optimal time delay $\Delta t_{\boldsymbol{bb'}}$,  effective weight $\widetilde{\Theta}_{\boldsymbol{bb'}}$ and baseline length $\bar{b}$.
+Plotted properties are optimal time delay $\Delta t_{\boldsymbol{bb'}}$,  effective weight $\widetilde{\Theta}_{\boldsymbol{bb'}}$, and baseline length $\bar{b}$.
 Only those points with $\widetilde{\Theta}_{\boldsymbol{bb'}}>0.01$ are shown (the weight for the top class of equivalent pair is normalized to 1).
-Only one baseline length is shown since all top contributing pairs have very similar lengths as expected.
-The scatter plots are shown with transparency so that darker regions indicate degeneracies.
-Scatter points of the 3 arrays overlap, in order indicated by the legend.
+Scatter points of the three arrays overlap, in the order indicated by the legend.
 }
@@
 The points at $\Delta t=0$ are the equivalent baselines.
-Note HERA arrays have fewer data points with high $\Delta t_{\boldsymbol{bb'}}$, due to their comparatively narrow beam.
-Also note that all near-equivalent pairs for HERA arrays appear at $\widetilde{\Theta}_{\boldsymbol{bb'}}\lesssim 0.4$.
+HERA arrays have fewer data points with high $\Delta t_{\boldsymbol{bb'}}$, due to their comparatively narrow beam.
+All near-equivalent pairs for HERA arrays appear at $\widetilde{\Theta}_{\boldsymbol{bb'}}\lesssim 0.4$.
 
@@
 HERA arrays do not exhibit this trend here because of a selection effect.
-Only baseline pairs with $\widetilde{\Theta}_{\boldsymbol{bb'}}>0.01$ are shown.
 Shorter HERA baselines require much longer $\Delta t$ to overlap, partly because of the hexagonal structure requires $60\degree$ of rotation to overlap baselines, partly because of the smaller primary beam.
 Thus most short HERA baseline pairs are not shown in this figure.
+Only baseline pairs with $\widetilde{\Theta}_{\boldsymbol{bb'}}>0.01$ are shown.
 
@@
 The wide structures spanning the whole range of $\bar{b}$ are the near-equivalent baseline pairs.
-We see again for HERA arrays all near-equivalent pairs appear at $\widetilde{\Theta}_{\boldsymbol{bb'}}\lesssim 0.4$.
+We see that for HERA arrays, all near-equivalent pairs appear at $\widetilde{\Theta}_{\boldsymbol{bb'}}\lesssim 0.4$.
 For PAPER-128, the existence of short near-equivalent pairs with high multiplicity partially fills the high $\widetilde{\Theta}_{\boldsymbol{bb'}}$ region.
@@
 The trend is particularly obvious in the linear structure for the HERA arrays, which are the equivalent baselines.
-Since they all have the unity $\Theta_{\boldsymbol{bb'}}$ by definition, the linearly decreasing trend of $\widetilde{\Theta}_{\boldsymbol{bb'}}$ is a direct measure of the baseline multiplicity structures of the HERA array configurations.
+All having unity $\Theta_{\boldsymbol{bb'}}$ by definition, the linearly decreasing trend of $\widetilde{\Theta}_{\boldsymbol{bb'}}$ is a direct measure of the baseline multiplicity structures of the HERA array configurations.
 
@@
 This is because they are longer baselines with lower multiplicity.
-In the end these baseline classes still lead to high contributions to total sensitivity (Fig.
-\ref{fig:osens}) because there are a lot more such baseline pairs for HERA.
+In the end, these baseline classes still lead to high contributions to total sensitivity (Fig.
+\ref{fig:osens}) because there are many more such baseline pairs for HERA.
 
@@
 
-Clearly not all baseline pairs give significant contribution to sensitivity.
+As Fig.
+\ref{fig:pairplot} suggests, not all baseline pairs contribute significantly to sensitivity.
 Due to the large number of baseline pairs, it may not be computationally feasible to include all pairs.
 Having quantified the sensitivity from a given pair of baselines, we study the cumulative sensitivity of the array depending on which baseline pairs we include.
-Evidently we should prefer the pairs with larger $\widetilde{\Theta}_{\boldsymbol{bb'}}$.
+By construction, we prefer the pairs with larger $\widetilde{\Theta}_{\boldsymbol{bb'}}$.
 In Fig.
-\ref{fig:osens} we plot $\rho$ against the minimum $\widetilde{\Theta}_{\boldsymbol{bb'}}$.
-$\rho(\widetilde{\Theta}_{min})$ is the sensitivity of the array when baseline-pairs that have $\widetilde{\Theta}_{\boldsymbol{bb'}}>\widetilde{\Theta}_{min}$ are included.
+\ref{fig:osens}, we plot fractional sensitivity $\rho$ as a function of $\widetilde{\Theta}_{min}$, the minimum cutoff for $\widetilde{\Theta}_{\boldsymbol{bb'}}$.
+In other words $\rho(\widetilde{\Theta}_{min})$ is the sensitivity when all baseline pairs with $\widetilde{\Theta}_{\boldsymbol{bb'}}>\widetilde{\Theta}_{min}$ are included.
+
+Note that here we normalized sensitivity for each array as the fraction of the total sensitivity of an array, i.e.
+when all baseline pairs are used.
+The plot therefore does not compare absolute sensitivity across different arrays.
+We see as expected that in all cases using the nearly equivalent baselines lead to more and more significant improvements with lower $\widetilde{\Theta}_{min}$, or in other words when more baseline pairs are used.
+We point out that PAPER results until now \citep{Ali2015, paper32} only used the top contributing baseline pairs.
+In other words these results only explored sensitivity $\rho(\widetilde{\Theta}_{min}\sim 1)$.
+
 The dashed lines represent the values when only the equivalent baseline-pairs are used.
-We see as expected that in all cases using the nearly equivalent baselines lead to more and more significant improvements with lower $\widetilde{\Theta}_{min}$, or in other words when more baseline pairs are used.
-The small HERA37, with no gap (like in HERA350) or short nearly equivalent baselines (like in PAPER 128), will not benefit much from the nearly equivalent baselines.
+The small and unfractured HERA-37, with no gap (like in HERA-350) or short nearly equivalent baselines (like in PAPER 128), will not benefit much from the nearly equivalent baselines.
 Near-equivalent baselines make an increasingly significant contribution to the total sensitivity as the number of array elements are increased.
-PAPER-128 is designed with highly redundant nearly equivalent baselines, and thus  these baselines start contributing at higher $\widetilde{\Theta}_{min}$, but the gapped HERA configurations will benefit even more from nearly equivalent baselines at low $\widetilde{\Theta}_{min}$ due to the presence of more classes of such pairs.
-Note that here we normalized sensitivity for each array as the fraction of the total sensitivity when all baseline pairs are used.
-The plot therefore does not compare absolute sensitivity across different arrays.
+PAPER-128 is designed with highly redundant nearly equivalent baselines, and thus  these baselines start contributing at higher $\widetilde{\Theta}_{min}$.
+The fractured HERA configurations will benefit even more from nearly equivalent baselines at low $\widetilde{\Theta}_{min}$ due to the presence of more classes of such pairs.
 The stepwise pattern seen in almost all cases are characteristic of a regular grid; as we step to lower $\widetilde{\Theta}_{\boldsymbol{bb'}}$ large groups of baseline pair classes get included in ``batch''.
-Thus the ``gapless'' HERA37, with more regularity in its antenna configurations than the larger counterparts, exhibit more step patterns in fractional sensitivity.
+Thus the ``unfractured'' HERA-37, with more regularity in its antenna configurations than the larger counterparts, exhibit more step patterns in fractional sensitivity.
 %\begin{centered}
@@
 
-One benefit of delay-transform based power spectrum estimation is the ability to isolation foreground in Fourier space.
+One benefit of delay-transform based power-spectrum estimation is the ability to isolate foregrounds in Fourier space.
 We show that our proposed estimator does not lead to extra foreground leakage into the EoR window.
 We then apply our estimator to a variety array configurations.
-We show that for different configurations of PAPER and HERA, the incorporation of partial redundancy accounts for $20\%$ to $60\%$ of the array's sensitivity.
-The proposed power spectrum estimator retains benefits of delay-transform techniques for isolating foregrounds and avoiding systematics and enables one to unlock the full sensitivity of a redundant radio interferometer while simultaneously enjoying the benefits of delay-transform techniques for isolating foregrounds and avoiding systematics.
+We show that for different configurations of PAPER and HERA, the incorporation of partial redundancy accounts for $20\%$ to $60\%$ of the array's total sensitivity.
+The proposed power-spectrum estimator retains benefits of delay-transform techniques for isolating foregrounds and avoiding systematics and enables one to unlock the full sensitivity of a redundant radio interferometer while simultaneously enjoying the benefits of delay-transform techniques for isolating foregrounds and avoiding systematics.
 

### 2017-07-22 Rotation paper/draft_working.tex
Added lots of citations to chromaticity section

@@
 A crucial concern when measuring the $21\,\textrm{cm}$ power spectrum is the possibility of astrophysical foreground contamination.
-Being orders of magnitude larger than the cosmological signal, even small residuals from an imperfect isolation of these foregrounds can masquerade as a false detection.
-Recent literature treatments of this problem have shown that if one combines the mathematical properties of interferometric measurement with the empirical fact that foregrounds are spectrally smooth, a clean separation of foregrounds and cosmological signal can in principle be achieved in harmonic space \citep{wedge1, wedge2}.
-These studies suggest that foregrounds preferentially appear in harmonic modes of the sky corresponding to fluctuations that are angularly fine but spectrally smooth, a region in harmonic space colloquially known as ``the wedge".
-One's foreground mitigation strategy can then consist mainly of avoiding those modes by working in the complementary region known as ``the EoR window".
+Because foregrounds are orders of magnitude larger than the cosmological signal, even small residuals from an imperfect isolation of these foregrounds can masquerade as a false detection.
+Recent literature treatments of this problem have shown that if one combines the mathematical properties of interferometric measurement with the empirical fact that foregrounds are spectrally smooth, a clean separation of foregrounds and cosmological signal can be achieved in harmonic space.
+These studies suggest that foregrounds preferentially appear in harmonic modes of the sky corresponding to fluctuations that are angularly fine but spectrally smooth, a region in harmonic space colloquially known as ``the wedge" \citep{Datta2010,Vedantham2012,Morales2012,delay-transform,Trott2012,Thyagarajan2013,pober_et_al2013b,dillon_et_al2014,Hazelton2013,Thyagarajan_et_al2015a,Thyagarajan_et_al2015b,wedge1, wedge2,chapman_et_al2016,pober_et_al2016,seo_and_hirata2016,jensen_et_al2016,kohn_et_al2016}.
+A foreground mitigation strategy can then consist mainly of avoiding those modes by working in the complementary region known as ``the EoR window".
 However, for this strategy to be viable, it is necessary to show that one's data analysis pipeline does not corrupt the separation between the wedge and the EoR window.
@@
 As discussed in previous sections, the combination of information from nearly equivalent baselines for wide-field telescopes is inherently a three-dimensional problem.
-Although the estimator of Section \ref{sec:method} does not have such limitations, to discuss foreground signature with possible curved sky effects, we must employ a set of basis functions where angular fluctuations are encoded by spherical harmonics, and radial fluctuations by spherical Bessel functions.
+Although the estimator of Section \ref{sec:method} does not have such limitations, to discuss foreground signatures with possible curved sky effects, we must employ a set of basis functions where angular fluctuations are encoded by spherical harmonics, and radial fluctuations by spherical Bessel functions.
 
-\cite{sphere} discusses in detail power spectrum analyses using spherical harmonics.
+\cite{sphere} discusses in detail power-spectrum analyses using spherical harmonics.
 In such as basis, modes in the sky are indexed by the magnitude $k$ of the wavevector $\mathbf{k}$ and spherical harmonic indices $(\ell, m)$.
@@
 \begin{equation}
-W_{\ell}(k; \mathbf{b}, \mathbf{b}^\prime , \tau) = \frac{2 k^2}{\pi \Theta} \sum_m g_{\ell m} (k ; \mathbf{b}, \tau) g_{\ell m}^{\psi*} (k ; \mathbf{b}^\prime, \tau).
+W_{\ell}(k; \mathbf{b}, \mathbf{b}^\prime , \tau) \equiv \frac{2 k^2}{\pi \Theta} \sum_m g_{\ell m} (k ; \mathbf{b}, \tau) g_{\ell m}^{\psi*} (k ; \mathbf{b}^\prime, \tau).
 \end{equation}
@@
 Figure \ref{fig:multiblwindows} shows four example window functions.
-The two window functions in the top plot are for equivalent baseline power spectrum estimators, i.e., where one squares the delay spectrum from a single baseline and normalizes the result.
+The two window functions in the top plot are for equivalent baseline power-spectrum estimators, i.e., where one squares the delay spectrum from a single baseline and normalizes the result.
 The window function localized at low $\ell$ is for the baseline \{2,0\} of around $30\,\textrm{m}$ in length, while the window function localized at high $\ell$ is for \{12,0\}, roughly $180\,\textrm{m}$ in length.

### 2017-07-22 Rotation paper/draft_working.tex

--- 2017-07-21 Rotation paper/draft_working.tex
@@
 If $\psi_{\nu}=0$ in  Eq.
-\eqref{eq:Thetanu}, $\Theta_{\nu}$ at the peak of correlation is in general complex, and often far from real.
+\eqref{eq:Thetanu}, $\Theta_{\nu}$ at the peak of correlation ( as a function of time) is generally complex, and often far from real.
 Furthermore, this phase of peak correlation is inevitably frequency dependent.
-This frequency dependence of the phase would lead to destructive interference when we integrate over frequency, unless we correct $\Theta_{\nu}$ by an extra phase $\psi_\nu$.
+This frequency dependence would lead to destructive interference when we integrate over frequency, unless we correct $\Theta_{\nu}$ by a phase $\psi_\nu$.
 
@@
 By default the correlators of a drift-scan array phase the two visibilities both to zenith at the same time.
-When they are cross-multiplied with a time lag, the visibilities must be rephased before delay transform to account for the movement of the phase center.
-The effect of the phase thus roughly corresponds in a shift of the delay mode measured, and we should expect $\psi_{\nu}$ to be in first order a linear function of $\nu$:
+When they are cross-multiplied with a time lag, the visibilities must be rephased before the delay transform to account for the movement of the phase center.
+The effect of the phase thus roughly corresponds in a shift of the delay mode measured, and we should expect $\psi_{\nu}$ to be approximately a linear function of $\nu$:
 \[
@@
 In Fig.
-\ref{fig:phi_nu} we show the phases of $\Theta_\nu$ at the optimal time offset or, in other words, when $|\Theta_\nu|$ is maximized.
+\ref{fig:phi_nu} we show the phases of $\Theta_\nu$ at the optimal time offset, i.e.
+when $|\Theta_\nu|$ is maximized.
 The phases are shown for the baseline pair \{2,0:2,1\} of PAPER-128.
@@
 The need for rephasing can thus be understood as a symptom of the underlying spatial decoherence; coherence at the phase center would not extend to the entire beam pattern (see Section \ref{sec:visual}).
-Thus while integrating over frequencies and sky-direction in Eq.
-\eqref{eq:Theta} there are necessarily sky-directions that do not add coherently.
+While integrating over frequencies and sky-direction in Eq.
+\eqref{eq:Theta}, there are necessarily sky-directions that do not add coherently.
 This decoherence across the beam pattern cannot be removed and plays a fundamental role in determining how much sensitivity one can recover from nearly equivalent baselines.
@@
 \caption{Planned Hydrogen Epoch of Reionization Array antenna configurations.
-HERA 37 is expected to complete and start collecting data in summer of 2017, and the other three are planned configurations in the next phases.
+HERA 37 is expected to complete and start collecting data in August of 2017, and the other three are planned configurations in the next phases.
 For HERA350, only the 320 elements in the core are shown.

### 2017-07-22 Rotation paper/draft_save.tex
incorporated Aaron's last round comments

--- 2017-07-22 Rotation paper/draft_working.tex
@@
 A crucial concern when measuring the $21\,\textrm{cm}$ power spectrum is the possibility of astrophysical foreground contamination.
-Because foregrounds are orders of magnitude larger than the cosmological signal, even small residuals from an imperfect isolation of these foregrounds can masquerade as a false detection.
-Recent literature treatments of this problem have shown that if one combines the mathematical properties of interferometric measurement with the empirical fact that foregrounds are spectrally smooth, a clean separation of foregrounds and cosmological signal can be achieved in harmonic space.
-These studies suggest that foregrounds preferentially appear in harmonic modes of the sky corresponding to fluctuations that are angularly fine but spectrally smooth, a region in harmonic space colloquially known as ``the wedge" \citep{Datta2010,Vedantham2012,Morales2012,delay-transform,Trott2012,Thyagarajan2013,pober_et_al2013b,dillon_et_al2014,Hazelton2013,Thyagarajan_et_al2015a,Thyagarajan_et_al2015b,wedge1, wedge2,chapman_et_al2016,pober_et_al2016,seo_and_hirata2016,jensen_et_al2016,kohn_et_al2016}.
-A foreground mitigation strategy can then consist mainly of avoiding those modes by working in the complementary region known as ``the EoR window".
+Being orders of magnitude larger than the cosmological signal, even small residuals from an imperfect isolation of these foregrounds can masquerade as a false detection.
+Recent literature treatments of this problem have shown that if one combines the mathematical properties of interferometric measurement with the empirical fact that foregrounds are spectrally smooth, a clean separation of foregrounds and cosmological signal can in principle be achieved in harmonic space \citep{wedge1, wedge2}.
+These studies suggest that foregrounds preferentially appear in harmonic modes of the sky corresponding to fluctuations that are angularly fine but spectrally smooth, a region in harmonic space colloquially known as ``the wedge".
+One's foreground mitigation strategy can then consist mainly of avoiding those modes by working in the complementary region known as ``the EoR window".
 However, for this strategy to be viable, it is necessary to show that one's data analysis pipeline does not corrupt the separation between the wedge and the EoR window.
@@
 As discussed in previous sections, the combination of information from nearly equivalent baselines for wide-field telescopes is inherently a three-dimensional problem.
-Although the estimator of Section \ref{sec:method} does not have such limitations, to discuss foreground signatures with possible curved sky effects, we must employ a set of basis functions where angular fluctuations are encoded by spherical harmonics, and radial fluctuations by spherical Bessel functions.
+Although the estimator of Section \ref{sec:method} does not have such limitations, to discuss foreground signature with possible curved sky effects, we must employ a set of basis functions where angular fluctuations are encoded by spherical harmonics, and radial fluctuations by spherical Bessel functions.
 
-\cite{sphere} discusses in detail power-spectrum analyses using spherical harmonics.
+\cite{sphere} discusses in detail power spectrum analyses using spherical harmonics.
 In such as basis, modes in the sky are indexed by the magnitude $k$ of the wavevector $\mathbf{k}$ and spherical harmonic indices $(\ell, m)$.
@@
 \begin{equation}
-W_{\ell}(k; \mathbf{b}, \mathbf{b}^\prime , \tau) \equiv \frac{2 k^2}{\pi \Theta} \sum_m g_{\ell m} (k ; \mathbf{b}, \tau) g_{\ell m}^{\psi*} (k ; \mathbf{b}^\prime, \tau).
+W_{\ell}(k; \mathbf{b}, \mathbf{b}^\prime , \tau) = \frac{2 k^2}{\pi \Theta} \sum_m g_{\ell m} (k ; \mathbf{b}, \tau) g_{\ell m}^{\psi*} (k ; \mathbf{b}^\prime, \tau).
 \end{equation}
@@
 Figure \ref{fig:multiblwindows} shows four example window functions.
-The two window functions in the top plot are for equivalent baseline power-spectrum estimators, i.e., where one squares the delay spectrum from a single baseline and normalizes the result.
+The two window functions in the top plot are for equivalent baseline power spectrum estimators, i.e., where one squares the delay spectrum from a single baseline and normalizes the result.
 The window function localized at low $\ell$ is for the baseline \{2,0\} of around $30\,\textrm{m}$ in length, while the window function localized at high $\ell$ is for \{12,0\}, roughly $180\,\textrm{m}$ in length.

### 2017-07-22 Rotation paper/draft_working.tex
incorporated Aaron's last round comments

--- 2017-07-22 Rotation paper/draft_working.tex
@@
 Baselines that are slightly different in length and orientation
-``rotate into'' each other at a time delay.
+``rotate into'' each other at a time offset.
 Baselines of different lengths and orientations therefore contain partially coherent information.
@@
 for all baselines.
-Power-spectrum measurements are typically taken from $\sim10$\,MHz centered around the corresponding redshift of interest (e.g.
-150\,MHz for redshift $z\approx9.5$), thus imposing a sharp drop-off on $\phi(\nu)$.
+Power-spectrum measurements are typically taken from $\sim0.01$\,GHz centered around the corresponding redshift of interest (e.g.
+0.15\,GHz for redshift $z\approx9.5$), thus imposing a sharp drop-off on $\phi(\nu)$.
 
@@
 \]
-where $r$ is the magnitude of the vector $\bold{r}$, $\nu_{21}=1420$\,MHz is the 21\,cm transition rest frequency, $\nu_{0}$ is
+where $r$ is the magnitude of the vector $\bold{r}$, $\nu_{21}=1.420$\,GHz is the 21\,cm transition rest frequency, $\nu_{0}$ is
 a reference central frequency with corresponding redshift $z_{0}$, $H_0$ is the current-day Hubble constant,  
@@
 \footnote{For many cases, the linear-order interpretation is sufficient.
-We keep the general term to reserve the option to determine its exact form numerically.
+We keep the general term to reserve the option to determine its exact form numerically as this information comes at no additional computational cost.
 }
@@
 
-Roughly speaking, Eqs. \eqref{eq:opp} through \eqref{eq:Thetanu} tell us that the product of visibilities at a time delay is proportional to the power spectrum times the Fourier
+Roughly speaking, Eqs. \eqref{eq:opp} through \eqref{eq:Thetanu} tell us that the product of visibilities at a time offset is proportional to the power spectrum times the Fourier
 transform of the cross multiplied beam patterns. As a check, when applied to equivalent baselines,
@@
 \eqref{eq:opp} and Eq.
-\eqref{eq:Theta} we can, for any given pair of baseline classes and time delay, estimate the degree of redundancy, here represented by $\Theta$.
+\eqref{eq:Theta} we can, for any given pair of baseline classes and time offsets, estimate the degree of redundancy, here represented by $\Theta$.
 This allows us to achieve our goals stated in the introduction: to identify
@@
 In the drift-scan case, we indeed see a linear relation corresponding to a delay of $\Delta\tau\approx15.7$\,ns.
+This delay corresponds to the light travel time from one antenna to another of the rotated baseline.
 In the tracking case, since the phase center is fixed to the sky, only second order effects are observed.
@@
 
-In this section we delve into visualizing and testing our method, and explore sensitivity contributions of various baseline classes for PAPER and HERA array configurations. 
+In this section we delve into visualizations and analyses of our method, and explore sensitivity contributions of various baseline classes for PAPER and HERA array configurations. 
 \subsection{Visualization \label{sec:visual}}
+
+\begin{figure*}[h!]
+\includegraphics[width=\textwidth]{BeamFringe}
+
+\caption{PAPER-128 beam-fringe patterns and their point-spread functions on the $uv$-plane for a frequency of $0.15$\,GHz.
+Selected baselines have 30-meter East-West separations, and 0- and 4-meter North-South separations, respectively.
+The top panels show the real parts of the beam-fringe patterns of the baselines (left and middle), and their conjugate product (right).
+The beam-fringe values are normalized such that the peak of the original beam is unity.
+The bottom-left and middle panels show the peak-normalized $uv$ point spread of the two beam-fringe patterns.
+The product of the beam spreads displayed in the bottom-right panel shows that power concentrated at $u\approx15$ is recovered.
+}
+\label{fig:beamfringe}
+\end{figure*}
 
@@
 To gain some visual intuition of the formalism in Section \ref{sec:method}, we show explicit beam fringe patterns of the two baselines and their interactions.
-For this section, we relax into the flat sky approximation so that $(l,m,\sqrt{1-l^2-m^2})\equiv\hat{s}$ and examine a single frequency of $150$\,MHz.
+For this section, we relax into the flat sky approximation so that $(l,m,\sqrt{1-l^2-m^2})\equiv\hat{s}$ and examine a single frequency of $0.15$\,GHz.
 
 At the time offset of maximum redundancy, we expect two baselines to have to same fringe pattern in both frequency and phase.
-Due to the time delay, however, the beam centers would be slightly shifted with respect to each other.
+Due to the time offset, however, the beam centers would be slightly shifted with respect to each other.
 In the top panels of Fig.
@@
 The top right panel show their conjugate product $g_{b}^{*}g_{b'}e^{-i\psi_\nu} = B^{*}(\hat{\boldsymbol{s}},\nu)B(\boldsymbol{\Gamma}\hat{\boldsymbol{s}},\nu)\exp\left[i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma}\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right]$, where we have chosen global phase $\psi_{\nu}$ such that the peak of the conjugate product is real.
-The product of the beam fringe patterns in top right shows that the fringes
+The product of the beam fringe patterns in the top right shows that the fringes
  cancel out as we expect.
 Comparing with Eq.
-\eqref{eq:Thetanu}, we see that $\Theta_\nu=\int d\Omega g_{b}^{*}g_{b'}$ is the sum over the image shown in the top right panel.
-Thus we show that at the optimal time offset the fringe patterns of the two baselines cancel out, leading to a substantial weight $\Theta_\nu$.
-The frequency-summed weight $\Theta$ is constructed with rephasing as described in Section \ref{sec:rephs} so that the conjugate product (top right panel) for different frequencies align in phase.
+\eqref{eq:Thetanu}, we see that $\Theta_\nu=\int d\Omega g_{b}^{*}g_{b'}e^{-i\psi_\nu}$ is the sum over the image shown in the top right panel.
+At the optimal time offset, the fringe patterns of the two baselines cancel out, leading to a substantial weight $\Theta_\nu$.
+The frequency-summed weight $\Theta$ is constructed with rephasing as described in Section \ref{sec:rephs} so that the conjugate product (top-right panel) for different frequencies align in phase.
 
-Similarly, we show the correlation of the two baselines in terms of their $uv$ beam spread overlap.
-The bottom left and middle panels are (peak-normalized) instantaneous $uv$ coverage of two baselines ($\widetilde{g}_{b}$ and $\widetilde{g}_{b'}$) while the bottom right panel is their product.
-We see from the bottom right panel that the cross-multiplied baselines recover power concentrated at $u\approx15$.
-The product of the $uv$ beams spread shown in the bottom right panel should not be confused with the Fourier transform of the beam-fringe conjugate product shown in the top right panel.
-
-\begin{figure*}[h!]
-\includegraphics[width=\textwidth]{BeamFringe}
-
-\caption{PAPER-128 beam-fringe patterns and their point-spread functions on the $uv$-plane for a frequency of $150$\,MHz.
-Selected baselines have 30 meter East-West separation, and 0 and 4 meter North-South separation, respectively.
-The top panels show the real parts of the beam-fringe patterns of the baselines (left and middle), and their conjugate product (right).
-The beam-fringe values are normalized such that the peak of the original beam is unity.
-The bottom left and middle panels show the peak-normalized $uv$ beam power of the two beam-fringe patterns.
-The product of the beam spreads displayed in the bottom right panel shows that power concentrated at $u\approx15$ is recovered.
-}
-\label{fig:beamfringe}
-\end{figure*}
+Similarly, we show the correlation of the two baselines in terms of the overlap of their point-spread functions in the $uv$-plane.
+The bottom-left and middle panels are the peak-normalized instantaneous $uv$ coverage of two baselines ($\widetilde{g}_{b}$ and $\widetilde{g}_{b'}$) while the bottom right panel is their product.
+We see from the bottom right panel that the cross-multiplied baselines' recovered power is concentrated at $u\approx15$.
+The product of the $uv$ point-spread functions shown in the bottom-right panel should not be confused with the Fourier transform of the beam-fringe conjugate product shown in the top-right panel.
 
@@
 For numerical convenience we assumed that the bandpass $\phi_\nu$ (see Eq.
-\eqref{eq:B}) takes the form of a half-period sine curve that peaks at $150\,\textrm{\,MHz}$ and goes to zero $5\,\textrm{\,MHz}$ on either side of the peak.
+\eqref{eq:B}) takes the form of a half-period sine curve that peaks at $0.15\,\textrm{\,GHz}$ and goes to zero $5\,\textrm{\,MHz}$ on either side of the peak.
 
@@
 \ref{fig:sensplot}
-we see for example \{1,0\} has higher multiplicity than \{2,0\},
+we see for example that \{1,0\} has higher multiplicity than \{2,0\},
 or \{1,1\}. Assuming that each equivalent baseline delivers
@@
 that if we add visibility measurements of all the equivalent baselines,
-we get a factor of $\sqrt{M}$ reduction in noise
+we get a factor of $\sqrt{M}$ reduction in the noise
 level $\sigma_N$ of the visibility.
@@
 by the multiplicity factor.
-Points that have zero time delay are the equivalent baseline pairs and their weighted correlation values simply reflect the multiplicity factor.
-For clarity of presentation we have ``folded over'' the negative time delays and combined baseline pairs that are mirror images.
+Points that have zero optimal time offset are the equivalent baseline pairs and their weighted correlation values simply reflect the multiplicity factor.
+For clarity of presentation we have ``folded over'' the negative time offsets and combined baseline pairs that are mirror images.
 
@@
 \caption{Pairplots of the top contributing baseline pairs in three arrays.
-Plotted properties are optimal time delay $\Delta t_{\boldsymbol{bb'}}$,  effective weight $\widetilde{\Theta}_{\boldsymbol{bb'}}$, and baseline length $\bar{b}$.
+Plotted properties are optimal time offset $\Delta t_{\boldsymbol{bb'}}$ (in hours),  effective weight $\widetilde{\Theta}_{\boldsymbol{bb'}}$, and baseline length $\bar{b}$ (in meters).
 Only those points with $\widetilde{\Theta}_{\boldsymbol{bb'}}>0.01$ are shown (the weight for the top class of equivalent pair is normalized to 1).
@@
 \section{Conclusion}
-The 21\,cm brightness temperature is a direct tracer of reionization.
-Recently a surge of efforts in radio astronomy aims to measure the 21\,cm brightness temperature power spectrum of reionization.
-Measurement of the faint and diffuse signal requires high sensitivity, thus arrays are designed with multiple copies of the same baselines to enable repeated measurements of the same Fourier modes.
+Upcoming arrays aiming at measuring the 21\,cm power spectrum, including HERA and SKA-low, feature increasing number of antennas on regular grids to maximize coherent information.
+Having efficient pipelines to extract full sensitivity of these arrays is thus a pressing necessity.
 
@@
 \sigma_P^2 &= \langle P^2\rangle -\langle P \rangle^2,\\
-&\propto \langle \frac{(V_{1S}+V_{1N})^2 (V_{2S}+V_{2N})^2}{\Theta^2} \rangle - \langle \frac{(V_{1S}+V_{1N}) (V_{2S}+V_{2N})}{\Theta} \rangle ^2,\\
-&= \frac{1}{\Theta_{12}^2} \left( V_{1S}^2\sigma_{V2}^2+V_{2S}^2\sigma_{V1}^2+\langle V_{1N}^2 V_{2N}^2\rangle\right), \\
+&\propto \langle \frac{(V_{1S}+V_{1N})^2 (V_{2S}+V_{2N})^2}{\Theta^2} \rangle - \langle \frac{(V_{1S}+V_{1N}) (V_{2S}+V_{2N})}{\Theta} \rangle ^2\\
+&= \frac{1}{\Theta_{12}^2} \left( V_{1S}^2\sigma_{V2}^2+V_{2S}^2\sigma_{V1}^2+\langle V_{1N}^2 V_{2N}^2\rangle\right)\\
 &= \frac{1}{\Theta_{12}^2} \left[ V_{S}^2(\sigma_{V2}^2+\sigma_{V1}^2) + \sigma_{V1}^2 \sigma_{V2}^2\right], 
@@
 \begin{aligned}
-\sigma_P^2 & \propto  \frac{\sigma_0^4}{\Theta_{12}^2} \left[ \rho_0 \left(\frac{1}{M_1}+\frac{1}{M_2} \right) + \frac{1}{M_1 M_2}\right], \\
+\sigma_P^2 & \propto  \frac{\sigma_0^4}{\Theta_{12}^2} \left[ \rho_0 \left(\frac{1}{M_1}+\frac{1}{M_2} \right) + \frac{1}{M_1 M_2}\right]\\
 &\propto \frac{1}{\widetilde{\Theta}_{12}^2}.

### 2017-07-22 Rotation paper/draft_working.tex
incorporated Aaron's last round comments
few corrections

@@
 \caption{Planned Hydrogen Epoch of Reionization Array antenna configurations.
-HERA 37 is expected to complete and start collecting data in summer of 2017, and the other three are planned configurations in the next phases.
-For HERA350, only the 320 elements in the core are shown.
+HERA-37 is expected to complete and start collecting data in August of 2017, and the other three are planned configurations in the next phases.
+For HERA-350, only the 320 elements in the core are shown.
 }

### 2017-07-22 Rotation paper/draft_working.tex
save

@@
 We then apply our estimator to a variety array configurations.
-We show that for different configurations of PAPER and HERA, the incorporation of partial redundancy accounts for $20\%$ to $60\%$ of the array's total sensitivity.
+We show that for arrays with increasing number of antennas, partially redundant baselines make up a more significant portion of the array's total sensitivity.
 The proposed power-spectrum estimator retains benefits of delay-transform techniques for isolating foregrounds and avoiding systematics and enables one to unlock the full sensitivity of a redundant radio interferometer while simultaneously enjoying the benefits of delay-transform techniques for isolating foregrounds and avoiding systematics.

### 2017-07-22 Rotation paper/draft_working.tex
Fixed errors in bilbiography

@@
 \graphicspath{{../figures/}}
+\newcommand{\pasa}{PASA}
 
@@
 regarding fundamental physics of inflation, neutrino mass, and the phenomenology
-of the first stars and galaxies (e.g. \citealt{LiuOpticalDepth, Liu2016b, Mao2008, DEw21cm, Bull2015, Oyama20131186}).
+of the first stars and galaxies (e.g., \citealt{LiuOpticalDepth, Liu2016b, Mao2008, DEw21cm, Bull2015, Oyama20131186}).
 
-Observational studies of reionization, including Gunn-Peterson measurements of quasi-stellar objects \citep{Fan2006} and Cosmic Microwave Background temperature and anisotropy measurements (CMB; \citealt{Planck2016}), the Kinetic Sunyaev-Zeldo'vich effect \citep{ZahnKSZ, GeorgeKSZ, kszpatchy} and Lyman alpha galaxy observations \citep{Rhoads2017, Rhoads2012, mcquinnLyA} have given us indications of the rough time-frame of reionization, but only limited constraints on the finer spatial and temporal structures.
+Observational studies of reionization, including Gunn-Peterson measurements of quasi-stellar objects \citep{Fan2006} and Cosmic Microwave Background temperature and anisotropy measurements (CMB; \citealt{Planck2016}), the kinetic Sunyaev-Zel'dovich effect \citep{ZahnKSZ, GeorgeKSZ, kszpatchy} and Lyman alpha galaxy observations \citep{Rhoads2017, Rhoads2012, mcquinnLyA} have given us indications of the rough time frame of reionization, but only limited constraints on the finer spatial and temporal structures.
 A surge of recent radio-astronomical experiments of reionization focus on measuring the ``spin-flip'' transition of neutral
@@
 Current-generation instruments include the Donald C. Backer Precision Array for Probing
-the Epoch of Reionization (PAPER; \citealt{Ali2015,paper32}), Murchison
-Widefield Array (MWA; \citealt{Bowman2013, Tingay2013}), Low Frequency Array (LOFAR; \citealt{LOFAR}). Next-generation instruments include the Hydrogen Epoch of Reionization
-Array (HERA; e.g. \citealt{HERA,HERAconfiguration,HERABEAM1,HERADISH2}), which is under construction, 
+the Epoch of Reionization (PAPER; \citealt{Ali2015,paper32}), the Murchison
+Widefield Array (MWA; \citealt{Bowman2013, Tingay2013}), and the Low Frequency Array (LOFAR; \citealt{LOFAR}). Next-generation instruments include the Hydrogen Epoch of Reionization
+Array (HERA; e.g. \citealt{HERA,JoshAntPos,HERABEAM1,HERADISH2}), which is under construction, 
 and the Square Kilometer Array Low (SKA-low; e.g. \citealt{skalow2015} and references therein), which is currently in planning stages.
@@
 To satisfy the sensitivity needs, modern arrays are large (ranging upwards from 100 elements) and are typically of the drift-scan (static-pointing) type to limit cost.
-To further improve sensitivity, experiments such as PAPER and HERA feature multiple copies of the same baselines to repeatedly measure the same Fourier signal\citep{first-paper}.
+To further improve sensitivity, experiments such as PAPER and HERA feature multiple copies of the same baselines to repeatedly measure the same Fourier signal \citep{first-paper}.
 
@@
 This is the basis of earth-rotation synthesis.
-While imaging based power-spectrum pipelines naturally include all redundancy information, visibility-based pipelines so far only cross multiply fully redundant baselines, i.e.
-baselines of the same length and orientation.
+While imaging based power-spectrum pipelines naturally include all redundancy information, visibility-based pipelines so far only cross multiply fully redundant baselines, i.e., baselines of the same length and orientation.
 
@@
 The gaps are designed so as to improve $uv$ coverage and ease calibration without compromising sensitivity, but also produces many more nearly equivalent baselines than a pure hexagonal layout.
-The motivations behind the designs are explained in \cite{HERAconfiguration}.
+The motivations behind the designs are explained in \cite{JoshAntPos}.
 Compared to PAPER-128, the hexagonal pattern of HERA lack short baselines oriented close to each other, and thus we expect to see only longer nearly equivalent baselines with high correlations.

### 2017-07-22 Rotation paper/draft_working.tex
Reconfigured equation and figure placements; fixed typos

@@
 
-The epoch of reionization represents the last key
+The Epoch of Reionization (EoR) represents the last key
 stage of our Universe's early evolution. The study of this event stands at
@@
 The Earth's rotation causes the baselines in a drift-scan array to pick up different modes of the sky with time.
-Rotation-synthesis makes use of the rotation-induced $uv$ coverage map to form an image.
+Rotation synthesis makes use of the rotation-induced $uv$ coverage map to form an image.
 With visibility-based pipelines, the same information can be extracted.
@@
 In the bottom panel, the two baselines marked in red are an example of an equivalent pair.
-We denote a equivalency-class of baselines in the PAPER grid by their separations, in this case  \{2,0\}, for the
+We denote a equivalency class of baselines in the PAPER grid by their separations, in this case  \{2,0\}, for the
 antennas are separated by 2 units east and 0 units north. Similarly,
 the baselines marked in blue and yellow are respectively examples
-of \{2,1\} and \{2, -1\}.
-Note that \{2,0\} and \{-2, 0\} are the same class of baselines and should not be counted twice. Antennas in purely north-south baselines
+of \{2,1\} and $\{2, -1\}$.
+Note that \{2,0\} and $\{-2, 0\}$ are the same class of baselines and should not be counted twice. Antennas in purely north-south baselines
 are close (4m), and hence these baselines are not suitable
@@
 equivalent of which are 
-\{2, 0\}, \{2, 1\} and \{2, -1\}.
+\{2, 0\}, \{2, 1\} and $\{2, -1\}$.
 There, baselines are cross-multiplied within each class.
@@
 In general, track-crossings are
-not accurate enough for time offset determination, nor can it give estimate of the degree of redundancy.
+not accurate enough to determine the optimal time offset for the cross multiplication of visibilities, nor can it provide an estimate of the degree of redundancy.
 For drift-scan arrays, even a hypothetical crossing in $uvw$ space would not imply perfect redundancy.
@@
 \]
-Inverting for $\nu$:
+Inverting for $\nu$ gives
 \begin{equation}
@@
 \end{equation}
-
-In the thin-shell limit, we can write:
+This provides us with a mapping between the observational frequency and the cosmological distance in the radial direction. For the angular directions, we have in the thin-shell limit
 \begin{equation}
@@
 \eqref{eq:Vis1} resembles a Fourier transform along the angular directions, it is natural to perform a further Fourier transform along the radial, or equivalently, frequency axis.
-The transform, called delay-transform \citep{PB2009}, has gained popularity recently since it has been shown that foregrounds are isolated in delay space\footnote{We discuss foreground isolation in detail in Section \ref{sec:chromaticity}.}.
-We define the delay-transformed visibility as:
+The transform, called the delay transform \citep{PB2009}, has gained popularity recently since it has been shown that foregrounds are isolated in delay space.\footnote{We discuss foreground isolation in detail in Section \ref{sec:chromaticity}.} We define the delay-transformed visibility as:
 \small
 \begin{equation}
-\begin{aligned}V(\boldsymbol{b},\tau) & =\int d\nu V_{\nu}(\boldsymbol{b})\phi(\nu)\exp\left(2\pi i\nu\tau\right)\\
+\begin{aligned}V(\boldsymbol{b},\tau) & \equiv \int d\nu V_{\nu}(\boldsymbol{b})\phi(\nu)\exp\left(2\pi i\nu\tau\right)\\
  & =\int d\Omega d\nu B(\hat{\boldsymbol{s}},\nu)T(\hat{\boldsymbol{s}},\nu)\exp\left[-2\pi i\nu\left(\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}-\tau\right)\right]. 
@@
 \normalsize
-where $d\nu=-dr/Y$ and $d^{3}r=-X^{2}Yd\Omega d\nu$. 
+\!\!where $d\nu=-dr/Y$ and $d^{3}r=-X^{2}Yd\Omega d\nu$. 
 We have written $\nu_{r}$ as a reminder that $\nu$ and $r$ are related
@@
 in the topocentric frame with the three-dimensional rotation operator $\boldsymbol{\Gamma}$:
-
 \[
-\begin{aligned}V_{\psi}(\boldsymbol{b'},\tau) & = \int\frac{d^{3}r}{X^{2}Y}B(\boldsymbol{\Gamma}\boldsymbol{r})T(\boldsymbol{r})\\
-& \qquad \qquad  \exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\boldsymbol{\Gamma}\hat{\boldsymbol{r}}-\tau\right)\nu_{r}-i\psi_\nu\right],\end{aligned}
+\begin{aligned}V_{\psi}(\boldsymbol{b'},\tau) & = \int\frac{d^{3}r}{X^{2}Y}B(\boldsymbol{\Gamma}\boldsymbol{r})T(\boldsymbol{r}) e^{-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\boldsymbol{\Gamma}\hat{\boldsymbol{r}}-\tau\right)\nu_{r}-i\psi_\nu},\end{aligned}
 \]
-where we introduced a frequency-dependent phase $\psi_{\nu}$, which to linear-order in $\nu$ corresponds to phasing the two visibilities to the same phase center.
-\footnote{For many cases, the linear-order interpretation is sufficient.
-We keep the general term to reserve the option to determine its exact form numerically as this information comes at no additional computational cost.
-}
+%\[
+%\begin{aligned}V_{\psi}(\boldsymbol{b'},\tau) & = \int\frac{d^{3}r}{X^{2}Y}B(\boldsymbol{\Gamma}\boldsymbol{r})T(\boldsymbol{r}) \times\\
+%& \qquad \qquad  \exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\boldsymbol{\Gamma}\hat{\boldsymbol{r}}-\tau\right)\nu_{r}-i\psi_\nu\right],\end{aligned}
+%\]
+where we introduced a frequency-dependent phase $\psi_{\nu}$, which to linear-order in $\nu$ one would typically pick to correspond to the re-phasing of the two visibilities to the same phase center.\footnote{For many cases, the linear-order interpretation is sufficient.
+We keep the general term to reserve the option to determine its exact form numerically as this information comes at no additional computational cost.}
 
 With implicit bounds of integrals from $-\infty$ to $\infty$, we have:
-\begin{widetext}
-\begin{equation}
-\begin{aligned} \langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle
- & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\\
- & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\boldsymbol{\kappa})T(\boldsymbol{\kappa'})\rangle e^{-i(\boldsymbol{\kappa}\cdot \boldsymbol{r}-\boldsymbol{\kappa'}\cdot\boldsymbol{r'})}\right)B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\\
- & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\boldsymbol{\kappa}\cdot(\boldsymbol{r}-\boldsymbol{r'})}\right)B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\\
- & \approx P(k_{\bar{b}, \tau})\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2} \delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'}) \Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\\
- & = P(k_{\bar{b}, \tau})\int \frac{d^{3}r}{(X^2Y)^2}B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r})\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma} \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right],
-\end{aligned}
+\begin{eqnarray}
+ &&\langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle \nonumber \\
+ && =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}) \nonumber \\
+ &&  =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\boldsymbol{\kappa})T(\boldsymbol{\kappa'})\rangle e^{-i(\boldsymbol{\kappa}\cdot \boldsymbol{r}-\boldsymbol{\kappa'}\cdot\boldsymbol{r'})}\right) \nonumber \\
+ && \qquad  \times B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\nonumber \\
+ && =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\boldsymbol{\kappa}\cdot(\boldsymbol{r}-\boldsymbol{r'})}\right) \nonumber \\
+ && \qquad \times B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}) \nonumber \\
+ && \approx P(k_{\bar{b}, \tau})\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2} \delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'}) \Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}) \nonumber \\
+ && = P(k_{\bar{b}, \tau})\int \frac{d^{3}r}{(X^2Y)^2}B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r}) e^{-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma} \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}}, \qquad
 \label{eq:main}
-\end{equation}
-
+\end{eqnarray}
 where $\delta_D$ is the Dirac delta-function and
-\begin{equation}
-\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\equiv\exp\left[i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\boldsymbol{\Gamma}\hat{\boldsymbol{r}}'\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)-i\psi_{\nu}\right].
-\end{equation}
-
-\end{widetext}
-
-The third equality of Eq.(\ref{eq:main}) follows from assuming the translational invariance of statistics of the 21\,cm field. 
+\begin{eqnarray}
+\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\equiv &\exp\left[i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\boldsymbol{\Gamma}\hat{\boldsymbol{r}}'\right)\right] \nonumber \\
+& \times \exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)-i\psi_{\nu}\right].
+\end{eqnarray}
+%
+%\begin{widetext}
+%\begin{equation}
+%\begin{aligned} \langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle
+% & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\\
+% & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\boldsymbol{\kappa})T(\boldsymbol{\kappa'})\rangle e^{-i(\boldsymbol{\kappa}\cdot \boldsymbol{r}-\boldsymbol{\kappa'}\cdot\boldsymbol{r'})}\right)B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\\
+% & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\boldsymbol{\kappa}\cdot(\boldsymbol{r}-\boldsymbol{r'})}\right)B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\\
+% & \approx P(k_{\bar{b}, \tau})\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2} \delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'}) \Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\\
+% & = P(k_{\bar{b}, \tau})\int \frac{d^{3}r}{(X^2Y)^2}B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r})\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma} \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right],
+%\end{aligned}
+%\label{eq:main}
+%\end{equation}
+%
+%where $\delta_D$ is the Dirac delta-function and
+%\begin{equation}
+%\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\equiv\exp\left[i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\boldsymbol{\Gamma}\hat{\boldsymbol{r}}'\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)-i\psi_{\nu}\right].
+%\end{equation}
+%
+%\end{widetext}
+The third equality of Eq.(\ref{eq:main}) follows from assuming the translational invariance of the statistics of the 21\,cm field. 
 For the inequality in the fourth line, we have assumed that the  3D power spectrum varies negligibly over the $k$-space of interest and have thus pulled out of the integral the power spectrum centered at
@@
 where $\bar{b}$ is approximately the mean of $b$ and $b'$. 
-In Section \ref{sec:chromaticity} we will examine the exact form of the Fourier-space footprint probed as well as its effect on foreground isolation.  
+In Section \ref{sec:chromaticity} we will examine the exact form of the Fourier-space footprint that is being probed, as well as its effect on foreground isolation.  
 %Since $\boldsymbol{\Gamma}$ is a sky rotation, it does not affect $\nu$, hence we have taken $\nu_{r}$ outside the parenthesis in the phase term $\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma} \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)\right]$.
@@
 the power spectrum:
-
 \begin{equation}
 \begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle\\
- & \qquad \approx P(k_{\bar{b}, \tau})\int\frac{d\Omega d\nu}{X^{2}Y}B^{*}(\hat{\boldsymbol{s}},\nu)B(\boldsymbol{\Gamma}\hat{\boldsymbol{s}},\nu)\\
- & \qquad \qquad \qquad \qquad \exp\left[i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma}\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right].
+ & \approx P(k_{\bar{b}, \tau})\int\frac{d\Omega d\nu}{X^{2}Y}B^{*}(\hat{\boldsymbol{s}},\nu)B(\boldsymbol{\Gamma}\hat{\boldsymbol{s}},\nu) e^{i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma}\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}}.
  \end{aligned}
@@
 \end{equation}
-
 We can therefore form the power spectrum estimator for the baseline pair $\{\boldsymbol b: \boldsymbol  b'\}$:
-
 \begin{equation}
@@
 with
-
 \[
@@
 \]
-
 Notice that $\Theta$ has no dependence on $\tau$.
@@
 \subsection{Rephasing \label{sec:rephs}}
-If $\psi_{\nu}=0$ in  Eq.
-\eqref{eq:Thetanu}, $\Theta_{\nu}$ at the peak of correlation ( as a function of time) is generally complex, and often far from real.
+If $\psi_{\nu}$ were set to zero in  Eq.
+\eqref{eq:Thetanu}, $\Theta_{\nu}$ at the peak of correlation (as a function of time) is generally complex, and often far from real.
 Furthermore, this phase of peak correlation is inevitably frequency dependent.
@@
 
-To linear order in $\nu$, $\psi_\nu$ cancels the decoherence due to the two visibilities having different phase centers.
+Selecting $\psi_\nu$ to be a linear function of $\nu$ allows one to cancel
+%To linear order in $\nu$, $\psi_\nu$ cancels
+the decoherence due to the two visibilities having different phase centers.
 By default the correlators of a drift-scan array phase the two visibilities both to zenith at the same time.
@@
 \ref{fig:freqdiff} that although the amplitude of correlations match up for all time offsets, the phase would only locally match.
-The need for rephasing can thus be understood as a symptom of the underlying spatial decoherence; coherence at the phase center would not extend to the entire beam pattern (see Section \ref{sec:visual}).
+The need for rephasing can thus be understood as a symptom of the underlying spatial decoherence; coherence at the phase center does not extend to the entire beam pattern (see Section \ref{sec:visual}).
 While integrating over frequencies and sky-direction in Eq.
@@
 \cite{sphere} discusses in detail power-spectrum analyses using spherical harmonics.
-In such as basis, modes in the sky are indexed by the magnitude $k$ of the wavevector $\mathbf{k}$ and spherical harmonic indices $(\ell, m)$.
+In such a basis, modes in the sky are indexed by the magnitude $k$ of the wavevector $\mathbf{k}$ and spherical harmonic indices $(\ell, m)$.
 In other words, one defines
@@
 The window function localized at low $\ell$ is for the baseline \{2,0\} of around $30\,\textrm{m}$ in length, while the window function localized at high $\ell$ is for \{12,0\}, roughly $180\,\textrm{m}$ in length.
-The window functions for near-equivalent baseline estimates are shown in the bottom plot, for \{2,0:2,1\} and \{12,0:12,1\} at low- and high-$\bar{b}$, respectively.
+The window functions for near-equivalent baseline estimates are shown in the bottom plot, for \{2,0:2,1\} and \{12,0:12,1\} at low and high $\bar{b}$, respectively.
 In all cases, the window functions that we show are for $\tau = 703\,\textrm{ns}$, and are generated using primary beam models for PAPER.
@@
 
-To determine that actual relative contribution to sensitivity of these
+To determine the actual relative contribution to sensitivity of these
 baseline pairs, we also have to account for the multiplicities of
-these baseline classes, i.e. the number of antenna pairs with the
+these baseline classes, i.e., the number of antenna pairs with the
 same length and orientation.
@@
 we get a factor of $\sqrt{M}$ reduction in the noise
-level $\sigma_N$ of the visibility.
-The sensitivity contribution of $\{m,n\}$, cross multiplied with $\{m',n'\}$  thus roughly scales as $\sqrt{\left((16-|m|)(7-|n|)(16-|m'|)(7-|n'|)\right)}=\sqrt{MM'}$.
+level $\sigma_N$ of the visibility. The sensitivity contribution of $\{m,n\}$, cross multiplied with $\{m',n'\}$  thus roughly scales as $\sqrt{(16-|m|)(7-|n|)(16-|m'|)(7-|n'|)}=\sqrt{MM'}$.
 For cross-multiplications of nearly equivalent baselines of
@@
 
-The scaling in Eq. \eqref{eq:sensul} is rough for simplicity of motivation. As we derive in Appendix \ref{sec:appB}, this weight should be corrected by a factor proportional to $\rho_0$:
+The scaling in Eq. \eqref{eq:sensul} was an approximate one for simplicity of motivation. As we derive in Appendix \ref{sec:appB}, this weight should be corrected by a factor proportional to $\rho_0$:
 
@@
 Assuming a reionization signal of $\sim 30\,\text{mK}$, observation centered at 0.15\,GHz ($z=8.5$), and 120 days of integration with PAPER antennas, we have roughly
-(see Eq.(20) in \cite{first-paper})
+(see Eq.(20) in \citealt{first-paper})
 \begin{equation}
@@
 The plot therefore does not compare absolute sensitivity across different arrays.
-We see as expected that in all cases using the nearly equivalent baselines lead to more and more significant improvements with lower $\widetilde{\Theta}_{min}$, or in other words when more baseline pairs are used.
-We point out that PAPER results until now \citep{Ali2015, paper32} only used the top contributing baseline pairs.
-In other words these results only explored sensitivity $\rho(\widetilde{\Theta}_{min}\sim 1)$.
+We see as expected that in all cases, using the nearly equivalent baselines leads to increasingly significant improvements as $\widetilde{\Theta}_{min}$ is lowered, or in other words, when more baseline pairs are used.
+We point out that PAPER results until now \citep{Ali2015, paper32} only used the top contributing baseline pairs, essentially only exploiting the sensitivity of the $\rho(\widetilde{\Theta}_{min}\sim 1)$ scenario.
 
@@
 In this section we link our results from Section \ref{sec:method} to the traditional views of rotation synthesis and in particular $uv$ tracks.
-in describing the relative motion of the baseline relative to the sky, we have in Section \ref{sec:method} rotated the direction coordinate $\hat{s}$ with $\boldsymbol{\Gamma}$.
-The typical convention of rotation synthesis has rotation operates on $\boldsymbol{b'}$ instead of $\hat{\boldsymbol{s}}$, in which case Eq. \eqref{eq:Theta} takes the equivalent form
+In describing the relative motion of the baseline relative to the sky, we have in Section \ref{sec:method} rotated the direction coordinate $\hat{s}$ with $\boldsymbol{\Gamma}$.
+The typical convention of rotation synthesis has the rotation operator acting on $\boldsymbol{b'}$ instead of $\hat{\boldsymbol{s}}$, in which case Eq. \eqref{eq:Theta} takes the equivalent form
 \begin{equation}
@@
 \end{equation}
-
 Track-crossing corresponds to vanishing of the exponent
@@
 \eqref{eq:trackcross} vanishes and $\psi_\nu$ is zero for all $\nu$.
-In general, second-order effects due to the presence of $w$ is observed for the tracking measurements as in Fig.
+In general, second-order effects due to the presence of $w$ are observed for the tracking measurements as in Fig.
 \ref{fig:phi_nu}.
@@
 \label{sec:appB}
-In this appendix we give a brief derivation of the effective weight $\widetilde{\Theta}$ quoted in Section \ref{sec:sensitivity}.
-We combine the power spectrum measurements from distinct baseline classese by inverse variance weighting.
-We separate the visibility and power spectrum into signal and noise contributions:
+In this appendix we provide a derivation of the effective weight $\widetilde{\Theta}$ quoted in Section \ref{sec:sensitivity}.
+In doing so, we assume that the power spectrum measurements from distinct baseline classes are combined by inverse variance weighting.
+To begin, we separate the visibility and power spectrum into signal and noise contributions:
 \begin{equation}
@@
 
-The variance of $P$ constructed with visibilities $V_1$ and $V_2$ from two baseline classes can be estimated\footnote{We assume all noise terms to be independent for simplicity, in practice the correlation of different measurements ifrom equivalent baselines are alleviated by grouping the baselines in the class and the days of observation, as in \cite{Ali2015}}:
+The variance of $P$ constructed with visibilities $V_1$ and $V_2$ from two baseline classes can be estimated:
+%\footnote{We assume all noise terms to be independent for simplicity; in practice, the correlation of different measurements from equivalent baselines are alleviated by grouping the baselines in the class and the days of observation, as in \cite{Ali2015}}
 \begin{equation}
@@
 \end{equation}
-where in the second last line we have substituted in the visibility noise variance. In the final line we used Wick's theorem and the fact that the signal from the two visibilities are equal.
+where in the second last line we have substituted in the visibility noise variance, and have assumed independent noise between baselines.
+In the final line we used Wick's theorem and the fact that the signal from the two visibilities are equal.
 
-Recall from the discussion on multiplicities we can write
+Recall from the discussion on multiplicities that we can write
 \begin{equation}

### 2017-07-24 Rotation paper/draft_working.tex
save

@@
-%\documentclass[preprint2,numberedappendix,tighten,twocolappendix]{aastex6}  % USE THIS TO MAKE BIB, THEN FORMAT USING EMULATEAPJ
+%\documentclass[preprint2,numberedappendix,twocolappendix]{aastex6}  % USE THIS TO MAKE BIB, THEN FORMAT USING EMULATEAPJ
 \documentclass[twocolumn,apj,numberedappendix]{emulateapj}
@@
 \newcommand{\pasa}{PASA}
-
+\usepackage[backref,breaklinks,colorlinks,citecolor=blue]{hyperref}
 \begin{document}
@@
 Radio interferometers designed to measure the cosmological 21\,cm power spectrum require high sensitivity.
-Several modern low-frequency interferometers feature drift-scan antennas placed on regular grid to maximize the number of instantaneously coherent (redundant) measurements.
+Several modern low-frequency interferometers feature drift-scan antennas placed on a regular grid to maximize the number of instantaneously coherent (redundant) measurements.
 However, even for such maximum-redundancy arrays, significant sensitivity comes through partial coherence between baselines.
@@
 of \{2,1\} and $\{2, -1\}$.
-Note that \{2,0\} and $\{-2, 0\}$ are the same class of baselines and should not be counted twice. Antennas in purely north-south baselines
+Note that \{2,0\} and $\{-2, 0\}$ are the same class of baselines and should not be included twice in power-spectrum measurements. Antennas in purely north-south baselines
 are close (4m), and hence these baselines are not suitable
@@
 not accurate enough to determine the optimal time offset for the cross multiplication of visibilities, nor can it provide an estimate of the degree of redundancy.
-For drift-scan arrays, even a hypothetical crossing in $uvw$ space would not imply perfect redundancy.
+Furthermore, for drift-scan arrays, even a hypothetical crossing in $uvw$ space would not imply perfect redundancy.
 This will become evident in the next section, after we develop a more general formalism that accounts for the point spread function of primary.
@@
 \]
-where $r$ is the magnitude of the vector $\bold{r}$, $\nu_{21}=1.420$\,GHz is the 21\,cm transition rest frequency, $\nu_{0}$ is
+where $r$ is the magnitude of the vector $\bold{r}$, with $z$ its corresponding redshift. $\nu_{21}=1.420$\,GHz is the 21\,cm transition rest frequency, $\nu_{0}$ is
 a reference central frequency with corresponding redshift $z_{0}$, $H_0$ is the current-day Hubble constant,  
@@
 \[
-E(z)=\sqrt{\Omega_{m}(1+z)^{3}+\Omega_{\Lambda}}.
+E(z)=\sqrt{\Omega_{m}(1+z)^{3}+\Omega_{\Lambda}},
 \]
-Inverting for $\nu$ gives
+with $\Omega_{m}$ and $\Omega_{\Lambda}$ being respectively the normalized matter and dark energy density.
+
+Inverting Eq. \eqref{eq:r} for $\nu$ gives
 \begin{equation}
@@
 \caption{PAPER-128 beam-fringe patterns and their point-spread functions on the $uv$-plane for a frequency of $0.15$\,GHz.
-Selected baselines have 30-meter East-West separations, and 0- and 4-meter North-South separations, respectively.
-The top panels show the real parts of the beam-fringe patterns of the baselines (left and middle), and their conjugate product (right).
+Selected baselines have 30-meter East-West separations.
+The first baseline, shown in the first column, is purely East-West, while the second baseline, shown in the middle column, has 4-meter North-South separation and has been rotated with a time offset of 0.78 hours.
+Panels in the top row show the real parts of the beam-fringe patterns of the baselines (left and middle), and their conjugate product (right).
 The beam-fringe values are normalized such that the peak of the original beam is unity.
@@
 At the optimal time offset, the fringe patterns of the two baselines cancel out, leading to a substantial weight $\Theta_\nu$.
-The frequency-summed weight $\Theta$ is constructed with rephasing as described in Section \ref{sec:rephs} so that the conjugate product (top-right panel) for different frequencies align in phase.
+The frequency-summed weight $\Theta$ is constructed with rephasing as described in Section \ref{sec:rephs} so that the conjugate product (top-right panel) for different frequencies align in phase at their peaks.
+Note it is not possible to align the phase of $g_{b}^{*}g_{b'}e^{-i\psi_\nu}$ for all $(l,m,\nu)$ due to spatial decoherence across the beam as discussed at the end of Section \ref{sec:rephs}.
 
@@
 \caption{Planned Hydrogen Epoch of Reionization Array antenna configurations.
-HERA-37 is expected to complete and start collecting data in August of 2017, and the other three are planned configurations in the next phases.
+HERA-37 is expected to complete and start collecting data in August of 2017, while the other three are planned configurations in the upcoming phases.
 For HERA-350, only the 320 elements in the core are shown.
@@
 we show $\Theta_{\boldsymbol{bb'}}$ and $\Delta t_{\boldsymbol{bb'}}$ for a variety of baseline combinations. Baseline pairs that are mirror images of each other 
-give the same amount of redundancy ($\Theta_{\boldsymbol{bb'}}$), with the opposite time offset, as expected from symmetry. 
+give approximately the same amount of redundancy ($\Theta_{\boldsymbol{bb'}}$), with the opposite time offset, as expected from symmetry. 
 For example, \{1,0:1,1\} is mirror image of \{1,0:1,-1\} and these two pairs of baselines
@@
 tend to have higher correlations. In other words, correlation peaks
-that are closer to zero time offset are higher. This is expected for two reasons;
+with closer to zero time offset are higher. This is expected for two reasons;
 one is that the longer the time offset of maximum redundancy, the more the sky has moved with respect
@@
 Shorter HERA baselines require much longer $\Delta t$ to overlap, partly because of the hexagonal structure requires $60\degree$ of rotation to overlap baselines, partly because of the smaller primary beam.
-Thus most short HERA baseline pairs are not shown in this figure.
-Only baseline pairs with $\widetilde{\Theta}_{\boldsymbol{bb'}}>0.01$ are shown.
+Thus most short HERA baseline pairs are not shown in this figure, as they do not meet the selection criterion $\widetilde{\Theta}_{\boldsymbol{bb'}}>0.01$.
 
@@
 
-\item The top sensitivity-contributing near-equivalent pair in each array is normalized to the same $\Theta_{\boldsymbol{bb'}}$ (not shown), but those of HERA arrays have much lower $\widetilde{\Theta}_{\boldsymbol{bb'}}$ than the top pair of PAPER-128.
+\item The top sensitivity-contributing near-equivalent pair in each array is normalized to have the same $\Theta_{\boldsymbol{bb'}}$ (not shown), but those of HERA arrays have much lower $\widetilde{\Theta}_{\boldsymbol{bb'}}$ than the top pair of PAPER-128.
 This is because they are longer baselines with lower multiplicity.
@@
 We show that for arrays with increasing number of antennas, partially redundant baselines make up a more significant portion of the array's total sensitivity.
-The proposed power-spectrum estimator retains benefits of delay-transform techniques for isolating foregrounds and avoiding systematics and enables one to unlock the full sensitivity of a redundant radio interferometer while simultaneously enjoying the benefits of delay-transform techniques for isolating foregrounds and avoiding systematics.
+The proposed power-spectrum estimator enables one to unlock the full sensitivity of a redundant radio interferometer while retaining the benefits of delay-transform techniques for isolating foregrounds and avoiding systematics.
 
@@
 %\nocite{*}
-\bibliography{draft_working}
+\bibliography{draft_final}
 

### 2017-07-24 Rotation paper/draft_working.tex
save

@@
 \newcommand{\pasa}{PASA}
-\usepackage[backref,breaklinks,colorlinks,citecolor=blue]{hyperref}
+\usepackage[backref,breaklinks,colorlinks,citecolor=cyan]{hyperref}
 \begin{document}
@@
 To gain some visual intuition of the formalism in Section \ref{sec:method}, we show explicit beam fringe patterns of the two baselines and their interactions.
-For this section, we relax into the flat sky approximation so that $(l,m,\sqrt{1-l^2-m^2})\equiv\hat{s}$ and examine a single frequency of $0.15$\,GHz.
+For this section, we let $(l,m,\sqrt{1-l^2-m^2})\equiv\hat{s}$ and examine a single frequency of $0.15$\,GHz.
 
@@
 %\nocite{*}
-\bibliography{draft_final}
+\bibliography{draft_working}
 

### 2017-07-24 Rotation paper/draft_working.tex
save

@@
 \]
-where $\Delta\tau$ is a shift in delay corresponding to the phase center movement.
+where $\Delta\tau$ is a shift in delay corresponding to the phase center movement, more precisely the light travel time from one antenna to another of the rotated baseline (relative to the phase center).
 
@@
 In the drift-scan case, we indeed see a linear relation corresponding to a delay of $\Delta\tau\approx15.7$\,ns.
-This delay corresponds to the light travel time from one antenna to another of the rotated baseline.
 In the tracking case, since the phase center is fixed to the sky, only second order effects are observed.

### 2017-07-24 Rotation paper/draft_working.tex
save again

--- 2017-07-24 Rotation paper/draft_working.tex
@@
 \newcommand{\pasa}{PASA}
-\usepackage[backref,breaklinks,colorlinks,citecolor=cyan]{hyperref}
+\usepackage[breaklinks,colorlinks,citecolor=cyan]{hyperref}
 \begin{document}
@@
 where $\bold{k}$ is the cosmological wavenumber, $\delta_D^3$ is the three-dimensional Dirac delta function, and $\widetilde{T}(\boldsymbol{k})$ is the three-dimensional Fourier transform of the brightness temperature field $T(\bold{r})$, with $\bold r$ being the cosmological position coordinate.
-We begin by relating the observational coordinates $\hat{\boldsymbol{s}}$
+We begin by relating the observational coordinates\footnote{The observational direction $\hat{\boldsymbol{s}}$ is equal to the cosmological direction $\hat{\boldsymbol{r}}$.
+We shall nevertheless keep the notations separate to highlight the conceptual differences and switch between the symbols depending on the context.
+} $\hat{\boldsymbol{s}}$
 and $\nu$, to
@@
 \]
-with $\Omega_{m}$ and $\Omega_{\Lambda}$ being respectively the normalized matter and dark energy density.
-
-Inverting Eq. \eqref{eq:r} for $\nu$ gives
+with $\Omega_{m}$ and $\Omega_{\Lambda}$ being respectively the normalized matter and dark energy density. Inverting Eq. \eqref{eq:r} for $\nu$ gives
 \begin{equation}
@@
 \]
-where $\Delta\tau$ is a shift in delay corresponding to the phase center movement.
+where $\Delta\tau$ is a shift in delay corresponding to the phase center movement, i.e.
+\[
+\Delta \tau = \frac{\boldsymbol{b'}}{c}\left(\boldsymbol{\Gamma}\hat{\boldsymbol{s}}-\hat{\boldsymbol{s}}\right). 
+\]
 

### 2017-10-13 Rotation paper/draft_submission1.tex
addressed referee comments 1,2,3,6,7

--- 2017-07-24 Rotation paper/draft_working.tex
@@
 
-With implicit bounds of integrals from $-\infty$ to $\infty$, we have:
+With implicit bounds of integrals from $-\infty$ to $\infty$, we have
 \begin{eqnarray}
@@
  && =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\boldsymbol{\kappa}\cdot(\boldsymbol{r}-\boldsymbol{r'})}\right) \nonumber \\
- && \qquad \times B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}) \nonumber \\
- && \approx P(k_{\bar{b}, \tau})\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2} \delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'}) \Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}) \nonumber \\
- && = P(k_{\bar{b}, \tau})\int \frac{d^{3}r}{(X^2Y)^2}B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r}) e^{-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma} \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}}, \qquad
-\label{eq:main}
+ && \qquad \times B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}),
  \end{eqnarray}
-where $\delta_D$ is the Dirac delta-function and
+ where 
  \begin{eqnarray}
@@
 \end{eqnarray}
+The third equality of Eq.(\ref{eq:main}) follows from assuming the translational invariance of the statistics of the 21\,cm field.
+Continuing, we may make the assumption that the 3D power spectrum varies negligibly over the $k$-space of interest, thus allowing us to pull out of the integral the power spectrum centered at
+\[
+k_{\boldsymbol{b}, \tau} \equiv 2\pi \sqrt{\left(\frac{\tau}{Y}\right)^2 + \left(\frac{\bar{b}}{\lambda X}\right)^2},
+\label{eq:kbtau}
+\]
+where $\bar{b}$ is approximately the mean of $b$ and $b'$. This gives 
+ \begin{eqnarray}
+  &&\langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle \nonumber \\
+ && \approx P(k_{\bar{b}, \tau})\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2} \delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'}) \Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}) \nonumber \\
+ && = P(k_{\bar{b}, \tau})\!\!\int\!\!
+\frac{d^{3}r}{(X^2Y)^2}B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r}) e^{-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma} \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}}, \qquad
+\label{eq:main}
+\end{eqnarray}
+where $\delta_D$ is the Dirac delta-function, and notice that the phase factor $\exp\left[-i2\pi\tau\left(\nu-\nu'\right)\right]$
+drops out in the end.
+In factoring out the power spectrum from the integral, one is essentially making the approximation that the estimator probes only a single Fourier mode.
+In Section \ref{sec:chromaticity} we will relax this assumption and examine the exact form of the Fourier-space footprint that is being probed, as well as its effect on foreground isolation.
+
 %
@@
 %\end{widetext}
-The third equality of Eq.(\ref{eq:main}) follows from assuming the translational invariance of the statistics of the 21\,cm field. 
-For the inequality in the fourth line, we have assumed that the  3D power spectrum varies negligibly over the $k$-space of interest and have thus pulled out of the integral the power spectrum centered at
-\[
-k_{\boldsymbol{b}, \tau} \equiv 2\pi \sqrt{\left(\frac{\tau}{Y}\right)^2 + \left(\frac{\bar{b}}{\lambda X}\right)^2},
-\label{eq:kbtau}
-\]
-where $\bar{b}$ is approximately the mean of $b$ and $b'$. 
-In Section \ref{sec:chromaticity} we will examine the exact form of the Fourier-space footprint that is being probed, as well as its effect on foreground isolation.  
 %Since $\boldsymbol{\Gamma}$ is a sky rotation, it does not affect $\nu$, hence we have taken $\nu_{r}$ outside the parenthesis in the phase term $\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma} \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)\right]$.
-Notice that the phase factor $\exp\left[-i2\pi\tau\left(\nu-\nu'\right)\right]$
-drops out in the end.
 
@@
 
+\begin{figure}[t]
+\includegraphics[width=1\linewidth]{phi_nu}
+
+\caption{Frequency dependent peak phases for PAPER-128 baseline pair \{2,0:2,1\}.
+Both the drift-scan case and a hypothetical tracking baseline are shown in solid lines.
+The first-order, linear effects are shown in dashed lines.
+We have also fixed the global phase in both cases to 0 at 0.15\,GHz.
+The drift-scan case exhibits linear behavior and conforms well to the first order effect of a shift in delay space due to the movement of zenith.
+}
+\label{fig:phi_nu}
+\end{figure}
+
 Selecting $\psi_\nu$ to be a linear function of $\nu$ allows one to cancel
@@
 
-\begin{figure}[H]
-\includegraphics[width=1\linewidth]{phi_nu}
+\begin{figure}[t]
+\includegraphics[width=1.1\linewidth]{rephs_rb}
 
-\caption{Frequency dependent peak phases for PAPER-128 baseline pair \{2,0:2,1\}.
-Both the drift-scan case and a hypothetical tracking baseline are shown in solid lines.
-The first-order, linear effects are shown in dashed lines.
-We have also fixed the global phase in both cases to 0 at 0.15\,GHz.
-The drift-scan case exhibits linear behavior and conforms well to the first order effect of a shift in delay space due to the movement of zenith.
+\caption{Comparisons of the peak phases of two different frequencies.
+Specifically shown are the real parts of $\Theta_{\nu}$.
+First and third panels shows equivalent baselines, second and fourth show a pair of nearly equivalents.
+The top two panels have zero phase shift ($\psi_{\nu}=0$), and the bottom two are rephased to a time offset of 0.78 hours.
+The first and last panels thus show a coherently rephased series that would add constructively at the respective peaks.
 }
-\label{fig:phi_nu}
+\label{fig:freqdiff}
 \end{figure}
@@
 This decoherence across the beam pattern cannot be removed and plays a fundamental role in determining how much sensitivity one can recover from nearly equivalent baselines.
-
-\begin{figure}[H]
-\includegraphics[width=1.1\linewidth]{rephs_rb}
-
-\caption{Comparisons of the peak phases of two different frequencies.
-Specifically shown are the real parts of $\Theta_{\nu}$.
-First and third panels shows equivalent baselines, second and fourth show a pair of nearly equivalents.
-The top two panels have zero phase shift ($\psi_{\nu}=0$), and the bottom two are rephased to a time offset of 0.78 hours.
-The first and last panels thus show a coherently rephased series that would add constructively at the respective peaks.
-}
-\label{fig:freqdiff}
-\end{figure}
 
@@
 
+\begin{figure}[h]
+\includegraphics[width=1.1\linewidth]{multiblwindows}
+\caption{Sample window functions for delay-spectrum estimates of the power spectrum explored in this paper.
+The top plot shows window functions for a traditional equivalent-baseline delay-spectrum estimate, where a single baseline's delay spectrum is squared.
+The bottom plot shows the window functions resulting from nearly equivalent baselines.
+In each plot, the window function on the right corresponds to the window function for baselines of length $b \sim 180\,\textrm{m}$ in the East West direction, while the window function on the left is for $b\sim 30\,\textrm{m}$.
+The general extent of the window functions for the two-baseline estimator are seen to be roughly the same as for the single-baseline estimator, suggesting that the former is just as good as the latter in keeping the EoR window free of foregrounds.}
+\label{fig:multiblwindows}
+\end{figure}
+
 Figure \ref{fig:multiblwindows} shows four example window functions.
@@
 The nature of its angular fluctuations are therefore inherently different from its radial
-
-\begin{figure}[h]
-\includegraphics[width=1.1\linewidth]{multiblwindows}
-\caption{Sample window functions for delay-spectrum estimates of the power spectrum explored in this paper.
-The top plot shows window functions for a traditional equivalent-baseline delay-spectrum estimate, where a single baseline's delay spectrum is squared.
-The bottom plot shows the window functions resulting from nearly equivalent baselines.
-In each plot, the window function on the right corresponds to the window function for baselines of length $b \sim 180\,\textrm{m}$ in the East West direction, while the window function on the left is for $b\sim 30\,\textrm{m}$.
-The general extent of the window functions for the two-baseline estimator are seen to be roughly the same as for the single-baseline estimator, suggesting that the former is just as good as the latter in keeping the EoR window free of foregrounds.}
-\label{fig:multiblwindows}
-\end{figure}
 

### 2017-10-13 Rotation paper/draft_working.tex
correct typo

@@
 This technique avoids many systematics associated with combining data from different baselines and tracks the native sampling of the interferometer.
-An example of the visibility-based pipeline is presented in \cite{Ali2015}, which provides power spectrum
+An example of the visibility-based pipeline was presented in \cite{Ali2015}, which provided power spectrum
 measurements with the 64-element version of PAPER (henceforth as PAPER-64).
@@
 We shall use the 128-element PAPER array (henceforth referred to as PAPER-128) to motivate our formalism and demonstrate our method, and extend our results to several HERA configurations in Section \ref{sec:arrconf}.
-The PAPER array is located in the Karoo desert in South Africa (30:43:17.5
+The PAPER array \textbf{was} located in the Karoo desert in South Africa (30:43:17.5
 S, 21:25:41.8 E). The layout pattern is shown
@@
 \ref{fig:AntPos}.
-The array consists of a 112-element core in a rectangular grid, and 16 ``outriggers'' used primarily to aid calibration.
+The array \textbf{consisted} of a 112-element core in a rectangular grid, and 16 ``outriggers'' used primarily to aid calibration.
 In the bottom panel, the two baselines marked in red are an example of an equivalent pair.
@@
 On the other hand, the small North-South separation means that classes such as \{2,0\} and  \{2,1\} are expected to be near-equivalent.
-The PAPER-64 analysis of \cite{Ali2015} uses three classes of baselines, the PAPER-128
-equivalent of which are 
+The PAPER-64 analysis of \cite{Ali2015} \textbf{used} three classes of baselines, the PAPER-128
+equivalent of which \textbf{were} 
 \{2, 0\}, \{2, 1\} and $\{2, -1\}$.
-There, baselines are cross-multiplied within each class.
+There, baselines \textbf{were} cross-multiplied within each class.
 This paper provides the method for inter-class multiplications.
@@
 \begin{eqnarray}
-\hat{P}(k) &=& \frac{ \langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau) \rangle}{\Theta} \nonumber \\
+\hat{P}(k_{\bar{b}, \tau}) &=& \frac{ \langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau) \rangle}{\Theta} \nonumber \\
 &=& \sum_\ell \int \!dk\, W_{\ell}(k; \mathbf{b}, \mathbf{b}^\prime , \tau) P(k),
@@
 If the window function is substantially non-zero at low-$k$ modes---which is where foreground emission typically resides, given its spectral smoothness---then the EoR window will be contaminated by foregrounds.
+
+\begin{figure}[h]
+\includegraphics[width=1.1\linewidth]{multiblwindows}
+\caption{Sample window functions for delay-spectrum estimates of the power spectrum explored in this paper.
+The top plot shows window functions for a traditional equivalent-baseline delay-spectrum estimate, where a single baseline's delay spectrum is squared.
+The bottom plot shows the window functions resulting from nearly equivalent baselines.
+In each plot, the window function on the right corresponds to the window function for baselines of length $b \sim 180\,\textrm{m}$ in the East West direction, while the window function on the left is for $b\sim 30\,\textrm{m}$.
+The general extent of the window functions for the two-baseline estimator are seen to be roughly the same as for the single-baseline estimator, suggesting that the former is just as good as the latter in keeping the EoR window free of foregrounds.}
+\label{fig:multiblwindows}
+\end{figure}
 
@@
 
-First we can average the visibilities of the equivalent baselines. Since the core of PAPER-128 is a 16 by 7 antenna configuration, there
+First we can average the visibilities of the equivalent baselines. Since the core of PAPER-128 was a 16 by 7 antenna configuration, there
 are $M\equiv(16-|m|)\times(7-|n|)$ copies of the baseline class $\{m,n\}$. This means
@@
 Plotted properties are optimal time offset $\Delta t_{\boldsymbol{bb'}}$ (in hours),  effective weight $\widetilde{\Theta}_{\boldsymbol{bb'}}$, and baseline length $\bar{b}$ (in meters).
-Only those points with $\widetilde{\Theta}_{\boldsymbol{bb'}}>0.01$ are shown (the weight for the top class of equivalent pair is normalized to 1).
+\textbf{Shown on the diagonal are histograms of given quantities for the different arrays.
+On the off-diagonals we show the the scatter plots of the three quantities against each other.} Only those points with $\widetilde{\Theta}_{\boldsymbol{bb'}}>0.01$ are shown (the weight for the top class of equivalent pair is normalized to 1).
 Scatter points of the three arrays overlap, in the order indicated by the legend.
@@
 The properties shown are effective weight $\widetilde{\Theta}_{\boldsymbol{bb'}}$, optimal time offset $\Delta t_{\boldsymbol{bb'}}$, and average baseline length $\bar{b}$.
-We only show three of the mentioned arrays for visual clarity.
+\textbf{We show the histograms on the diagonals and scatter plots on the off-diagonals.} We only show three of the mentioned arrays for visual clarity.
 The other two results are qualitatively similar.
@@
 PAPER-128 is designed with highly redundant nearly equivalent baselines, and thus  these baselines start contributing at higher $\widetilde{\Theta}_{min}$.
-The fractured HERA configurations will benefit even more from nearly equivalent baselines at low $\widetilde{\Theta}_{min}$ due to the presence of more classes of such pairs.
-The stepwise pattern seen in almost all cases are characteristic of a regular grid; as we step to lower $\widetilde{\Theta}_{\boldsymbol{bb'}}$ large groups of baseline pair classes get included in ``batch''.
+\textbf{Incidentally, as we have seen in the scatter plot of $\widetilde{\Theta}_{\boldsymbol{bb'}}$ vs.
+$\bar{b}$ in Fig.
+\ref{fig:pairplot}, these are also the shorter baselines contributing to lower $k$-modes.} The fractured HERA configurations will benefit even more from nearly equivalent baselines at low $\widetilde{\Theta}_{min}$ due to the presence of more classes of such pairs.
+\textbf{However, compared to PAPER, HERA's boost in sensitivity mostly attributes to the longer baselines and thus high $k$-modes (see Fig.
+\ref{fig:pairplot}).
+} The stepwise pattern seen in almost all cases are characteristic of a regular grid; as we step to lower $\widetilde{\Theta}_{\boldsymbol{bb'}}$ large groups of baseline pair classes get included in ``batch''.
 Thus the ``unfractured'' HERA-37, with more regularity in its antenna configurations than the larger counterparts, exhibit more step patterns in fractional sensitivity.

### 2017-10-13 Rotation paper/draft_working.tex
Clarified discussion on window functions

@@
 This suggests that both equivalent- and near-equivalent-baseline delay-spectrum estimators are good estimators of the power spectrum.
-At higher $\ell$, the window functions for both estimators become elongated in $k$, with stronger tails towards the low $k$ modes where foregrounds reside.
-This is the phenomenology of the wedge, where at fine angular scales the foreground leakage to higher $k$ modes are more pronounced.
+\textbf{For window functions that are centered on higher $\ell$ modes, the window functions for both estimators become elongated in $k$, with stronger tails towards the low $k$ modes.
+Since the low $k$ modes are where the foregrounds reside, measurements of the high $\ell$ modes therefore mix in more foreground power.} This is the phenomenology of the wedge, where at fine angular scales the foreground leakage to higher $k$ modes are more pronounced.
 Importantly, however, we note that the window functions for the near-equivalent-baseline estimator are no more elongated than for the equivalent-baseline estimator.
-One may thus conclude that the our proposed near-equivalent-baseline estimator does not result in extra foreground contamination of the EoR window.
+One may thus conclude that our proposed near-equivalent-baseline estimator does not result in extra foreground contamination of the EoR window.
 

### 2017-10-18 Rotation paper/draft_working.tex
addressed all points and added references

@@
 In fact, most sensitivity forecasts to date do include partial redundancy \citep{Pobersens, HERA, JoshAntPos}.
-Recently, \cite{wterm} proposed a visibility-based approach to produce power spectrum measurements using partially redundant baselines in sky-tracking measurements.
-Arrays capable of such tracking include MWA and LOFAR.
-Our work introduces a two-baseline power spectrum estimator that can be applied to both tracking-capable arrays and drift-scan telescopes such as PAPER, HERA, and potentially SKA-low.
+\textbf{A number of methods to produce power spectrum measurements using partially redundant baselines had been studied in recent years.
+A power-spectrum estimator based on visibilities gridded in the $uv$-plane was introduced in \cite{Choudhury2014Estimators, Choudhury2016Taperingb, Choudhury2016Taperinga}, and compared to a simple pair-baseline estimator in \cite{Choudhury2014Estimators}.
+Due to considerations of computational complexity, the authors favored grid-based estimator over pair-baseline estimators despite the higher accuracy of the latter.
+\cite{wterm} proposed another approach based on gridded visibilities in sky-tracking measurements.
+Arrays capable of tracking include MWA and LOFAR, but not drift-scan telescopes such as PAPER, HERA, and potentially SKA-low.
+\cite{CTrott} studied three observing strategies (tracking, zenith drift scan, and drift and shift) with a pair-visibility based approach and concluded that zenith drift scan was the preferred observing strategy.
+In this paper, we introduce a pair-baseline power spectrum estimator that can be applied to both tracking-capable arrays and drift scan only arrays.
+Our formalism also provides a way to identify the pairs to cross-correlate, overcoming the previously considered computational disadvantage as compared to grid-based estimators.
+Furthermore, compared to the pair-baseline estimators studied in \cite{Choudhury2014Estimators} and \cite{CTrott}, our analysis of the sensitivities indicate the importance of the three dimensional nature of the sky, resulting in us directly addressing curved sky effects in our formalism.
+}
 
@@
 \[
-\Delta \tau = \frac{\boldsymbol{b'}}{c}\left(\boldsymbol{\Gamma}\hat{\boldsymbol{s}}-\hat{\boldsymbol{s}}\right). 
+\Delta \tau = \frac{\boldsymbol{b'}}{c}\cdot\left(\boldsymbol{\Gamma}\hat{\boldsymbol{s}}-\hat{\boldsymbol{s}}\right). 
 \]

### 2017-10-19 Rotation paper/draft_working.tex
Minor language edits to referee revisions

@@
 In fact, most sensitivity forecasts to date do include partial redundancy \citep{Pobersens, HERA, JoshAntPos}.
-\textbf{A number of methods to produce power spectrum measurements using partially redundant baselines had been studied in recent years.
+\textbf{A number of methods to produce power spectrum measurements using partially redundant baselines has been studied in recent years.
 A power-spectrum estimator based on visibilities gridded in the $uv$-plane was introduced in \cite{Choudhury2014Estimators, Choudhury2016Taperingb, Choudhury2016Taperinga}, and compared to a simple pair-baseline estimator in \cite{Choudhury2014Estimators}.
@@
 \cite{wterm} proposed another approach based on gridded visibilities in sky-tracking measurements.
-Arrays capable of tracking include MWA and LOFAR, but not drift-scan telescopes such as PAPER, HERA, and potentially SKA-low.
-\cite{CTrott} studied three observing strategies (tracking, zenith drift scan, and drift and shift) with a pair-visibility based approach and concluded that zenith drift scan was the preferred observing strategy.
-In this paper, we introduce a pair-baseline power spectrum estimator that can be applied to both tracking-capable arrays and drift scan only arrays.
-Our formalism also provides a way to identify the pairs to cross-correlate, overcoming the previously considered computational disadvantage as compared to grid-based estimators.
-Furthermore, compared to the pair-baseline estimators studied in \cite{Choudhury2014Estimators} and \cite{CTrott}, our analysis of the sensitivities indicate the importance of the three dimensional nature of the sky, resulting in us directly addressing curved sky effects in our formalism.
+%Arrays capable of tracking include MWA and LOFAR, but not drift-scan telescopes such as PAPER, HERA, and potentially SKA-low.
+\cite{CTrott} studied three observing strategies (tracking, drift scan, and a combination of the two) and developed a pair-visibility based approach to power spectrum estimation.
+%and concluded that zenith drift scan was the preferred observing strategy.
+In this paper, we extend the above works and introduce a non-gridded \emph{baseline}-pair power spectrum estimator that can be applied to both tracking-capable arrays and drift scan-only arrays.
+Our formalism also provides a way to identify which baseline pairs to cross-correlate, overcoming the aforementioned computational disadvantage of baseline-based estimators relative to grid-based estimators.
+Furthermore, compared to
+%the pair-baseline estimators studied in
+\cite{Choudhury2014Estimators} and \cite{CTrott}, our analysis 
+more explicitly treats the curved nature of the sky as well as its spectral dimension.
+%of the sensitivities indicate the importance of the three dimensional nature of the sky, resulting in us directly addressing curved sky effects in our formalism. 
 }
@@
 introduce some terminology that may differ from what is commonly found in the literature. 
-
-We make the distinction between a \textit{baseline}, which corresponds to two specific antennas, and a \textit{class of baselines}, which refers to all baselines of the same length and orientation in a given array.
-Baselines of the same class are traditionally called
-``redundant baselines'', because they measure the same Fourier mode
-in the sky.
-We shall call baselines in the same class \textit{equivalent baselines}, and reserve the \textit{redundancy} of two baselines to mean a variable function of the relative time offsets of their visibility time series.
-With this terminology, equivalent
-baselines are fully redundant with each other simultaneously at all
-times. Non-equivalent baselines also have partial redundancy, and the redundancy can be maximized 
-by shifting their time series by a relative time offset.
-
 \begin{figure}[H]
@@
 \end{figure}
+
+We make the distinction between a \textit{baseline}, which corresponds to two specific antennas, and a \textit{class of baselines}, which refers to all baselines of the same length and orientation in a given array.
+Baselines of the same class are traditionally called
+``redundant baselines'', because they measure the same Fourier mode
+in the sky.
+We shall call baselines in the same class \textit{equivalent baselines}, and reserve the \textit{redundancy} of two baselines to mean a variable function of the relative time offsets of their visibility time series.
+With this terminology, equivalent
+baselines are fully redundant with each other simultaneously at all
+times. Non-equivalent baselines also have partial redundancy, and the redundancy can be maximized 
+by shifting their time series by a relative time offset.
 
@@
 \subsection{Formalism \label{sec:formalism}}
-\label{sec:formalism}
 In this section we formulate a power spectrum estimator from
@@
 PAPER-128 is designed with highly redundant nearly equivalent baselines, and thus  these baselines start contributing at higher $\widetilde{\Theta}_{min}$.
-\textbf{Incidentally, as we have seen in the scatter plot of $\widetilde{\Theta}_{\boldsymbol{bb'}}$ vs.
+\textbf{As seen in the scatter plot of $\widetilde{\Theta}_{\boldsymbol{bb'}}$ vs.
 $\bar{b}$ in Fig.
-\ref{fig:pairplot}, these are also the shorter baselines contributing to lower $k$-modes.} The fractured HERA configurations will benefit even more from nearly equivalent baselines at low $\widetilde{\Theta}_{min}$ due to the presence of more classes of such pairs.
-\textbf{However, compared to PAPER, HERA's boost in sensitivity mostly attributes to the longer baselines and thus high $k$-modes (see Fig.
+\ref{fig:pairplot}, these are also the shorter baselines, which tend to contribute sensitivity to lower $k$-modes.} The fractured HERA configurations will benefit even more from nearly equivalent baselines at low $\widetilde{\Theta}_{min}$ due to the presence of more classes of such pairs.
+\textbf{However, compared to PAPER, HERA's boost in sensitivity is mostly attributable to the longer baselines---and thus high $k$---modes (see Fig.
 \ref{fig:pairplot}).

### 2017-11-28 Rotation paper/revision_1.tex
save

--- 2017-10-19 Rotation paper/draft_working.tex
@@
 %Arrays capable of tracking include MWA and LOFAR, but not drift-scan telescopes such as PAPER, HERA, and potentially SKA-low.
-\cite{CTrott} studied three observing strategies (tracking, drift scan, and a combination of the two) and developed a pair-visibility based approach to power spectrum estimation.
+\cite{CTrott} studied three observing strategies (tracking, drift scan, and a combination of the two) and developed a pair-visibility based approach to coherently combine visibilities gridded on the $uv$-plane for power spectrum estimation.
 %and concluded that zenith drift scan was the preferred observing strategy.
@@
 \ref{fig:pairplot}, these are also the shorter baselines, which tend to contribute sensitivity to lower $k$-modes.} The fractured HERA configurations will benefit even more from nearly equivalent baselines at low $\widetilde{\Theta}_{min}$ due to the presence of more classes of such pairs.
-\textbf{However, compared to PAPER, HERA's boost in sensitivity is mostly attributable to the longer baselines---and thus high $k$---modes (see Fig.
+\textbf{However, compared to PAPER, HERA's boost in sensitivity is mostly attributable to the longer baselines---and thus high $k$-modes (see Fig.
 \ref{fig:pairplot}).

### 2018-02-15 Rotation paper/draft_working.tex
removing boldtexts

@@
 In fact, most sensitivity forecasts to date do include partial redundancy \citep{Pobersens, HERA, JoshAntPos}.
-\textbf{A number of methods to produce power spectrum measurements using partially redundant baselines has been studied in recent years.
+A number of methods to produce power spectrum measurements using partially redundant baselines has been studied in recent years.
 A power-spectrum estimator based on visibilities gridded in the $uv$-plane was introduced in \cite{Choudhury2014Estimators, Choudhury2016Taperingb, Choudhury2016Taperinga}, and compared to a simple pair-baseline estimator in \cite{Choudhury2014Estimators}.
@@
 %of the sensitivities indicate the importance of the three dimensional nature of the sky, resulting in us directly addressing curved sky effects in our formalism.
-}
 
@@
 We shall use the 128-element PAPER array (henceforth referred to as PAPER-128) to motivate our formalism and demonstrate our method, and extend our results to several HERA configurations in Section \ref{sec:arrconf}.
-The PAPER array \textbf{was} located in the Karoo desert in South Africa (30:43:17.5
+The PAPER array was located in the Karoo desert in South Africa (30:43:17.5
 S, 21:25:41.8 E). The layout pattern is shown
@@
 \ref{fig:AntPos}.
-The array \textbf{consisted} of a 112-element core in a rectangular grid, and 16 ``outriggers'' used primarily to aid calibration.
+The array consisted of a 112-element core in a rectangular grid, and 16 ``outriggers'' used primarily to aid calibration.
 In the bottom panel, the two baselines marked in red are an example of an equivalent pair.
@@
 On the other hand, the small North-South separation means that classes such as \{2,0\} and  \{2,1\} are expected to be near-equivalent.
-The PAPER-64 analysis of \cite{Ali2015} \textbf{used} three classes of baselines, the PAPER-128
-equivalent of which \textbf{were} 
+The PAPER-64 analysis of \cite{Ali2015} used three classes of baselines, the PAPER-128
+equivalent of which were
 \{2, 0\}, \{2, 1\} and $\{2, -1\}$.
-There, baselines \textbf{were} cross-multiplied within each class.
+There, baselines were cross-multiplied within each class.
 This paper provides the method for inter-class multiplications.
@@
  && \qquad \times B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}),
+ \label{eq:main0}
  \end{eqnarray}
@@
 \end{eqnarray}
-The third equality of Eq.(\ref{eq:main}) follows from assuming the translational invariance of the statistics of the 21\,cm field.
+The third equality of Eq.
+(\ref{eq:main0}) follows from assuming the translational invariance of the statistics of the 21\,cm field.
 Continuing, we may make the assumption that the 3D power spectrum varies negligibly over the $k$-space of interest, thus allowing us to pull out of the integral the power spectrum centered at
@@
 This suggests that both equivalent- and near-equivalent-baseline delay-spectrum estimators are good estimators of the power spectrum.
-\textbf{For window functions that are centered on higher $\ell$ modes, the window functions for both estimators become elongated in $k$, with stronger tails towards the low $k$ modes.
-Since the low $k$ modes are where the foregrounds reside, measurements of the high $\ell$ modes therefore mix in more foreground power.} This is the phenomenology of the wedge, where at fine angular scales the foreground leakage to higher $k$ modes are more pronounced.
+For window functions that are centered on higher $\ell$ modes, the window functions for both estimators become elongated in $k$, with stronger tails towards the low $k$ modes.
+Since the low $k$ modes are where the foregrounds reside, measurements of the high $\ell$ modes therefore mix in more foreground power.
+This is the phenomenology of the wedge, where at fine angular scales the foreground leakage to higher $k$ modes are more pronounced.
 Importantly, however, we note that the window functions for the near-equivalent-baseline estimator are no more elongated than for the equivalent-baseline estimator.
@@
 Plotted properties are optimal time offset $\Delta t_{\boldsymbol{bb'}}$ (in hours),  effective weight $\widetilde{\Theta}_{\boldsymbol{bb'}}$, and baseline length $\bar{b}$ (in meters).
-\textbf{Shown on the diagonal are histograms of given quantities for the different arrays.
-On the off-diagonals we show the the scatter plots of the three quantities against each other.} Only those points with $\widetilde{\Theta}_{\boldsymbol{bb'}}>0.01$ are shown (the weight for the top class of equivalent pair is normalized to 1).
+On the diagonals we show histograms of given quantities for the different arrays, with the vertical axes linear in scale.
+On the off-diagonals we show the the scatter plots of the three quantities against each other.
+Only those points with $\widetilde{\Theta}_{\boldsymbol{bb'}}>0.01$ are shown (the weight for the top class of equivalent pair is normalized to 1).
 Scatter points of the three arrays overlap, in the order indicated by the legend.
@@
 The properties shown are effective weight $\widetilde{\Theta}_{\boldsymbol{bb'}}$, optimal time offset $\Delta t_{\boldsymbol{bb'}}$, and average baseline length $\bar{b}$.
-\textbf{We show the histograms on the diagonals and scatter plots on the off-diagonals.} We only show three of the mentioned arrays for visual clarity.
+We show the histograms on the diagonals and scatter plots on the off-diagonals.
+We only show three of the mentioned arrays for visual clarity.
 The other two results are qualitatively similar.
@@
 PAPER-128 is designed with highly redundant nearly equivalent baselines, and thus  these baselines start contributing at higher $\widetilde{\Theta}_{min}$.
-\textbf{As seen in the scatter plot of $\widetilde{\Theta}_{\boldsymbol{bb'}}$ vs.
+As seen in the scatter plot of $\widetilde{\Theta}_{\boldsymbol{bb'}}$ vs.
 $\bar{b}$ in Fig.
-\ref{fig:pairplot}, these are also the shorter baselines, which tend to contribute sensitivity to lower $k$-modes.} The fractured HERA configurations will benefit even more from nearly equivalent baselines at low $\widetilde{\Theta}_{min}$ due to the presence of more classes of such pairs.
-\textbf{However, compared to PAPER, HERA's boost in sensitivity is mostly attributable to the longer baselines---and thus high $k$---modes (see Fig.
+\ref{fig:pairplot}, these are also the shorter baselines, which tend to contribute sensitivity to lower $k$-modes.
+The fractured HERA configurations will benefit even more from nearly equivalent baselines at low $\widetilde{\Theta}_{min}$ due to the presence of more classes of such pairs.
+However, compared to PAPER, HERA's boost in sensitivity is mostly attributable to the longer baselines---and thus high $k$-modes (see Fig.
 \ref{fig:pairplot}).
-} The stepwise pattern seen in almost all cases are characteristic of a regular grid; as we step to lower $\widetilde{\Theta}_{\boldsymbol{bb'}}$ large groups of baseline pair classes get included in ``batch''.
+The stepwise pattern seen in almost all cases are characteristic of a regular grid; as we step to lower $\widetilde{\Theta}_{\boldsymbol{bb'}}$ large groups of baseline pair classes get included in ``batch''.
 Thus the ``unfractured'' HERA-37, with more regularity in its antenna configurations than the larger counterparts, exhibit more step patterns in fractional sensitivity.
